### 2019-02-04 histogram.tex
Added a few lines

%% bare_jrnl.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% see http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[journal]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[journal]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{booktabs}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
% Axel Sommerfeldt. This package may be useful when used in conjunction with 
% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
% submissions have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.4.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/pkg/endfloat
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Histogram of Gradient Orientations of Signal Plots applied to P300 Detection}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{Rodrigo~Ramele,~\IEEEmembership{Member,~IEEE,}
        Ana~Julia~Villar,
        and~Juan~Miguel~Santos% <-this % stops a space
\thanks{R. Ramele, A.J.Villar and J.M.Santos are with the Department
of  Computer Engineering, Instituto Tecn{o'}gico de Buenos Aires (ITBA), Ciudad de Buenos Aires,  Argentina e-mail: rramele@itba.edu.ar}% <-this % stops a space
\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
\markboth{IEEE Transactions on Biomedical Engineering}%
{Ramele \MakeLowercase{\textit{et al.}}: Histogram of Gradient Orientations of Signal Plots}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2015 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
The analysis of Electroencephalographic (EEG) signals is of ulterior importance to aid in the diagnosis of mental disease and to increase our understanding of the brain.  Traditionally, clinical EEG has been analyzed in terms of temporal waveforms, looking at rhythms in spontaneous activity, subjectively identifying troughs and peaks in Event-Related Potentials (ERP), or by studying graphoelements in pathological sleep stages.
Additionally, the discipline of Brain Computer Interfaces requires new methods to decode patterns from non-invasive EEG signals.  This field is developing alternative communication pathways to transmit volitional information from the Central Nervous System. The technology could potentially enhance the quality of life of patients affected by neurodegenerative disorders and other mental illness. This work mimics what electroencephalographers have been doing clinically, visually inspecting and categorizing phenomena within the EEG by the extraction of features from images of signal plots.  These features are constructed based on the calculation of histograms of oriented gradients from pixels around the signal plot. It aims to provide a new objective framework to analyze, characterize and classify EEG signal waveforms. The feasibility of the method is outlined by detecting the P300, an ERP elicited by the oddball paradigm of rare events, and implementing an offline P300-based BCI Speller.  The validity of the proposal is shown by offline processing a public dataset of Amyotrophic Lateral Sclerosis (ALS) patients and an own dataset of healthy subjects.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
IEEE, IEEEtran, journal, \LaTeX, paper, template.
\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by the IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.

%The EEG is traditionally analyzed in terms of temporal waveforms at certain channels, looking at power of rhythms in the spontaneous EEG, at amplitude and latency of the peaks and troughs in event- related potentials (ERPs), or at particular grapho-elements in patho- logical or sleep stages.

\IEEEPARstart{A}{lthough} recent advances in neuroimagining techniques, particularly radio-nuclear and radiological scanning methods \cite{Schomer2010}, have diminished the prospects of the traditional Electroencephalography (EEG), the advent and development of digitized devices has impelled for a revamping of this hundred years old technology.  Their versatility, ease of use, temporal resolution, ease of development and production, and its proliferation as consumer devices, are pushing EEG to become the de-facto non invasive portable or ambulatory method to access and harness brain information~\cite{DeVos2014}.

A key contribution to this expansion has been the field of Brain Computer Interfaces (BCI)~\cite{WolpawJonathanR2012} which is the pursuit of the development of a new channel of communication particularly aimed to persons affected by neurodegenerative diseases.

One noteworthy aspect of this novel communication channel is the ability to transmit information from the Central Nervous System (CNS) to a computer device and from there use that information to control a wheelchair~\cite{Carlson2013}, as input to a speller application~\cite{Guger2009a}, in a Virtual Reality environment~\cite{Lotte2013} or as aiding tool in a rehabilitation procedure~\cite{Jure2016}.  The holly grail of BCI is to implement a new complete and alternative pathway to restore lost locomotion~\cite{WolpawJonathanR2012}.

EEG signals are remarkably complex and have been characterized as a multichannel non-stationary stochastic process.  Additionally, they have high variability between different subjects and even between different moments for the same subject, requiring adaptive and co-adaptive calibration and learning procedures~\cite{Clerc}.  Hence, this imposes an outstanding challenge that is necessary to overcome in order to extract information from raw EEG signals.

%Moreover, EEG markers~\cite{Clerc} that can be used to  transmit volitional information are limited, and each one of them has a particular combination of appropriate methods to decode them. Inevitably, it is necessary to implement  distinct and specialized algorithmic methods, to filter the signal, enhance its Signal to Noise Ratio (SNR), and try to determine some meaning out of it.

BCI has gained mainstream public awareness with worldwide challenge competitions like Cybathlon~\cite{Riener2014,cybathlon2} and even been broadcasted during the inauguration ceremony of the 2014 Soccer World Cup.  New developments have overcome the out-of-the-lab high-bar and they are starting to be used in real world environments~\cite{Guger2017,Huggins2016}.  However, they still lack the necessary robustness, and its performance is well behind any other method of human computer interaction, including any kind of detection of residual muscular movement~\cite{Clerc}.

A few works have explored the idea of exploiting the signal waveform to analyze the EEG signal.  In~\cite{Alvarado-Gonzalez2016} an approach based on Slope Horizontal Chain Code is presented, whereas in~\cite{Yamaguchi2009} a similar procedure was implemented based on Mathematical Morphological Analysis.  The seminal work of Bandt-Pompe Permutation Entropy~\cite{Berger2017} also explores succinctly this idea as a basis to establish the time series ordinal patterns.  In the article~\cite{Ramele2016},  the authors introduce a method for classification of rhythmic EEG events like Visual Occipital Alpha Waves  and Motor Imagery Rolandic Central $\mu$ Rhythms using the Histogram of Gradient Orientations of signal plots.  Inspired in that work, we propose a novel application of the developed method to classify and describe transient events, particularly the P300 Event Related Potential.  
The proposed approach is based on the waveform analysis of the shape of the EEG signal.  The signal is drawn on a bidimensional image plot, vector gradients of pixels around the plot are obtained, and with them, the histogram of their orientations is calculated.  This histogram is a direct representation of the waveform of the signal. The method is built by mimicking what regularly electroencephalographers have been performing for almost a century as it is described in~\cite{Hartman2005}: visually inspecting raw signal plots.

This paper reports a method to, (1) describe a procedure to capture the shape of a waveform of an ERP component, the P300, using histograms of gradient orientations extracted from images of signal plots, and (2) outline the way in which this procedure can be used to implement an P300-Based BCI Speller application. Its validity is verified by offline processing two datasets, one of data from ALS patients and another one from data of healthy subjects.

This article unfolds as follows: Section~\ref{Feature} is dedicated to explain the Feature Extraction method based on Histogram of Gradient Orientations of the Signal Plot: Section~\ref{Pipeline} shows the preprocessing pipeline,  Section~\ref{Plot}  describes the image generation of the signal plot, Section~\ref{SIFT}  presents the feature extraction procedure while  Section~\ref{Classification}  introduces the Speller Matrix Letter Identification procedure.  In Section~\ref{Protocol}, the experimental protocol is expounded. Section~\ref{Results} shows the results of applying the proposed technique.  In the final Section~\ref{discussion}  we expose our remarks, conclusions and future work.

The P300~\cite{Farwell1988,Knuth2006} is a positive deflection of the EEG signal which occurs around $300$ ms after the onset of a rare and deviant stimulus that the subject is expected to attend.  It is produced under the oddball paradigm~\cite{WolpawJonathanR2012} and it is consistent across different subjects. It has a lower amplitude  ($\pm 5 \mu V $) compared to basal EEG activity, reaching a Signal to Noise Ratio (SNR) of around $-15$ db estimated based on the amplitude of the P300 response signal divided by the standard deviation of the background EEG activity~\cite{Hu2010}.  This signal can be used to implement a speller application by means of a Speller Matrix~\cite{Farwell1988}. This matrix is composed of 6 rows and 6 columns of numbers and letters. The subject can focus on one character of the matrix. Figure~\ref{fig:p300matrix} shows an example of the Speller Matrix used in the OpenVibe open source software~\cite{Renard2010}, where the flashes of rows and columns provide the deviant stimulus required to elicit this physiological response.   Each time a row or a column that contains the desired letter flashes, the corresponding synchronized EEG signal should also contain the P300 signature and by detecting it, the selected letter can be identified.

\subsection{Feature Extraction from Signal Plots} \label{Feature}

In this section, the signal preprocessing, the method for generating images from signal plots, the feature extraction procedure and the Speller Matrix identification are described.  Figure~\ref{fig:classification} shows a scheme of the entire process.

\subsubsection{Preprocessing Pipeline} \label{Pipeline}

The data obtained by the capturing device is digitalized and a multichannel EEG signal is constructed.

%A trial, as defined by the BCI2000 platform~\cite{Schalk2004}, is every attempt to select a letter from the speller.

%It is composed of signal segments $S_{i}^l$ corresponding to $k_a$ repetitions of flashes of 6 rows and $k_a =10$ repetitions of flashes of 6 columns of the matrix, yielding 120 repetitions.

The $6$ rows and $6$ columns of the Speller Matrix are intensified providing the visual stimulus.  The number of a row or column is a location. A sequence of twelve randomly permuted locations $l$ conform an intensification sequence. The whole set of twelve intensifications is repeated $k_a$ times.

%The multichannel EEG signal is processed on a channel by channel basis.

\begin{itemize}
\item \textbf{Signal Enhancement}: This stage consists of the enhancement of the SNR of the P300 pattern above the level of basal EEG. The pipeline starts by applying a notch filter to the raw digital signal, a $4$th degree $10$ Hz lowpass Butterworth filter and finally a decimation with a Finite Impulse Response (FIR) filter of order $30$ from the original sampling frequency down to $16$ Hz \cite{Krusienski2006}.
\item \textbf{Artifact Removal}: For every complete sequence of $12$ intensifications of $6$ rows and $6$ columns, a basic artifact elimination procedure is implemented by removing the entire sequence when any signal deviates above/bellow $ \pm 70 \mu V $.
\item \textbf{Segmentation}: For each of the $12$ intensifications of one intensification sequence,  a segment $S_{i}^l$  of a window of $t_{max} $ seconds of the multichannel signal is extracted, starting from the stimulus onset, corresponding to each row/column intensification $l$ and to the intensification sequence $i$. As intensifications are permuted in a random order, the segments are rearranged corresponding to row flickering, labeled 1-6, whereas those corresponding to column flickering are labeled 7-12.  Two of these segments should contain the P300 ERP signature time-locked to the flashing stimulus, one for the row, and one for the column.
\item \textbf{Signal Averaging}: \label{Average}  The P300 ERP is deeply buried under basal EEG so the standard approach to identify it is by point-to-point averaging the time-locked stacked signal segments.  Hence the values which are not related to, and not time-locked to the onset of the stimulus are canceled out~\cite{Liang2008}.

This last step determines the operation of any P300 Speller.  In order to obtain an improved signal in terms of its SNR,  repetitions of the sequence of row/column intensification are necessary.  And, at the same time, as long as more repetitions are needed, the ability to transfer information faster is diminished, so there is a trade-off that must be acutely determined.

The procedure to obtain the point-to-point averaged signal goes as follows:

\begin{enumerate}
\item \label{paso1}Highlight randomly the rows and columns from the matrix.  There is one row and one column that should match the letter selected by the subject.
\item  \label{paso2} Repeat step~\ref{paso1} $k_a$ times, obtaining the $1 \leq l \leq 12$ segments $S_1^l(n,c),\dots,S_{k_a}^l(n,c)$, of the EEG signal where the variables $1 \leq n \leq n_{max}$ and $1 \leq c \leq C$  correspond to sample points and channel, respectively. The parameter $C$ is the number of available EEG channels whereas $n_{max}=F_s \  t_{max}$ is the segment length and $F_s$ is the sampling frequency.  The parameter $k_a$ is the number of repetitions of intensifications and it is an input parameter of the algorithm.
\item \label{paso3} Compute the Ensemble Average by
\begin{equation}
x^l(n,c)= \frac{1}{k_a}\sum_{i=1}^{k_a}S_i^l(n,c) 
\label{averaging}
\end{equation}  
for $1 \leq n \leq n_{max}$ and for the channels $1 \leq c \leq C$.  This provide an averaged signal $x^l(n,c)$ for the twelve locations $ 1 \leq l \leq 12$.
\end{enumerate}
\end{itemize}

\subsubsection{Signal Plotting} \label{Plot}

Averaged signal segments are standardized and scaled for $1 \leq n \leq n_{max}$ and $1 \leq c \leq C$ by

\begin{equation}
\tilde{x}^l(n,c) = \left \lfloor{ \gamma \; \frac{( x^l(n,c) - \bar{x}^l(c)  )}{ \hat{\sigma}^l(c) } }\right \rfloor
\label{eq:standarizedaverages}
\end{equation}

\noindent where $\gamma > 0$ is an input parameter of the algorithm and  it is related to the image scale. In addition, $ x^l(n,c) $ is the point-to-point averaged multichannel EEG signal for the sample point $n$ and for channel $c$. Lastly, $$\bar{x}^l(c) =\frac{1}{n_{max}}\sum_{n=1}^{n_{max}}x^l(n,c)$$ and $$ \hat{\sigma}^l(c) = (\frac{1}{n_{max}-1}\sum_{n=1}^{n_{max}}(x^l(n,c)-\bar{x}^l(c))^2 )^{\frac{1}{2}}$$ are the mean and estimated standard deviation of $x^l(n,c), 1 \leq n \leq n_{max}$, for each channel $c$.

Consequently, a binary image $I^{(l,c)}$ is constructed according to

\begin{equation}
I^{(l,c)}(z_1,z_2) = \left\{ \begin{array}{rl}
255 & \text{if} \,  z_1 = \gamma \  n; \! z_2 = \tilde{x}^l(n,c) + z^l(c) \\
0   & \mbox{otherwise}
\end{array}\right.
\label{eq:images}
\end{equation}

\noindent with $255$ being white and representing the signal's value location and $0$ for black which is the background contrast, conforming a black-and-white plot of the signal.  Pixel arguments $ (z_1,z_2) \in \mathbb{N} \times \mathbb{N}$ iterate over the width (based on the length of the signal segment) and height (based on the peak-to-peak amplitude) of the newly created image with $1 \leq n \leq n_{max}$ and $1 \leq c \leq C$.  The value $z^l(c)$ is the image vertical position where the signal's zero value has to be situated in order to fit the entire signal within the image for each channel c:

\begin{equation}
z^l(c) = \left \lfloor{ \frac{\max_{n} \tilde{x}^l(n,c)  - \min_{n} \tilde{x}^l(n,c) }{2} }\right \rfloor -   \left \lfloor{ \frac{\max_{n} \tilde{x}^l(n,c)  + \min_{n} \tilde{x}^l(n,c)}{ 2} }\right \rfloor
\label{eq:zerolevel}
\end{equation}

\noindent where the minimization and maximization are carried out for $n$ varying between ${1 \leq n\leq n_{max}}$, and $ \lfloor \cdot  \rfloor $ denote the rounding to the smaller nearest integer of the number.

In order to complete the plot $I^{(l,c)}$ from the pixels, the Bresenham \cite{Bresenham1965,Ramele2016} algorithm is used to interpolate straight lines between each pair of  consecutive pixels.

\subsubsection{Feature Extraction: Histogram of Gradient Orientations}
\label{SIFT}

For each generated image $I^{(l,c)}$, a keypoint $\mathbf{p_k}$ is placed on a pixel $(x_{p_k}, y_{p_k})$ over the image plot and a window around the keypoint is considered. A local image patch of size $X_p \times X_p$ pixels is constructed by dividing the window in $16$ blocks of size $3s$ each one,  where $s$ is the scale of the local patch and it is an input parameter of the algorithm. It is arranged in a $4 \times 4$ grid and the pixel $ \mathbf{p_k}$ is the patch center, thus $X_p = 12s $ pixels.

A local representation of the signal shape within the patch can be described by obtaining the gradient orientations on each of the $16$ blocks $B_{i,j}$  with $ 0 \leq i,j \leq 3$ and creating a histogram of gradients.  This technique is based on Lowe's SIFT~\cite{Lowe2004} method, and it is biomimetically inspired in how the visual cortex detects shapes by analyzing orientations~\cite{cogprints561}.   In order to calculate the histogram, the interval $[0,360]$ of possible angles is divided in $8$ bins, each one of $45$ degrees.

 Hence, for each spatial bin $ 0 \leq i,j \leq 3$, corresponding to the indexes of each block $B_{i,j}$,  the orientations are accumulated in a  $3$-dimensional histogram $h$ through the following equation:

\begin{equation}
 h(\theta,i,j) = 3 s \sum_{\mathbf{p} \in I^{(l,c)}} w_\mathrm{ang}(\angle J(\mathbf{p}) - \theta)\, w_{ij}\left(\frac{\mathbf{p} - \mathbf{p_k}}{3 s}\right)\, |J(\mathbf{p})|
\label{eq:histogram}
\end{equation}

\noindent  where $\mathbf{p}$ is a pixel from the image $I^{(l,c)}$,  $\theta$ is the angle bin with $ \theta \in \{0, 45, 90, 135, 180, 225, 270, 315\} $,  $ |J(\mathbf{p})| $ is the norm of the gradient vector in the pixel $\mathbf{p}$ and it is computed using finite differences and $\angle J(\mathbf{p}) $ is the angle of the gradient vector.  The scalar $ w_\mathrm{ang}(\cdot) $  and vector $ w_{ij}(\cdot) $ functions are linear interpolations used by~\cite{Lowe2004} and \cite{Vedaldi2010} to provide a weighting contribution to eight adjacent bins.  They are calculated as

\begin{equation}
 w_{ij}(\mathbf{v}) = w( v_x - x_i ) w( v_y - y_j ) 
\label{eq:ij}
\end{equation}

\noindent with $ 0 \leq i,j \leq 3$ and

\begin{equation}
 w_\mathrm{ang}(\alpha) = \sum_{r = -1 }^{1} w \bigg( \frac{8\alpha}{2\pi} + 8r \bigg)
\label{eq:wang}
\end{equation}

\noindent where $x_i$ and $y_i$ are the spatial bin centers located in $ x_i,y_j \in \{-\frac{3}{2},-\frac{1}{2},\frac{1}{2},\frac{3}{2}\} $, $\mathbf{v} = ( v_x, v_y ) $ is a vector variable and $\alpha$ a scalar variable.  On the other hand, $r$ is an integer that can vary freely between $ \left[ -1,1 \right]$ which allows the argument $\alpha$ to be unconstrained in terms of its values in radians. The interpolating function $w(\cdot)$ is defined as $ w(z) = \max(0,|z|-1)$.

These binning functions conform a trilinear interpolation that has a combined effect of sharing the contribution of each oriented gradient between their eight adjacent bins in a tridimensional cube in the histogram space, and zero everywhere else.

Lastly, the fixed value of $ 3 $ is a magnification factor which corresponds to the number of pixels per each block when $s = 1$.  As the patch has  $16$ blocks and  $8$ bin angles are considered, for each location $l$ and channel $c$ a feature called \textit{descriptor} $\mathbf{d}^{(l,c)}$  of $128$ dimension is obtained. 
%It can be observed that the histogram is computed by multiplying by $ |J(\mathbf{p})| $, so the method considers both, the magnitude and the orientation of the gradient vector.

Figure~\ref{fig:sampledescriptor} shows an example of a patch and a scheme of the histogram computation. In (A) a plot of the signal and the patch centered around the keypoint is shown. In (B) the possible orientations on each patch are illustrated.  Only the upper-left four blocks are visible.  The first eight orientations of the first block, are labeled from $1$ to $8$ clockwise. The orientations of the second block $ B_{1,2} $ are labeled from $9$ to $16$.  This labeling continues left-to-right, up-down until the eight orientations for all the sixteen blocks are assigned. They form the corresponding descriptor $\mathbf{d}$ of $128$ coordinates. Finally, in (C) an enlarged image plot is shown where the oriented gradient vector for each pixel can be seen.

\subsubsection{Speller Matrix letter Identification}
\label{Classification}

\paragraph{P300 ERP Extraction}
Segments corresponding to row flickering are labeled 1-6, whereas those corresponding to column flickering are labeled 7-12.  The extraction process has the following steps:

\begin{itemize}
%\setcounter{enumi}{3}

\item \textbf{Step A:}\label{pasoa} First highlight rows and columns from the matrix in a random permutation order and obtain the Ensemble Average as detailed in steps~\ref{paso1}, \ref{paso2} and \ref{paso3} in Section \ref{Average}.
\item \textbf{Step B:}\label{paso4} Plot the signals $\tilde{x}^l(n,c)$,  $1 \leq n \leq n_{max}$, $1 \leq c \leq C $,  according Section~\ref{Plot} in order to generate the images $I^{(l,c)}$ for rows and columns $1 \leq l \leq 12$.

\item \textbf{Step C:} Obtain the descriptors $ \mathbf{d}^{(l,c)}$ for rows and columns from $I^{(l,c)}$  in accordance to the method described in Section~\ref{SIFT}.

\end{itemize}

\paragraph{Calibration}

A trial, as defined by the BCI2000 platform~\cite{Schalk2004}, is every attempt to select just one letter from the speller.  A set of trials is used for calibration and once the calibration is complete it can be used to identify new letters from new trials.

During the calibration phase, two descriptors $ \mathbf{d}^{(l,c)}$ are extracted for each available channel, corresponding to the locations $l$ of a selection of one previously instructed letter from the set of calibration trials.  These descriptors are the P300 templates, grouped together in a template set called $ T^c $.   The set is constructed using the steps described in Section \ref{Average} and the steps A, B and C of the P300 ERP extraction process.

Additionally, the best performing channel, $bpc$ is identified based on the the channel where the best Character Recognition Rate is obtained.

\paragraph{Letter identification}

In order to identify the selected letter, the template set $T^{bpc}$ is used as a database.  Thus, new descriptors are computed and they are compared against the descriptors belonging to the calibration template set $T^{bpc}$.

\begin{itemize}

\item \textbf{Step D:} Match to the calibration template $T^{bpc}$ by computing

\begin{equation}
\hat{row} = \arg \min_{l \in \{1,\dots,6\}} \sum_{q \in N_T(\mathbf{d}^{(l,bpc)})}^{} {\left\lVert q -  \mathbf{d}^{(l,bpc)} \right\rVert}  ^{2}
\label{eq:multiclassificationrow}
\end{equation}

\noindent and

\begin{equation}
\hat{col} = \arg \min_{l \in \{7,\dots,12\}} \sum_{q \in N_T(\mathbf{d}^{(l,bpc)})}^{} {\left\lVert q -  \mathbf{d}^{(l,bpc)} \right\rVert} ^{2}
\label{eq:multiclassificationcol}
\end{equation}

\noindent where $N_T(\mathbf{d}^{(l,bpc)})$  is defined as $N_T(\mathbf{d}^{(l,bpc)}) = \{\mathbf{d} \in T^{bpc} / $  is the k-nearest neighbor of $ \mathbf{d}^{(l,bpc)} \}$ for the best performing channel.  This set is obtained by sorting all the elements in $T^{bpc}$ based on distances between them and $\mathbf{d}^{(l,bpc)}$, choosing the $k$ with smaller values, with $k$ a parameter of the algorithm.  This procedure is based on the k-NBNN  algorithm~\cite{Boiman2008}.

\end{itemize}
By computing the aforementioned equations, the letter of the matrix can be determined from the intersection of the row $ \hat{row} $ and column $ \hat{col} $. 
Figure~\ref{fig:classification} shows a scheme of this process.

\subsection{Subsection Heading Here}
Subsection text here.

% needed in second column of first page if using \IEEEpubid
%\IEEEpubidadjcol

\subsubsection{Subsubsection Heading Here}
Subsubsection text here.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

% Note that the IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% the IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively. 
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

\section{Conclusion}
The conclusion goes here.

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

\appendices
\section{Proof of the First Zonklar Equation}
Appendix one text goes here.

% you can choose not to have a title for an appendix
% if you want by leaving the argument blank
\section{}
Appendix two text goes here.

% use section* for acknowledgment
\section*{Acknowledgment}

The authors would like to thank...

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{IEEEabrv,histogram}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
%\begin{thebibliography}{1}

%\bibitem{IEEEhowto:kopka}
%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

%\end{thebibliography}

% biography section
% 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

\begin{IEEEbiography}{Michael Shell}
Biography text here.
\end{IEEEbiography}

% if you will not have a photo at all:
\begin{IEEEbiographynophoto}{John Doe}
Biography text here.
\end{IEEEbiographynophoto}

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

\begin{IEEEbiographynophoto}{Jane Doe}
Biography text here.
\end{IEEEbiographynophoto}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}

### 2019-02-04 histogram.tex
Up with the basic version almost finished with our paper adapted to IEEE Transactions

@@
 \ifCLASSINFOpdf
-  % \usepackage[pdftex]{graphicx}
+\usepackage[pdftex]{graphicx}
+\usepackage{subfig}
+
   % declare the path(s) where your graphic files are
@@
 \begin{IEEEkeywords}
-IEEE, IEEEtran, journal, \LaTeX, paper, template.
+electroencephalography, histogram of gradient orientations, brain-computer interfaces, P300, SIFT, amyotrophic lateral sclerosis, naive-bayes near neighbours, waveforms
 \end{IEEEkeywords}
@@
 
+\section{Materials and Methods}
+
 The P300~\cite{Farwell1988,Knuth2006} is a positive deflection of the EEG signal which occurs around $300$ ms after the onset of a rare and deviant stimulus that the subject is expected to attend.
@@
 
+\begin{figure}[h!]
+\centering
+\includegraphics[width=8cm]{openvibep300matrix.png}
+\caption{Example of the $6 \times 6$ Speller Matrix used in the study obtained from the OpenVibe software.  Rows and columns flash in random permutations.}
+\label{fig:p300matrix}
+\end{figure}
+
 \subsection{Feature Extraction from Signal Plots} \label{Feature}
@@
 Figure~\ref{fig:classification} shows a scheme of the entire process.
+
+\begin{figure*}[!t]
+\centering
+\subfloat[Classification]{\includegraphics[width=15cm]{classificationgraph.pdf}}
+\caption{For each column and row, an averaged, standardized and scaled signal $\tilde{x}^l(n,c)$ is obtained from the segments $S_i^l$  corresponding to the $k_a$ intensification sequences with $ 1 \leq i \leq k_a $ and location $l$ varying between $1$ and $12$.
+From the averaged signal, the image $I^{(l,c)}$ of the signal plot is generated and each descriptor is computed.
+By comparing each descriptor against the set of templates, the P300 ERP can be detected, and finally the desired letter from the matrix can be inferred.}
+\label{fig:classification}
+\end{figure*}
 
@@
 The pipeline starts by applying a notch filter to the raw digital signal, a $4$th degree $10$ Hz lowpass Butterworth filter and finally a decimation with a Finite Impulse Response (FIR) filter of order $30$ from the original sampling frequency down to $16$ Hz \cite{Krusienski2006}.
-\item \textbf{Artifact Removal}: For every complete sequence of $12$ intensifications of $6$ rows and $6$ columns, a basic artifact elimination procedure is implemented by removing the entire sequence when any signal deviates above/bellow $ \pm 70 \mu V $.
+\item \textbf{Artifact Removal}: For every complete sequence of $12$ intensifications of $6$ rows and $6$ columns, a basic artifact elimination procedure is implemented by removing the entire sequence when any signal deviates  $ \pm 70 \mu V $.
 \item \textbf{Segmentation}: For each of the $12$ intensifications of one intensification sequence,  a segment $S_{i}^l$  of a window of $t_{max} $ seconds of the multichannel signal is extracted, starting from the stimulus onset, corresponding to each row/column intensification $l$ and to the intensification sequence $i$.
@@
 \item \label{paso1}Highlight randomly the rows and columns from the matrix.  There is one row and one column that should match the letter selected by the subject.
-\item  \label{paso2} Repeat step~\ref{paso1} $k_a$ times, obtaining the $1 \leq l \leq 12$ segments $S_1^l(n,c),\dots,S_{k_a}^l(n,c)$, of the EEG signal where the variables $1 \leq n \leq n_{max}$ and $1 \leq c \leq C$  correspond to sample points and channel, respectively.
+\item  \label{paso2} Repeat step~\ref{paso1} $k_a$ times, obtaining the segments $S_1^l(n,c),\dots,S_{k_a}^l(n,c)$, with $1 \leq l \leq 12$, of the EEG signal where the variables $1 \leq n \leq n_{max}$ and $1 \leq c \leq C$  correspond to sample points and channel, respectively.
 The parameter $C$ is the number of available EEG channels whereas $n_{max}=F_s \  t_{max}$ is the segment length and $F_s$ is the sampling frequency.
@@
 I^{(l,c)}(z_1,z_2) = \left\{ \begin{array}{rl}
-255 & \text{if} \,  z_1 = \gamma \  n; \! z_2 = \tilde{x}^l(n,c) + z^l(c) \\
+255 & \text{if} \,  z_1 = \gamma \   n \quad  \text{and} \quad  z_2 = \tilde{x}^l(n,c) + z^l(c) \\
 0   & \mbox{otherwise}
@@
 
-\begin{equation}
-z^l(c) = \left \lfloor{ \frac{\max_{n} \tilde{x}^l(n,c)  - \min_{n} \tilde{x}^l(n,c) }{2} }\right \rfloor -   \left \lfloor{ \frac{\max_{n} \tilde{x}^l(n,c)  + \min_{n} \tilde{x}^l(n,c)}{ 2} }\right \rfloor
+%\begin{equation}
+%z^l(c) = \left \lfloor{ \frac{\max_{n} \tilde{x}^l(n,c)  - \min_{n} \tilde{x}^l(n,c) }{2} }\right \rfloor -   //
+%\left \lfloor{ \frac{\max_{n} \tilde{x}^l(n,c)  + \min_{n} \tilde{x}^l(n,c)}{ 2} }\right \rfloor
+%\label{eq:zerolevel}
+%\end{equation}
+
+\begin{align*}
+z^l(c) = \left \lfloor{ \min_{n} \tilde{x}^l(n,c)  }\right \rfloor
 \label{eq:zerolevel}
-\end{equation}
+\end{align*}
 
 \noindent where the minimization and maximization are carried out for $n$ varying between ${1 \leq n\leq n_{max}}$, and $ \lfloor \cdot  \rfloor $ denote the rounding to the smaller nearest integer of the number.
+This value represents the length of the positive peak of the signal because the signal is inverted before the plot is constructed.
 
@@
 
-\noindent  where $\mathbf{p}$ is a pixel from the image $I^{(l,c)}$,  $\theta$ is the angle bin with $ \theta \in \{0, 45, 90, 135, 180, 225, 270, 315\} $,  $ |J(\mathbf{p})| $ is the norm of the gradient vector in the pixel $\mathbf{p}$ and it is computed using finite differences and $\angle J(\mathbf{p}) $ is the angle of the gradient vector.
+\noindent  where $\mathbf{p}$ is a pixel from the image $I^{(l,c)}$,  $\theta$ is the angle bin with $ \theta $  belonging to $ \{0, 45, 90, 135, 180, 225, 270, 315\} $,  $ |J(\mathbf{p})| $ is the norm of the gradient vector in the pixel $\mathbf{p}$ and it is computed using finite differences and $\angle J(\mathbf{p}) $ is the angle of the gradient vector.
 The scalar $ w_\mathrm{ang}(\cdot) $  and vector $ w_{ij}(\cdot) $ functions are linear interpolations used by~\cite{Lowe2004} and \cite{Vedaldi2010} to provide a weighting contribution to eight adjacent bins.
@@
 
-\subsubsection{Speller Matrix letter Identification}
-\label{Classification}
-
-\paragraph{P300 ERP Extraction}
-Segments corresponding to row flickering are labeled 1-6, whereas those corresponding to column flickering are labeled 7-12.  The extraction process has the following steps:
-
-\begin{itemize}
-%\setcounter{enumi}{3}
-
-\item \textbf{Step A:}\label{pasoa} First highlight rows and columns from the matrix in a random permutation order and obtain the Ensemble Average as detailed in steps~\ref{paso1}, \ref{paso2} and \ref{paso3} in Section \ref{Average}.
-\item \textbf{Step B:}\label{paso4} Plot the signals $\tilde{x}^l(n,c)$,  $1 \leq n \leq n_{max}$, $1 \leq c \leq C $,  according Section~\ref{Plot} in order to generate the images $I^{(l,c)}$ for rows and columns $1 \leq l \leq 12$.
-
-\item \textbf{Step C:} Obtain the descriptors $ \mathbf{d}^{(l,c)}$ for rows and columns from $I^{(l,c)}$  in accordance to the method described in Section~\ref{SIFT}.
-
-\end{itemize}
-
-\paragraph{Calibration}
-
-A trial, as defined by the BCI2000 platform~\cite{Schalk2004}, is every attempt to select just one letter from the speller.
-A set of trials is used for calibration and once the calibration is complete it can be used to identify new letters from new trials.
-
-During the calibration phase, two descriptors $ \mathbf{d}^{(l,c)}$ are extracted for each available channel, corresponding to the locations $l$ of a selection of one previously instructed letter from the set of calibration trials.
-These descriptors are the P300 templates, grouped together in a template set called $ T^c $.
-The set is constructed using the steps described in Section \ref{Average} and the steps A, B and C of the P300 ERP extraction process.
-
-Additionally, the best performing channel, $bpc$ is identified based on the the channel where the best Character Recognition Rate is obtained.
-
-\paragraph{Letter identification}
-
-In order to identify the selected letter, the template set $T^{bpc}$ is used as a database.
-Thus, new descriptors are computed and they are compared against the descriptors belonging to the calibration template set $T^{bpc}$.
-
-\begin{itemize}
-
-\item \textbf{Step D:} Match to the calibration template $T^{bpc}$ by computing
-
-\begin{equation}
-\hat{row} = \arg \min_{l \in \{1,\dots,6\}} \sum_{q \in N_T(\mathbf{d}^{(l,bpc)})}^{} {\left\lVert q -  \mathbf{d}^{(l,bpc)} \right\rVert}  ^{2}
-\label{eq:multiclassificationrow}
-\end{equation}
-
-\noindent and
-
-\begin{equation}
-\hat{col} = \arg \min_{l \in \{7,\dots,12\}} \sum_{q \in N_T(\mathbf{d}^{(l,bpc)})}^{} {\left\lVert q -  \mathbf{d}^{(l,bpc)} \right\rVert} ^{2}
-\label{eq:multiclassificationcol}
-\end{equation}
-
-\noindent where $N_T(\mathbf{d}^{(l,bpc)})$  is defined as $N_T(\mathbf{d}^{(l,bpc)}) = \{\mathbf{d} \in T^{bpc} / $  is the k-nearest neighbor of $ \mathbf{d}^{(l,bpc)} \}$ for the best performing channel.
-This set is obtained by sorting all the elements in $T^{bpc}$ based on distances between them and $\mathbf{d}^{(l,bpc)}$, choosing the $k$ with smaller values, with $k$ a parameter of the algorithm.
-This procedure is based on the k-NBNN  algorithm~\cite{Boiman2008}.
-
-\end{itemize}
-By computing the aforementioned equations, the letter of the matrix can be determined from the intersection of the row $ \hat{row} $ and column $ \hat{col} $. 
-Figure~\ref{fig:classification} shows a scheme of this process.
-
-\subsection{Subsection Heading Here}
-Subsection text here.
-
-% needed in second column of first page if using \IEEEpubid
-%\IEEEpubidadjcol
-
-\subsubsection{Subsubsection Heading Here}
-Subsubsection text here.
-
-% An example of a floating figure using the graphicx package.
-% Note that \label must occur AFTER (or within) \caption.
-% For figures, \caption should occur after the \includegraphics.
-% Note that IEEEtran v1.7 and later has special internal code that
-% is designed to preserve the operation of \label within \caption
-% even when the captionsoff option is in effect. However, because
-% of issues like this, it may be the safest practice to put all your
-% \label just after \caption rather than within \caption{}.
-%
-% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
-% option should be used if it is desired that the figures are to be
-% displayed while in draft mode.
-%
-%\begin{figure}[!t]
+%\begin{figure}[h!]
 %\centering
-%\includegraphics[width=2.5in]{myfigure}
-% where an .eps filename suffix will be assumed under latex, 
-% and a .pdf suffix will be assumed for pdflatex; or what has been declared
-% via \DeclareGraphicsExtensions.
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
+%\includegraphics[width=12cm]{gradients.png}\label{samplegradients}
+%\caption{ (A) Example of a plot of the signal, a keypoint and the corresponding patch.
+(B) A scheme of the orientation's histogram computation.
+Only the upper-left four blocks are visible.
+The first eight orientations of the first block, are labeled from $1$ to $8$ clockwise.
+The orientation of the second block $ B_{1,2} $ is labeled from $9$ to $16$.
+This labeling continues left-to-right, up-down until the eight orientations for all the sixteen blocks are assigned.
+They form the corresponding descriptor of $128$ coordinates.
+The length of each arrow represent the value of the histogram on each direction for each block.
+(C) Vector field of oriented gradients.
+Each pixel is assigned an orientation and magnitude calculated  using finite differences.
+}
+%\label{fig:sampledescriptor}
 %\end{figure}
 
-% Note that the IEEE typically puts floats only at the top, even when this
-% results in a large percentage of a column being occupied by floats.
-
-% An example of a double column floating figure using two subfigures.
-% (The subfig.sty package must be loaded for this to work.)
-% The subfigure \label commands are set within each subfloat command,
-% and the \label for the overall figure must come after \caption.
-% \hfil is used as a separator to get equal spacing.
-% Watch out that the combined width of all the subfigures on a 
-% line do not exceed the text width or a line break will occur.
-%
+\begin{figure*}[!t]
+\centering
+\subfloat[Case I]{\includegraphics[width=15cm]{gradients.png}}
+\caption{ (A) Example of a plot of the signal, a keypoint and the corresponding patch.
+(B) A scheme of the orientation's histogram computation.
+Only the upper-left four blocks are visible.
+The first eight orientations of the first block, are labeled from $1$ to $8$ clockwise.
+The orientation of the second block $ B_{1,2} $ is labeled from $9$ to $16$.
+This labeling continues left-to-right, up-down until the eight orientations for all the sixteen blocks are assigned.
+They form the corresponding descriptor of $128$ coordinates.
+The length of each arrow represent the value of the histogram on each direction for each block.
+(C) Vector field of oriented gradients.
+Each pixel is assigned an orientation and magnitude calculated  using finite differences.
+}
+\label{fig:sampledescriptor}
+\end{figure*}
+
 %\begin{figure*}[!t]
@@
 %\end{figure*}
-%
-% Note that often IEEE papers with subfigures do not employ subfigure
-% captions (using the optional argument to \subfloat[]), but instead will
-% reference/describe all of them (a), (b), etc., within the main caption.
-% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
-% labels, the optional argument to \subfloat must be present. If a
-% subcaption is not desired, just leave its contents blank,
-% e.g., \subfloat[].
-
-% An example of a floating table. Note that, for IEEE style tables, the
-% \caption command should come BEFORE the table and, given that table
-% captions serve much like titles, are usually capitalized except for words
-% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
-% and up, which are usually not capitalized unless they are the first or
-% last word of the caption. Table text will default to \footnotesize as
-% the IEEE normally uses this smaller font for tables.
-% The \label must come after \caption as always.
-%
-%\begin{table}[!t]
-%% increase table row spacing, adjust to taste
-%\renewcommand{\arraystretch}{1.3}
-% if using array.sty, it might be a good idea to tweak the value of
-% \extrarowheight as needed to properly center the text within the cells
-%\caption{An Example of a Table}
-%\label{table_example}
-%\centering
-%% Some packages, such as MDW tools, offer better commands for making tables
-%% than the plain LaTeX2e tabular which is used here.
-%\begin{tabular}{|c||c|}
-%\hline
-%One & Two\\
-%\hline
-%Three & Four\\
-%\hline
-%\end{tabular}
-%\end{table}
-
-% Note that the IEEE does not put floats in the very first column
-% - or typically anywhere on the first page for that matter. Also,
-% in-text middle ("here") positioning is typically not used, but it
-% is allowed and encouraged for Computer Society conferences (but
-% not Computer Society journals). Most IEEE journals/conferences use
-% top floats exclusively. 
-% Note that, LaTeX2e, unlike IEEE journals/conferences, places
-% footnotes above bottom floats. This can be corrected via the
-% \fnbelowfloat command of the stfloats package.
-
-\section{Conclusion}
-The conclusion goes here.
+
+\subsubsection{Speller Matrix letter Identification}
+\label{Classification}
+
+\paragraph{P300 ERP Extraction}
+Segments corresponding to row flickering are labeled 1-6, whereas those corresponding to column flickering are labeled 7-12.  The extraction process has the following steps:
+
+\begin{itemize}
+%\setcounter{enumi}{3}
+
+\item \textbf{Step A:}\label{pasoa} First highlight rows and columns from the matrix in a random permutation order and obtain the Ensemble Average as detailed in steps~\ref{paso1}, \ref{paso2} and \ref{paso3} in Section \ref{Average}.
+\item \textbf{Step B:}\label{paso4} Plot the signals $\tilde{x}^l(n,c)$,  $1 \leq n \leq n_{max}$, $1 \leq c \leq C $,  according Section~\ref{Plot} in order to generate the images $I^{(l,c)}$ for rows and columns $1 \leq l \leq 12$.
+
+\item \textbf{Step C:} Obtain the descriptors $ \mathbf{d}^{(l,c)}$ for rows and columns from $I^{(l,c)}$  in accordance to the method described in Section~\ref{SIFT}.
+
+\end{itemize}
+
+\paragraph{Calibration}
+
+A trial, as defined by the BCI2000 platform~\cite{Schalk2004}, is every attempt to select just one letter from the speller.
+A set of trials is used for calibration and once the calibration is complete it can be used to identify new letters from new trials.
+
+During the calibration phase, two descriptors $ \mathbf{d}^{(l,c)}$ are extracted for each available channel, corresponding to the locations $l$ of a selection of one previously instructed letter from the set of calibration trials.
+These descriptors are the P300 templates, grouped together in a template set called $ T^c $.
+The set is constructed using the steps described in Section \ref{Average} and the steps A, B and C of the P300 ERP extraction process.
+
+Additionally, the best performing channel, $bpc$ is identified based on the the channel where the best Character Recognition Rate is obtained.
+
+\paragraph{Letter identification}
+
+In order to identify the selected letter, the template set $T^{bpc}$ is used as a database.
+Thus, new descriptors are computed and they are compared against the descriptors belonging to the calibration template set $T^{bpc}$.
+
+\begin{itemize}
+
+\item \textbf{Step D:} Match to the calibration template $T^{bpc}$ by computing
+
+\begin{equation}
+\hat{row} = \arg \min_{l \in \{1,\dots,6\}} \sum_{q \in N_T(\mathbf{d}^{(l,bpc)})}^{} {\left\lVert q -  \mathbf{d}^{(l,bpc)} \right\rVert}  ^{2}
+\label{eq:multiclassificationrow}
+\end{equation}
+
+\noindent and
+
+\begin{equation}
+\hat{col} = \arg \min_{l \in \{7,\dots,12\}} \sum_{q \in N_T(\mathbf{d}^{(l,bpc)})}^{} {\left\lVert q -  \mathbf{d}^{(l,bpc)} \right\rVert} ^{2}
+\label{eq:multiclassificationcol}
+\end{equation}
+
+\noindent where $N_T(\mathbf{d}^{(l,bpc)})$  is defined as $N_T(\mathbf{d}^{(l,bpc)}) = \{\mathbf{d} \in T^{bpc} / $  is the k-nearest neighbor of $ \mathbf{d}^{(l,bpc)} \}$ for the best performing channel.
+This set is obtained by sorting all the elements in $T^{bpc}$ based on distances between them and $\mathbf{d}^{(l,bpc)}$, choosing the $k$ with smaller values, with $k$ a parameter of the algorithm.
+This procedure is based on the k-NBNN  algorithm~\cite{Boiman2008}.
+
+\end{itemize}
+By computing the aforementioned equations, the letter of the matrix can be determined from the intersection of the row $ \hat{row} $ and column $ \hat{col} $. 
+Figure~\ref{fig:classification} shows a scheme of this process.
+
+\section{Experimental Protocol} \label{Protocol}
+
+To verify the validity of the proposed framework and method, the public dataset 008-2014~\cite{Riccio2013} published on the BNCI-Horizon website~\cite{Brunner2014} by  IRCCS Fondazione Santa Lucia, is used.
+Additionally, an own dataset with the same experimental conditions is generated.
+Both of them are utilized to perform an offline BCI Simulation to decode the spelled words from the provided signals.
+
+The algorithm is implemented using  VLFeat~\cite{Vedaldi2010} Computer Vision libraries on MATLAB V2014a (Mathworks Inc., Natick, MA, USA).
+Furthermore, for a sake of reproducibility, the code of the algorithm has been made available at: https://bitbucket.org/itba/hist.
+
+In the following sections the characteristics of the datasets and parameters of the identification algorithm are described.
+
+\subsection{P300 ALS Public Dataset} \label{ALSDataset}
+
+The experimental protocol used to generate this dataset is explained in~\cite{Riccio2013} but can be summarized as follows:  8 subjects with confirmed diagnoses but on different stages of ALS disease, were recruited and accepted to perform the experiments.
+The Visual P300 detection task designed for this experiment consisted of spelling 7 words of 5 letters each, using the traditional P300 Speller Matrix~\cite{Farwell1988}.
+The flashing of rows and columns provide the deviant stimulus required to elicit this physiological response.
+The first 3 words are used for calibration and the remaining 4 words, for testing with visual feedback.
+A trial is every attempt to select a letter from the speller.
+It is composed of signal segments corresponding to $k_a =10$ repetitions of flashes of 6 rows and $k_a =10$ repetitions of flashes of 6 columns of the matrix, yielding 120 repetitions.
+Flashing of a row or a column is performed for 0.125 s, following by a resting period (i.e.
+inter-stimulus interval) of the same length.
+After 120 repetitions an inter-trial pause is included before resuming with the following letter.
+
+The recorded dataset was sampled at 256 Hz and it consisted of a scalp multichannel EEG signal for electrode channels Fz, Cz, Pz, Oz, P3, P4, PO7 and PO8, identified according to the 10-20 International System,  for each one of the 8 subjects.
+The recording device was a research-oriented digital EEG device (g.Mobilab, g.Tec, Austria) and the data acquisition and stimuli delivery were handled by the BCI2000 open source software~\cite{Schalk2004}.
+
+In order to assess and verify the identification of the P300 response, subjects are instructed to perform a copy-spelling task.
+They have to fix their attention to successive letters for copying a previously determined set of words, in contrast to a free-running operation of the speller where each user decides on its own what letter to choose.
+
+\subsection{P300 for healthy subjects}
+
+We replicate the same experiment on healthy subjects using a wireless digital EEG device (g.Nautilus, g.Tec, Austria).
+The experimental conditions are the same as those used for the previous dataset, as detailed in section~\ref{ALSDataset}.
+The produced dataset is available in a public online repository~\cite{owndataset}.
+
+Participants are recruited voluntarily and the experiment is conducted anonymously in accordance with the Declaration of Helsinki published by the World Health Organization.
+No monetary compensation is handed out and all participants agree and sign a written informed consent.
+This study is approved by the \textit{Departamento de Investigaci{\'o}n y Doctorado, Instituto Tecnol{\'o}gico de Buenos Aires (ITBA)}.
+All healthy subjects have normal or corrected-to-normal vision and no history of neurological disorders.
+The experiment is performed with 8 subjects, 6 males, 2 females, 6 right-handed, 2 left-handed, average age 29.00 years, standard deviation  11.56 years, range 20-56 years.
+
+EEG data is collected in a single recording session.
+Participants are seated in a comfortable chair, with their vision aligned to a computer screen located one meter in front of them.
+The handling and processing of the data and stimuli is conducted by the OpenVibe platform~\cite{Renard2010}.
+
+Gel-based active electrodes (g.LADYbird, g.Tec, Austria) are used on the same positions Fz, Cz, Pz, Oz, P3,P4, PO7 and PO8.
+Reference is set to the right ear lobe and ground is preset as the AFz position.
+Sampling frequency is slightly different, and is set to 250 Hz, which is the closest possible to the one used with the other dataset.
+
+%Fz, Cz, P3, Pz, P4, PO7, PO8 and Oz.
+
+%8 gel-based active electrodes (g.LADYbird) + g.LADYbird (GND) + g.GAMMAearclip (REF) C3, Cz, C4, CPz, P3, Pz, P4, POz, GND: AFz, REF: right ear
+
+\subsection{Parameters}
+
+The patch size is $X_P = 12s \times 12s$ pixels, where $s$ is the scale of the local patch and it is an input parameter of the algorithm.
+The P300 event can have a span of $400$ ms and its amplitude can reach $ 10 \mu V $~\cite{Rao2013}.
+Hence it is necessary to utilize a signal segment of size $t_{max} = 1$ second and a size patch $X_P$ that could capture an entire transient event.
+With this purpose in consideration, the $s$ value election is essential.
+
+%necesitamos definir el valor de s en función de los parámetros de la señal, de modo tal que el parche cubra el evento completo.  
+We propose the Equations~\ref{eq:mapping2} and~\ref{eq:mapping1} to compute the scale value in horizontal and vertical directions, respectively. 
+\begin{equation}
+s_x = \frac{ \gamma \;  \lambda \  F_s}{12}
+\label{eq:mapping2}
+\end{equation}
+
+\begin{equation}
+s_y= \frac{\gamma \; \Delta \mu V}{12} 
+\label{eq:mapping1}
+\end{equation}
+
+\noindent where $ \lambda $ is the length in seconds covered by the patch, $ F_s $ is the sampling frequency of the EEG signal (downsampled to 16 Hz) and  $\Delta  \mu V $ corresponds to the amplitude in microvolts that can be covered by the height of the patch.
+The geometric structure of the patch forces a squared configuration, then we discerned that by using $ s =s_x =s_y = 3 $ and $ \gamma = 4 $,  the local patch and the descriptor can identify events of 9 $ \mu V $ of amplitude, with a span of $ \lambda = 0.56$ seconds.
+This also determines that $ 1 $ pixel represents $ \frac{1}{\gamma}= \frac{1}{4} \mu V $ on the vertical direction and $\frac{1}{F_s \ \gamma}=\frac{1}{64}$ seconds on the horizontal direction.
+The keypoints  $\mathbf{p_k}$  are located at $ (x_{p_k}, y_{p_k} )= ( 0.55 F_s \ \gamma, z^l(c) )= (35,  z^l(c)) $ for the corresponding channel $c$ and location $l$ (see Equation~\ref{eq:zerolevel}).
+In this way the whole transient event is captured.
+Figure~\ref{fig:patchgeometry} shows a patch of a signal plot covering the complete amplitude (vertical direction) and the complete span of the signal event (horizontal direction).
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=10cm]{patchgeometry.pdf}
+\caption{The scale of local patch is selected in order to capture the whole transient event.
+The size of the patch is $X_p \times X_p$ pixels.
+The vertical size consists of $4$ blocks of size $3 s_y$ pixels which is high enough as to contain the signal $\Delta  \mu V $, the peak-to-peak amplitude of the transient event.
+The horizontal size includes $4$ blocks  of $3 s_x$ and covers the entire duration in seconds of the transient signal event, $ \lambda $.
+}
+\label{fig:patchgeometry}
+\end{figure}
+
+Lastly, the number of channels $C$ is equal to $8$ for both datasets, and the number of intensification sequences $k_a$ is fixed to $10$.
+The parameter $k$ used to construct the set $N_T(\mathbf{d}^{(l,c)})$ is assigned to $k=7$, which was found empirically to achieve better results.
+In addition, the norm used on  Equations \ref{eq:multiclassificationrow} and \ref{eq:multiclassificationcol} is the cosine norm, and descriptors are normalized to $ \left[ -1, 1 \right] $.
+
+\section{Results} \label{Results}
+\label{section:results}
+
+Table~\ref{tab:resultsals} shows the results of applying the Histogram of Gradient Orientations (HIST) algorithm to the subjects of the public dataset of ALS patients.
+The percentage of correctly spelled letters is calculated while performing an offline BCI Simulation.
+From the seven words for each subject, the first three are used for calibration, and the remaining four are used for testing.
+The best performing channel  $bpc$ is informed as well.
+The target ratio is $1:36$; hence theoretical chance level is $2.8\%$.
+It can be observed that the best performance of the letter identification method is reached in a dissimilar channel depending on the subject being studied.
+
+\begin{table}[htb]
+\caption{Character recognition rates for the public dataset of ALS patients using the Histogram of Gradient (HIST) calculated from  single-channel plots.
+Performance rates using single-channel signals with the SVM classifier are shown for comparison.
+The best performing channel $bpc$ for each method is visualized}
+\centering
+%% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
+\begin{tabular}{c|cc|cc}
+\toprule
+\textbf{Participant}	&  $bpc$ 	&  HIST &  $bpc$	&  Single Channel SVM \\
+\midrule
+1     &     Cz   &   $35\%$    &  Cz   & $15\%$   \\
+2     &     Fz   &   $85\%$      &  PO8   & $25\%$   \\
+3     &     Cz   &   $25\%$    &  Fz   & $5\%$   \\
+4     &     PO8 &   $55\%$   &  Oz   & $5\%$    \\
+5     &     PO7 &   $40\%$    &  P3   & $25\%$   \\
+6     &     PO7 &   $60\%$  &  PO8   & $20\%$    \\
+7     &     PO8 &   $80\%$   &  Fz   & $30\%$     \\
+8     &     PO7 &   $95\%$     &  PO7   & $85\%$ \\
+
+%\bottomrule
+\end{tabular}
+\label{tab:resultsals}
+\end{table}
+
+Table~\ref{tab:resultsals} and~\ref{tab:resultsown} show for comparison the obtained performance rates using single-channel signals with the Support Vector Machine (SVM)~\cite{Scholkopf2001} classifier.
+This method is configured to use a linear kernel.
+The best performing channel, where the best letter identification rate was achieved, is also depicted.
+
+\begin{table}[htb]
+\caption{Character recognition rates for the own dataset of healthy subjects using the Histogram of Gradient (HIST) calculated from  single-channel plots.
+Performance rates using single-channel signals with the SVM classifier are shown for comparison.
+The best performing channel $bpc$ for each method is visualized.}
+\centering
+%% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
+\begin{tabular}{c|cc|cc}
+\toprule
+\textbf{Participant}	&  $bpc$	&  HIST &  $bpc$	&  Single Channel SVM \\
+\midrule
+1     &     Oz   &   $40\%$  &  Cz   &  $10\%$    \\
+2     &     PO7   &   $30\%$      &  Cz   & $5\%$   \\
+3     &     P4   &   $40\%$    &  P3   & $10\%$    \\
+4     &     P4 &   $45\%$    &  P4   & $35\%$     \\
+5     &     P4 &   $60\%$  &  P3   & $10\%$     \\
+6     &     Pz &   $50\%$ &  P4   & $25\%$     \\
+7     &     PO7 &   $70\%$  &  P3   & $30\%$     \\
+8     &     P4 &   $50\%$    &  PO7   & $10\%$    \\
+
+%\bottomrule
+\end{tabular}
+\label{tab:resultsown}
+\end{table}
+
+%The spelled words are \textit{GATTO}, \textit{MENTE}, \textit{VIOLA} and \textit{REBUS}.
+
+The Information Transfer Rate (ITR), or Bit Transfer Rate (BTR), in the case of reactive BCIs~\cite{WolpawJonathanR2012}  depends on the amount of signal averaging required to transmit a valid and robust selection.
+Figure~\ref{fig:performance} shows the performance curves for varying intensification sequences for the subjects included in the dataset of ALS patients.
+It can be noticed that the percentage of correctly identified letters depends on the number of intensification sequences that are used to obtain the averaged signal.
+Moreover, when the number of intensification sequences tend to 1, which corresponds to single-intensification character recognition, the performance is reduced.
+As mentioned before, the SNR of the P300 obtained from only one segment of the intensification sequence is very low and the shape of its P300 component is not very well defined.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=8cm]{performance.eps}
+\caption{Performance curves for the eight subjects included in the dataset of ALS patients.  Three out of eight subjects achieved the necessary performance to implement a valid P300 speller.}
+\label{fig:performance}
+\end{figure}
+
+In Table~\ref{tab:resultsown} the results obtained for 8 healthy subjects are shown.
+It can be observed that the performance is above chance level.
+It was verified that HIST method has an improved performance at letter identification than SVM that process the signals on a channel by channel strategy (Wilcoxon signed-rank test, $p =  0.004$ for both datasets).
+
+%In Tables~\ref{tab:resultsals} and~\ref{tab:resultsown} results for character recognition rates using single channel signals with the SVM~\cite{Scholkopf2001}  classification algorithm are also shown.
+This algorithm was configured to use a linear kernel.
+The best performing channel where the best letter identification rate was obtained is also depicted.
+
+%The PE algorithm, which is also devised on a time-domain description of the waveform, was implemented according to \cite{Unakafova2013} and its parameters were adjusted as stated by \cite{Zanin2012}, with an \textit{order} of $2$ and a \textit{sliding window} of size $10$.
+
+Tables~\ref{tab:resultsalsswlda} and~\ref{tab:resultsownswlda} are presented in order to compare the performance of the HIST method versus a multichannel version of the Stepwise Linear Discriminant Analysis (SWLDA) and SVM classification algorithms for both datasets.
+The feature was formed by concatenating all the channels~\cite{Krusienski2006}.
+SWLDA is the methodology proposed by the ALS dataset's publisher.
+Since authors \cite{Riccio2013} did not report the Character Recognition Rate obtained for this dataset, we replicate their procedure and include the performance obtained with the SWLDA algorithm at letter identification.
+It was verified for the dataset of ALS patients that it has similar performance  against other methods like SWLDA or SVM, which use a multichannel feature (Quade test with $p=0.55$) whereas for the dataset of healthy subjects significant differences were found (Quade test with $p=0.02$) where only the HIST method achieved a different performance than SVM (with multiple comparisons, significant difference of level $0.05$).
+
+\begin{table}[htb]
+\caption{Character recognition rates and the best performing channel $bpc$ for the public dataset of ALS patients using the Histogram of Gradient (HIST) (repeated here for comparison purposes).
+Performance rates obtained by SWLDA and SVM classification algorithms with a multichannel concatenated feature.}
+\centering
+%% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
+\begin{tabular}{c|cc|c|c}
+\toprule
+%\textbf{Participant}	&  \textbf{BPC}	& \multicolumn{2}{c}{Character Recognition Rates}\\
+%\cline{1-5} \\
+\textbf{Participant}	&  $bpc$	&  HIST & Multichannel SWLDA & Multichannel SVM \\
+                                    &  for HIST        &           &                                       &   \\
+\midrule
+1     &     Cz   &   $35\%$  & $45\%$  & $40\%$\\
+2     &     Fz   &   $85\%$  & $30\%$   & $50\%$   \\
+3     &     Cz   &   $25\%$  & $65\%$ & $55\%$   \\
+4     &     PO8 &   $55\%$ & $40\%$  & $50\%$   \\
+5     &     PO7 &   $40\%$ & $35\%$  & $45\%$   \\
+6     &     PO7 &   $60\%$ &  $35\%$  & $70\%$   \\
+7     &     PO8 &   $80\%$ & $60\%$   & $35\%$   \\
+8     &     PO7 &   $95\%$  & $90\%$   & $95\%$  \\
+
+%\bottomrule
+\end{tabular}
+\label{tab:resultsalsswlda}
+\end{table}
+
+\begin{table}[htb]
+\caption{Character recognition rates and the best performing channel $bpc$ for the own dataset of healthy subjects using the Histogram of Gradient (HIST) (repeated here for comparison purposes).
+Performance rates obtained by SWLDA and SVM classification algorithms with a multichannel concatenated feature.}
+\centering
+%% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
+\begin{tabular}{c|cc|c|c}
+\toprule
+%\textbf{Participant}	&  \textbf{BPC}	& \multicolumn{2}{c}{Character Recognition Rates}\\
+%\cline{1-5} \\
+\textbf{Participant}	&  $bpc$ 	&  HIST & Multichannel SWLDA & Multichannel SVM  \\
+                                    &  for HIST        &           &                                       &   \\
+\midrule
+1     &     Oz   &     $40\%$  &     $65\%$  &     $40\%$ \\
+2     &     PO7   &     $30\%$ &   $15\%$  &     $10\%$ \\
+3     &     P4   &     $40\%$ &     $50\%$  &     $25\%$ \\
+4     &     P4   &     $45\%$ &     $40\%$  &     $20\%$ \\
+5     &     P4   &      $60\%$ &    $30\%$  &     $20\%$ \\
+6     &     Pz   &      $50\%$ &    $35\%$  &     $30\%$ \\
+7     &     PO7   &      $70\%$ &  $25\%$  &     $30\%$ \\
+8     &     P4   &      $50\%$ &    $35\%$  &     $20\%$ \\
+
+%\bottomrule
+\end{tabular}
+\label{tab:resultsownswlda}
+\end{table}
+
+%It was verified for the dataset of ALS patients that it has similar performance  against other methods like SWLDA or SVM, which use a multichannel feature (Quade test with $p=0.55$) whereas for the dataset of healthy subjects significant differences where found (Quade test with $p=0.02$) where the HIST method achieved a better performance than SVM (with multiple comparisons, significant difference of level $0.05$).
+
+%\subsection{Occipital Channels}
+
+The P300 ERP  consists of two overlapping components: the P3a and P3b, the former with frontocentral distribution while the later stronger on centroparietal region~\cite{Polich2007}.
+Hence, the standard practice is to find the stronger response on the central channel Cz~\cite{Riccio2013}.
+However, \cite{Krusienski2006} show that the response may also arise in occipital regions.
+We found that by analyzing only the waveforms, occipital channels PO8 and PO7 show higher performances for some subjects.
+
+%\subsection{Stability of the P300 shape}
+
+As subjects have varying \textit{latencies} and \textit{amplitudes} of their P300 components, they also have a varying stability of the \textit{shape} of the generated ERP \cite{Nam2010}.
+Figure~\ref{fig:p300templates} shows 10 sample P300 templates patches for patients 8 and 3 from the dataset of ALS patients.
+It can be discerned that in coincidence with the performance results, the P300 signature is more clear and consistent for subject 8 (A) while for subject 3 (B) the characteristic pattern is more difficult to perceive.
+
+\begin{figure*}[!t]
+\centering
+\subfloat[Case I]{\includegraphics[width=15cm]{subject.png}\label{subject8}}
+\caption{Ten sample P300 template patches for subjects 8 (A) and 3 (B) of the ALS Dataset.  Downward deflection is positive polarity. }
+\label{fig:p300templates}
+\end{figure*}
+
+Additionally, the stability of the P300 component waveform has been extensively studied in patients with ALS \cite{SellersandEmanuelDonchin2006,TomohiroMadarame2008,Nijboer2009,Mak2012,McCane2015} where it was found that these patients have a stable P300 component, which were also sustained across different sessions.
+In line with these results we do not find evidence of a difference in terms of the performance obtained by analyzing the waveforms (HIST) for the group of patients with ALS and the healthy group of volunteers (Mann-Whitney U Test, $p=0.46$).
+Particularly, the best performance is obtained for a subject from the ALS dataset for which, based on visual observation, the shape of they P300 component is consistently identified.
+
+%\subsection{Descriptor Space and classification method}
+
+It is important to remark that when applied to binary images obtained from signal plots, the feature extraction method described in Section \ref{SIFT} generates sparse descriptors.
+Under this subspace we found that using the cosine metric yielded a significant performance improvement.
+On the other hand, the unary classification scheme based on the NBNN algorithm proved very beneficial for the P300 Speller Matrix.
+This is due to the fact that this approach solves the unbalance dataset problem which is inherent to the oddball paradigm~\cite{Tibon2015}.
+
+%Using the same feature but with classification methods SVM, feed forward Neural Networks and SWLDA  common in BCI Research achieved a reduced performance.
+
+\section{Discussion}
+\label{discussion}
+
+%In this paper, a new unsupervised method to enhance evoked response by target stimuli in an oddball paradigm was presented.
+Only given the time indexes of rows/columns intensifications, the proposed algorithm estimates the main components of the P300 subspace by providing the best SNR.
+It was shown to efficiently improve the quality of the evoked responses by taking into account the signal and the noise, as opposed to principal component analysis, which only considers the signal.
+Using this method to enhance P300 subspace before the BCI classification task speeds up the BCI since less words are required to train the spatial filters and the linear classifier, given a certain percentage of good symbol prediction.
+Moreover, using this spatial enhancement significantly reduces the dimension of the feature vector used to predict words.
+
+%For both datasets, the experimental protocol uses a very short inter-stimulus interval which has the potential to increase the ITR but at the same time it reduces the amplitude of the P300 response, hence it may be more difficult to detect it~\cite{Rao2013}.
+It is known that ISI alters the P300 amplitude and may affect the chance to detect the ERP.
+
+%In the case of the P300 response, the oddball paradigm requires that one of the stimuli be infrequent.
+Hence this forces the data to be unbalanced~\cite{Tibon2015}.
+At the same time, the NBNN method suffers from biased classification on unbalanced classes~\cite{Fornoni2014}.
+%Para solucionar este problema,
+
+Among other applications of Brain Computer Interfaces, the goal of the discipline is to provide communication assistance to people affected by neuro-degenerative diseases, who are the most likely population to benefit from BCI systems and EEG processing and analysis.
+
+In this work, a method to extract an objective metric from the waveform of the plots of EEG signals is presented.
+Its usage to implement a valid P300-Based BCI Speller application is expounded.
+Additionally, its validity is evaluated using a public dataset of ALS patients and an own dataset of healthy subjects.
+
+%The method works on a channel by channel basis; in this way the best performing channel can be identified and used it to reduce the number of required EEG electrodes, leading to the development of more ergonomic capturing device.
+
+It was verified that this method has an improved performance at letter identification than other methods that process the signals on a channel by channel strategy, and it even has a comparable performance against other methods like SWLDA or SVM, which uses a multichannel feature.
+Furthermore, this method has the advantage that shapes of waveforms can be analyzed in an objective way.
+We observed that the shape of the P300 component is more stable in occipital channels, where the performance for identifying letters is higher.
+We additionally verified that ALS P300 signatures are stable in comparison to those of healthy subjects.
+
+%Further work should be conducted over larger samples to cross-check the validity of these results.
+
+We believe that the use of descriptors based on histogram of gradient orientation, presented in this work, can also be utilized for deriving a shape metric in the space of the P300 signals which can complement other metrics based on time-domain as those defined by~\cite{Mak2012}.
+It is important to notice that the analysis of waveform shapes is usually performed in a qualitative approach based on visual inspection~\cite{SellersandEmanuelDonchin2006}, and a complementary methodology which offer a quantitative metric will be beneficial to these routinely analysis of the waveform of ERPs.
+
+%and, based on this idea, we wanted to complement the methodology with a cuantitative and objective sight
+
+\subsection{Conclusion}
+
+The goal of this work is to answer the question if a P300 component could be solely determined by inspecting automatically their waveforms.
+We conclude affirmatively, though two very important issues still remain:
+
+First, the stability of the P300 in terms of its shape is crucial: the averaging procedure, montages, the signal to noise ratio and spatial filters all of them are non-physiological factors that affect the stability of the shape of the P300 ERP.
+We tested a preliminary approach to assess if the morphological shape of the P300 of the averaged signal can be stabilized by applying different alignments of the stacked segments (see Figure~\ref{fig:classification}) and we verified that there is a better performance when a correct segment alignment is applied.
+We applied Dynamic Time Warping (DTW)~\cite{Casarotto2005} to automate the alignment procedure but we were unable to find a substantial improvement.
+Further work to study the stability of the shape of the P300 signature component needs to be addressed.
+
+The second problem is the amplitude variation of the P300.
+We propose a solution by standardizing the signal, shown in Equation~\ref{eq:standarizedaverages}.
+It has the effect of normalizing the peak-to-peak amplitude, moderating its variation.
+It has also the advantage of reducing noise that was not reduced by the averaging procedure.
+It is important to remark that the averaged signal variance depends on the number of segments used to compute it \cite{van2006signal}.
+The standardizing process converts the signal to unit signal variance which makes it independent of the number $k_a$ of signals averaged.
+Although this is initially an advantageous approach, the standardizing process reduces the amplitude of any significant P300 complex diminishing its automatic interpretation capability.
+
+In our opinion, the best benefit of the presented method is that a closer collaboration of the field of BCI with physicians can be fostered \cite{Chavarriaga2017}, since this procedure intent to imitate human visual observation.
+Automatic classification of patterns in EEG that are specifically identified by their shapes like K-Complex, Vertex Waves, Positive Occipital Sharp Transient~\cite{Hartman2005} are a prospect future work to be considered.
+We are currently working in unpublished material analyzing K-Complex components that could eventually provide  assistance to physicians to locate these EEG patterns, specially in long recording periods, frequent in sleep research~\cite{Michel2012}.
+Additionally, it can be used for artifact removal which is performed on many occasions by visually inspecting signals.
+This is due to the fact that the descriptors are a direct representation of the shape of signal waveforms.
+In line with these applications,  it can be used to build a database~\cite{Chavarriaga2017} of quantitative representations of waveforms and improve atlases~\cite{Hartman2005}, which are currently based on qualitative descriptions of signal shapes.
 
@@
 
-\appendices
-\section{Proof of the First Zonklar Equation}
-Appendix one text goes here.
-
-% you can choose not to have a title for an appendix
-% if you want by leaving the argument blank
-\section{}
-Appendix two text goes here.
-
 % use section* for acknowledgment
-\section*{Acknowledgment}
-
-The authors would like to thank...
+\section{Acknowledgments}
+This project was supported by the ITBACyT-15 funding program issued by ITBA University from Buenos Aires, Argentina.
+
+\section*{Conflict of Interest Statement}
+The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.
 

### 2019-02-05 histogram.tex
Upload a new refined version of the manuscript prepared for this journal

@@
 \thanks{R. Ramele, A.J.Villar and J.M.Santos are with the Department
-of  Computer Engineering, Instituto Tecn{o'}gico de Buenos Aires (ITBA), Ciudad de Buenos Aires,  Argentina e-mail: rramele@itba.edu.ar}% <-this % stops a space
-\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}
+of  Computer Engineering, Instituto Tecn{\'o}gico de Buenos Aires (ITBA), Ciudad de Buenos Aires,  Argentina e-mail: rramele@itba.edu.ar}% <-this % stops a space
+\thanks{Manuscript received ----; revised ---.}}
 
@@
 \begin{abstract}
+250 words objective, methods, results, conclusion and significance
 The analysis of Electroencephalographic (EEG) signals is of ulterior importance to aid in the diagnosis of mental disease and to increase our understanding of the brain.
@@
 In addition, $ x^l(n,c) $ is the point-to-point averaged multichannel EEG signal for the sample point $n$ and for channel $c$.
-Lastly, $$\bar{x}^l(c) =\frac{1}{n_{max}}\sum_{n=1}^{n_{max}}x^l(n,c)$$ and $$ \hat{\sigma}^l(c) = (\frac{1}{n_{max}-1}\sum_{n=1}^{n_{max}}(x^l(n,c)-\bar{x}^l(c))^2 )^{\frac{1}{2}}$$ are the mean and estimated standard deviation of $x^l(n,c), 1 \leq n \leq n_{max}$, for each channel $c$.
+Lastly, $$\bar{x}^l(c) =\frac{1}{n_{max}}\sum_{n=1}^{n_{max}}x^l(n,c)$$ and $$ \hat{\sigma}^l(c) = \Bigg\{ \frac{1}{n_{max}-1}\sum_{n=1}^{n_{max}} \left[  x^l(n,c)-\bar{x}^l(c) \right] ^2 \Bigg\}  ^{\frac{1}{2}}$$ are the mean and estimated standard deviation of $x^l(n,c), 1 \leq n \leq n_{max}$, for each channel $c$.
 
@@
 I^{(l,c)}(z_1,z_2) = \left\{ \begin{array}{rl}
-255 & \text{if} \,  z_1 = \gamma \   n \quad  \text{and} \quad  z_2 = \tilde{x}^l(n,c) + z^l(c) \\
+255 & \text{if} \,  z_1 = \gamma \   n \;  \text{and} \;  z_2 = \tilde{x}^l(n,c) + z^l(c) \\
 0   & \mbox{otherwise}
@@
 
-\begin{align*}
+\begin{equation}
 z^l(c) = \left \lfloor{ \min_{n} \tilde{x}^l(n,c)  }\right \rfloor
 \label{eq:zerolevel}
-\end{align*}
+\end{equation}
 
@@
 It can be observed that the performance is above chance level.
-It was verified that HIST method has an improved performance at letter identification than SVM that process the signals on a channel by channel strategy (Wilcoxon signed-rank test, $p =  0.004$ for both datasets).
+It was verified that HIST method has an improved performance at letter identification than SVM that process the signals on a channel by channel strategy (Wilcoxon signed-rank test, $p\_value =  0.004$ for both datasets).
 
@@
 Since authors \cite{Riccio2013} did not report the Character Recognition Rate obtained for this dataset, we replicate their procedure and include the performance obtained with the SWLDA algorithm at letter identification.
-It was verified for the dataset of ALS patients that it has similar performance  against other methods like SWLDA or SVM, which use a multichannel feature (Quade test with $p=0.55$) whereas for the dataset of healthy subjects significant differences were found (Quade test with $p=0.02$) where only the HIST method achieved a different performance than SVM (with multiple comparisons, significant difference of level $0.05$).
+It was verified for the dataset of ALS patients that it has similar performance  against other methods like SWLDA or SVM, which use a multichannel feature (Quade test with $p\_value=0.55$) whereas for the dataset of healthy subjects significant differences were found (Quade test with $p\_value=0.02$) where only the HIST method achieved a different performance than SVM (multiple comparisons significant difference of level $0.05$).
 
@@
 \subfloat[Case I]{\includegraphics[width=15cm]{subject.png}\label{subject8}}
-\caption{Ten sample P300 template patches for subjects 8 (A) and 3 (B) of the ALS Dataset.  Downward deflection is positive polarity. }
+\caption{Ten sample P300 template patches for subjects 8 (A) and 3 (B) of the ALS Dataset.
+Downward deflection is positive polarity.
+The P300 signature waveform is more clearly and consistent characterized for subject 8, whereas for subject 3 the characteristic patterns is more difficult to spot.
+}
 \label{fig:p300templates}
@@
 Additionally, the stability of the P300 component waveform has been extensively studied in patients with ALS \cite{SellersandEmanuelDonchin2006,TomohiroMadarame2008,Nijboer2009,Mak2012,McCane2015} where it was found that these patients have a stable P300 component, which were also sustained across different sessions.
-In line with these results we do not find evidence of a difference in terms of the performance obtained by analyzing the waveforms (HIST) for the group of patients with ALS and the healthy group of volunteers (Mann-Whitney U Test, $p=0.46$).
+In line with these results we do not find evidence of a difference in terms of the performance obtained by analyzing the waveforms (HIST) for the group of patients with ALS and the healthy group of volunteers (Mann-Whitney U Test, $p\_value=0.46$).
 Particularly, the best performance is obtained for a subject from the ALS dataset for which, based on visual observation, the shape of they P300 component is consistently identified.
@@
 % your own custom macro containing the \includegraphics command to make things
-% simpler here.)
+% simpler here.)\begin{IEEEbiographynophoto}
 %\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
@@
 
-\begin{IEEEbiography}{Michael Shell}
-Biography text here.
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{ramele}}]{Rodrigo Ramele}
+currently works at the Centro de Inteligencia Computacional (CIC), ITBA University, Buenos Aires, Argentina.
+Rodrigo does research in Brain Computer Interfaces, Artificial Intelligence, Signal Processing and Robotics.
 \end{IEEEbiography}
 
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{villar}}]{Ana Julia Villar}
+was graduated in applied mathematics and got her Ph.D.degree in Robust Statistic at the  Universidad de Buenos Aires, Argentina.
+She is currently a professor at the Instituto Tecn{\'o}gico de Buenos Aires.
+Dr.
+Villar has authored or co-authored several papers in the fields of Brain Computer Interface (BCI), Reinforcement Learning and Robust Statistic.
+Her research interests include statistical machine learning, robust statistics and their applications to BCI.
+\end{IEEEbiography}
+
 % if you will not have a photo at all:
-\begin{IEEEbiographynophoto}{John Doe}
-Biography text here.
-\end{IEEEbiographynophoto}
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{santos}}]{Juan Miguel Santos}
+is the Director of the Centro de Inteligencia Computacional (CiC) of ITBA University in Buenos Aires, Argentina, where he works in three different areas: robotics development, reinforcement learning and brain-computer interfaces.
+In the first two lines he works since 1993 developing and constructing robots, national and international robotics conferences and international peer-review journals.
+Currently developing a new development line based on SLAM to autonomous robots.
+\end{IEEEbiography}
 
@@
 %\newpage
-
-\begin{IEEEbiographynophoto}{Jane Doe}
-Biography text here.
-\end{IEEEbiographynophoto}
 

### 2019-08-26 gsync.tex
Modifico el paper para empezar a documetnar toda la informacion para el algoritmo de sincronizacion de seniales

--- 2019-02-04 histogram.tex
@@
 % Do not put math or special symbols in the title.
-\title{Histogram of Gradient Orientations of Signal Plots applied to P300 Detection}
+\title{A Signal Averaging Procedure based on Gambini's Algorithm}
 %
@@
 \author{Rodrigo~Ramele,~\IEEEmembership{Member,~IEEE,}
-        Ana~Julia~Villar,
+        Juliana~Gambini,
         and~Juan~Miguel~Santos% <-this % stops a space
-\thanks{R. Ramele, A.J.Villar and J.M.Santos are with the Department
+\thanks{R. Ramele, J.Gambini and J.M.Santos are with the Department
 of  Computer Engineering, Instituto Tecn{o'}gico de Buenos Aires (ITBA), Ciudad de Buenos Aires,  Argentina e-mail: rramele@itba.edu.ar}% <-this % stops a space
@@
 \markboth{IEEE Transactions on Biomedical Engineering}%
-{Ramele \MakeLowercase{\textit{et al.}}: Histogram of Gradient Orientations of Signal Plots}
+{Ramele \MakeLowercase{\textit{et al.}}: A Signal Averaging Procedure based on Gambini's Algorithm}
 % The only time the second header will appear is for the odd numbered pages
@@
 \begin{abstract}
-The analysis of Electroencephalographic (EEG) signals is of ulterior importance to aid in the diagnosis of mental disease and to increase our understanding of the brain.
-Traditionally, clinical EEG has been analyzed in terms of temporal waveforms, looking at rhythms in spontaneous activity, subjectively identifying troughs and peaks in Event-Related Potentials (ERP), or by studying graphoelements in pathological sleep stages.
-Additionally, the discipline of Brain Computer Interfaces requires new methods to decode patterns from non-invasive EEG signals.
-This field is developing alternative communication pathways to transmit volitional information from the Central Nervous System.
-The technology could potentially enhance the quality of life of patients affected by neurodegenerative disorders and other mental illness.
-This work mimics what electroencephalographers have been doing clinically, visually inspecting and categorizing phenomena within the EEG by the extraction of features from images of signal plots.
-These features are constructed based on the calculation of histograms of oriented gradients from pixels around the signal plot.
-It aims to provide a new objective framework to analyze, characterize and classify EEG signal waveforms.
-The feasibility of the method is outlined by detecting the P300, an ERP elicited by the oddball paradigm of rare events, and implementing an offline P300-based BCI Speller.
-The validity of the proposal is shown by offline processing a public dataset of Amyotrophic Lateral Sclerosis (ALS) patients and an own dataset of healthy subjects.
+EEG
+BCI
+Signal Averaging
+Solutions
+What lies behind.
+
 \end{abstract}
@@
 
-A few works have explored the idea of exploiting the signal waveform to analyze the EEG signal.
-In~\cite{Alvarado-Gonzalez2016} an approach based on Slope Horizontal Chain Code is presented, whereas in~\cite{Yamaguchi2009} a similar procedure was implemented based on Mathematical Morphological Analysis.
-The seminal work of Bandt-Pompe Permutation Entropy~\cite{Berger2017} also explores succinctly this idea as a basis to establish the time series ordinal patterns.
-In the article~\cite{Ramele2016},  the authors introduce a method for classification of rhythmic EEG events like Visual Occipital Alpha Waves  and Motor Imagery Rolandic Central $\mu$ Rhythms using the Histogram of Gradient Orientations of signal plots.
-Inspired in that work, we propose a novel application of the developed method to classify and describe transient events, particularly the P300 Event Related Potential.
-The proposed approach is based on the waveform analysis of the shape of the EEG signal.
-The signal is drawn on a bidimensional image plot, vector gradients of pixels around the plot are obtained, and with them, the histogram of their orientations is calculated.
-This histogram is a direct representation of the waveform of the signal.
-The method is built by mimicking what regularly electroencephalographers have been performing for almost a century as it is described in~\cite{Hartman2005}: visually inspecting raw signal plots.
+A few works have analyzed the problem of syncrhonizing different signals to average them and to enhance their SNR.
 
@@
 In the final Section~\ref{discussion}  we expose our remarks, conclusions and future work.
-
-The P300~\cite{Farwell1988,Knuth2006} is a positive deflection of the EEG signal which occurs around $300$ ms after the onset of a rare and deviant stimulus that the subject is expected to attend.
-It is produced under the oddball paradigm~\cite{WolpawJonathanR2012} and it is consistent across different subjects.
-It has a lower amplitude  ($\pm 5 \mu V $) compared to basal EEG activity, reaching a Signal to Noise Ratio (SNR) of around $-15$ db estimated based on the amplitude of the P300 response signal divided by the standard deviation of the background EEG activity~\cite{Hu2010}.
-This signal can be used to implement a speller application by means of a Speller Matrix~\cite{Farwell1988}.
-This matrix is composed of 6 rows and 6 columns of numbers and letters.
-The subject can focus on one character of the matrix.
-Figure~\ref{fig:p300matrix} shows an example of the Speller Matrix used in the OpenVibe open source software~\cite{Renard2010}, where the flashes of rows and columns provide the deviant stimulus required to elicit this physiological response.
-Each time a row or a column that contains the desired letter flashes, the corresponding synchronized EEG signal should also contain the P300 signature and by detecting it, the selected letter can be identified.
 

### 2019-08-26 gsync.tex

--- 2019-02-05 histogram.tex
@@
 \author{Rodrigo~Ramele,~\IEEEmembership{Member,~IEEE,}
-        Ana~Julia~Villar,
+        Juliana~Gambini,
         and~Juan~Miguel~Santos% <-this % stops a space
-\thanks{R. Ramele, A.J.Villar and J.M.Santos are with the Department
-of  Computer Engineering, Instituto Tecn{\'o}gico de Buenos Aires (ITBA), Ciudad de Buenos Aires,  Argentina e-mail: rramele@itba.edu.ar}% <-this % stops a space
-\thanks{Manuscript received ----; revised ---.}}
+\thanks{R. Ramele, J.Gambini and J.M.Santos are with the Department
+of  Computer Engineering, Instituto Tecn{o'}gico de Buenos Aires (ITBA), Ciudad de Buenos Aires,  Argentina e-mail: rramele@itba.edu.ar}% <-this % stops a space
+\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}
 
@@
 \begin{abstract}
-250 words objective, methods, results, conclusion and significance
-The analysis of Electroencephalographic (EEG) signals is of ulterior importance to aid in the diagnosis of mental disease and to increase our understanding of the brain.
-Traditionally, clinical EEG has been analyzed in terms of temporal waveforms, looking at rhythms in spontaneous activity, subjectively identifying troughs and peaks in Event-Related Potentials (ERP), or by studying graphoelements in pathological sleep stages.
-Additionally, the discipline of Brain Computer Interfaces requires new methods to decode patterns from non-invasive EEG signals.
-This field is developing alternative communication pathways to transmit volitional information from the Central Nervous System.
-The technology could potentially enhance the quality of life of patients affected by neurodegenerative disorders and other mental illness.
-This work mimics what electroencephalographers have been doing clinically, visually inspecting and categorizing phenomena within the EEG by the extraction of features from images of signal plots.
-These features are constructed based on the calculation of histograms of oriented gradients from pixels around the signal plot.
-It aims to provide a new objective framework to analyze, characterize and classify EEG signal waveforms.
-The feasibility of the method is outlined by detecting the P300, an ERP elicited by the oddball paradigm of rare events, and implementing an offline P300-based BCI Speller.
-The validity of the proposal is shown by offline processing a public dataset of Amyotrophic Lateral Sclerosis (ALS) patients and an own dataset of healthy subjects.
+EEG
+BCI
+Signal Averaging
+Solutions
+What lies behind.
 \end{abstract}

### 2020-02-14 alpibot.tex
Upload a new version of a paper

--- 2019-08-26 gsync.tex
@@
 % Do not put math or special symbols in the title.
-\title{A Signal Averaging Procedure based on Gambini's Algorithm}
+\title{Pulmonary rehabilitation Robot Follower}
 %
@@
 
-\author{Rodrigo~Ramele,~\IEEEmembership{Member,~IEEE,}
-        Juliana~Gambini,
+\author{Luciano Bianchi, Esteban Buzniak, Rodrigo~Ramele,~\IEEEmembership{Member,~IEEE,}
         and~Juan~Miguel~Santos% <-this % stops a space
-\thanks{R. Ramele, J.Gambini and J.M.Santos are with the Department
+\thanks{R. Ramele  and J.M.Santos are with the Department
 of  Computer Engineering, Instituto Tecn{o'}gico de Buenos Aires (ITBA), Ciudad de Buenos Aires,  Argentina e-mail: rramele@itba.edu.ar}% <-this % stops a space
@@
 \markboth{IEEE Transactions on Biomedical Engineering}%
-{Ramele \MakeLowercase{\textit{et al.}}: A Signal Averaging Procedure based on Gambini's Algorithm}
+{Bianchi \MakeLowercase{\textit{et al.}} Pulmonary rehabilitation Robot Follower}
 % The only time the second header will appear is for the odd numbered pages
@@
 \begin{abstract}
-EEG
-BCI
-Signal Averaging
-Solutions
-What lies behind.
 
@@
 \begin{IEEEkeywords}
-IEEE, IEEEtran, journal, \LaTeX, paper, template.
+robotics, tethered, HOT, COPD
 \end{IEEEkeywords}
@@
 
-\IEEEPARstart{A}{lthough} recent advances in neuroimagining techniques, particularly radio-nuclear and radiological scanning methods \cite{Schomer2010}, have diminished the prospects of the traditional Electroencephalography (EEG), the advent and development of digitized devices has impelled for a revamping of this hundred years old technology.
-Their versatility, ease of use, temporal resolution, ease of development and production, and its proliferation as consumer devices, are pushing EEG to become the de-facto non invasive portable or ambulatory method to access and harness brain information~\cite{DeVos2014}.
+\IEEEPARstart{A}{lthough} re
 
-A key contribution to this expansion has been the field of Brain Computer Interfaces (BCI)~\cite{WolpawJonathanR2012} which is the pursuit of the development of a new channel of communication particularly aimed to persons affected by neurodegenerative diseases.
+First describe the problem.  Cite the japanese work and the work similar to what we were doing that is in the book.
 
-One noteworthy aspect of this novel communication channel is the ability to transmit information from the Central Nervous System (CNS) to a computer device and from there use that information to control a wheelchair~\cite{Carlson2013}, as input to a speller application~\cite{Guger2009a}, in a Virtual Reality environment~\cite{Lotte2013} or as aiding tool in a rehabilitation procedure~\cite{Jure2016}.
-The holly grail of BCI is to implement a new complete and alternative pathway to restore lost locomotion~\cite{WolpawJonathanR2012}.
+Describe the robot.  How it works, the hardware software, python based the devices and tools used and other architectural elements. Add a graph of how it works.
 
-EEG signals are remarkably complex and have been characterized as a multichannel non-stationary stochastic process.
-Additionally, they have high variability between different subjects and even between different moments for the same subject, requiring adaptive and co-adaptive calibration and learning procedures~\cite{Clerc}.
-Hence, this imposes an outstanding challenge that is necessary to overcome in order to extract information from raw EEG signals.
+Describe the PVC factoring process that Esteban created.  The materials and the name of the CNC that they are using to manufacture everything.  Add the components.
 
-%Moreover, EEG markers~\cite{Clerc} that can be used to  transmit volitional information are limited, and each one of them has a particular combination of appropriate methods to decode them.
-Inevitably, it is necessary to implement  distinct and specialized algorithmic methods, to filter the signal, enhance its Signal to Noise Ratio (SNR), and try to determine some meaning out of it.
+Describe the active spring using scrapped DVD motors.
 
-BCI has gained mainstream public awareness with worldwide challenge competitions like Cybathlon~\cite{Riener2014,cybathlon2} and even been broadcasted during the inauguration ceremony of the 2014 Soccer World Cup.
-New developments have overcome the out-of-the-lab high-bar and they are starting to be used in real world environments~\cite{Guger2017,Huggins2016}.
-However, they still lack the necessary robustness, and its performance is well behind any other method of human computer interaction, including any kind of detection of residual muscular movement~\cite{Clerc}.
+Control strategies: describe both of them.
 
-A few works have analyzed the problem of syncrhonizing different signals to average them and to enhance their SNR.
+Results: 
+show the results of the control strategy using Webots and also the same strategy with the real robot.
 
-This paper reports a method to, (1) describe a procedure to capture the shape of a waveform of an ERP component, the P300, using histograms of gradient orientations extracted from images of signal plots, and (2) outline the way in which this procedure can be used to implement an P300-Based BCI Speller application.
-Its validity is verified by offline processing two datasets, one of data from ALS patients and another one from data of healthy subjects.
+Discuss:
+Describe how everything worked.   Which one is the better and describe how they worked.
 
-This article unfolds as follows: Section~\ref{Feature} is dedicated to explain the Feature Extraction method based on Histogram of Gradient Orientations of the Signal Plot: Section~\ref{Pipeline} shows the preprocessing pipeline,  Section~\ref{Plot}  describes the image generation of the signal plot, Section~\ref{SIFT}  presents the feature extraction procedure while  Section~\ref{Classification}  introduces the Speller Matrix Letter Identification procedure.
-In Section~\ref{Protocol}, the experimental protocol is expounded.
-Section~\ref{Results} shows the results of applying the proposed technique.
-In the final Section~\ref{discussion}  we expose our remarks, conclusions and future work.
+\section{Materials and Methods}
 
-\subsection{Feature Extraction from Signal Plots} \label{Feature}
+\section{Experimental Protocol} \label{Protocol}
 
-In this section, the signal preprocessing, the method for generating images from signal plots, the feature extraction procedure and the Speller Matrix identification are described.
-Figure~\ref{fig:classification} shows a scheme of the entire process.
+To verify the validity of the proposed framework and method,
 
-\subsubsection{Preprocessing Pipeline} \label{Pipeline}
+The experimental protocol used to
 
-The data obtained by the capturing device is digitalized and a multichannel EEG signal is constructed.
+The recorded dataset was sampled at 256 Hz and it consisted of a scalp multichannel EEG signal for electrode channels Fz, Cz, Pz, Oz, P3, P4, PO7 and PO8, identified according to the 10-20 International System,  for each one of the 8 subjects.
+The recording device was a research-oriented digital EEG device (g.Mobilab, g.Tec, Austria) and the data acquisition and stimuli delivery were handled by the BCI2000 open source software~\cite{Schalk2004}.
 
-%A trial, as defined by the BCI2000 platform~\cite{Schalk2004}, is every attempt to select a letter from the speller.
+In order to assess
 
-%It is composed of signal segments $S_{i}^l$ corresponding to $k_a$ repetitions of flashes of 6 rows and $k_a =10$ repetitions of flashes of 6 columns of the matrix, yielding 120 repetitions.
+\section{Results} \label{Results}
+\label{section:results}
 
-The $6$ rows and $6$ columns of the Speller Matrix are intensified providing the visual stimulus.
-The number of a row or column is a location.
-A sequence of twelve randomly permuted locations $l$ conform an intensification sequence.
-The whole set of twelve intensifications is repeated $k_a$ times.
+\begin{figure}[h!]
+\centering
+\includegraphics[width=8cm]{performance.eps}
+\caption{Performance curves for the eight subjects included in the dataset of ALS patients.  Three out of eight subjects achieved the necessary performance to implement a valid P300 speller.}
+\label{fig:performance}
+\end{figure}
 
-%The multichannel EEG signal is processed on a channel by channel basis.
+\begin{figure*}[!t]
+\centering
+\subfloat[Case I]{\includegraphics[width=15cm]{subject.png}\label{subject8}}
+\caption{Ten sample P300 template patches for subjects 8 (A) and 3 (B) of the ALS Dataset.
+Downward deflection is positive polarity.
+The P300 signature waveform is more clearly and consistent characterized for subject 8, whereas for subject 3 the characteristic patterns is more difficult to spot.
+}
+\label{fig:p300templates}
+\end{figure*}
 
-\begin{itemize}
-\item \textbf{Signal Enhancement}: This stage consists of the enhancement of the SNR of the P300 pattern above the level of basal EEG.
-The pipeline starts by applying a notch filter to the raw digital signal, a $4$th degree $10$ Hz lowpass Butterworth filter and finally a decimation with a Finite Impulse Response (FIR) filter of order $30$ from the original sampling frequency down to $16$ Hz \cite{Krusienski2006}.
-\item \textbf{Artifact Removal}: For every complete sequence of $12$ intensifications of $6$ rows and $6$ columns, a basic artifact elimination procedure is implemented by removing the entire sequence when any signal deviates above/bellow $ \pm 70 \mu V $.
-\item \textbf{Segmentation}: For each of the $12$ intensifications of one intensification sequence,  a segment $S_{i}^l$  of a window of $t_{max} $ seconds of the multichannel signal is extracted, starting from the stimulus onset, corresponding to each row/column intensification $l$ and to the intensification sequence $i$.
-As intensifications are permuted in a random order, the segments are rearranged corresponding to row flickering, labeled 1-6, whereas those corresponding to column flickering are labeled 7-12.
-Two of these segments should contain the P300 ERP signature time-locked to the flashing stimulus, one for the row, and one for the column.
-\item \textbf{Signal Averaging}: \label{Average}  The P300 ERP is deeply buried under basal EEG so the standard approach to identify it is by point-to-point averaging the time-locked stacked signal segments.
-Hence the values which are not related to, and not time-locked to the onset of the stimulus are canceled out~\cite{Liang2008}.
+\section{Discussion}
+\label{discussion}
 
-This last step determines the operation of any P300 Speller.
-In order to obtain an improved signal in terms of its SNR,  repetitions of the sequence of row/column intensification are necessary.
-And, at the same time, as long as more repetitions are needed, the ability to transfer information faster is diminished, so there is a trade-off that must be acutely determined.
+\subsection{Conclusion}
 
-The procedure to obtain the point-to-point averaged signal goes as follows:
-
-\begin{enumerate}
-\item \label{paso1}Highlight randomly the rows and columns from the matrix.  There is one row and one column that should match the letter selected by the subject.
-\item  \label{paso2} Repeat step~\ref{paso1} $k_a$ times, obtaining the $1 \leq l \leq 12$ segments $S_1^l(n,c),\dots,S_{k_a}^l(n,c)$, of the EEG signal where the variables $1 \leq n \leq n_{max}$ and $1 \leq c \leq C$  correspond to sample points and channel, respectively.
-The parameter $C$ is the number of available EEG channels whereas $n_{max}=F_s \  t_{max}$ is the segment length and $F_s$ is the sampling frequency.
-The parameter $k_a$ is the number of repetitions of intensifications and it is an input parameter of the algorithm.
-\item \label{paso3} Compute the Ensemble Average by
-\begin{equation}
-x^l(n,c)= \frac{1}{k_a}\sum_{i=1}^{k_a}S_i^l(n,c) 
-\label{averaging}
-\end{equation}  
-for $1 \leq n \leq n_{max}$ and for the channels $1 \leq c \leq C$.  This provide an averaged signal $x^l(n,c)$ for the twelve locations $ 1 \leq l \leq 12$.
-\end{enumerate}
-\end{itemize}
-
-\subsubsection{Signal Plotting} \label{Plot}
-
-Averaged signal segments are standardized and scaled for $1 \leq n \leq n_{max}$ and $1 \leq c \leq C$ by
-
-\begin{equation}
-\tilde{x}^l(n,c) = \left \lfloor{ \gamma \; \frac{( x^l(n,c) - \bar{x}^l(c)  )}{ \hat{\sigma}^l(c) } }\right \rfloor
-\label{eq:standarizedaverages}
-\end{equation}
-
-\noindent where $\gamma > 0$ is an input parameter of the algorithm and  it is related to the image scale.
-In addition, $ x^l(n,c) $ is the point-to-point averaged multichannel EEG signal for the sample point $n$ and for channel $c$.
-Lastly, $$\bar{x}^l(c) =\frac{1}{n_{max}}\sum_{n=1}^{n_{max}}x^l(n,c)$$ and $$ \hat{\sigma}^l(c) = (\frac{1}{n_{max}-1}\sum_{n=1}^{n_{max}}(x^l(n,c)-\bar{x}^l(c))^2 )^{\frac{1}{2}}$$ are the mean and estimated standard deviation of $x^l(n,c), 1 \leq n \leq n_{max}$, for each channel $c$.
-
-Consequently, a binary image $I^{(l,c)}$ is constructed according to
-
-\begin{equation}
-I^{(l,c)}(z_1,z_2) = \left\{ \begin{array}{rl}
-255 & \text{if} \,  z_1 = \gamma \  n; \! z_2 = \tilde{x}^l(n,c) + z^l(c) \\
-0   & \mbox{otherwise}
-\end{array}\right.
-\label{eq:images}
-\end{equation}
-
-\noindent with $255$ being white and representing the signal's value location and $0$ for black which is the background contrast, conforming a black-and-white plot of the signal.
-Pixel arguments $ (z_1,z_2) \in \mathbb{N} \times \mathbb{N}$ iterate over the width (based on the length of the signal segment) and height (based on the peak-to-peak amplitude) of the newly created image with $1 \leq n \leq n_{max}$ and $1 \leq c \leq C$.
-The value $z^l(c)$ is the image vertical position where the signal's zero value has to be situated in order to fit the entire signal within the image for each channel c:
-
-\begin{equation}
-z^l(c) = \left \lfloor{ \frac{\max_{n} \tilde{x}^l(n,c)  - \min_{n} \tilde{x}^l(n,c) }{2} }\right \rfloor -   \left \lfloor{ \frac{\max_{n} \tilde{x}^l(n,c)  + \min_{n} \tilde{x}^l(n,c)}{ 2} }\right \rfloor
-\label{eq:zerolevel}
-\end{equation}
-
-\noindent where the minimization and maximization are carried out for $n$ varying between ${1 \leq n\leq n_{max}}$, and $ \lfloor \cdot  \rfloor $ denote the rounding to the smaller nearest integer of the number.
-
-In order to complete the plot $I^{(l,c)}$ from the pixels, the Bresenham \cite{Bresenham1965,Ramele2016} algorithm is used to interpolate straight lines between each pair of  consecutive pixels.
-
-\subsubsection{Feature Extraction: Histogram of Gradient Orientations}
-\label{SIFT}
-
-For each generated image $I^{(l,c)}$, a keypoint $\mathbf{p_k}$ is placed on a pixel $(x_{p_k}, y_{p_k})$ over the image plot and a window around the keypoint is considered.
-A local image patch of size $X_p \times X_p$ pixels is constructed by dividing the window in $16$ blocks of size $3s$ each one,  where $s$ is the scale of the local patch and it is an input parameter of the algorithm.
-It is arranged in a $4 \times 4$ grid and the pixel $ \mathbf{p_k}$ is the patch center, thus $X_p = 12s $ pixels.
-
-A local representation of the signal shape within the patch can be described by obtaining the gradient orientations on each of the $16$ blocks $B_{i,j}$  with $ 0 \leq i,j \leq 3$ and creating a histogram of gradients.
-This technique is based on Lowe's SIFT~\cite{Lowe2004} method, and it is biomimetically inspired in how the visual cortex detects shapes by analyzing orientations~\cite{cogprints561}.
-In order to calculate the histogram, the interval $[0,360]$ of possible angles is divided in $8$ bins, each one of $45$ degrees.
-
- Hence, for each spatial bin $ 0 \leq i,j \leq 3$, corresponding to the indexes of each block $B_{i,j}$,  the orientations are accumulated in a  $3$-dimensional histogram $h$ through the following equation:
-
-\begin{equation}
- h(\theta,i,j) = 3 s \sum_{\mathbf{p} \in I^{(l,c)}} w_\mathrm{ang}(\angle J(\mathbf{p}) - \theta)\, w_{ij}\left(\frac{\mathbf{p} - \mathbf{p_k}}{3 s}\right)\, |J(\mathbf{p})|
-\label{eq:histogram}
-\end{equation}
-
-\noindent  where $\mathbf{p}$ is a pixel from the image $I^{(l,c)}$,  $\theta$ is the angle bin with $ \theta \in \{0, 45, 90, 135, 180, 225, 270, 315\} $,  $ |J(\mathbf{p})| $ is the norm of the gradient vector in the pixel $\mathbf{p}$ and it is computed using finite differences and $\angle J(\mathbf{p}) $ is the angle of the gradient vector.
-The scalar $ w_\mathrm{ang}(\cdot) $  and vector $ w_{ij}(\cdot) $ functions are linear interpolations used by~\cite{Lowe2004} and \cite{Vedaldi2010} to provide a weighting contribution to eight adjacent bins.
-They are calculated as
-
-\begin{equation}
- w_{ij}(\mathbf{v}) = w( v_x - x_i ) w( v_y - y_j ) 
-\label{eq:ij}
-\end{equation}
-
-\noindent with $ 0 \leq i,j \leq 3$ and
-
-\begin{equation}
- w_\mathrm{ang}(\alpha) = \sum_{r = -1 }^{1} w \bigg( \frac{8\alpha}{2\pi} + 8r \bigg)
-\label{eq:wang}
-\end{equation}
-
-\noindent where $x_i$ and $y_i$ are the spatial bin centers located in $ x_i,y_j \in \{-\frac{3}{2},-\frac{1}{2},\frac{1}{2},\frac{3}{2}\} $, $\mathbf{v} = ( v_x, v_y ) $ is a vector variable and $\alpha$ a scalar variable.
-On the other hand, $r$ is an integer that can vary freely between $ \left[ -1,1 \right]$ which allows the argument $\alpha$ to be unconstrained in terms of its values in radians.
-The interpolating function $w(\cdot)$ is defined as $ w(z) = \max(0,|z|-1)$.
-
-These binning functions conform a trilinear interpolation that has a combined effect of sharing the contribution of each oriented gradient between their eight adjacent bins in a tridimensional cube in the histogram space, and zero everywhere else.
-
-Lastly, the fixed value of $ 3 $ is a magnification factor which corresponds to the number of pixels per each block when $s = 1$.
-As the patch has  $16$ blocks and  $8$ bin angles are considered, for each location $l$ and channel $c$ a feature called \textit{descriptor} $\mathbf{d}^{(l,c)}$  of $128$ dimension is obtained.
-%It can be observed that the histogram is computed by multiplying by $ |J(\mathbf{p})| $, so the method considers both, the magnitude and the orientation of the gradient vector.
-
-Figure~\ref{fig:sampledescriptor} shows an example of a patch and a scheme of the histogram computation.
-In (A) a plot of the signal and the patch centered around the keypoint is shown.
-In (B) the possible orientations on each patch are illustrated.
-Only the upper-left four blocks are visible.
-The first eight orientations of the first block, are labeled from $1$ to $8$ clockwise.
-The orientations of the second block $ B_{1,2} $ are labeled from $9$ to $16$.
-This labeling continues left-to-right, up-down until the eight orientations for all the sixteen blocks are assigned.
-They form the corresponding descriptor $\mathbf{d}$ of $128$ coordinates.
-Finally, in (C) an enlarged image plot is shown where the oriented gradient vector for each pixel can be seen.
-
-\subsubsection{Speller Matrix letter Identification}
-\label{Classification}
-
-\paragraph{P300 ERP Extraction}
-Segments corresponding to row flickering are labeled 1-6, whereas those corresponding to column flickering are labeled 7-12.  The extraction process has the following steps:
-
-\begin{itemize}
-%\setcounter{enumi}{3}
-
-\item \textbf{Step A:}\label{pasoa} First highlight rows and columns from the matrix in a random permutation order and obtain the Ensemble Average as detailed in steps~\ref{paso1}, \ref{paso2} and \ref{paso3} in Section \ref{Average}.
-\item \textbf{Step B:}\label{paso4} Plot the signals $\tilde{x}^l(n,c)$,  $1 \leq n \leq n_{max}$, $1 \leq c \leq C $,  according Section~\ref{Plot} in order to generate the images $I^{(l,c)}$ for rows and columns $1 \leq l \leq 12$.
-
-\item \textbf{Step C:} Obtain the descriptors $ \mathbf{d}^{(l,c)}$ for rows and columns from $I^{(l,c)}$  in accordance to the method described in Section~\ref{SIFT}.
-
-\end{itemize}
-
-\paragraph{Calibration}
-
-A trial, as defined by the BCI2000 platform~\cite{Schalk2004}, is every attempt to select just one letter from the speller.
-A set of trials is used for calibration and once the calibration is complete it can be used to identify new letters from new trials.
-
-During the calibration phase, two descriptors $ \mathbf{d}^{(l,c)}$ are extracted for each available channel, corresponding to the locations $l$ of a selection of one previously instructed letter from the set of calibration trials.
-These descriptors are the P300 templates, grouped together in a template set called $ T^c $.
-The set is constructed using the steps described in Section \ref{Average} and the steps A, B and C of the P300 ERP extraction process.
-
-Additionally, the best performing channel, $bpc$ is identified based on the the channel where the best Character Recognition Rate is obtained.
-
-\paragraph{Letter identification}
-
-In order to identify the selected letter, the template set $T^{bpc}$ is used as a database.
-Thus, new descriptors are computed and they are compared against the descriptors belonging to the calibration template set $T^{bpc}$.
-
-\begin{itemize}
-
-\item \textbf{Step D:} Match to the calibration template $T^{bpc}$ by computing
-
-\begin{equation}
-\hat{row} = \arg \min_{l \in \{1,\dots,6\}} \sum_{q \in N_T(\mathbf{d}^{(l,bpc)})}^{} {\left\lVert q -  \mathbf{d}^{(l,bpc)} \right\rVert}  ^{2}
-\label{eq:multiclassificationrow}
-\end{equation}
-
-\noindent and
-
-\begin{equation}
-\hat{col} = \arg \min_{l \in \{7,\dots,12\}} \sum_{q \in N_T(\mathbf{d}^{(l,bpc)})}^{} {\left\lVert q -  \mathbf{d}^{(l,bpc)} \right\rVert} ^{2}
-\label{eq:multiclassificationcol}
-\end{equation}
-
-\noindent where $N_T(\mathbf{d}^{(l,bpc)})$  is defined as $N_T(\mathbf{d}^{(l,bpc)}) = \{\mathbf{d} \in T^{bpc} / $  is the k-nearest neighbor of $ \mathbf{d}^{(l,bpc)} \}$ for the best performing channel.
-This set is obtained by sorting all the elements in $T^{bpc}$ based on distances between them and $\mathbf{d}^{(l,bpc)}$, choosing the $k$ with smaller values, with $k$ a parameter of the algorithm.
-This procedure is based on the k-NBNN  algorithm~\cite{Boiman2008}.
-
-\end{itemize}
-By computing the aforementioned equations, the letter of the matrix can be determined from the intersection of the row $ \hat{row} $ and column $ \hat{col} $. 
-Figure~\ref{fig:classification} shows a scheme of this process.
-
-\subsection{Subsection Heading Here}
-Subsection text here.
-
-% needed in second column of first page if using \IEEEpubid
-%\IEEEpubidadjcol
-
-\subsubsection{Subsubsection Heading Here}
-Subsubsection text here.
-
-% An example of a floating figure using the graphicx package.
-% Note that \label must occur AFTER (or within) \caption.
-% For figures, \caption should occur after the \includegraphics.
-% Note that IEEEtran v1.7 and later has special internal code that
-% is designed to preserve the operation of \label within \caption
-% even when the captionsoff option is in effect. However, because
-% of issues like this, it may be the safest practice to put all your
-% \label just after \caption rather than within \caption{}.
-%
-% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
-% option should be used if it is desired that the figures are to be
-% displayed while in draft mode.
-%
-%\begin{figure}[!t]
-%\centering
-%\includegraphics[width=2.5in]{myfigure}
-% where an .eps filename suffix will be assumed under latex, 
-% and a .pdf suffix will be assumed for pdflatex; or what has been declared
-% via \DeclareGraphicsExtensions.
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure}
-
-% Note that the IEEE typically puts floats only at the top, even when this
-% results in a large percentage of a column being occupied by floats.
-
-% An example of a double column floating figure using two subfigures.
-% (The subfig.sty package must be loaded for this to work.)
-% The subfigure \label commands are set within each subfloat command,
-% and the \label for the overall figure must come after \caption.
-% \hfil is used as a separator to get equal spacing.
-% Watch out that the combined width of all the subfigures on a 
-% line do not exceed the text width or a line break will occur.
-%
-%\begin{figure*}[!t]
-%\centering
-%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
-%\label{fig_first_case}}
-%\hfil
-%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
-%\label{fig_second_case}}
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure*}
-%
-% Note that often IEEE papers with subfigures do not employ subfigure
-% captions (using the optional argument to \subfloat[]), but instead will
-% reference/describe all of them (a), (b), etc., within the main caption.
-% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
-% labels, the optional argument to \subfloat must be present. If a
-% subcaption is not desired, just leave its contents blank,
-% e.g., \subfloat[].
-
-% An example of a floating table. Note that, for IEEE style tables, the
-% \caption command should come BEFORE the table and, given that table
-% captions serve much like titles, are usually capitalized except for words
-% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
-% and up, which are usually not capitalized unless they are the first or
-% last word of the caption. Table text will default to \footnotesize as
-% the IEEE normally uses this smaller font for tables.
-% The \label must come after \caption as always.
-%
-%\begin{table}[!t]
-%% increase table row spacing, adjust to taste
-%\renewcommand{\arraystretch}{1.3}
-% if using array.sty, it might be a good idea to tweak the value of
-% \extrarowheight as needed to properly center the text within the cells
-%\caption{An Example of a Table}
-%\label{table_example}
-%\centering
-%% Some packages, such as MDW tools, offer better commands for making tables
-%% than the plain LaTeX2e tabular which is used here.
-%\begin{tabular}{|c||c|}
-%\hline
-%One & Two\\
-%\hline
-%Three & Four\\
-%\hline
-%\end{tabular}
-%\end{table}
-
-% Note that the IEEE does not put floats in the very first column
-% - or typically anywhere on the first page for that matter. Also,
-% in-text middle ("here") positioning is typically not used, but it
-% is allowed and encouraged for Computer Society conferences (but
-% not Computer Society journals). Most IEEE journals/conferences use
-% top floats exclusively. 
-% Note that, LaTeX2e, unlike IEEE journals/conferences, places
-% footnotes above bottom floats. This can be corrected via the
-% \fnbelowfloat command of the stfloats package.
-
-\section{Conclusion}
-The conclusion goes here.
+The goal of this work is to verify if a tethered robot to implement a following scheme on a patient during a pulmonary rehabilitation procedure is too simplistic to be a factual implementation.
 
@@
 
-\appendices
-\section{Proof of the First Zonklar Equation}
-Appendix one text goes here.
+% use section* for acknowledgment
+\section{Acknowledgments}
+This project was supported by ALPI organization.
 
-% you can choose not to have a title for an appendix
-% if you want by leaving the argument blank
-\section{}
-Appendix two text goes here.
-
-% use section* for acknowledgment
-\section*{Acknowledgment}
-
-The authors would like to thank...
+\section*{Conflict of Interest Statement}
+The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.
 
@@
 % argument is your BibTeX string definitions and bibliography database(s)
-\bibliography{IEEEabrv,histogram}
+\bibliography{IEEEabrv,gsync}
 %
@@
 
-\begin{IEEEbiography}{Michael Shell}
-Biography text here.
-\end{IEEEbiography}
-
-% if you will not have a photo at all:
-\begin{IEEEbiographynophoto}{John Doe}
-Biography text here.
-\end{IEEEbiographynophoto}
+%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{ramele}}]{Rodrigo Ramele}
+%currently works at the Centro de Inteligencia Computacional (CIC), ITBA University, Buenos Aires, Argentina.
+Rodrigo does research in Brain Computer Interfaces, Artificial Intelligence, Signal Processing and Robotics.
+%\end{IEEEbiography}
+%
+%% if you will not have a photo at all:
+%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{santos}}]{Juan Miguel Santos}
+%is the Director of the Centro de Inteligencia Computacional (CiC) of ITBA University in Buenos Aires, Argentina, where he works in three different areas: robotics development, reinforcement learning and brain-computer interfaces.
+In the first two lines he works since 1993 developing and constructing robots, national and international robotics conferences and international peer-review journals.
+Currently developing a new development line based on SLAM to autonomous robots.
+%\end{IEEEbiography}
 

### 2020-02-14 alpibot.tex
Upload a new version of this paper based on the ALPI bot solution.

@@
 
-\author{Luciano Bianchi, Esteban Buzniak, Rodrigo~Ramele,~\IEEEmembership{Member,~IEEE,}
-        and~Juan~Miguel~Santos% <-this % stops a space
+\author{Bianchi Luciano, Buzniak Esteban, Ramele~Rodrigo,~\IEEEmembership{Member,~IEEE,}
+        and Santos~Juan~Miguel% <-this % stops a space
 \thanks{R. Ramele  and J.M.Santos are with the Department
@@
 \begin{abstract}
-
+Patients that suffer Chronic Obstructive Pulmonary Disease (COPD) underwent a procedure called Pulmonary Rehabilitation that helps them to improve disease prognosis.
+During Pulmonary Rehabilitation procedures patients require external oxygen assistance.
+The oxygen tank cannot be carried by the patient and some external assistance is required.
+In this work a basic robot follower is proposed to carry the oxygen tank based on a differential tethered scheme.
+Two algorithms are proposed in a very basic configuration.
+BLABLABLABLABLABLA
 \end{abstract}
@@
 \begin{IEEEkeywords}
-robotics, tethered, HOT, COPD
+robotics, tethered, COPD
 \end{IEEEkeywords}
@@
 
-Describe the PVC factoring process that Esteban created.  The materials and the name of the CNC that they are using to manufacture everything.  Add the components.
+Describe the PVC manufacturing process that Esteban created.  The materials and the name of the CNC that they are using to manufacture everything.  Add the components.
 
@@
 
+The purpose of this section is to evaluate the feasibility of both control strategies, their engineering issues.
+
 \begin{figure}[h!]
 \centering
-\includegraphics[width=8cm]{performance.eps}
-\caption{Performance curves for the eight subjects included in the dataset of ALS patients.  Three out of eight subjects achieved the necessary performance to implement a valid P300 speller.}
+\includegraphics[width=8cm]{images/alpibot1.jpg}
+\caption{This robot prototype.}
 \label{fig:performance}
@@
 \centering
-\subfloat[Case I]{\includegraphics[width=15cm]{subject.png}\label{subject8}}
+\subfloat[Case I]{\includegraphics[width=15cm]{images/alpibot2.jpg}\label{subject8}}
 \caption{Ten sample P300 template patches for subjects 8 (A) and 3 (B) of the ALS Dataset.
@@
 \label{discussion}
+
+\subsection{Cliical Assesment} 
+\label{section:assesment}
 

### 2020-02-17 alpibot.tex
Iterating.  Added more intro and references.

@@
 % \documentclass[journal]{../sty/IEEEtran}
+
+\usepackage[utf8]{inputenc}
+\usepackage{hyperref}
+\usepackage{algorithm2e} % For algorithms
+\usepackage{amsmath, amssymb, amsthm} % Mathematical packages
+\hypersetup{hidelinks=true}
 
@@
 BLABLABLABLABLABLA
+
+This work is inspired in () but with a different setting in mind, particularly for Pulmonary Rehabilitation procedures.
+
+First describe the problem.  Cite the japanese work and the work similar to what we were doing that is in the book.
+
+Describe the robot.  How it works, the hardware software, python based the devices and tools used and other architectural elements. Add a graph of how it works.
+
+Describe the PVC manufacturing process that Esteban created.  The materials and the name of the CNC that they are using to manufacture everything.  Add the components.
+
+Describe the active spring using scrapped DVD motors.
+
+Control strategies: describe both of them.
+
+Results: 
+show the results of the control strategy using Webots and also the same strategy with the real robot.
+
+Discuss:
+Describe how everything worked.   Which one is the better and describe how they worked.
 \end{abstract}
@@
 
-\IEEEPARstart{A}{lthough} re
+\IEEEPARstart{C}{hronic} Obstructive Pulmonary Disease (COPD) is an umbrella term that describe several pulmonary affections characterized by a eskeletomuscular atrophy.
+In order to carve these after effects a treatment called Pulmonary Rehabilitation is regularly performed on patients.
+During these treatments patients underwent walking procedures and physical exercises under the supervision of physical therapist.
+These patients are characterized by a severe low saturation illness and they require effective oxygen supply, particularly when performing physical activity.
+Hence, patients require to carry with them an oxygen tank to assist in the oxygenotherapy.
+However, their own condition prevent them with the ability to precisely carry the often bulky external tank.
+This situation pertain as a realistic solution to provide two physical therapist, with one of them carrying the oxygen tank, or to devise a customization to wheelchairs to allow them to carry the oxygen tank.
+This ended up sometimes being a cumbersome situation for the physical therapist because he had to push and maneuver the chain with the oxygent tank, and at the same time take care of the patient.
 
-First describe the problem.  Cite the japanese work and the work similar to what we were doing that is in the book.
+These scenario may be aggravated by the fact that the Rehabilitation Gym may be sometimes crowded with different patients, physiologist, physical therapists.
 
-Describe the robot.  How it works, the hardware software, python based the devices and tools used and other architectural elements. Add a graph of how it works.
+Robotics applications can offer a solution in this scenario.
 
-Describe the PVC manufacturing process that Esteban created.  The materials and the name of the CNC that they are using to manufacture everything.  Add the components.
+COPD treatment for patients has been tackled from many perspectives.
+An extensive body of research deals with COPD patients from the point of view of telemedicine applications to track biological signals for patients in home care scenarios.
 
-Describe the active spring using scrapped DVD motors.
+Tethered robots, on the other hand, have been extensively research in robotics communities.
+They offered a very simple solution to some common navigation problems, and they can be very effective on robot-to robot interaction or while interacting with humans in Human Robot Interfaces, collaborative robotics~\cite{Rekleitis2001,Hirata2009}.
 
-Control strategies: describe both of them.
+Differential tethered robots are so widespread.  The work presented by \cite{Ortlieb2016} offers an excellent scheme though aimed to a different purpose.
 
-Results: 
-show the results of the control strategy using Webots and also the same strategy with the real robot.
+REFERENCE TO RAS MAgazine and ONU . \cite{Khamis2019}
 
-Discuss:
-Describe how everything worked.   Which one is the better and describe how they worked.
+This work follows the line of \cite{Endo2015}
 
@@
 
+In order to prioritize several iterations that could be verified by key stakeholders, we developed a simple robotic prototype constructed on Internet of Robotic Things~\cite{Simoens2018}.
+This discipline is part of what is being called 4th industrial revolution and is reshaping many aspects of global manufacturing, including robotic and healthcare.
+
+\subsection{Solution Design}
+
+The medical community is frequently skeptical to technological solutions for medical personel (REF).
+At the same time, development process as design thinking and other similar tendencies prioritize rapid development, prototyping that can bring quickly feedback from real users about the drawbacks and improving opportunities.
+In this context, it is important to perform rapid prototyping.
+
+\subsection{Hardware}
+
+\subsection{Software Components}
+
+\subsection{Control Strategy}
+
+In order to verify the feasibility of the system, two simple control strategies are devised.  The first one is called ...
+
 \section{Experimental Protocol} \label{Protocol}
+
+This section describes the experimental protocol used to evaluate the performance of the proposed solution.
+
+Several metrics are proposed to evaluate the performance.  They are.
+
+First the simulatiom is sdescribed and later the evaluation on the prototyped rrobotic platform is described.
+
+\subsection{Simulation}
+
+\subsection{Real world}
 
@@
 The recorded dataset was sampled at 256 Hz and it consisted of a scalp multichannel EEG signal for electrode channels Fz, Cz, Pz, Oz, P3, P4, PO7 and PO8, identified according to the 10-20 International System,  for each one of the 8 subjects.
-The recording device was a research-oriented digital EEG device (g.Mobilab, g.Tec, Austria) and the data acquisition and stimuli delivery were handled by the BCI2000 open source software~\cite{Schalk2004}.
+The recording device was a research-oriented digital EEG device (g.Mobilab, g.Tec, Austria) and the data acquisition and stimuli delivery were handled by the BCI2000 open source software~\cite{Na2009}.
 
@@
 
-\subsection{Cliical Assesment} 
+\subsection{Clinical Assessment} 
 \label{section:assesment}
 
-\subsection{Conclusion}
+\subsection{Conclusion and Future Work}
 
@@
 \bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-\bibliography{IEEEabrv,gsync}
+\bibliography{alpibot}
+
 %

### 2020-02-19 alpibot.tex
Upload a new version with more references on the introduction.

@@
 
-Robotics applications can offer a solution in this scenario.
+Robotics applications can offer a solution in this scenario.  Ground Service Robot in a person following scenario.
+
+UGVs
 
@@
 Tethered robots, on the other hand, have been extensively research in robotics communities.
-They offered a very simple solution to some common navigation problems, and they can be very effective on robot-to robot interaction or while interacting with humans in Human Robot Interfaces, collaborative robotics~\cite{Rekleitis2001,Hirata2009}.
+They offered a very simple solution to some common navigation problems, and they can be very effective on robot-to robot interaction or while interacting with humans in Human Robot Interfaces, collaborative robotics~\cite{Rekleitis2001,Hirata2009,Ferrin2010}.
+
+Person follower \cite{Islam2019}
 
@@
 
-The medical community is frequently skeptical to technological solutions for medical personel (REF).
+The medical community is frequently skeptical to technological solutions for medical personel~\cite{Gaggioli2006,Fasola2012,Cherubini2008,Wolpaw2012BraincomputerIP}.
 At the same time, development process as design thinking and other similar tendencies prioritize rapid development, prototyping that can bring quickly feedback from real users about the drawbacks and improving opportunities.

### 2020-02-24 alpibot.tex
Upload a little bit more of info.

@@
 
+%\usepackage{fancyhdr}
+%\usepackage{array}
+%\usepackage[rightcaption]{sidecap}
+%\usepackage{wrapfig}
+%\usepackage{float}
+%\usepackage[labelfont=bf]{caption} % bold text for captions
+%\usepackage[para]{threeparttable} % fancy tables, check these before you use them
+%\usepackage{url}
+%\usepackage[table,xcdraw]{xcolor}
+\usepackage{tikz}
+%\usepackage{pgfplots}
+%\pgfplotsset{width=10cm,compat=1.9}
+%\usepgfplotslibrary{external}
+%\tikzexternalize
+%\usepackage{svg}
+
 % Some very useful LaTeX packages include:
@@
 \ifCLASSINFOpdf
-\usepackage[pdftex]{graphicx}
+%\usepackage[pdftex]{graphicx}
 \usepackage{subfig}
@@
 
+The proposed solution is a tethered robot follower, using a pair of strings.
+The following mechanism is implemented by means of measuring encoders on two different reels located according to figure~\ref{fig:components}.
+
+The driving mechanism proposed is a differential wheel robot.
+
 \subsection{Hardware}
+
+In order to achieve quick prototyping cycles, we used extensively tools and modules to the IoRT wave.  Frames are constructed from aluminum extrusions produced by Makeblock (Shenzhen, China).
+
+Motors are in-wheel configuration and they provide optical encoding.
+A microcontroller is used to implement motor controlling and encoder processing.
+This is based on the popular Arduino platform.
+On top of it the Adafruit Shield v2.
+This Arduino bridge can be used to control simultaneously four DC motors.
+
+The microcontroller is connected through serial connection to the main SBC RaspberryPi Model 3B+.
+The main controlling algorithm run on this board, it connects to wifi and provides telemetry and the abiilty to receive remote commands by means a very simple UDP command interface.
+
+Reels were designed from PVC extrusions based on figure.
+They are glued to regular DC motors scavenged from old compact discs.
+Hardware specifications are described on table \ref{tab:specifications}.
+
+\begin{figure}[]
+    \centering
+    \tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt        
+    \begin{tikzpicture}[x=0.50pt,y=0.50pt,yscale=-1,xscale=1]
+    %uncomment if require: \path (0,1774); %set diagram left start at 0, and has height of 1774
+    %Shape: Rectangle [id:dp3163749047300972] 
+    \draw   (341,1289.29) -- (356.46,1289.29) -- (356.46,1484.1) -- (341,1484.1) -- cycle ;
+    %Straight Lines [id:da006917544597343617] 
+    \draw [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][line width=0.75]    (309.46,1464.29) -- (126.19,1383.9) ;
+
+    %Straight Lines [id:da6556986586628986] 
+    \draw [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 230; green, 164; blue, 164 }  ,fill opacity=1 ][line width=0.75]    (310.46,1310.29) -- (126.19,1383.9) ;
+
+    %Shape: Rectangle [id:dp7070304476532343] 
+    \draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ] (309.5,1289) -- (388,1289) -- (388,1324.67) -- (309.5,1324.67) -- cycle ;
+    %Shape: Ellipse [id:dp4674581553313458] 
+    \draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ] (110.01,1383.9) ..
+controls (110.01,1374.53) and (117.25,1366.94) ..
+(126.19,1366.94) ..
+controls (135.12,1366.94) and (142.37,1374.53) ..
+(142.37,1383.9) ..
+controls (142.37,1393.28) and (135.12,1400.87) ..
+(126.19,1400.87) ..
+controls (117.25,1400.87) and (110.01,1393.28) ..
+(110.01,1383.9) -- cycle ;
+    %Shape: Rectangle [id:dp46270765707902994] 
+    \draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ] (308.5,1448.5) -- (387,1448.5) -- (387,1484.17) -- (308.5,1484.17) -- cycle ;
+    %Shape: Circle [id:dp1556138070188774] 
+    \draw  [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ,fill opacity=1 ] (497.83,1383.83) ..
+controls (497.83,1379.51) and (501.34,1376) ..
+(505.67,1376) ..
+controls (509.99,1376) and (513.5,1379.51) ..
+(513.5,1383.83) ..
+controls (513.5,1388.16) and (509.99,1391.67) ..
+(505.67,1391.67) ..
+controls (501.34,1391.67) and (497.83,1388.16) ..
+(497.83,1383.83) -- cycle ;
+    %Straight Lines [id:da9759075394788362] 
+    \draw    (362.11,1349.34) -- (505.67,1383.83) ;
+
+    %Straight Lines [id:da0292648785573405] 
+    \draw    (362.11,1423.58) -- (429.74,1404.86) -- (505.67,1383.83) ;
+
+    %Rounded Rect [id:dp19413777874284577] 
+    \draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 184; green, 233; blue, 134 }  ,fill opacity=1 ] (336,1423.58) ..
+controls (336,1422.16) and (337.16,1421) ..
+(338.58,1421) -- (359.53,1421) ..
+controls (360.96,1421) and (362.11,1422.16) ..
+(362.11,1423.58) -- (362.11,1431.34) ..
+controls (362.11,1432.76) and (360.96,1433.92) ..
+(359.53,1433.92) -- (338.58,1433.92) ..
+controls (337.16,1433.92) and (336,1432.76) ..
+(336,1431.34) -- cycle ;
+    %Rounded Rect [id:dp9015992902460356] 
+    \draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 184; green, 233; blue, 134 }  ,fill opacity=1 ] (336,1341.58) ..
+controls (336,1340.16) and (337.16,1339) ..
+(338.58,1339) -- (359.53,1339) ..
+controls (360.96,1339) and (362.11,1340.16) ..
+(362.11,1341.58) -- (362.11,1349.34) ..
+controls (362.11,1350.76) and (360.96,1351.92) ..
+(359.53,1351.92) -- (338.58,1351.92) ..
+controls (337.16,1351.92) and (336,1350.76) ..
+(336,1349.34) -- cycle ;
+
+    % Text Node
+    \draw  [draw opacity=0][fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ]  (478.5,1350) -- (535.5,1350) -- (535.5,1374) -- (478.5,1374) -- cycle  ;
+    \draw (507,1362) node    {$subject$};
+    % Text Node
+    \draw (302,1344) node    {$left\ reel$};
+    % Text Node
+    \draw (299,1425) node    {$right\ reel$};
+    % Text Node
+    \draw (125,1414) node    {$castor\ wheel$};
+    % Text Node
+    \draw (422,1348) node  [rotate=-12.55]  {$left\ thread$};
+    % Text Node
+    \draw (423,1422) node  [rotate=-344.79]  {$right\ thread$};
+    % Text Node
+    \draw (348,1275) node    {$left\ wheel$};
+    % Text Node
+    \draw (348,1496) node    {$right\ wheel$};
+    \end{tikzpicture}
+    \caption{Components of the robotic vehicle and following mechanism}
+    \label{fig:components}
+\end{figure}{}
 
@@
 
+We replicated the experiments performed by Endo but used a different curve.
+
 Several metrics are proposed to evaluate the performance.  They are.
@@
 \subsection{Simulation}
+
+A model of the proposed design was first built on Webots application.  The threading mechanism was implemented using virtual threads (REF).
 

### 2020-04-10 alpibot.tex
Subo un monton de info.  Parte de la intro, subo todas las imagenes de los resultados, agrego las tablas, agrego el clinical assessment y parte de la discusión.  Falta pero es un avance.

@@
 \hypersetup{hidelinks=true}    
+\usepackage{algorithmic}
 
@@
 % Do not put math or special symbols in the title.
-\title{Pulmonary rehabilitation Robot Follower}
+\title{A Control Strategy for a Tethered Follower Robot for Pulmonary Rehabilitation}
 %
@@
 \markboth{IEEE Transactions on Biomedical Engineering}%
-{Bianchi \MakeLowercase{\textit{et al.}} Pulmonary rehabilitation Robot Follower}
+{Bianchi \MakeLowercase{\textit{et al.}} Controller Design for a Tethered Follower Robot for Pulmonary Rehabilitation}
 % The only time the second header will appear is for the odd numbered pages
@@
 Two algorithms are proposed in a very basic configuration.
-BLABLABLABLABLABLA
-
-This work is inspired in () but with a different setting in mind, particularly for Pulmonary Rehabilitation procedures.
-
-First describe the problem.  Cite the japanese work and the work similar to what we were doing that is in the book.
-
-Describe the robot.  How it works, the hardware software, python based the devices and tools used and other architectural elements. Add a graph of how it works.
-
-Describe the PVC manufacturing process that Esteban created.  The materials and the name of the CNC that they are using to manufacture everything.  Add the components.
-
-Describe the active spring using scrapped DVD motors.
-
-Control strategies: describe both of them.
-
-Results: 
-show the results of the control strategy using Webots and also the same strategy with the real robot.
-
-Discuss:
-Describe how everything worked.   Which one is the better and describe how they worked.
 \end{abstract}
@@
 
-\IEEEPARstart{C}{hronic} Obstructive Pulmonary Disease (COPD) is an umbrella term that describe several pulmonary affections characterized by a eskeletomuscular atrophy.
-In order to carve these after effects a treatment called Pulmonary Rehabilitation is regularly performed on patients.
-During these treatments patients underwent walking procedures and physical exercises under the supervision of physical therapist.
-These patients are characterized by a severe low saturation illness and they require effective oxygen supply, particularly when performing physical activity.
-Hence, patients require to carry with them an oxygen tank to assist in the oxygenotherapy.
+\IEEEPARstart{C}{hronic} Obstructive Pulmonary Disease (COPD) is an umbrella term that describe several pulmonary affections characterized by a eskeletomuscular atrophy~\cite{Kocsis2016,Wu2012}.
+In order to carve these after effects a Pulmonary Rehabilitation procedure is a viable treatment for these patients.
+During these procedures patients underwent controlled walking activities and physical exercises under the supervision of a physical therapist.
+These patients are characterized by a severe low saturation illness and they require effective oxygen supply, particularly when performing physical activity~\cite{Celli2014}.
+Hence, patients require to carry with them an oxygen tank for the oxygenotherapy assistance.
 However, their own condition prevent them with the ability to precisely carry the often bulky external tank.
-This situation pertain as a realistic solution to provide two physical therapist, with one of them carrying the oxygen tank, or to devise a customization to wheelchairs to allow them to carry the oxygen tank.
-This ended up sometimes being a cumbersome situation for the physical therapist because he had to push and maneuver the chain with the oxygent tank, and at the same time take care of the patient.
+This situation entails a pragmatic solution to provide an additional physical therapist solely for the purpose of carrying the oxygen tank.
+Alternatively, they require to devise a customized wheelchair to allow them to carry the oxygen tank on top of it, leading to a cumbersome situation for the physical therapist because s/he had to push and maneuver the wheelchair, and at the same time, take care of the patient and perform the rehabilitation treatment.
 
-These scenario may be aggravated by the fact that the Rehabilitation Gym may be sometimes crowded with different patients, physiologist, physical therapists.
+The scenario may be aggravated by the fact that this procedure is performed on a rehabilitation gym that could be potentially crowded with several patients, physiologist and physical therapists.
 
-Robotics applications can offer a solution in this scenario.  Ground Service Robot in a person following scenario.
+An alternative solution is to use an assitive ground service robot~\cite{Neto2015} to carry the oxygen tank, while following the patient in a leader-follower configuration.
+There are two reasons that support the initial viability of this idea.
+First,  the rehabilitation gym is a constrained environment where this problem can be tackled by an Unmanned Ground Vehicle (UGV).
+On the other hand, the range of movements performed by the patient during the Rehabilitation Procedure is highly predictable by the treatment.
+At the same time,  the global robotic research community persuade the development of robotic affordable solutions to the social and health-related worldwide problems~\cite{Khamis2019}.
 
-UGVs
+For the implementation of the leader-follower strategy, several solutions have been proposed, including SLAM solutions, vision-based systems or based on electromagnetical beacons~\cite{Islam2019}.
+The work presented by \cite{Ortlieb2016} explores a differential tethered robotic system to perform camera-based gait analysis of the leader.
+For COPD Pulmonary Rehabilitaiton procedures, the patient is already umbilicaly linked to the oxygen tank via the breathing cannula.
+Hence, a robotic solution can exploit this circumstance to perform the \textit{following} mechanism based on a tethered controller.
+Tethered robots have been extensively researched in robotics~\cite{Ahn2006}.
+They offer a very simple solution to some common navigation problems, and they can be very effective in robot-to robot interaction, collaborative robotics, or while interacting with humans in Human Robot Interfaces~\cite{Rekleitis2001,Hirata2009,Ferrin2010}.
 
-COPD treatment for patients has been tackled from many perspectives.
-An extensive body of research deals with COPD patients from the point of view of telemedicine applications to track biological signals for patients in home care scenarios.
+On the other hand, several assistance devices for COPD treatments have been proposed.
+Particularly relevant are novel telemedice applications to enhance complementary rehabiliation exersice at home that can track biological markers for those patients~\cite{Wu2012}.
+The work presented here follows the line pointed by \cite{Endo2015}.
+Authors studied the use of a single thread tethered follower robot for home oxygen therapy, and compared two different control algorithms and their effectiveness to mimic the leader path and to avoid obstacles.
+However their approach, focuses on the usage of the device exclusively for home therapy, and not within the context of a Pulmonary Rehabilitation procedure performed by medical personnel on medical installations.
 
-Tethered robots, on the other hand, have been extensively research in robotics communities.
-They offered a very simple solution to some common navigation problems, and they can be very effective on robot-to robot interaction or while interacting with humans in Human Robot Interfaces, collaborative robotics~\cite{Rekleitis2001,Hirata2009,Ferrin2010}.
-
-Person follower \cite{Islam2019}
-
-Differential tethered robots are so widespread.  The work presented by \cite{Ortlieb2016} offers an excellent scheme though aimed to a different purpose.
-
-REFERENCE TO RAS MAgazine and ONU . \cite{Khamis2019}
-
-This work follows the line of \cite{Endo2015}
+Hence, this work aims to provide an initial study on the controlling mechanism of a differential tethered robot that implements the leader-follower configuration on a Pulmonary Rehabilitation procedure.
+This document unrolls as follows.
+Section~\ref{materials} describes poses the problem and the solution design, Section~\ref{experimental} document the experimental protocol to perform the solution assessment on a simulation and on a real world scenario.
+Results and Discussions are described in~\ref{results} and \ref{discussion}.
+The Clinical Assessment performed jointly with medical personnel is tackled in Section~\ref{clinical}.
+Finally, conclusions are exposed in the remaining section~\ref{conclusion}.
 
 \section{Materials and Methods}
+\label{materials}
 
@@
 In order to achieve quick prototyping cycles, we used extensively tools and modules to the IoRT wave.  Frames are constructed from aluminum extrusions produced by Makeblock (Shenzhen, China).
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=8cm]{images/alpibot1.jpg}
+\caption{This robot prototype.}
+\label{fig:performance}
+\end{figure}
 
@@
 
+\begin{figure*}[!t]
+\centering
+\subfloat[Case I]{\includegraphics[width=15cm]{images/alpibot2.jpg}\label{subject8}}
+\caption{Robot prototype built to evaluate the real world scenario of described following mechanisms for the tethered robot.}
+\label{fig:p300templates}
+\end{figure*}
+
+\subsection{Active Reel Spring}
+
+As previously mentioned, an active spring mechanism is also put in place to keep the threads tense at all times and measure the distance to the subject precisely.
+This active spring mechanism can be implemented by just activating both reel motors at all times to keep the threads tense.
+However, in order to extend the useful life of the reel motors, and to save battery, an algorithm to intelligently activate and deactivate the motors was developed.
+
+The algorithm works as follows:
+
+\begin{enumerate}
+    \item While wheels are moving, retract reels.
+    \item If wheels stop moving, wait for \textit{reel wait time} seconds, then retract reels. 
+    \item Retract reels until the reel encoders values have not changed during \textit{reel retract time} seconds.
+    \item If wheels started moving or the encoder values have changed while retracting, start the \textit{reel retract time} countdown again.
+\end{enumerate}{}
+
+%Here is the pseudocode that implements that logic. This function is executed in every run loop in the Arduino module:
+
+%\begin{algorithm}[]
+%\label{alg:active_spring}
+%\caption{Active spring algorithm}
+%\begin{algorithmic}
+%
+%\STATE $currentReelPulses \gets readPulses()$
+%
+%\IF{$retracting$}
+%    \IF{$lastReelPulses != currentReelPulses$}
+%    \STATE $lastReelStart \gets currentTime()$
+%    \ENDIF
+%    \IF{\NOT $ wheelsMoving $ \AND $ currentTime() - lastReelStart > rrt$}
+%    \STATE $retracting \gets \FALSE$
+%    \STATE $lastReelEnd \gets currentTime()$
+%    \ENDIF
+%\ELSE
+%    \IF{$wheelsMoving$}
+%        \STATE $retracting \gets \TRUE$
+%        \STATE $lastReelStart \gets currentTime()$
+%    \ENDIF
+%    \IF{$ currentTime() - lastReelEnd > rwt $}
+%        \STATE $retracting \gets \TRUE$
+%        \STATE $lastReelStart \gets currentTime()$
+%    \ENDIF
+%\ENDIF
+%
+%\STATE $lastReelPulses \gets currentReelPulses$
+%
+%\IF{$retracting$}
+%    \STATE \textit{Activate reel motors}
+%\ELSE
+%    \STATE \textit{Deactivate reel motors}
+%\ENDIF
+%\end{algorithmic}
+%\end{algorithm}
+
 \subsection{Software Components}
@@
 
-In order to verify the feasibility of the system, two simple control strategies are devised.  The first one is called ...
+Two simple control strategies are proposed and evaluated.
+The first one is called Follow-the-thread and the second strategy is Rotate-And-Go.
+The Adafruit Shield controller provides a very stable output signal, provided battery are kept above reference.
+Hence, wheels motor control is open-loop.
 
-\section{Experimental Protocol} \label{Protocol}
+\section{Experimental Protocol} \label{experimental}
 
 This section describes the experimental protocol used to evaluate the performance of the proposed solution.
+The Pulmonary Rehabilitation procedure consists on a series of walking activities aimed to promote patient muscular recovery and well being~\cite{Wu2012}.
+They are slow pace motions following a specific trajectory on a rehabilitation gym.
 
-We replicated the experiments performed by Endo but used a different curve.
+In order to standardize the procedure~\cite{Sprunk2016}, the \textit{Lemniscate of Gerono} is used as desired trajectory, a curve shaped like an $\infty$ symbol, described by the Equation~\ref{eq:lemniscate}:
 
-Several metrics are proposed to evaluate the performance.  They are.
+\begin{equation}
+\begin{array}{ll}  
+ x( \phi ) = a \; cos( \phi )     \\
+ y( \phi ) = a \; cos ( \phi )   \; sin ( \phi )    \\ 
+ \textit{where}  \; \phi \in \lbrace -\pi , \pi  \rbrace
+\end{array}
+\label{eq:lemniscate}
+\end{equation}
 
-First the simulatiom is sdescribed and later the evaluation on the prototyped rrobotic platform is described.
+\noindent where $ \phi $ is the free parameter, and $x$ and $y$ determine the shape of the trajectory on the navigation plane.
+
+The reason  this shape was chosen is because it combines different kinds of trajectories that we want to test the vehicle in: long straight segments, sharp and soft curves, all in one single shape.
+Similar curves are also used in other proposed experiments in \cite{Neto2015,Endo2015}.
+Regarding metrics, three are proposed to evaluate the performance.
+They are
+
+\begin{itemize}
+    \item \textit{Normal path deviation}: the subject trajectory is divided into small segments and then the normal distance to the robot trajectory is calculated for each of those segments.
+Path deviation is relevant to evaluate how closely the robot mimics the leader path, which is the ultimate goal of the robotic vehicle.
+    \item \textit{Maximum path deviation}: the maximum \textit{normal path deviation} registered during an experiment.
+    \item \textit{Total path deviation}: The area under the curve resulting from the the \textit{normal path deviation} over the length travelled by the leader.
+    \item \textit{Robot-leader distance}: The euclidean distance between the robot and the leader, at any point in time.
+This is particularly important since the robot has a limited amount of thread available, so if the leader uses all the available thread, it will start dragging the robot and damaging the following mechanism.
+It is a scenario that must not happen under any circumstance, as it can also be dangerous for a potential patient using the device.
+\end{itemize}{}
+
+In this work, a \textit{following behavior} is considered satisfactory if its maximum path deviation is less than 0.75 m and the robot-leader distance never exceeds 1.5 m~\cite{MunozCeballos2010}.
+
+First the simulation is described and later the evaluation on the robotic prototype is detailed.
 
@@
 
-A model of the proposed design was first built on Webots application.  The threading mechanism was implemented using virtual threads (REF).
+A model of the proposed design was first built on Webots application.
+The threading mechanism was implemented using virtual threads~\cite{Rekleitis2001}.
+The leader traveled according to a predefined path with constant velocity, followiing the lemniscata trajectory.
+
+The simulation is also useful to study the effects of the different constants in each strategy on the movement of the robot.
 
@@
 
-To verify the validity of the proposed framework and method,
+%To verify the validity of the proposed framework and method,
+%The experimental protocol used to 
+%In order to assess
 
-The experimental protocol used to
-
-The recorded dataset was sampled at 256 Hz and it consisted of a scalp multichannel EEG signal for electrode channels Fz, Cz, Pz, Oz, P3, P4, PO7 and PO8, identified according to the 10-20 International System,  for each one of the 8 subjects.
-The recording device was a research-oriented digital EEG device (g.Mobilab, g.Tec, Austria) and the data acquisition and stimuli delivery were handled by the BCI2000 open source software~\cite{Na2009}.
-
-In order to assess
-
-\section{Results} \label{Results}
-\label{section:results}
-
-The purpose of this section is to evaluate the feasibility of both control strategies, their engineering issues.
+A real world experiment was performed, begging to maintain the same conditions implemented on the simulation environment.
+A lemniscata of $  a = 1 m $ was drawn on the floor on the visual capturing room.
 
@@
 \centering
-\includegraphics[width=8cm]{images/alpibot1.jpg}
-\caption{This robot prototype.}
-\label{fig:performance}
+\includegraphics[width=8cm]{images/capture.png}
+\caption{Trajectory for the Follow-The-Thread control strategy.}
+\label{fig:distance_sim}
 \end{figure}
 
-\begin{figure*}[!t]
+\section{Results} \label{Results}
+\label{results}
+
+Simulation results for both control strategies are shown on Figures 1 and 2.
+
+Results metrics for the simualtions are shown on table 1.
+
+\begin{center}
+\begin{tabular}{ |c|c|c|c| }
+\hline
+$c_v$ & $c_{\alpha}$ & Max path deviation [m] & Area under \textit{npd} curve \\
+\hline
+10  &   15  & 0.3614 & 2.0651\\
+15  &   5  & 0.4325 & 2.055\\
+\textbf{15}  &   \textbf{10}  & \textbf{0.2188} & \textbf{1.0902}\\
+15  &   15  & 0.2891 & 1.5059\\
+5  &   20  & 0.5733 & 3.7289\\
+\hline
+\end{tabular}
+\captionof{table}{Normal path deviation and maximum path deviation for different Follow the Thread constants.}
+\label{tab:ft_npd_table}
+\end{center}
+
+\begin{center}
+\begin{tabular}{ |c|c|c|c|c| }
+\hline
+$c_v$ & $c_r$ & $Dt_{off}$ [m] & Max path deviation [m] & Area under \textit{npd} curve \\
+\hline
+10  &   20  &   0.1  & 0.4310 & 1.6380\\
+20  &   20  &   0.05  & 0.7775 & 3.1139\\
+\textbf{20}  &  \textbf{20}  & \textbf{0.1} & \textbf{0.4123} & \textbf{0.9872}\\
+20  &   35  &   0.1  & 0.4143 & 1.4820\\
+20  &   5  &  0.05  & 0.7815 & 3.0892\\
+35  &   20  &   0.1  & 0.6337 & 1.6190\\
+\hline
+\end{tabular}
+\captionof{table}{Normal path deviation and maximum path deviation for different Rotate and Go constants.}
+\label{tab:ft_npd_table}
+\end{center}
+
+Additionally, speed profiles can be observed from both trajectories.
+
+\begin{figure}[h!]
 \centering
-\subfloat[Case I]{\includegraphics[width=15cm]{images/alpibot2.jpg}\label{subject8}}
-\caption{Ten sample P300 template patches for subjects 8 (A) and 3 (B) of the ALS Dataset.
-Downward deflection is positive polarity.
-The P300 signature waveform is more clearly and consistent characterized for subject 8, whereas for subject 3 the characteristic patterns is more difficult to spot.
-}
-\label{fig:p300templates}
-\end{figure*}
+\includegraphics[width=8cm]{images/ft_a2_n1500_cv15_cr15.eps}
+\caption{Trajectory for the Follow-The-Thread control strategy.}
+\label{fig:distance_sim}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=8cm]{images/rgv_cv20_cr20_dr10.eps}
+\caption{Trajectory for the Rotate-And-Go control strategy.}
+\label{fig:distance_sim}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=8cm]{images/ft_cv15_cr10_leader_robot_speed.eps}
+\caption{Follow-The-Thread.}
+\label{fig:distance_sim}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=8cm]{images/rg_cv20_cr20_bv10_dtoff10_leader_robot_speed.eps}
+\caption{Speed profiles for Rotate-And-Go.}
+\label{fig:distance_sim}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=8cm]{images/ft_vs_rg_dist.eps}
+\caption{Distance between robot and leader [m]}
+\label{fig:distance_sim}
+\end{figure}
+
+Results for the real world experiment can be seen on Figures 3 and 4.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=8cm]{images/ft1cap.eps}
+\caption{Trajectories for the Follow-The-Thread control strategy.}
+\label{fig:distance_sim}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=8cm]{images/rg2cap.eps}
+\caption{Distance between robot and leader [m]}
+\label{fig:distance_sim}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=8cm]{images/ft1cap_leader_robot_speed.eps}
+\caption{Follow the thread speed profiles}
+\label{fig:distance_sim}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=8cm]{images/rg2cap_leader_robot_speed.eps}
+\caption{Rotate and Go speed profiles}
+\label{fig:distance_sim}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=8cm]{images/cap_ft_vs_rg_dist.eps}
+\caption{Rotate and Go speed profiles}
+\label{fig:distance_sim}
+\end{figure}
+
+\begin{figure} 
+    \centering
+  \subfloat[a\label{1a}]{%
+       \includegraphics[width=0.45\linewidth]{images/cap_ft_vs_rg_dist.eps}}
+    \hfill
+  \subfloat[b\label{1b}]{%
+        \includegraphics[width=0.45\linewidth]{images/cap_ft_vs_rg_dist.eps}}
+    \\
+  \subfloat[c\label{1c}]{%
+        \includegraphics[width=0.45\linewidth]{images/cap_ft_vs_rg_dist.eps}}
+    \hfill
+  \subfloat[d\label{1d}]{%
+        \includegraphics[width=0.45\linewidth]{images/cap_ft_vs_rg_dist.eps}}
+  \caption{(a), (b) Some examples from CIFAR-10 \cite{4}. The objects in     
+        single-label images are usually roughly aligned.(c),(d) However, the 
+        assumption of object alignment is not valid for multi-label
+        images. Also note the partial visibility and occlusion
+        between objects in the multi-label images.}
+  \label{fig1} 
+\end{figure}
+
+\begin{center}
+\begin{tabular}{ |c|c|c|c| }
+\hline
+$c_v$ & $c_{\alpha}$ & Max path deviation [m] & Area under \textit{npd} curve \\
+\hline
+25  &   20  & 0.3876 & 1.9761\\
+25  &   35  & 0.4672 & 2.3528\\
+\hline
+\end{tabular}
+\captionof{table}{Normal path deviation and maximum path deviation in motion capture experiments using Follow the Thread.}
+\label{tab:cap_ft_npd_table}
+\end{center}
+
+\begin{center}
+\begin{tabular}{ |c|c|c|c|c| }
+\hline
+$c_v$ & $c_r$ & $Dt_{off}$ & Max path deviation [m] & Area under \textit{npd} curve \\
+\hline
+30  &   35  & 0.04  & 0.4116 & 2.8309\\
+30  &   35  & 0.08 & 0.3739 & 2.0367\\
+\hline
+
+\end{tabular}
+\captionof{table}{Normal path deviation and maximum path deviation in motion capture experiments using Rotate and Go.}
+\label{tab:cap_rg_npd_table}
+\end{center}
 
@@
 \subsection{Clinical Assessment} 
-\label{section:assesment}
+\label{clinical}
+
+No amount of metrics are enough to evaluate if the robot is a viable solution for this problem or not, without the input and the evaluation of the people that are going to physically make use of it.
+
+As this project is a joint research between ALPI and ITBA, we invited 4 people from the ALPI staff, 3 doctors and one engineer, to the motion capture laboratory at ITBA, where we made a live demonstration of the robot working and following a moving person.
+
+In the demonstration, we described how the robot worked, how it was built and how to operate it. Both control strategies were explained, along with the main superficial differences between them.
+
+Afterwards, we invited the doctors to use the robot themselves, simulating they were the patient being followed.
+We let the doctors switch between the two control strategies to evaluate both of them, and made different tests, one walking along path drawn in the floor (figure \ref{fig:cap_lab_lemniscate}), and another one walking freely in the space available in the lab.
+We let them use the robot as long as they wanted to, to get a general idea of how it behaved, and how it could be used in the rehabilitation process.
+After all evaluations were finished, we answered questions and discussed various aspects of the vehicle prototype with the ALPI staff, and then proceeded to ask them a survey to document their experience with the robot, get their expert opinion on how the two strategies compared against each other, and what other improvements were needed in order to deliver a fully usable product.
+
+\begin{table*}[t]
+\begin{center}
+\begin{tabular}[!t]{|c|c|}
+\hline
+Question & Avg. answer \\
+\hline
+How would you qualify, from 1 to 5, your overall experience with Follow the Thread? (1:Bad, 5:Excellent)  & 5.0\\
+\hline
+How safe would a patient be, from 1 to 5, being followed by the robot using the Follow the Thread strategy? (1:Very unsafe, 5:Very safe)  & 4.25\\
+\hline
+How would you qualify, from 1 to 5, your overall experience with Rotate and Go? (1:Bad, 5:Excellent)  & 3.5\\
+\hline
+How safe would a patient be, from 1 to 5, being followed by the robot using the Rotate and Go strategy? (1:Very unsafe, 5:Very safe) & 3.25\\
+\hline
+\end{tabular}
+\captionof{table}{Answers to survey questions.}
+\label{tab:alpi_q_table_1}
+\end{center}
+\end{table*}
+
+According to their answers, and the discussion we had after testing the robot, the general opinion was that the \textit{Follow the Thread} strategy was safer and more convenient for the task.
+In the survey, when asked \textit{Which of the two strategies is more effective at following the patient in a rehabilitation exercise?}, all 4 people responded that \textit{Follow the Thread} is "much better".
+
+The main concern with the \textit{Rotate and Go} strategy was that having to wait for the robot to rotate before moving forward might be unsafe, as the patient could move away from it and compromise the cannula connecting him or her to the oxygen tank.
+This issue was identified during our own tests, and was not mentioned when explaining the following mechanism to the doctors, to avoid skewing them.
+They independently identified this problem, and emphasized that it could be a great source of discomfort for the patient.
+
+Another aspect that was value from the \textit{Follow the Thread} strategy is that since its smoother movement, with no sudden stops or accelerations, was favorable for the stability of the robot if it were to carry the heavy oxygen tank.
+
+Two needed security measures were also brought up by the ALPI team.
+Firstly, the need to add some mechanism for obstacle avoidance.
+They mentioned the need to have sensor to detect if the robot was about to hit something (specially the patient), and stop immediately, apart from what the control strategy indicated.
+Secondly, they recognized that some patients have very weak stability, and might fall down or take a step back, towards the robot, so it should be able to automatically move away from the patient, in order not to become another obstacle for him or her.
+"Tener en cuenta la distancia a la que se encuentra el robot por si el paciente llega a dar un paso atrás - para evitar riesto de caída'" and "Estaría bueno pensar en algún método de emergencia por caída'' were two quotes related to this issues, taken from the surveys.
+
+In order to have more information for the next steps in the development of the robot, we asked for their advise to design the mechanism to attach the threads to the patient being followed.
+Two ideas were proposed: a belt strapped to the patient waist, or a clasp that would also to the clothes of the patient, also near its waistline.
+The waist is a good attachment point, since it is relatively stable when the patient moves, compared to its hands or legs, that may make sudden movements and confuse the robot sensors.
 
 \subsection{Conclusion and Future Work}
+\label{conclusion}
 
 The goal of this work is to verify if a tethered robot to implement a following scheme on a patient during a pulmonary rehabilitation procedure is too simplistic to be a factual implementation.
+
+From the practical experiments we showed that the both algorithms gave reasonable following performance in task of following the leader along a lemniscate-shaped path.
+This research is that this performance was achieved using very low cost components.
+It was proven that a reasonable following behaviour could be accomplished with a simple mechanism, a characteristic that significantly lowers the price of a device like this, putting it within reach of many medical institutions.
+
+Each control strategy has its advantages, but according various metrics described in this work, the \textit{Follow and Turn} strategy had a more desirable behaviour, as it tended to follow the leader from a closer distance at all times, while moving in a smooth and predictable way.
+
+We understood that no metric is enough to evaluate how appropriate the robotic vehicle can be for a real life scenario.
+Therefore, we gathered feedback from healthcare professionals from ALPI, who provided us with invaluable feedback to evaluate our solution.
+Over all, they also valued the Follow and Turn strategy as being the safer and more effective one.
+Most importantly, they also validated the research made and approved the direction of the project.
+They proposed a series of improvements and next steps, and after seeing the prototype in action, believed that we were close to a real, working solution.
+
+It is essential to involve stakeholders such as patients, doctors, nurses, and any other professionals involved in the rehabilitation process early in the design roadmap.
+They are the ones who understand the problem better than anyone else, and will be the end users of a product like this.
+
+The staff at ALPI was enthusiastic about helping in the development of the robot, and their encouragement and support were the key reasons we could take this project forward.
+This work not only made progress from a technical standpoint, but could also become the foundation of a long-term relationship of collaboration and research between ALPI and ITBA.
+This is also the start of a long-term project at ITBA.
+One that can go on to become a complete, real world solution that can improve the quality of life of many people through the use of technology and engineering.
+
+\subsection{Future Work}
+A following mechanism has been designed and implemented in a hardware prototype.
+The mechanism, along with the two control algorithms, has shown promise to be a viable solution to implement in a functioning robot that ALPI, or any other medical institution, can use in their every day exercises with COPD patients.
+
+The next steps for this project would be to take this prototype and build a bigger version, with motors powerful enough to carry a real oxygen tank.
+
+Further research is also needed to develop a full fledged solution: 
+\begin{itemize}
+    \item Power the motor wheels with a \textit{closed-loop and more efficient controller}, to have a more precise control over their speed.
+    \item Redesign the active spring mechanism in the reels to avoid overheating.
+In our tests, the small electric motors in each reel got very hot after just 20 minutes of use.
+The algorithm we designed to use them more efficiently (\ref{alg:active_spring}) was enough to keep them in manageable temperatures (under $70^0C$), which was enough for testing the prototype, but would not be appropriate for a useful robot.
+Not only because the heat produced can be dangerous, but also because the motors can overheat and break their insulation in regular use.
+    \item An easy and safe interaction between the patient, the operator and the robot.
+How to communicate the state of the robot to the operator, how to control and manipulate the robot in an effective and user-friendly way.
+    \item Safety measures to keep the patient and the doctor or nurse safe when using the robot.
+Not only safe from the robot movement, but also from its electronic components, which can be dangerous when used near the pure oxygen tank if they are not protected properly.
+    \item Investigate and design mechanisms for obstacle avoidance.
+This need was emphasized by the doctors from ALPI.
+The robot should have mechanisms in place to deal with emergency situations, and under no circumstance it can hit the patient or the doctor operating it.
+    \item Achieve a battery autonomy that makes the robot useful throughout a complete pulmonary rehabilitation exercise.
+It is crucial for its usefulness to be able to hold a charge for this period of time, along with the ability to quickly swap batteries if the vehicle will be continually used with different patients.
+\end{itemize}{}
 

### 2020-04-10 alpibot.tex
Agrego los agradecimientos.

@@
         and Santos~Juan~Miguel% <-this % stops a space
-\thanks{R. Ramele  and J.M.Santos are with the Department
+\thanks{E. Buzniak, R. Ramele  and J.M.Santos are with the Department
 of  Computer Engineering, Instituto Tecn{o'}gico de Buenos Aires (ITBA), Ciudad de Buenos Aires,  Argentina e-mail: rramele@itba.edu.ar}% <-this % stops a space
-\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}
+\thanks{Manuscript received April 19, 2020; revised August 26, 2015.}}
 
@@
 \section{Acknowledgments}
-This project was supported by ALPI organization.
+This project is part of a joint collaboration between ALPI organization and the ITBA University.
+Authors would like to thank thoughtfully for their initiative, support given by the ALPI organization.
+Specially to Natalia Nerina Meda, Soledad Suriá, Eduardo Etcheverry and Sergio Carlos Franco, for their invaluable help and enthusiasm to move this project forward.
 

### 2020-04-14 alpibot.tex
Upload a new version.  Still some work ahead but the general layout is already here.

@@
         and Santos~Juan~Miguel% <-this % stops a space
-\thanks{E. Buzniak, R. Ramele  and J.M.Santos are with the Department
-of  Computer Engineering, Instituto Tecn{o'}gico de Buenos Aires (ITBA), Ciudad de Buenos Aires,  Argentina e-mail: rramele@itba.edu.ar}% <-this % stops a space
+\thanks{E. Buzniak, R. Ramele  and J.M.Santos are with the CiC Laboratory at the Department
+of  Computer Engineering, Instituto Tecnológico de Buenos Aires (ITBA), Ciudad de Buenos Aires,  Argentina e-mail: rramele@itba.edu.ar}% <-this % stops a space
 \thanks{Manuscript received April 19, 2020; revised August 26, 2015.}}
@@
 
-\IEEEPARstart{C}{hronic} Obstructive Pulmonary Disease (COPD) is an umbrella term that describe several pulmonary affections characterized by a eskeletomuscular atrophy~\cite{Kocsis2016,Wu2012}.
-In order to carve these after effects a Pulmonary Rehabilitation procedure is a viable treatment for these patients.
+\IEEEPARstart{C}{hronic} Obstructive Pulmonary Disease (COPD) is an umbrella term that describes several pulmonary affections.
+They are characterized as a slowly progressive condition marked by airflow limitation, being cigarette smoking the main etiologic factor~\cite{MacNee2005}.
+This pathology presents an eskeletomuscular atrophy~\cite{Kocsis2016,Wu2012} and in order to carve these after effects a Pulmonary Rehabilitation procedure is a viable treatment for these patients.
 During these procedures patients underwent controlled walking activities and physical exercises under the supervision of a physical therapist.
-These patients are characterized by a severe low saturation illness and they require effective oxygen supply, particularly when performing physical activity~\cite{Celli2014}.
+These patients present a severe low saturation illness and they require effective oxygen supply, particularly when performing physical activity~\cite{Celli2014}.
 Hence, patients require to carry with them an oxygen tank for the oxygenotherapy assistance.
@@
 This situation entails a pragmatic solution to provide an additional physical therapist solely for the purpose of carrying the oxygen tank.
-Alternatively, they require to devise a customized wheelchair to allow them to carry the oxygen tank on top of it, leading to a cumbersome situation for the physical therapist because s/he had to push and maneuver the wheelchair, and at the same time, take care of the patient and perform the rehabilitation treatment.
+Alternatively, they are required to improvise a customized wheelchair to allow them to carry the oxygen tank on top of it, leading to a cumbersome situation for the physical therapist because s/he had to push and maneuver the wheelchair, and at the same time, take care of the patient and perform the rehabilitation treatment.
 
@@
 On the other hand, the range of movements performed by the patient during the Rehabilitation Procedure is highly predictable by the treatment.
-At the same time,  the global robotic research community persuade the development of robotic affordable solutions to the social and health-related worldwide problems~\cite{Khamis2019}.
-
-For the implementation of the leader-follower strategy, several solutions have been proposed, including SLAM solutions, vision-based systems or based on electromagnetical beacons~\cite{Islam2019}.
+At the same time,  the global robotic research community looks forward for the development of robotic affordable solutions to the social and health-related worldwide problems~\cite{Khamis2019}.
+
+For the implementation of the leader-follower strategy, several solutions have been proposed, including Simultaneos Localization And Mapping (SLAM) alternatives, vision-based systems or based on electromagnetical beacons~\cite{Islam2019}.
 The work presented by \cite{Ortlieb2016} explores a differential tethered robotic system to perform camera-based gait analysis of the leader.
@@
 On the other hand, several assistance devices for COPD treatments have been proposed.
-Particularly relevant are novel telemedice applications to enhance complementary rehabiliation exersice at home that can track biological markers for those patients~\cite{Wu2012}.
-The work presented here follows the line pointed by \cite{Endo2015}.
+Particularly relevant are novel telemedice applications to enhance complementary rehabiliation exercise at home that can track biological markers for patients~\cite{Wu2012}.
+The work presented here follows the line established by \cite{Endo2015}.
 Authors studied the use of a single thread tethered follower robot for home oxygen therapy, and compared two different control algorithms and their effectiveness to mimic the leader path and to avoid obstacles.
-However their approach, focuses on the usage of the device exclusively for home therapy, and not within the context of a Pulmonary Rehabilitation procedure performed by medical personnel on medical installations.
+However, their approach focuses on the usage of the device exclusively for home therapy, and not within the context of a Pulmonary Rehabilitation procedure performed by medical personnel on medical facilities.
 
 Hence, this work aims to provide an initial study on the controlling mechanism of a differential tethered robot that implements the leader-follower configuration on a Pulmonary Rehabilitation procedure.
-This document unrolls as follows.
-Section~\ref{materials} describes poses the problem and the solution design, Section~\ref{experimental} document the experimental protocol to perform the solution assessment on a simulation and on a real world scenario.
-Results and Discussions are described in~\ref{results} and \ref{discussion}.
+To do so, this document unrolls as follows.
+Section~\ref{materials} poses the problem and the solution design.
+Section~\ref{experimental} documents the experimental protocol to perform the solution assessment on a simulation and on a real world scenario.
+Results and Discussions are described in~\ref{results} and~\ref{discussion}.
 The Clinical Assessment performed jointly with medical personnel is tackled in Section~\ref{clinical}.
@@
 
-In order to prioritize several iterations that could be verified by key stakeholders, we developed a simple robotic prototype constructed on Internet of Robotic Things~\cite{Simoens2018}.
-This discipline is part of what is being called 4th industrial revolution and is reshaping many aspects of global manufacturing, including robotic and healthcare.
+%This discipline is part of what is being called 4th industrial revolution and is reshaping many aspects of global manufacturing, including robotic and healthcare.
+
+To be effective, any technological solution for the medical community requires active involvement of key stakeholders: physicians, care-givers, patients and their families~\cite{Gaggioli2006,Fasola2012,Cherubini2008,Wolpaw2012BraincomputerIP,Salvini2015}.
+Hence, design methodologies that allow rapid prototyping can bring quickly feedback from real users about drawbacks or opportunities for improvements.
+
+Looking forward to achieve this goal, a basic robotic configuration is designed that allows the implementation of the  tethered controlling mechanism, while keeping away other necessary features that will be the focus of future iterations.
+This design is first simulated in a simulation environment, and later, a basic hardware prototype based on Internet of Robotic Things~\cite{Simoens2018}  is built to verify the design guidelines and assumptions on a real world scenario.
 
@@
 
-The medical community is frequently skeptical to technological solutions for medical personel~\cite{Gaggioli2006,Fasola2012,Cherubini2008,Wolpaw2012BraincomputerIP}.
-At the same time, development process as design thinking and other similar tendencies prioritize rapid development, prototyping that can bring quickly feedback from real users about the drawbacks and improving opportunities.
-In this context, it is important to perform rapid prototyping.
-
-The proposed solution is a tethered robot follower, using a pair of strings.
-The following mechanism is implemented by means of measuring encoders on two different reels located according to figure~\ref{fig:components}.
-
-The driving mechanism proposed is a differential wheel robot.
+The proposed solution is a Differential-Wheeled Robot (DWR) tethered to the followed subject with two threads ending at a single point attached to the subject waist, back or hand.
+In the same axis as the two front wheels, the robot has two reels separated by a certain length from which these threads come.
+As the subject moves away from the robot, the reels release thread so that the patient does not physically drag the vehicle.
+When the opposite happens, and the vehicle gets closer to the subject, an active spring mechanism driven by electric motors move each reel to retract the thread.
+The threads need to be tense at all times so that the encoders in each reel can be used to continuously measure the distance between the subject and the reel as devised in Figure~\ref{fig:components}.
+
+Encoders in each reel measure the difference in length of each thread compared to its initial position.
+This difference in length is the input for the control algorithm.
+Using the encoder, the difference in length of the thread can be measured with this formula:
+
+\begin{equation}
+\label{eq:encoder_len}
+L_d = pulses . \frac{2\pi r}{ppr}
+\end{equation}{}
+
+The variable $r$ is the radius of the reel and $ppr$ is the pulses per revolution (resolution) of the encoder.  The initial position of the threads is configurable.
 
@@
 
-In order to achieve quick prototyping cycles, we used extensively tools and modules to the IoRT wave.  Frames are constructed from aluminum extrusions produced by Makeblock (Shenzhen, China).
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=8cm]{images/alpibot1.jpg}
-\caption{This robot prototype.}
-\label{fig:performance}
-\end{figure}
-
-Motors are in-wheel configuration and they provide optical encoding.
-A microcontroller is used to implement motor controlling and encoder processing.
-This is based on the popular Arduino platform.
-On top of it the Adafruit Shield v2.
-This Arduino bridge can be used to control simultaneously four DC motors.
-
-The microcontroller is connected through serial connection to the main SBC RaspberryPi Model 3B+.
-The main controlling algorithm run on this board, it connects to wifi and provides telemetry and the abiilty to receive remote commands by means a very simple UDP command interface.
-
-Reels were designed from PVC extrusions based on figure.
-They are glued to regular DC motors scavenged from old compact discs.
-Hardware specifications are described on table \ref{tab:specifications}.
+Frames are constructed from aluminum extrusions produced by Makeblock (Shenzhen, China).
 
@@
 
+Motors are in-wheel configuration and they provide optical encoding.
+A microcontroller is used to implement motor controlling and encoder processing.
+This is based on the popular Arduino platform.
+On top of it the Adafruit Shield v2.
+This Arduino bridge can be used to control simultaneously four DC motors.
+
+The microcontroller is connected through serial connection to the main SBC RaspberryPi Model 3B+.
+The main controlling algorithm run on this board, it connects to wifi and provides telemetry and the abiilty to receive remote commands by means a very simple UDP command interface.
+
+Reels were designed from PVC extrusions based on figure.  They are glued to regular DC motors scavenged from old compact discs.  
+%Hardware specifications are described on table \ref{tab:specifications}.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=8cm]{images/alpibot1.jpg}
+\caption{This robot prototype.}
+\label{fig:performance}
+\end{figure}
+
 \begin{figure*}[!t]
@@
 \subfloat[Case I]{\includegraphics[width=15cm]{images/alpibot2.jpg}\label{subject8}}
-\caption{Robot prototype built to evaluate the real world scenario of described following mechanisms for the tethered robot.}
+\caption{Robot prototype built to evaluate the real world scenario of the following mechanisms for the tethered robot.}
 \label{fig:p300templates}
@@
 
-As previously mentioned, an active spring mechanism is also put in place to keep the threads tense at all times and measure the distance to the subject precisely.
-This active spring mechanism can be implemented by just activating both reel motors at all times to keep the threads tense.
-However, in order to extend the useful life of the reel motors, and to save battery, an algorithm to intelligently activate and deactivate the motors was developed.
+As previously mentioned, an active spring mechanism is also put in place to keep the threads tense.
+However, in order to extend the useful life of the reel motors, and to save battery, an algorithm to  activate and deactivate the motors is developed.
 
@@
 
+\subsubsection{Follow the Thread}
+
+This control strategy is similar to the one presented in \cite{Ortlieb2016}.
+It is based on the idea that the difference in length between both threads provides a quasi-linear function of the relative angle between the subject and the vehicle orientation and, additionally, the mean of both thread length allows to approximate the relative distance between the robot and the subject.
+
+\begin{equation}
+v_{tar} = c_v (\frac{D_l + D_r}{2} - d_0)
+\end{equation}
+
+\begin{equation}
+vleft_{tar} = v_{tar} + c_{\alpha} (D_l - D_r)
+\end{equation}
+
+\begin{equation}
+vright_{tar} = v_{tar} - c_{\alpha} (D_l - D_r)
+\end{equation}
+
+To stop the vehicle completely when it is close to its expected position, an additional condition is added:
+
+\begin{equation}
+\textit{if  }{\frac{D_l + D_r}{2} < d_0}\textit{ then } vleft_{tar} = vright_{tar} = 0 
+\end{equation}
+
+$d_0$ is a constant offset. $c_v$ and $c_{\alpha}$ are constant coefficients used for calibration.
+
+\subsubsection{Rotate and Go}
+
+The \textit{Rotate and Go} algorithm divides the vehicle movement in two steps:
+
+1) Rotating the vehicle around the center point of the axis that connects its front wheels in order to aim at the subject.
+
+2) Go forward in a straight line until the vehicle is at the expected distance to the subject or the vehicle is not moving in the direction of the subject.
+
+The algorithm is detailed in Algorithm~\ref{alg:rotateandgoalg}.
+The variable $V_r$ is the speed at which the vehicle will rotate on its axis, and $V_f$ is the speed at which the vehicle will move forward once it can move on the subject's direction.
+
+\begin{algorithm}[]
+\caption{Rotate and Go algorithm}
+\begin{algorithmic}
+\STATE $Dt \gets D_l - D_r$
+\STATE $Dm \gets \frac{D_r+D_l}{2}$
+\STATE $V_r \gets c_r * (abs(Dt) - Dt_off) + base_{vr}$ 
+\STATE $V_f \gets c_v * (Dm - Dm_off)$
+
+\IF{$abs(Dt) > Dt_{off}$}
+\IF{$D_l > D_r$}
+\STATE $vleft_{tar} \gets V_r$
+\STATE $vright_{tar} \gets -V_r$
+\ELSE
+\STATE $vleft_{tar} \gets -V_r$
+\STATE $vright_{tar} \gets V_r$
+\ENDIF
+
+\ELSE
+\IF{$D_m > Dm_{off}$}
+\STATE $vright_{tar} \gets V_f$
+\STATE $vleft_{tar} \gets V_f$
+\ELSE
+\STATE $vright_{tar} \gets 0$
+\STATE $vleft_{tar} \gets 0$
+\ENDIF
+\ENDIF
+\STATE Return $vright_{tar}$ and $vleft_{tar}$.
+\end{algorithmic}
+\label{alg:rotateandgoalg}
+\end{algorithm}
+
 \section{Experimental Protocol} \label{experimental}
@@
 
-The reason  this shape was chosen is because it combines different kinds of trajectories that we want to test the vehicle in: long straight segments, sharp and soft curves, all in one single shape.
+The reason  this shape was chosen is because it combines different kinds of trajectories where the vehicle can be tested: long straight segments, sharp and soft curves, all in one single shape.
 Similar curves are also used in other proposed experiments in \cite{Neto2015,Endo2015}.
@@
 
-A real world experiment was performed, begging to maintain the same conditions implemented on the simulation environment.
-A lemniscata of $  a = 1 m $ was drawn on the floor on the visual capturing room.
+A real world experiment was performed, pegging to the same conditions implemented on the simulation environment.
+A motion capture system is used to track the movement of a human leader along a predetermined path.
+The motion tracking system consists of an array of 16 OptiTrack (NaturalPoint Inc, Oregon, US) Flex 3 cameras, which measure the position of reflective markers with an accuracy of $\pm1$ a sampling rate of 100 Hz.
+The calibration and data collection was made using the Motive motion capture software.
+
+A tracked marker was placed in each side of the robot (on top of each thread reel).
+The human leader used his hand to grab the tip at which the two tethers were tied together.
+A third marker was placed in his hand, using a glove.
+The lemniscate of Gerono, used in the simulation, was drawn in the floor (\ref{fig:cap_lab_lemniscate}), and the human leader tried to move his hand following this shape as close as possible, with stable speed.
+
+The shape was drawn according to the shape described in equation \ref{eq:lemniscate_of_gerono}, using $a = 2$, in meters.
+
+The three markers allowed us to measure the trajectory of the robot and the leader, and then obtain the same metrics calculated in the simulation.
 
@@
 \includegraphics[width=8cm]{images/capture.png}
-\caption{Trajectory for the Follow-The-Thread control strategy.}
+\caption{Hardware prototype on the motion capture system and a testing subject holding the threads.
+The lemniscata of gerone was marked on the floor.
+The subject follows this track on the experiments performed.}
 \label{fig:distance_sim}
@@
 
-Simulation results for both control strategies are shown on Figures 1 and 2.
-
-Results metrics for the simualtions are shown on table 1.
+Simulation results for both control strategies are shown on Figure~\ref{fig:simulationresults}.
+Results metrics for the simulations are shown on Table~\ref{tab:simulationmetricsftt} for the Follow-The-Thread strategy, whereas metrics for Rotate-And-Go are shown on Table~\ref{tab:simulationmetricsrg}.
 
@@
 \hline
-$c_v$ & $c_{\alpha}$ & Max path deviation [m] & Area under \textit{npd} curve \\
+$c_v$ & $c_{\alpha}$ & m.p.d. & Area under \textit{npd} curve \\
 \hline
@@
 \end{tabular}
-\captionof{table}{Normal path deviation and maximum path deviation for different Follow the Thread constants.}
-\label{tab:ft_npd_table}
+\captionof{table}{Max path deviation m.p.d. [m] and Area under the normal path deviation for different Follow the Thread constants.}
+\label{tab:simulationmetricsftt}
 \end{center}
@@
 \hline
-$c_v$ & $c_r$ & $Dt_{off}$ [m] & Max path deviation [m] & Area under \textit{npd} curve \\
+$c_v$ & $c_r$ & $Dt_{off}$ & m.p.d. & Area under \textit{npd} curve \\
 \hline
@@
 \end{tabular}
-\captionof{table}{Normal path deviation and maximum path deviation for different Rotate and Go constants.}
-\label{tab:ft_npd_table}
+\captionof{table}{Max path deviation m.p.d. (in [m]) and area under normal path deviation for different Rotate and Go constants.}
+\label{tab:simulationmetricsrg}
 \end{center}
@@
 
-\begin{figure}[h!]
-\centering
-\includegraphics[width=8cm]{images/ft_a2_n1500_cv15_cr15.eps}
-\caption{Trajectory for the Follow-The-Thread control strategy.}
-\label{fig:distance_sim}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=8cm]{images/rgv_cv20_cr20_dr10.eps}
-\caption{Trajectory for the Rotate-And-Go control strategy.}
-\label{fig:distance_sim}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=8cm]{images/ft_cv15_cr10_leader_robot_speed.eps}
-\caption{Follow-The-Thread.}
-\label{fig:distance_sim}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=8cm]{images/rg_cv20_cr20_bv10_dtoff10_leader_robot_speed.eps}
-\caption{Speed profiles for Rotate-And-Go.}
-\label{fig:distance_sim}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=8cm]{images/ft_vs_rg_dist.eps}
-\caption{Distance between robot and leader [m]}
-\label{fig:distance_sim}
-\end{figure}
-
-Results for the real world experiment can be seen on Figures 3 and 4.
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=8cm]{images/ft1cap.eps}
-\caption{Trajectories for the Follow-The-Thread control strategy.}
-\label{fig:distance_sim}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=8cm]{images/rg2cap.eps}
-\caption{Distance between robot and leader [m]}
-\label{fig:distance_sim}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=8cm]{images/ft1cap_leader_robot_speed.eps}
-\caption{Follow the thread speed profiles}
-\label{fig:distance_sim}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=8cm]{images/rg2cap_leader_robot_speed.eps}
-\caption{Rotate and Go speed profiles}
-\label{fig:distance_sim}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=8cm]{images/cap_ft_vs_rg_dist.eps}
-\caption{Rotate and Go speed profiles}
-\label{fig:distance_sim}
-\end{figure}
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[width=8cm]{images/ft_a2_n1500_cv15_cr15.eps}
+%\caption{Trajectories of the leader and follower for the Follow-The-Thread control strategy.}
+%\label{fig:distance_sim}
+%\end{figure}
+%
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[width=8cm]{images/rgv_cv20_cr20_dr10.eps}
+%\caption{Trajectories of the leader and follower for Rotate-And-Go control strategy.}
+%\label{fig:distance_sim}
+%\end{figure}
+%
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[width=8cm]{images/ft_cv15_cr10_leader_robot_speed.eps}
+%\caption{Speed profiles of the leader and follower for Follow-The-Thread control strategy.}
+%\label{fig:distance_sim}
+%\end{figure}
+%
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[width=8cm]{images/rg_cv20_cr20_bv10_dtoff10_leader_robot_speed.eps}
+%\caption{Speed profiles of the leader and follower for Rotate-And-Go control strategy.}
+%\label{fig:distance_sim}
+%\end{figure}
+%
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[width=8cm]{images/ft_vs_rg_dist.eps}
+%\caption{Separation distance between robot and leader for both strategies [m].}
+%\label{fig:distance_sim}
+%\end{figure}
 
@@
   \subfloat[a\label{1a}]{%
-       \includegraphics[width=0.45\linewidth]{images/cap_ft_vs_rg_dist.eps}}
+       \includegraphics[width=0.45\linewidth]{images/ft_a2_n1500_cv15_cr15.eps}}
     \hfill
   \subfloat[b\label{1b}]{%
-        \includegraphics[width=0.45\linewidth]{images/cap_ft_vs_rg_dist.eps}}
+        \includegraphics[width=0.45\linewidth]{images/rgv_cv20_cr20_dr10.eps}}
     \\
   \subfloat[c\label{1c}]{%
-        \includegraphics[width=0.45\linewidth]{images/cap_ft_vs_rg_dist.eps}}
+        \includegraphics[width=0.45\linewidth]{images/ft_cv15_cr10_leader_robot_speed.eps}}
     \hfill
   \subfloat[d\label{1d}]{%
-        \includegraphics[width=0.45\linewidth]{images/cap_ft_vs_rg_dist.eps}}
-  \caption{(a), (b) Some examples from CIFAR-10 \cite{4}. The objects in     
-        single-label images are usually roughly aligned.(c),(d) However, the 
-        assumption of object alignment is not valid for multi-label
-        images. Also note the partial visibility and occlusion
-        between objects in the multi-label images.}
+        \includegraphics[width=0.45\linewidth]{images/rg_cv20_cr20_bv10_dtoff10_leader_robot_speed.eps}}
+    \\
+  \subfloat[d\label{1e}]{%
+        \includegraphics[width=0.85\linewidth]{images/ft_vs_rg_dist.eps}}
+  \caption{Simulation Results: Trajectories of the leader and follower for Follow-The-Thread (a) and Rotate-And-Go (b) . 
+        Speed profiles of the leader and follower for Follow-The-Thread (c) and Rotate and Go (d). 
+        (e) Separation distance between robot and leader for both strategies [m].}
+  \label{fig:simulationresults} 
+\end{figure}
+
+Results for the real world experiment can be seen on Figures 3 and 4.
+
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[width=8cm]{images/ft1cap.eps}
+%\caption{Trajectories for the Follow-The-Thread control strategy.}
+%\label{fig:distance_sim}
+%\end{figure}
+%
+%
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[width=8cm]{images/rg2cap.eps}
+%\caption{Distance between robot and leader [m]}
+%\label{fig:distance_sim}
+%\end{figure}
+%
+%
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[width=8cm]{images/ft1cap_leader_robot_speed.eps}
+%\caption{Follow the thread speed profiles}
+%\label{fig:distance_sim}
+%\end{figure}
+%
+%
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[width=8cm]{images/rg2cap_leader_robot_speed.eps}
+%\caption{Rotate and Go speed profiles}
+%\label{fig:distance_sim}
+%\end{figure}
+%
+%
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[width=8cm]{images/cap_ft_vs_rg_dist.eps}
+%\caption{Rotate and Go speed profiles}
+%\label{fig:distance_sim}
+%\end{figure}
+
+\begin{figure} 
+    \centering
+  \subfloat[a\label{1a}]{%
+       \includegraphics[width=0.45\linewidth]{images/ft1cap.eps}}
+    \hfill
+  \subfloat[b\label{1b}]{%
+        \includegraphics[width=0.45\linewidth]{images/rg2cap.eps}}
+    \\
+  \subfloat[c\label{1c}]{%
+        \includegraphics[width=0.45\linewidth]{images/ft1cap_leader_robot_speed.eps}}
+    \hfill
+  \subfloat[d\label{1d}]{%
+        \includegraphics[width=0.45\linewidth]{images/rg2cap_leader_robot_speed.eps}}
+    \\
+  \subfloat[d\label{1e}]{%
+        \includegraphics[width=0.85\linewidth]{images/cap_ft_vs_rg_dist.eps}}
+  \caption{Experimentation Results (a), (b) Trajectories of the leader and follower for Follow-The-Thread (a) and Rotate-And-Go (b) . 
+        (c), (d) Speed profiles of the leader and follower for Follow-The-Thread (c) and Rotate and Go (d). 
+        (e) Separation distance between robot and leader for both strategies [m].}
   \label{fig1} 
@@
 \hline
-$c_v$ & $c_{\alpha}$ & Max path deviation [m] & Area under \textit{npd} curve \\
+$c_v$ & $c_{\alpha}$ & m.p.d. [m] & Area under \textit{npd} curve \\
 \hline
@@
 \end{tabular}
-\captionof{table}{Normal path deviation and maximum path deviation in motion capture experiments using Follow the Thread.}
+\captionof{table}{Max path deviation m.p.d. and area under normal path deviation in motion capture experiments using Follow the Thread.}
 \label{tab:cap_ft_npd_table}
@@
 \hline
-$c_v$ & $c_r$ & $Dt_{off}$ & Max path deviation [m] & Area under \textit{npd} curve \\
+$c_v$ & $c_r$ & $Dt_{off}$ & m.p.d. [m] & Area under \textit{npd} curve \\
 \hline
@@
 \end{tabular}
-\captionof{table}{Normal path deviation and maximum path deviation in motion capture experiments using Rotate and Go.}
+\captionof{table}{Max path deviation m.p.d. and area under normal path deviation in motion capture experiments using Rotate and Go.}
 \label{tab:cap_rg_npd_table}
@@
 Firstly, the need to add some mechanism for obstacle avoidance.
-They mentioned the need to have sensor to detect if the robot was about to hit something (specially the patient), and stop immediately, apart from what the control strategy indicated.
+They mentioned the need to have sensors to detect if the robot was about to hit something (specially the patient), and stop immediately, apart from what the control strategy indicated.
 Secondly, they recognized that some patients have very weak stability, and might fall down or take a step back, towards the robot, so it should be able to automatically move away from the patient, in order not to become another obstacle for him or her.
-"Tener en cuenta la distancia a la que se encuentra el robot por si el paciente llega a dar un paso atrás - para evitar riesto de caída'" and "Estaría bueno pensar en algún método de emergencia por caída'' were two quotes related to this issues, taken from the surveys.
 
 In order to have more information for the next steps in the development of the robot, we asked for their advise to design the mechanism to attach the threads to the patient being followed.
-Two ideas were proposed: a belt strapped to the patient waist, or a clasp that would also to the clothes of the patient, also near its waistline.
-The waist is a good attachment point, since it is relatively stable when the patient moves, compared to its hands or legs, that may make sudden movements and confuse the robot sensors.
+Two ideas were proposed: a belt strapped to the patient waist, or a clasp tied to the clothes of the patient, also near its waistline.
+The waist is a good attachment point, since it is relatively more stable when the patient moves, compared to its hands or legs, that may make sudden movements and confuse the robot sensors.
 
@@
 
-The goal of this work is to verify if a tethered robot to implement a following scheme on a patient during a pulmonary rehabilitation procedure is too simplistic to be a factual implementation.
-
-From the practical experiments we showed that the both algorithms gave reasonable following performance in task of following the leader along a lemniscate-shaped path.
-This research is that this performance was achieved using very low cost components.
-It was proven that a reasonable following behaviour could be accomplished with a simple mechanism, a characteristic that significantly lowers the price of a device like this, putting it within reach of many medical institutions.
+From the practical experiments, it is verified that the both algorithms gave reasonable following performance in the task of following the leader along a lemniscate-shaped path.
+This reasonable following behaviour is accomplished with a simple mechanism, a characteristic that significantly keep the price of the device low, putting it within reach of many medical institutions.
 
@@
 
-We understood that no metric is enough to evaluate how appropriate the robotic vehicle can be for a real life scenario.
-Therefore, we gathered feedback from healthcare professionals from ALPI, who provided us with invaluable feedback to evaluate our solution.
-Over all, they also valued the Follow and Turn strategy as being the safer and more effective one.
-Most importantly, they also validated the research made and approved the direction of the project.
-They proposed a series of improvements and next steps, and after seeing the prototype in action, believed that we were close to a real, working solution.
-
-It is essential to involve stakeholders such as patients, doctors, nurses, and any other professionals involved in the rehabilitation process early in the design roadmap.
-They are the ones who understand the problem better than anyone else, and will be the end users of a product like this.
-
-The staff at ALPI was enthusiastic about helping in the development of the robot, and their encouragement and support were the key reasons we could take this project forward.
-This work not only made progress from a technical standpoint, but could also become the foundation of a long-term relationship of collaboration and research between ALPI and ITBA.
-This is also the start of a long-term project at ITBA.
-One that can go on to become a complete, real world solution that can improve the quality of life of many people through the use of technology and engineering.
+Insightful feedback is gathered from healthcare professionals from ALPI, who provided invaluable data to evaluate the solution.
+Over all, they highlight the Follow and Turn strategy as being the safer and more effective one.
+Most importantly, they also validated the research and were enthusiast about the direction of the project.
+They proposed a series of improvements and next steps after seeing the prototype in action.
+
+As described in the beginning, It is essential to involve stakeholders such as patients, doctors, nurses, and any other professionals involved in the rehabilitation process early in the design roadmap.
+They are the ones who understand the problem better than anyone else, and will be the end users of any developed product, as long as it is useful for them.
+
+The staff at ALPI was enthusiastic about helping in the development of the robot, and their encouragement and support are key reasons to take this project forward.
+There are yet many challenges ahead, but these initial results are promising about the viability to develop a real world solution that can improve the quality of life of many people through the use of technology and engineering.
 
 \subsection{Future Work}
-A following mechanism has been designed and implemented in a hardware prototype.
+%A following mechanism has been designed and implemented in a hardware prototype.
 The mechanism, along with the two control algorithms, has shown promise to be a viable solution to implement in a functioning robot that ALPI, or any other medical institution, can use in their every day exercises with COPD patients.
 
-The next steps for this project would be to take this prototype and build a bigger version, with motors powerful enough to carry a real oxygen tank.
-
-Further research is also needed to develop a full fledged solution: 
+The next steps for this project is to scale the design to be able to carry safely the oxygen tank.
+
+Further research is needed to develop a full fledged solution: 
 \begin{itemize}
@@
 In our tests, the small electric motors in each reel got very hot after just 20 minutes of use.
-The algorithm we designed to use them more efficiently (\ref{alg:active_spring}) was enough to keep them in manageable temperatures (under $70^0C$), which was enough for testing the prototype, but would not be appropriate for a useful robot.
+The active reel algorithm is enough to keep them in manageable temperatures (under $70^0C$), which is enough for testing the prototype, but it will not be appropriate for a useful robot.
 Not only because the heat produced can be dangerous, but also because the motors can overheat and break their insulation in regular use.
@@
 How to communicate the state of the robot to the operator, how to control and manipulate the robot in an effective and user-friendly way.
-    \item Safety measures to keep the patient and the doctor or nurse safe when using the robot.
-Not only safe from the robot movement, but also from its electronic components, which can be dangerous when used near the pure oxygen tank if they are not protected properly.
-    \item Investigate and design mechanisms for obstacle avoidance.
-This need was emphasized by the doctors from ALPI.
+    \item Safety measures to keep the patient and the care-giver safe when using the robot. Not only safe from the robot movement, but also from its electronic components.
+    \item An obstacle avoidance subsystem.
+This necessity is emphasized by the personnel from ALPI.
 The robot should have mechanisms in place to deal with emergency situations, and under no circumstance it can hit the patient or the doctor operating it.
@@
 This project is part of a joint collaboration between ALPI organization and the ITBA University.
-Authors would like to thank thoughtfully for their initiative, support given by the ALPI organization.
-Specially to Natalia Nerina Meda, Soledad Suriá, Eduardo Etcheverry and Sergio Carlos Franco, for their invaluable help and enthusiasm to move this project forward.
+Authors would like to thank thoughtfully for the initiative and support given by the ALPI organization, and to the Director of Rehabilitation Technology Dra.
+Mercedes Molinuevo.
+Authors would like to thank Natalia Nerina Meda, Soledad Suriá, Eduardo Etcheverry and Sergio Carlos Franco, for their invaluable help and enthusiasm to move this project forward.
 

### 2020-04-14 alpibot.tex
Adjustments on the abstract.

@@
 During Pulmonary Rehabilitation procedures patients require external oxygen assistance.
-The oxygen tank cannot be carried by the patient and some external assistance is required.
-In this work a basic robot follower is proposed to carry the oxygen tank based on a differential tethered scheme.
-Two algorithms are proposed in a very basic configuration.
+The oxygen tank cannot be carried by the patient and external assistance is required.
+This work provides an initial study on the controlling mechanism of a differential tethered robot that implements a leader-follower configuration to carry the oxygen tank for Pulmonary Rehabilitation procedures.
+Two alternative control strategies are proposed.
+Results are simulated and tested on a real prototype in a Motion Capture System.
 \end{abstract}

### 2020-04-15 alpibot.tex
Fixed Esteban's name.

@@
 
-\author{Bianchi Luciano, Buzniak Esteban, Ramele~Rodrigo,~\IEEEmembership{Member,~IEEE,}
+\author{Bianchi Luciano, Buzniak Esteban Alejandro, Ramele~Rodrigo,~\IEEEmembership{Member,~IEEE,}
         and Santos~Juan~Miguel% <-this % stops a space
-\thanks{E. Buzniak, R. Ramele  and J.M.Santos are with the CiC Laboratory at the Department
+\thanks{E. A. Buzniak, R. Ramele  and J.M.Santos are with the CiC Laboratory at the Department
 of  Computer Engineering, Instituto Tecnológico de Buenos Aires (ITBA), Ciudad de Buenos Aires,  Argentina e-mail: rramele@itba.edu.ar}% <-this % stops a space
-\thanks{Manuscript received April 19, 2020; revised August 26, 2015.}}
+\thanks{Manuscript received April 19, 2020; revised August 26, 2020.}}
 
@@
 % The paper headers
-\markboth{IEEE Transactions on Biomedical Engineering}%
+\markboth{IEEE Latin American Transactions}%
 {Bianchi \MakeLowercase{\textit{et al.}} Controller Design for a Tethered Follower Robot for Pulmonary Rehabilitation}
@@
 \begin{abstract}
-Patients that suffer Chronic Obstructive Pulmonary Disease (COPD) underwent a procedure called Pulmonary Rehabilitation that helps them to improve disease prognosis.
-During Pulmonary Rehabilitation procedures patients require external oxygen assistance.
+Patients that suffer Chronic Obstructive Pulmonary Disease (COPD) undergo a procedure called Pulmonary Rehabilitation that helps them to improve disease prognosis.
+During Pulmonary Rehabilitation procedures patients require oxygen assistance.
 The oxygen tank cannot be carried by the patient and external assistance is required.
@@
 During these procedures patients underwent controlled walking activities and physical exercises under the supervision of a physical therapist.
-These patients present a severe low saturation illness and they require effective oxygen supply, particularly when performing physical activity~\cite{Celli2014}.
+These patients present a severe low oxygen saturation illness and they require effective oxygen supply, particularly when performing physical activity~\cite{Celli2014}.
 Hence, patients require to carry with them an oxygen tank for the oxygenotherapy assistance.
@@
 \subsection{Solution Design}
-
-The proposed solution is a Differential-Wheeled Robot (DWR) tethered to the followed subject with two threads ending at a single point attached to the subject waist, back or hand.
-In the same axis as the two front wheels, the robot has two reels separated by a certain length from which these threads come.
-As the subject moves away from the robot, the reels release thread so that the patient does not physically drag the vehicle.
-When the opposite happens, and the vehicle gets closer to the subject, an active spring mechanism driven by electric motors move each reel to retract the thread.
-The threads need to be tense at all times so that the encoders in each reel can be used to continuously measure the distance between the subject and the reel as devised in Figure~\ref{fig:components}.
-
-Encoders in each reel measure the difference in length of each thread compared to its initial position.
-This difference in length is the input for the control algorithm.
-Using the encoder, the difference in length of the thread can be measured with this formula:
-
-\begin{equation}
-\label{eq:encoder_len}
-L_d = pulses . \frac{2\pi r}{ppr}
-\end{equation}{}
-
-The variable $r$ is the radius of the reel and $ppr$ is the pulses per revolution (resolution) of the encoder.  The initial position of the threads is configurable.
-
-\subsection{Hardware}
-
-Frames are constructed from aluminum extrusions produced by Makeblock (Shenzhen, China).
 
@@
 
-Motors are in-wheel configuration and they provide optical encoding.
-A microcontroller is used to implement motor controlling and encoder processing.
-This is based on the popular Arduino platform.
-On top of it the Adafruit Shield v2.
-This Arduino bridge can be used to control simultaneously four DC motors.
-
-The microcontroller is connected through serial connection to the main SBC RaspberryPi Model 3B+.
-The main controlling algorithm run on this board, it connects to wifi and provides telemetry and the abiilty to receive remote commands by means a very simple UDP command interface.
-
-Reels were designed from PVC extrusions based on figure.  They are glued to regular DC motors scavenged from old compact discs.  
-%Hardware specifications are described on table \ref{tab:specifications}.
+The proposed solution is a Differential-Wheeled Robot (DWR) tethered to the followed subject with two threads ending at a single point attached to the subject waist, back or hand.
+In the same axis as the two front wheels, the robot has two reels separated by a certain length from which these threads come.
+As the subject moves away from the robot, the reels release thread so that the patient does not physically drag the vehicle.
+When the opposite happens, and the vehicle gets closer to the subject, an active spring mechanism driven by electric motors move each reel to retract the thread.
+The threads need to be tense at all times so that the encoders in each reel can be used to continuously measure the distance between the subject and the reel as devised in Figure~\ref{fig:components}.
+
+% El detalle de los encoders se puede eliminar, es info común.
+Encoders in each reel measure the difference in length of each thread compared to its initial position.
+This difference in length is the input for the control algorithm.
+Using the encoder, the difference in length of the thread can be measured with this formula:
+
+\begin{equation}
+\label{eq:encoder_len}
+L_d = pulses . \frac{2\pi r}{ppr}
+\end{equation}{}
+
+The variable $r$ is the radius of the reel and $ppr$ is the pulses per revolution (resolution) of the encoder.  The initial position of the threads is configurable.
+
+\subsection{Hardware}
 
@@
 \includegraphics[width=8cm]{images/alpibot1.jpg}
-\caption{This robot prototype.}
-\label{fig:performance}
+\caption{Side-view of the robot prototype.  A differential wheel, the reel and the rear free castor wheel can be observed from the picture.}
+\label{fig:alpibot1}
 \end{figure}
+
+Frames are constructed from aluminum extrusions produced by Makeblock (Shenzhen, China).
+The prototype is a three-wheeled robot with two frontal differential drive wheels and a free castor wheel as a third point of contact on the back, as can be seen on Figure~\ref{fig:alpibot2}.
+
+Two motors Makeblock Optical Encoder Motor-25 9V/86 rpm are used on in-wheel configuration providing optical encoding.
+A microcontroller Arduino (Arduino LLC, Italy) Mega 2560 is used to implement the control loop and to provide encoder processing.
+On top of it an Adafruit (Adafruit, New York City) Motor Shield v2 bridge is used to drive the four DC motors, one for each wheel and one on each reel.
+The Arduino board is also connected to a Single Board Computer Raspberry Pi (Raspberry Pi Foundation, United Kingdom) 3B+ through serial connection on the USB port.
 
@@
 \subfloat[Case I]{\includegraphics[width=15cm]{images/alpibot2.jpg}\label{subject8}}
-\caption{Robot prototype built to evaluate the real world scenario of the following mechanisms for the tethered robot.}
-\label{fig:p300templates}
+\caption{Top view of the robot prototype.
+The SBC is shown on the center, alongside the Arduino Mega board.
+Both reels can be seen on the same vertical plane of the wheel axis.
+A power bank (white) is located on one side of the robot, and on the rear part, the motor power battery (black) is located.}
+\label{fig:alpibot2}
 \end{figure*}
+
+The Raspberry (Raspberry Pi Foundation, United Kingdom) board connects to a WiFi network and can receive remote commands to control the robot.
+It also broadcasts the telemetry data to any listening devices on the network.
+The control algorithms run in this board, which continuously communicates with the Arduino board to receive sensor data and to issue commands to move the motors or retract the reels.
+The algorithms are programmed in python 3.7, which allows the exact same code to run both the simulated and real world prototype.
+
+%The microcontroller is connected through serial connection to the main SBC RaspberryPi Model 3B+.
+The main controlling algorithm run on this board, it connects to wifi and provides telemetry and the abiilty to receive remote commands by means a very simple UDP command interface.
+
+The two reels are designed from PVC extrusions and are shown on Figure~\ref{fig:alpibot2}.
+They are attached to regular FA-12350 DC motors scavenged from old compact discs.
+Each reel is axially locked to inexpensive Ky-040 rotary encoder which provides around 20 pulses per revolution.
+%Hardware specifications are described on table \ref{tab:specifications}.
+
+The prototype can be seen on Figure~\ref{fig:alpibot1} and~\ref{fig:alpibot2}.
+It has two separate batteries, one powering the Raspberry Pi, the Arduino board and the encoder electronics, and the other powering the drive and reel motors.
+The first battery is a commercial power bank with a capacity of 10000 mA$\cdot$h, and an output of 3.1A (over two USB ports) at 5V.
+The motor battery is a set of 10 AAA nickel–metal hydride batteries (1.5V each), with a total output of 15V.
 
@@
 
-Two simple control strategies are proposed and evaluated.
+Two simple algorithmic control strategies are proposed and evaluated.
 The first one is called Follow-the-thread and the second strategy is Rotate-And-Go.
-The Adafruit Shield controller provides a very stable output signal, provided battery are kept above reference.
+%The Adafruit Shield controller provides a very stable output signal, provided battery are kept above reference.
 Hence, wheels motor control is open-loop.
@@
 v_{tar} = c_v (\frac{D_l + D_r}{2} - d_0)
+\label{eq:lemniscata}
 \end{equation}
@@
 $d_0$ is a constant offset. $c_v$ and $c_{\alpha}$ are constant coefficients used for calibration.
+
+\begin{figure}[]
+    \centering
+    \tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt
+
+    \begin{tikzpicture}[x=0.50pt,y=0.50pt,yscale=-1,xscale=1]
+    %uncomment if require: \path (0,2317); %set diagram left start at 0, and has height of 2317
+
+    %Shape: Rectangle [id:dp785217180323701] 
+    \draw   (332,1925.29) -- (347.46,1925.29) -- (347.46,2120.1) -- (332,2120.1) -- cycle ;
+    %Straight Lines [id:da8987873257842873] 
+    \draw [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][line width=0.75]    (300.46,2100.29) -- (117.19,2019.9) ;
+    %Straight Lines [id:da585150688871364] 
+    \draw [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 230; green, 164; blue, 164 }  ,fill opacity=1 ][line width=0.75]    (301.46,1946.29) -- (117.19,2019.9) ;
+    %Shape: Rectangle [id:dp8556526042433679] 
+    \draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ] (300.5,1925) -- (379,1925) -- (379,1960.67) -- (300.5,1960.67) -- cycle ;
+    %Shape: Ellipse [id:dp024503202773356936] 
+    \draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ] (101.01,2019.9) ..
+controls (101.01,2010.53) and (108.25,2002.94) ..
+(117.19,2002.94) ..
+controls (126.12,2002.94) and (133.37,2010.53) ..
+(133.37,2019.9) ..
+controls (133.37,2029.28) and (126.12,2036.87) ..
+(117.19,2036.87) ..
+controls (108.25,2036.87) and (101.01,2029.28) ..
+(101.01,2019.9) -- cycle ;
+    %Shape: Rectangle [id:dp4521863424396907] 
+    \draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ] (299.5,2084.5) -- (378,2084.5) -- (378,2120.17) -- (299.5,2120.17) -- cycle ;
+    %Rounded Rect [id:dp9798921511906625] 
+    \draw  [draw opacity=0][fill={rgb, 255:red, 184; green, 233; blue, 134 }  ,fill opacity=1 ] (327,2059.58) ..
+controls (327,2058.16) and (328.16,2057) ..
+(329.58,2057) -- (350.53,2057) ..
+controls (351.96,2057) and (353.11,2058.16) ..
+(353.11,2059.58) -- (353.11,2067.34) ..
+controls (353.11,2068.76) and (351.96,2069.92) ..
+(350.53,2069.92) -- (329.58,2069.92) ..
+controls (328.16,2069.92) and (327,2068.76) ..
+(327,2067.34) -- cycle ;
+    %Rounded Rect [id:dp969028519585216] 
+    \draw  [draw opacity=0][fill={rgb, 255:red, 184; green, 233; blue, 134 }  ,fill opacity=1 ] (327,1977.58) ..
+controls (327,1976.16) and (328.16,1975) ..
+(329.58,1975) -- (350.53,1975) ..
+controls (351.96,1975) and (353.11,1976.16) ..
+(353.11,1977.58) -- (353.11,1985.34) ..
+controls (353.11,1986.76) and (351.96,1987.92) ..
+(350.53,1987.92) -- (329.58,1987.92) ..
+controls (328.16,1987.92) and (327,1986.76) ..
+(327,1985.34) -- cycle ;
+    %Straight Lines [id:da8580549640030513] 
+    \draw [color={rgb, 255:red, 128; green, 128; blue, 128 }  ,draw opacity=1 ][line width=1.5]  [dash pattern={on 1.69pt off 2.76pt}]  (340.06,2063.46) -- (496.67,2020.83) ;
+    %Shape: Circle [id:dp1613016595870852] 
+    \draw  [draw opacity=0][fill={rgb, 255:red, 155; green, 155; blue, 155 }  ,fill opacity=1 ] (488.83,2020.83) ..
+controls (488.83,2016.51) and (492.34,2013) ..
+(496.67,2013) ..
+controls (500.99,2013) and (504.5,2016.51) ..
+(504.5,2020.83) ..
+controls (504.5,2025.16) and (500.99,2028.67) ..
+(496.67,2028.67) ..
+controls (492.34,2028.67) and (488.83,2025.16) ..
+(488.83,2020.83) -- cycle ;
+    %Straight Lines [id:da10232028406547378] 
+    \draw [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][line width=0.75]    (338.75,1943.33) -- (402.58,1942.95) ;
+    \draw [shift={(405.58,1942.93)}, rotate = 539.65] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;
+    %Straight Lines [id:da5543519847675118] 
+    \draw [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][line width=0.75]    (337.75,2103.33) -- (401.58,2102.95) ;
+    \draw [shift={(404.58,2102.93)}, rotate = 539.65] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;
+    %Straight Lines [id:da5830729933065488] 
+    \draw [color={rgb, 255:red, 128; green, 128; blue, 128 }  ,draw opacity=1 ][line width=1.5]  [dash pattern={on 1.69pt off 2.76pt}]  (340.06,1981.46) -- (496.67,2020.83) ;
+    %Straight Lines [id:da27552399356884627] 
+    \draw [color={rgb, 255:red, 128; green, 128; blue, 128 }  ,draw opacity=1 ][line width=1.5]  [dash pattern={on 1.69pt off 2.76pt}]  (340.06,1981.46) -- (488.5,1937) ;
+    %Straight Lines [id:da06167712807711301] 
+    \draw [color={rgb, 255:red, 128; green, 128; blue, 128 }  ,draw opacity=1 ][line width=1.5]  [dash pattern={on 1.69pt off 2.76pt}]  (340.06,2063.46) -- (543.5,1918.83) ;
+    %Straight Lines [id:da887785453759708] 
+    \draw [color={rgb, 255:red, 74; green, 144; blue, 226 }  ,draw opacity=1 ][line width=2.25]    (488.5,1937) -- (543.5,1918.83) ;
+    %Straight Lines [id:da7211287180458504] 
+    \draw [color={rgb, 255:red, 74; green, 144; blue, 226 }  ,draw opacity=1 ][line width=2.25]    (469.5,1973) -- (543.5,1918.83) ;
+    %Shape: Circle [id:dp7802267148020259] 
+    \draw  [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ,fill opacity=1 ] (535.67,1918.83) ..
+controls (535.67,1914.51) and (539.17,1911) ..
+(543.5,1911) ..
+controls (547.83,1911) and (551.33,1914.51) ..
+(551.33,1918.83) ..
+controls (551.33,1923.16) and (547.83,1926.67) ..
+(543.5,1926.67) ..
+controls (539.17,1926.67) and (535.67,1923.16) ..
+(535.67,1918.83) -- cycle ;
+
+    % Text Node
+    \draw (524,1960) node  [font=\large,color={rgb, 255:red, 74; green, 144; blue, 226 }  ,opacity=1 ]  {$ \begin{array}{l}
+    D_{r}\\
+    \end{array}$};
+    % Text Node
+    \draw (510,1910) node  [font=\large,color={rgb, 255:red, 74; green, 144; blue, 226 }  ,opacity=1 ,rotate=-358.27]  {$ \begin{array}{l}
+    D_{l}\\
+    \end{array}$};
+    % Text Node
+    \draw (496,2042) node  [color={rgb, 255:red, 128; green, 128; blue, 128 }  ,opacity=1 ]  {$subject\ start$};
+    % Text Node
+    \draw  [draw opacity=0][fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ]  (554.5,1904) -- (611.5,1904) -- (611.5,1928) -- (554.5,1928) -- cycle  ;
+    \draw (583,1916) node    {$subject$};
+    % Text Node
+    \draw (390.58,1925.93) node [anchor=west] [inner sep=0.75pt]  [font=\normalsize,color={rgb, 255:red, 0; green, 0; blue, 0 }  ,opacity=1 ]  {$vleft_{tar}$};
+    % Text Node
+    \draw (407.58,2100.93) node [anchor=west] [inner sep=0.75pt]  [font=\normalsize,color={rgb, 255:red, 0; green, 0; blue, 0 }  ,opacity=1 ]  {$vright_{tar}$};
+
+    \end{tikzpicture}
+    \caption{Following mechanism parameters}
+    \label{fig:follow_params}
+\end{figure}{}
 
@@
 
-In order to standardize the procedure~\cite{Sprunk2016}, the \textit{Lemniscate of Gerono} is used as desired trajectory, a curve shaped like an $\infty$ symbol, described by the Equation~\ref{eq:lemniscate}:
+In order to standardize the procedure~\cite{Sprunk2016}, the \textit{Lemniscate of Gerono} is used as desired trajectory, a curve shaped like an $\infty$ symbol, described by the Equations~\ref{eq:lemniscate}:
 
@@
 Similar curves are also used in other proposed experiments in \cite{Neto2015,Endo2015}.
-Regarding metrics, three are proposed to evaluate the performance.
-They are
+
+Regarding metrics, three are proposed to evaluate the performance.  They are:
 
@@
 The simulation is also useful to study the effects of the different constants in each strategy on the movement of the robot.
+The leader starts at the midpoint of the trajectory and completes a full complete circuit getting back to the initial position, while the robot follows its track.
+Four different configuration sets of $c_v$ and  $c_{\alpha}$ were tested for Follow-The-Thread, while 6 different configurations were tested for Rotate-And-Go.
 
@@
 A motion capture system is used to track the movement of a human leader along a predetermined path.
-The motion tracking system consists of an array of 16 OptiTrack (NaturalPoint Inc, Oregon, US) Flex 3 cameras, which measure the position of reflective markers with an accuracy of $\pm1$ a sampling rate of 100 Hz.
+The motion tracking system consists of an array of 16 OptiTrack (NaturalPoint Inc, Oregon, US) Flex 3 cameras, which measure the position of reflective markers with an accuracy of $\pm1$ [cm] at sampling rate of 100 Hz.
 The calibration and data collection was made using the Motive motion capture software.
@@
 A third marker was placed in his hand, using a glove.
-The lemniscate of Gerono, used in the simulation, was drawn in the floor (\ref{fig:cap_lab_lemniscate}), and the human leader tried to move his hand following this shape as close as possible, with stable speed.
-
-The shape was drawn according to the shape described in equation \ref{eq:lemniscate_of_gerono}, using $a = 2$, in meters.
-
-The three markers allowed us to measure the trajectory of the robot and the leader, and then obtain the same metrics calculated in the simulation.
+The lemniscate of Gerono, used in the simulation, was drawn in the floor, and the human leader tried to move his hand following this shape as close as possible, with stable speed.
+The shape was marked according to the shape described in Equation~\ref{eq:lemniscata}, using $a = 2$ [m].
+
+The three markers allowed us to measure the trajectory of both the robot and the leader, and then obtain the same metrics calculated in the simulation.
+Four experiments were performed for each set of parameter configurations.
+In this case, only two set of configuration were tested for each strategy.
 
@@
   \subfloat[a\label{1a}]{%
-       \includegraphics[width=0.45\linewidth]{images/ft_a2_n1500_cv15_cr15.eps}}
-    \hfill
+       \includegraphics[width=0.49\linewidth]{images/ft_a2_n1500_cv15_cr15.eps}}
+    \hspace*{-1.5em}
   \subfloat[b\label{1b}]{%
-        \includegraphics[width=0.45\linewidth]{images/rgv_cv20_cr20_dr10.eps}}
+        \includegraphics[width=0.49\linewidth]{images/rgv_cv20_cr20_dr10.eps}}
     \\
+    \vspace*{-1.5em}
   \subfloat[c\label{1c}]{%
-        \includegraphics[width=0.45\linewidth]{images/ft_cv15_cr10_leader_robot_speed.eps}}
-    \hfill
+        \includegraphics[width=0.49\linewidth]{images/ft_cv15_cr10_leader_robot_speed.eps}}
+    \hspace*{-1.5em}
   \subfloat[d\label{1d}]{%
-        \includegraphics[width=0.45\linewidth]{images/rg_cv20_cr20_bv10_dtoff10_leader_robot_speed.eps}}
+        \includegraphics[width=0.49\linewidth]{images/rg_cv20_cr20_bv10_dtoff10_leader_robot_speed.eps}}
     \\
+    \vspace*{-2.5em}
   \subfloat[d\label{1e}]{%
-        \includegraphics[width=0.85\linewidth]{images/ft_vs_rg_dist.eps}}
+        \includegraphics[width=0.99\linewidth]{images/ft_vs_rg_dist.eps}}
   \caption{Simulation Results: Trajectories of the leader and follower for Follow-The-Thread (a) and Rotate-And-Go (b) . 
@@
 
-Results for the real world experiment can be seen on Figures 3 and 4.
+Results for the real world experiment can be seen on Figure~\ref{fig:realworldresults}.
+Table~\ref{tab:cap_ft_npd_table} show the metrics of the Follow-The-Thread strategy, whereas Table~\ref{tab:cap_rg_npd_table} provides the metrics for the Rotate-And-Go approach.
 
@@
   \subfloat[a\label{1a}]{%
-       \includegraphics[width=0.45\linewidth]{images/ft1cap.eps}}
-    \hfill
+       \includegraphics[width=0.49\linewidth]{images/ft1cap.eps}}
+    \hspace*{-1.5em}
   \subfloat[b\label{1b}]{%
-        \includegraphics[width=0.45\linewidth]{images/rg2cap.eps}}
+        \includegraphics[width=0.49\linewidth]{images/rg2cap.eps}}
     \\
+    \vspace*{-1.5em}
   \subfloat[c\label{1c}]{%
-        \includegraphics[width=0.45\linewidth]{images/ft1cap_leader_robot_speed.eps}}
-    \hfill
+        \includegraphics[width=0.49\linewidth]{images/ft1cap_leader_robot_speed.eps}}
+    \hspace*{-1.5em}
   \subfloat[d\label{1d}]{%
-        \includegraphics[width=0.45\linewidth]{images/rg2cap_leader_robot_speed.eps}}
+        \includegraphics[width=0.49\linewidth]{images/rg2cap_leader_robot_speed.eps}}
     \\
+    \vspace*{-2.5em}
   \subfloat[d\label{1e}]{%
-        \includegraphics[width=0.85\linewidth]{images/cap_ft_vs_rg_dist.eps}}
-  \caption{Experimentation Results (a), (b) Trajectories of the leader and follower for Follow-The-Thread (a) and Rotate-And-Go (b) . 
-        (c), (d) Speed profiles of the leader and follower for Follow-The-Thread (c) and Rotate and Go (d). 
+        \includegraphics[width=0.99\linewidth]{images/cap_ft_vs_rg_dist.eps}}
+  \caption{Experimentation Results: Trajectories of the leader and follower for Follow-The-Thread (a) and Rotate-And-Go (b). 
+        Speed profiles of the leader and follower for Follow-The-Thread (c) and Rotate and Go (d). 
         (e) Separation distance between robot and leader for both strategies [m].}
-  \label{fig1} 
+  \label{fig:realworldresults} 
 \end{figure}
@@
 
-As this project is a joint research between ALPI and ITBA, we invited 4 people from the ALPI staff, 3 doctors and one engineer, to the motion capture laboratory at ITBA, where we made a live demonstration of the robot working and following a moving person.
-
-In the demonstration, we described how the robot worked, how it was built and how to operate it. Both control strategies were explained, along with the main superficial differences between them.
-
-Afterwards, we invited the doctors to use the robot themselves, simulating they were the patient being followed.
-We let the doctors switch between the two control strategies to evaluate both of them, and made different tests, one walking along path drawn in the floor (figure \ref{fig:cap_lab_lemniscate}), and another one walking freely in the space available in the lab.
-We let them use the robot as long as they wanted to, to get a general idea of how it behaved, and how it could be used in the rehabilitation process.
-After all evaluations were finished, we answered questions and discussed various aspects of the vehicle prototype with the ALPI staff, and then proceeded to ask them a survey to document their experience with the robot, get their expert opinion on how the two strategies compared against each other, and what other improvements were needed in order to deliver a fully usable product.
+ALPI is a non-profit civil association located in Buenos Aires, Argentina, that provides neuromotor rehabilitation for pediatric and adult patients.
+It was founded in 1943 with the main focus of treating children with poliomyelitis, and has since expanded to deal with all kinds of related diseases.
+
+Four professional care-givers from ALPI  were invited to test and evaluate the controlling strategy on the prototype.
+A live demonstration of the robot working and following a moving person was performed.
+
+In the demonstration, the robot design was outlined and an explanation was given on how the robot worked, how it was built and how to operate it.
+Both control strategies were expounded, along with the main superficial differences between them.
+
+Afterwards, health care professionals were invited to use the robot themselves, simulating they were the patient being followed.
+They were allowed to switch between the two control strategies to evaluate both of them, and made different tests, one walking along standardized path marked on the floor and another one walking freely along the available space of the Motion Capture Lab .
+They used the robot freely to get a general idea of how it behaved, and how it could be used in the rehabilitation process.
+After all evaluations were finished, various aspects of the vehicle prototype were discussed and then a survey was handed out to document their experience with the robot, get their expert opinion on how the two strategies compared against each other, and what other improvements were needed in order to deliver a fully usable product.
+Survey questions and their averaged numerical evaluations are provided in Table~\ref{tab:alpi_q_table_1}.
 
@@
 
-As described in the beginning, It is essential to involve stakeholders such as patients, doctors, nurses, and any other professionals involved in the rehabilitation process early in the design roadmap.
+As described in the beginning, it is essential to involve stakeholders such as patients, doctors, nurses, and any other professionals involved in the rehabilitation process early in the design roadmap.
 They are the ones who understand the problem better than anyone else, and will be the end users of any developed product, as long as it is useful for them.
@@
 Mercedes Molinuevo.
-Authors would like to thank Natalia Nerina Meda, Soledad Suriá, Eduardo Etcheverry and Sergio Carlos Franco, for their invaluable help and enthusiasm to move this project forward.
+Additionally, authors would also like to offer tremendous gratitude to Natalia Nerina Meda, Soledad Suriá, Eduardo Etcheverry and Sergio Carlos Franco, for the idea of this project and for their invaluable help and enthusiasm to move this project forward.
 

### 2020-04-18 alpibot.tex
Upload of a new version of the pdf.

@@
 On the other hand, several assistance devices for COPD treatments have been proposed.
-Particularly relevant are novel telemedice applications to enhance complementary rehabiliation exercise at home that can track biological markers for patients~\cite{Wu2012}.
+Particularly relevant are novel telemedice applications to enhance complementary rehabiliation exercise at home that can track biological markers for patients~\cite{Yang2018,Wu2012}.
 The work presented here follows the line established by \cite{Endo2015}.
@@
 
-The Raspberry (Raspberry Pi Foundation, United Kingdom) board connects to a WiFi network and can receive remote commands to control the robot.
-It also broadcasts the telemetry data to any listening devices on the network.
+The Raspberry Pi board connects to a WiFi network and can receive remote commands to control the robot.
+It also broadcasts telemetry data to any listening devices on the network.
 The control algorithms run in this board, which continuously communicates with the Arduino board to receive sensor data and to issue commands to move the motors or retract the reels.
@@
 
-\subsection{Software Components}
+%\subsection{Software Components}
 
@@
 It is based on the idea that the difference in length between both threads provides a quasi-linear function of the relative angle between the subject and the vehicle orientation and, additionally, the mean of both thread length allows to approximate the relative distance between the robot and the subject.
+They are described by Equations~\ref{eq:lemniscata}, \ref{eq:lemniscata2} and~\ref{eq:lemniscata3}.
 
@@
 vleft_{tar} = v_{tar} + c_{\alpha} (D_l - D_r)
+\label{eq:lemniscata2}
 \end{equation}
@@
 vright_{tar} = v_{tar} - c_{\alpha} (D_l - D_r)
+\label{eq:lemniscata3}
 \end{equation}
+
+\noindent where  $c_v$ and $c_{\alpha}$ are constant coefficients used for calibration.
+As shown in Figure~\ref{fig:follow_params}, reel encoder information is $D_l$ and $D_r$ for the left and right reel, and finally $vleft_{tar}$ are $vright_{tar}$ are the power values that are directly used to drive power to each wheel motor.
 
@@
 
-$d_0$ is a constant offset. $c_v$ and $c_{\alpha}$ are constant coefficients used for calibration.
+\noindent where $d_0$ is a constant offset that is used to customized the length of the thread where the robot does not move.
 
@@
 
-1) Rotating the vehicle around the center point of the axis that connects its front wheels in order to aim at the subject.
-
-2) Go forward in a straight line until the vehicle is at the expected distance to the subject or the vehicle is not moving in the direction of the subject.
-
-The algorithm is detailed in Algorithm~\ref{alg:rotateandgoalg}.
+\begin{itemize}
+\item Rotating the vehicle around the center point of the axis that connects its front wheels in order to aim at the subject.
+\item Go forward in a straight line until the vehicle is at the expected distance to the subject or the vehicle is not moving in the direction of the subject.
+\end{itemize}
+
+The procedure is detailed in Algorithm~\ref{alg:rotateandgoalg}.
 The variable $V_r$ is the speed at which the vehicle will rotate on its axis, and $V_f$ is the speed at which the vehicle will move forward once it can move on the subject's direction.
+The algorithm requires three parameters, $c_v$ and $c_r$ that regulates the coefficients for the forward and rotation movement, and an additional parameter $Dt_off$ that regulates the sensibility of the rotation movement.
 
@@
 
-A model of the proposed design was first built on Webots application.
+A model of the proposed design was first built on Webots~\cite{Michel2004} simulator.
 The threading mechanism was implemented using virtual threads~\cite{Rekleitis2001}.
@@
 A third marker was placed in his hand, using a glove.
-The lemniscate of Gerono, used in the simulation, was drawn in the floor, and the human leader tried to move his hand following this shape as close as possible, with stable speed.
+The lemniscate of Gerono, used in the simulation, was marked on the floor, and the human leader tried to move his hand following this shape as close as possible, with stable speed.
 The shape was marked according to the shape described in Equation~\ref{eq:lemniscata}, using $a = 2$ [m].
 
-The three markers allowed us to measure the trajectory of both the robot and the leader, and then obtain the same metrics calculated in the simulation.
+The three markers allowed to measure the trajectory of both the robot and the leader, and then obtain the same metrics calculated in the simulation.
 Four experiments were performed for each set of parameter configurations.

### 2020-04-21 alpibot.tex
Upload a new versio of the paper, with the Discussions sections completed.

@@
 \begin{IEEEkeywords}
-robotics, tethered, COPD
+SAR, IoRT, tethered, leader-follower, pulmonary rehabilitation, COPD
 \end{IEEEkeywords}
@@
 This pathology presents an eskeletomuscular atrophy~\cite{Kocsis2016,Wu2012} and in order to carve these after effects a Pulmonary Rehabilitation procedure is a viable treatment for these patients.
-During these procedures patients underwent controlled walking activities and physical exercises under the supervision of a physical therapist.
+During these procedures patients undergo controlled walking activities and physical exercises under the supervision of a physical therapist.
 These patients present a severe low oxygen saturation illness and they require effective oxygen supply, particularly when performing physical activity~\cite{Celli2014}.
@@
 
-On the other hand, several assistance devices for COPD treatments have been proposed.
+At the same time, several assistance devices for COPD treatments have been proposed.
 Particularly relevant are novel telemedice applications to enhance complementary rehabiliation exercise at home that can track biological markers for patients~\cite{Yang2018,Wu2012}.
@@
 This difference in length is the input for the control algorithm.
-Using the encoder, the difference in length of the thread can be measured with this formula:
+Using the encoder, the difference in length of the thread can be measured with Equation~\ref{eq:encoder_len}.
 
@@
 
-The variable $r$ is the radius of the reel and $ppr$ is the pulses per revolution (resolution) of the encoder.  The initial position of the threads is configurable.
+\noindent where the variable $r$ is the radius of the reel and $ppr$ is the pulses per revolution (resolution) of the encoder.  The initial position of the threads is configurable.
 
@@
 On top of it an Adafruit (Adafruit, New York City) Motor Shield v2 bridge is used to drive the four DC motors, one for each wheel and one on each reel.
-The Arduino board is also connected to a Single Board Computer Raspberry Pi (Raspberry Pi Foundation, United Kingdom) 3B+ through serial connection on the USB port.
+The Arduino board is also connected to a Single Board Computer (SBC) Raspberry Pi (Raspberry Pi Foundation, United Kingdom) 3B+ through serial connection on the USB port.
 
@@
 
-The Raspberry Pi board connects to a WiFi network and can receive remote commands to control the robot.
+The SBC connects to a WiFi network and can receive remote commands to control the robot.
 It also broadcasts telemetry data to any listening devices on the network.
@@
 \begin{itemize}
-    \item \textit{Normal path deviation}: the subject trajectory is divided into small segments and then the normal distance to the robot trajectory is calculated for each of those segments.
+    \item \textit{Normal path deviation, n.p.d.}: the subject trajectory is divided into small segments and then the normal distance to the robot trajectory is calculated for each of those segments.
 Path deviation is relevant to evaluate how closely the robot mimics the leader path, which is the ultimate goal of the robotic vehicle.
-    \item \textit{Maximum path deviation}: the maximum \textit{normal path deviation} registered during an experiment.
-    \item \textit{Total path deviation}: The area under the curve resulting from the the \textit{normal path deviation} over the length travelled by the leader.
-    \item \textit{Robot-leader distance}: The euclidean distance between the robot and the leader, at any point in time.
-This is particularly important since the robot has a limited amount of thread available, so if the leader uses all the available thread, it will start dragging the robot and damaging the following mechanism.
-It is a scenario that must not happen under any circumstance, as it can also be dangerous for a potential patient using the device.
+    \item \textit{Total path deviation, t.p.d.}: The area under the curve resulting from the the \textit{normal path deviation} over the length traveled by the leader.
+    \item \textit{Maximum path deviation, m.p.d.}: the maximum \textit{normal path deviation} registered during an experiment.    
+    \item \textit{Robot-leader distance, r.l.d.}: The euclidean distance between the robot and the leader, at any point in time.
+This is particularly important since the robot has a limited amount of thread available, so if the leader uses all the available thread, it will start dragging the robot and damaging the following mechanis, and overall it may rise the possibility of disconnecting the breathing oxygen cannula.
+This is a scenario that must not happen under any circumstance, as it can also be dangerous for a potential patient using the device.
 \end{itemize}{}
@@
 The threading mechanism was implemented using virtual threads~\cite{Rekleitis2001}.
-The leader traveled according to a predefined path with constant velocity, followiing the lemniscata trajectory.
+The leader traveled according to a predefined path with constant velocity, followiing the lemniscate trajectory.
 
@@
 \caption{Hardware prototype on the motion capture system and a testing subject holding the threads.
-The lemniscata of gerone was marked on the floor.
+The lemniscate of gerone was marked on the floor.
 The subject follows this track on the experiments performed.}
@@
 \hline
-$c_v$ & $c_{\alpha}$ & m.p.d. & Area under \textit{npd} curve \\
+$c_v$ & $c_{\alpha}$ & m.p.d. & t.p.d. \\
 \hline
@@
 \end{tabular}
-\captionof{table}{Max path deviation m.p.d. [m] and Area under the normal path deviation for different Follow the Thread constants.}
+\captionof{table}{Max path deviation m.p.d. [m] and Total path deviation t.p.d. for different Follow the Thread constants.}
 \label{tab:simulationmetricsftt}
@@
 \hline
-$c_v$ & $c_r$ & $Dt_{off}$ & m.p.d. & Area under \textit{npd} curve \\
+$c_v$ & $c_r$ & $Dt_{off}$ & m.p.d. & t.p.d.\\
 \hline
@@
 \end{tabular}
-\captionof{table}{Max path deviation m.p.d. (in [m]) and area under normal path deviation for different Rotate and Go constants.}
+\captionof{table}{Max path deviation m.p.d. (in [m]) and total path deviation t.p.d. for different Rotate and Go constants.}
 \label{tab:simulationmetricsrg}
@@
     \centering
-  \subfloat[a\label{1a}]{%
+  \subfloat[ \label{1a}]{%
        \includegraphics[width=0.49\linewidth]{images/ft_a2_n1500_cv15_cr15.eps}}
     \hspace*{-1.5em}
-  \subfloat[b\label{1b}]{%
+  \subfloat[ \label{1b}]{%
         \includegraphics[width=0.49\linewidth]{images/rgv_cv20_cr20_dr10.eps}}
@@
     \vspace*{-1.5em}
-  \subfloat[c\label{1c}]{%
+  \subfloat[ \label{1c}]{%
         \includegraphics[width=0.49\linewidth]{images/ft_cv15_cr10_leader_robot_speed.eps}}
     \hspace*{-1.5em}
-  \subfloat[d\label{1d}]{%
+  \subfloat[ \label{1d}]{%
         \includegraphics[width=0.49\linewidth]{images/rg_cv20_cr20_bv10_dtoff10_leader_robot_speed.eps}}
@@
     \vspace*{-2.5em}
-  \subfloat[d\label{1e}]{%
+  \subfloat[ \label{1e}]{%
         \includegraphics[width=0.99\linewidth]{images/ft_vs_rg_dist.eps}}
@@
     \centering
-  \subfloat[a\label{1a}]{%
+  \subfloat[ \label{1a}]{%
        \includegraphics[width=0.49\linewidth]{images/ft1cap.eps}}
     \hspace*{-1.5em}
-  \subfloat[b\label{1b}]{%
+  \subfloat[ \label{1b}]{%
         \includegraphics[width=0.49\linewidth]{images/rg2cap.eps}}
@@
     \vspace*{-1.5em}
-  \subfloat[c\label{1c}]{%
+  \subfloat[ \label{1c}]{%
         \includegraphics[width=0.49\linewidth]{images/ft1cap_leader_robot_speed.eps}}
     \hspace*{-1.5em}
-  \subfloat[d\label{1d}]{%
+  \subfloat[ \label{1d}]{%
         \includegraphics[width=0.49\linewidth]{images/rg2cap_leader_robot_speed.eps}}
@@
     \vspace*{-2.5em}
-  \subfloat[d\label{1e}]{%
+  \subfloat[ \label{1e}]{%
         \includegraphics[width=0.99\linewidth]{images/cap_ft_vs_rg_dist.eps}}
@@
 \hline
-$c_v$ & $c_{\alpha}$ & m.p.d. [m] & Area under \textit{npd} curve \\
+$c_v$ & $c_{\alpha}$ & m.p.d. [m] & t.p.d. \\
 \hline
@@
 \end{tabular}
-\captionof{table}{Max path deviation m.p.d. and area under normal path deviation in motion capture experiments using Follow the Thread.}
+\captionof{table}{Max path deviation m.p.d. and area under normal path deviation t.p.d. in motion capture experiments using Follow the Thread.}
 \label{tab:cap_ft_npd_table}
@@
 \hline
-$c_v$ & $c_r$ & $Dt_{off}$ & m.p.d. [m] & Area under \textit{npd} curve \\
+$c_v$ & $c_r$ & $Dt_{off}$ & m.p.d. [m] & t.p.d. \\
 \hline
@@
 \end{tabular}
-\captionof{table}{Max path deviation m.p.d. and area under normal path deviation in motion capture experiments using Rotate and Go.}
+\captionof{table}{Max path deviation m.p.d. and area under normal path deviation t.p.d. in motion capture experiments using Rotate and Go.}
 \label{tab:cap_rg_npd_table}
@@
 
+From the leader and robot trajectories in Figure~\ref{fig:simulationresults}(a,c), both strategies exhibits basic following behaviour.
+However, Rotate-And-Go on Figure~\ref{fig:simulationresults}(b,d) generates a more irregular path, due to the two stage movement algorithm.
+
+Regarding algorithms parameters, the simulation shows that for Follow-The-Thread, a low $c_{\alpha}$  means that the robot is slow to turn and makes wider turns, providing a smoother trajectory.
+On the other hand, for low $c_v$ values, the robot tends to lag behind the leader when it is going in a straight line.
+The highlighted values on Table~\ref{tab:simulationmetricsrg} show the best configuration found.
+For the Rotate-And-Go configuration, a high $c_r$ was important so the robot can turn quickly to point to the leader, but high values of $c_r$ and $c_v$ should be avoided, which may lead to dynamical issues with the inertia affecting the forward direction and missing the leader.
+Lastly, increasing the $Dt_{off}$ from 0.05 to 0.10 made the vehicle less prone to fall behind and results in a better following profile.
+
+%From the leader and robot trajectories in Figure \ref{fig:sim_ft_cv15_cr15}, we can see that the robot using Follow the Thread control exhibits basic following behaviour.
+Deviation occurs when the leader path turns with a tight radius.
+Comparing different simulations using different values of $c_{\alpha}$, we can also see that the robot tends to miss or cut corners when the leader turns, depending on the value of this constant (figures \ref{fig:sim_ft_cv15_cr5}, \ref{fig:sim_ft_cv15_cr10} and \ref{fig:sim_ft_cv15_cr15}).
+When $c_{\alpha}$ is low, the robot is slow to turn and makes wider turns, and when it is high, it can turn sharply as soon as the leader starts turning, so it will "cut" the curve.
+
+%Is was also found that with low $c_v$ values, the robot tends to lag behind the leader when it is going in a straight line, and as a consequence, when the leader turns, the follower is further away and has to make a sharper turn to keep following the leader.
+Again, a very low values of $c_v$ makes the follower "cut corners", as can be seen in figure \ref{fig:sim_ft_cv5_cr20}.
+
+%The rotateand go We see that this strategy also exhibits a following behaviour, although in a more irregular path, as expected, due to the two stage movement algorithm.
+In this case we evaluated the effects of the constants $c_v$, $c_r$ and $Dt_{off}$.
+
+%We found that it is important to have a high $c_r$ so the robot turns quickly to point to the leader, or otherwise, as the leader turns, the robot never stops rotating since it takes too much time to get to the point where it can go forward.
+Therefore, if the leader makes a long turn, the robot never stops rotating and stops following the leader (figure \ref{fig:rgv_cv20_cr5_dr05}).
+
+%However, a very high $c_r$ causes a different issue.
+When the robot finally points at the leader and starts going forward, if it was rotating too fast, the inertia of the wheels and the robot itself affect the forward movement of the vehicle, which can turn sideways as it tries to go forward in a straight line, diverting the robot from its intended path.
+A similar thing happens with a very high $c_v$: when the robot stops to rotate, its inertia prevents it from stopping completely and its path is affected as a consequence.
+This can be seen in figure \ref{fig:rgv_cv35_cr20_dr10}.
+
+%Lastly, we found that calibrating the $Dt_{off}$ parameter also has positive effects in the following performance, since it leaves more margin for the robot to go forward before rotating to point at the leader again.
+Increasing the $Dt_{off}$ from 0.05 to 0.10 made the vehicle less prone to fall behind and results in a better following (see figure \ref{fig:rg_dt_comp}).
+
+%Both behaviours are described by figures \ref{fig:ft_speed_1} and \ref{fig:rg_speed_1}, which show the speed of the leader and the robot over time.
+
+%When we measured the robot-leader distance over time, we see another significant between the two strategies.
+As mentioned before, using \textit{Rotate and Go}, as the robot rotates, the leader can get further away.
+Figures \ref{fig:ft_vs_rg_dist} and \ref{fig:ft_vs_rg_dist_hist} show the distance over time and the distribution of distances for each strategy, respectively.
+
+In line with the simulated results, a similar behaviour was found on the experiments performed inside the Motion Capture Lab, and the robot exhibits following behaviour for both control strategies as shown on Figure~\ref{fig:realworldresults}.
+Parameters had to be readjusted for the real world case, but the same situation occurs for both strategies.
+
+%We can see that, again, the robot exhibits following behaviour.
+The trajectories in the figures mentioned above show a similar behaviour to the one in the software simulation, for both control strategies.
+
+%As in the software simulation, we also evaluated the influence of the algorithm parameters in the robot movement, using the hardware prototype.
+Figures \ref{fig:ft1_cap_leader_robot} and \ref{fig:ft2_cap_leader_robot} show the robot and leader movement using different values for $c_{\alpha}$.
+As we have seen in the simulation, a higher $c_{\alpha}$ makes the robot tend to "cut corners" (figure \ref{fig:ft2_cap_leader_robot}).
+
+%For the \textit{Rotate and Go} strategy, we compared how different values of $Dt_{off}$ behaved.
+As we can see in figure \ref{fig:ft2_cap_leader_robot}, a higher $Dt_{off}$ makes the robot less sensitive to leader turns, so the robot has time to move forward and follow the shape of the curve more accurately.
+
+As expected, the \textit{Rotate and Go} strategy performs a \textit{stop and go} movement, since the vehicle completely stops when rotating to face the leader.
+This is shown on~\ref{fig:simulationresults}(c,d) as well as on the real experiment on Figure~\ref{fig:realworldresults}(c,d).
+The smoother movement of the \textit{Follow the Thread} algorithm can be desired by the followed subject, since it can be perceived as less violent or unexpected.
+%Additionally, when the vehicle stops to rotate, if the leader continues moving, when going forward, the robot might need to accelerate sharply to catch up with the leader, resulting in a sudden movement that can be disturbing for a medical patient or a doctor in the middle of a treatment.
+
+%As it was mentioned before, a smoother movement, with no sudden accelerations, can be desired in the environment where the robot is expected to function, so the \textit{Follow and Turn} strategy is better in this respect.
+
+%Another thing to note is that, since the robot has inertia when it tries to stop and rotate (in \textit{Rotate and Go}), its speed never gets to  zero as it rotates on its axis.
+This can also be an issue when following the leader, since the robot can be pushed off its course due to its own movement.
+
+%As in the simulation, we also see differences in the robot-leader distance over time.
+The \textit{Rotate and Go} strategy performs worse in this respect, reaching higher maximum distances and a higher average distance to the leader (table \ref{tab:cap_ft_vs_rg_dist_table}).
+
+Finally, regarding the Robot-leader distance r.l.d., the \textit{Rotate and Go} strategy results in a less stable distance (higher standard deviation), with higher maximum values, both on the simulated (Figure~\ref{fig:simulationresults}(e)) and on the real world scenario (Figure~\ref{fig:realworldresults}(e)).
+These values are also compared in Table \ref{tab:cap_ft_npd_table} and Table~\ref{tab:cap_rg_npd_table}.
+This can be specially problematic if we consider the cannula connecting the oxygen tank to the patient, as the cannula has a limited length.
+
 \subsection{Clinical Assessment} 
@@
 Afterwards, health care professionals were invited to use the robot themselves, simulating they were the patient being followed.
-They were allowed to switch between the two control strategies to evaluate both of them, and made different tests, one walking along standardized path marked on the floor and another one walking freely along the available space of the Motion Capture Lab .
+They were allowed to switch between the two control strategies to evaluate both of them, and made different tests, one walking along standardized path marked on the floor and another one walking freely along the available space of the Motion Capture Lab.
 They used the robot freely to get a general idea of how it behaved, and how it could be used in the rehabilitation process.
@@
 This project is part of a joint collaboration between ALPI organization and the ITBA University.
-Authors would like to thank thoughtfully for the initiative and support given by the ALPI organization, and to the Director of Rehabilitation Technology Dra.
+Authors would like to thank thoughtfully for the initiative and support given by ALPI, and to the Director of Rehabilitation Technology Dra.
 Mercedes Molinuevo.

### 2020-04-26 alpibot.tex

@@
 
-\author{Bianchi Luciano, Buzniak Esteban Alejandro, Ramele~Rodrigo,~\IEEEmembership{Member,~IEEE,}
+\author{Bianchi Luciano, Buniak Esteban Alejandro, Ramele~Rodrigo,~\IEEEmembership{Member,~IEEE,}
         and Santos~Juan~Miguel% <-this % stops a space
-\thanks{E. A. Buzniak, R. Ramele  and J.M.Santos are with the CiC Laboratory at the Department
+\thanks{E. A. Buniak, R. Ramele  and J.M.Santos are with the CIC Laboratory at the Department
 of  Computer Engineering, Instituto Tecnológico de Buenos Aires (ITBA), Ciudad de Buenos Aires,  Argentina e-mail: rramele@itba.edu.ar}% <-this % stops a space
@@
 The oxygen tank cannot be carried by the patient and external assistance is required.
-This work provides an initial study on the controlling mechanism of a differential tethered robot that implements a leader-follower configuration to carry the oxygen tank for Pulmonary Rehabilitation procedures.
+This work provides an initial study on the controlling mechanism of a differential tethered robot that implements a leader-follower configuration to carry the oxygen tank for these procedures.
 Two alternative control strategies are proposed.
@@
 They are characterized as a slowly progressive condition marked by airflow limitation, being cigarette smoking the main etiologic factor~\cite{MacNee2005}.
-This pathology presents an eskeletomuscular atrophy~\cite{Kocsis2016,Wu2012} and in order to carve these after effects a Pulmonary Rehabilitation procedure is a viable treatment for these patients.
+This pathology presents an eskeletomuscular atrophy~\cite{Kocsis2016,Wu2012}.
+In order to carve these after effects a Pulmonary Rehabilitation procedure is a viable treatment for patients.
 During these procedures patients undergo controlled walking activities and physical exercises under the supervision of a physical therapist.
-These patients present a severe low oxygen saturation illness and they require effective oxygen supply, particularly when performing physical activity~\cite{Celli2014}.
+Patients present a severe low oxygen saturation illness and they require effective oxygen supply, particularly when performing physical activity~\cite{Celli2014}.
 Hence, patients require to carry with them an oxygen tank for the oxygenotherapy assistance.
 However, their own condition prevent them with the ability to precisely carry the often bulky external tank.
-This situation entails a pragmatic solution to provide an additional physical therapist solely for the purpose of carrying the oxygen tank.
+This situation entails to find a pragmatic solution to avoid an additional physical therapist has to carry the oxygen tank.
 Alternatively, they are required to improvise a customized wheelchair to allow them to carry the oxygen tank on top of it, leading to a cumbersome situation for the physical therapist because s/he had to push and maneuver the wheelchair, and at the same time, take care of the patient and perform the rehabilitation treatment.
@@
 At the same time, several assistance devices for COPD treatments have been proposed.
-Particularly relevant are novel telemedice applications to enhance complementary rehabiliation exercise at home that can track biological markers for patients~\cite{Yang2018,Wu2012}.
+Particularly relevant are novel telemedicine~\cite{Banerjee2020} applications to enhance complementary rehabiliation exercise at home that can track biological markers for patients~\cite{Yang2018,Wu2012}.
 The work presented here follows the line established by \cite{Endo2015}.
-Authors studied the use of a single thread tethered follower robot for home oxygen therapy, and compared two different control algorithms and their effectiveness to mimic the leader path and to avoid obstacles.
+Authors studied the use of a single thread tethered follower robot for home oxygen therapy, and compared two different control algorithms and their effectiveness to mimic the leader trajectory and to avoid obstacles.
 However, their approach focuses on the usage of the device exclusively for home therapy, and not within the context of a Pulmonary Rehabilitation procedure performed by medical personnel on medical facilities.
 
-Hence, this work aims to provide an initial study on the controlling mechanism of a differential tethered robot that implements the leader-follower configuration on a Pulmonary Rehabilitation procedure.
+Hence, this work provides an initial study on the controlling mechanism of a differential tethered robot that implements the leader-follower configuration on a Pulmonary Rehabilitation procedure.
 To do so, this document unrolls as follows.
@@
 This difference in length is the input for the control algorithm.
-Using the encoder, the difference in length of the thread can be measured with Equation~\ref{eq:encoder_len}.
+Using the encoder, the difference in length for each thread can be measured with Equations~\ref{eq:encoder_len}.
 
@@
 \label{eq:encoder_len}
-L_d = pulses . \frac{2\pi r}{ppr}
+\begin{array}{ll}  
+D_l = pulses_{l} \; \frac{2\pi r}{ppr}  \\
+\\
+D_r = pulses_{r} \; \frac{2\pi r}{ppr}
+\end{array}
 \end{equation}{}
 
-\noindent where the variable $r$ is the radius of the reel and $ppr$ is the pulses per revolution (resolution) of the encoder.  The initial position of the threads is configurable.
+\noindent where pulses obtained for each reel encoder are $pulses_{l}$ and $pulses_{r}$ for the left and right encoder.
+This equation provides the estimated values for the left $D_l$ and right $D_r$  thread distance.
+The variable $r$ is the radius of the reel and $ppr$ is the pulses per revolution (resolution) of the encoder.
+The initial position of the threads is configurable.
 
@@
 The first battery is a commercial power bank with a capacity of 10000 mA$\cdot$h, and an output of 3.1A (over two USB ports) at 5V.
-The motor battery is a set of 10 AAA nickel–metal hydride batteries (1.5V each), with a total output of 15V.
+The motor battery is a set of 10 AAA nickel–metal hydride batteries (1.2V each), with a total output of 12V.
 
@@
 Two simple algorithmic control strategies are proposed and evaluated.
-The first one is called Follow-the-thread and the second strategy is Rotate-And-Go.
+The first one is called Follow-the-thread and the second strategy is Rotate and go.
 %The Adafruit Shield controller provides a very stable output signal, provided battery are kept above reference.
@@
 \noindent where  $c_v$ and $c_{\alpha}$ are constant coefficients used for calibration.
-As shown in Figure~\ref{fig:follow_params}, reel encoder information is $D_l$ and $D_r$ for the left and right reel, and finally $vleft_{tar}$ are $vright_{tar}$ are the power values that are directly used to drive power to each wheel motor.
+As shown in Figure~\ref{fig:follow_params},  $D_l$ and $D_r$ are thread distances obtained from encoder information from Equation~\ref{eq:encoder_len}, $v_{tar}$ is the estimated forward velocity for the target and finally $vleft_{tar}$ are $vright_{tar}$ are the power values that are directly used to drive power to each wheel motor.
 
@@
 
-\noindent where $d_0$ is a constant offset that is used to customized the length of the thread where the robot does not move.
+\noindent where $d_0$ is a constant offset that is used to customize the length of the thread where the robot does not move.
 
@@
 
-The \textit{Rotate and Go} algorithm divides the vehicle movement in two steps:
+The \textit{Rotate and go} algorithm divides the vehicle movement in two steps:
 
@@
 \item Rotating the vehicle around the center point of the axis that connects its front wheels in order to aim at the subject.
-\item Go forward in a straight line until the vehicle is at the expected distance to the subject or the vehicle is not moving in the direction of the subject.
+\item Go forward in a straight line until the vehicle is at the expected distance to the subject.
 \end{itemize}
@@
 The algorithm requires three parameters, $c_v$ and $c_r$ that regulates the coefficients for the forward and rotation movement, and an additional parameter $Dt_off$ that regulates the sensibility of the rotation movement.
+The constant parameter $Dm_off$ is similar to $d_0$, and is used to customize the length of the thread where the robot does not move.
 
 \begin{algorithm}[]
-\caption{Rotate and Go algorithm}
+\caption{Rotate and go algorithm}
 \begin{algorithmic}
@@
 \begin{itemize}
-    \item \textit{Normal path deviation, n.p.d.}: the subject trajectory is divided into small segments and then the normal distance to the robot trajectory is calculated for each of those segments.
-Path deviation is relevant to evaluate how closely the robot mimics the leader path, which is the ultimate goal of the robotic vehicle.
-    \item \textit{Total path deviation, t.p.d.}: The area under the curve resulting from the the \textit{normal path deviation} over the length traveled by the leader.
-    \item \textit{Maximum path deviation, m.p.d.}: the maximum \textit{normal path deviation} registered during an experiment.    
+    \item \textit{Normal trajectory deviation, n.t.d.}: the subject trajectory is divided into small segments and then the normal distance to the robot trajectory is calculated for each of those segments.
+Trajectory deviation is relevant to evaluate how closely the robot mimics the leader trajectory, which is the ultimate goal of the robotic vehicle.
+    \item \textit{Total trajectory deviation, t.t.d.}: The area under the curve resulting from the the \textit{normal trajectory deviation} over the length traveled by the leader.
+    \item \textit{Maximum trajectory deviation, m.t.d.}: the maximum \textit{normal trajectory deviation} registered during an experiment.    
     \item \textit{Robot-leader distance, r.l.d.}: The euclidean distance between the robot and the leader, at any point in time.
@@
 
-In this work, a \textit{following behavior} is considered satisfactory if its maximum path deviation is less than 0.75 m and the robot-leader distance never exceeds 1.5 m~\cite{MunozCeballos2010}.
+In this work, a \textit{following behavior} is considered satisfactory if its maximum trajectory deviation is less than 0.75 m and the robot-leader distance never exceeds 1.5 m~\cite{MunozCeballos2010}.
 
@@
 The threading mechanism was implemented using virtual threads~\cite{Rekleitis2001}.
-The leader traveled according to a predefined path with constant velocity, followiing the lemniscate trajectory.
+The leader traveled according to a predefined trajectory with constant velocity, following the lemniscate trajectory.
 
@@
 The leader starts at the midpoint of the trajectory and completes a full complete circuit getting back to the initial position, while the robot follows its track.
-Four different configuration sets of $c_v$ and  $c_{\alpha}$ were tested for Follow-The-Thread, while 6 different configurations were tested for Rotate-And-Go.
+Four different configuration sets of $c_v$ and  $c_{\alpha}$ were tested for Follow the thread, while 6 different configurations were tested for Rotate and go.
 
@@
 A real world experiment was performed, pegging to the same conditions implemented on the simulation environment.
-A motion capture system is used to track the movement of a human leader along a predetermined path.
+A motion capture system is used to track the movement of a human leader along a predetermined trajectory.
 The motion tracking system consists of an array of 16 OptiTrack (NaturalPoint Inc, Oregon, US) Flex 3 cameras, which measure the position of reflective markers with an accuracy of $\pm1$ [cm] at sampling rate of 100 Hz.
@@
 Simulation results for both control strategies are shown on Figure~\ref{fig:simulationresults}.
-Results metrics for the simulations are shown on Table~\ref{tab:simulationmetricsftt} for the Follow-The-Thread strategy, whereas metrics for Rotate-And-Go are shown on Table~\ref{tab:simulationmetricsrg}.
+Results metrics for the simulations are shown on Table~\ref{tab:simulationmetricsftt} for the Follow the thread strategy, whereas metrics for Rotate and go are shown on Table~\ref{tab:simulationmetricsrg}.
 
@@
 \hline
-$c_v$ & $c_{\alpha}$ & m.p.d. & t.p.d. \\
+$c_v$ & $c_{\alpha}$ & m.t.d. & t.t.d. \\
 \hline
@@
 \end{tabular}
-\captionof{table}{Max path deviation m.p.d. [m] and Total path deviation t.p.d. for different Follow the Thread constants.}
+\captionof{table}{Max trajectory deviation m.t.d. [m] and Total trajectory deviation t.t.d. for different Follow the thread constants.}
 \label{tab:simulationmetricsftt}
@@
 \hline
-$c_v$ & $c_r$ & $Dt_{off}$ & m.p.d. & t.p.d.\\
+$c_v$ & $c_r$ & $Dt_{off}$ & m.t.d. & t.t.d.\\
 \hline
@@
 \end{tabular}
-\captionof{table}{Max path deviation m.p.d. (in [m]) and total path deviation t.p.d. for different Rotate and Go constants.}
+\captionof{table}{Max trajectory deviation m.t.d. (in [m]) and total trajectory deviation t.p.d. for different Rotate and go constants.}
 \label{tab:simulationmetricsrg}
@@
 %\includegraphics[width=8cm]{images/ft_a2_n1500_cv15_cr15.eps}
-%\caption{Trajectories of the leader and follower for the Follow-The-Thread control strategy.}
+%\caption{Trajectories of the leader and follower for the Follow the thread control strategy.}
 %\label{fig:distance_sim}
@@
 %\includegraphics[width=8cm]{images/rgv_cv20_cr20_dr10.eps}
-%\caption{Trajectories of the leader and follower for Rotate-And-Go control strategy.}
+%\caption{Trajectories of the leader and follower for Rotate and go control strategy.}
 %\label{fig:distance_sim}
@@
 %\includegraphics[width=8cm]{images/ft_cv15_cr10_leader_robot_speed.eps}
-%\caption{Speed profiles of the leader and follower for Follow-The-Thread control strategy.}
+%\caption{Speed profiles of the leader and follower for Follow the thread control strategy.}
 %\label{fig:distance_sim}
@@
 %\includegraphics[width=8cm]{images/rg_cv20_cr20_bv10_dtoff10_leader_robot_speed.eps}
-%\caption{Speed profiles of the leader and follower for Rotate-And-Go control strategy.}
+%\caption{Speed profiles of the leader and follower for Rotate and go control strategy.}
 %\label{fig:distance_sim}
@@
         \includegraphics[width=0.99\linewidth]{images/ft_vs_rg_dist.eps}}
-  \caption{Simulation Results: Trajectories of the leader and follower for Follow-The-Thread (a) and Rotate-And-Go (b) . 
-        Speed profiles of the leader and follower for Follow-The-Thread (c) and Rotate and Go (d). 
+  \caption{Simulation Results: Trajectories of the leader and follower for Follow the thread (a) and Rotate and go (b) . 
+        Speed profiles of the leader and follower for Follow the thread (c) and Rotate and go (d). 
         (e) Separation distance between robot and leader for both strategies [m].}
@@
 Results for the real world experiment can be seen on Figure~\ref{fig:realworldresults}.
-Table~\ref{tab:cap_ft_npd_table} show the metrics of the Follow-The-Thread strategy, whereas Table~\ref{tab:cap_rg_npd_table} provides the metrics for the Rotate-And-Go approach.
+Table~\ref{tab:cap_ft_npd_table} show the metrics of the Follow the thread strategy, whereas Table~\ref{tab:cap_rg_npd_table} provides the metrics for the Rotate and go approach.
 
@@
 %\includegraphics[width=8cm]{images/ft1cap.eps}
-%\caption{Trajectories for the Follow-The-Thread control strategy.}
+%\caption{Trajectories for the Follow the thread control strategy.}
 %\label{fig:distance_sim}
@@
 %\includegraphics[width=8cm]{images/rg2cap_leader_robot_speed.eps}
-%\caption{Rotate and Go speed profiles}
+%\caption{Rotate and go speed profiles}
 %\label{fig:distance_sim}
@@
 %\includegraphics[width=8cm]{images/cap_ft_vs_rg_dist.eps}
-%\caption{Rotate and Go speed profiles}
+%\caption{Rotate and go speed profiles}
 %\label{fig:distance_sim}
@@
         \includegraphics[width=0.99\linewidth]{images/cap_ft_vs_rg_dist.eps}}
-  \caption{Experimentation Results: Trajectories of the leader and follower for Follow-The-Thread (a) and Rotate-And-Go (b). 
-        Speed profiles of the leader and follower for Follow-The-Thread (c) and Rotate and Go (d). 
+  \caption{Experimentation Results: Trajectories of the leader and follower for Follow the thread (a) and Rotate and go (b). 
+        Speed profiles of the leader and follower for Follow the thread (c) and Rotate and go (d). 
         (e) Separation distance between robot and leader for both strategies [m].}
@@
 \hline
-$c_v$ & $c_{\alpha}$ & m.p.d. [m] & t.p.d. \\
+$c_v$ & $c_{\alpha}$ & m.t.d. [m] & t.t.d. \\
 \hline
@@
 \end{tabular}
-\captionof{table}{Max path deviation m.p.d. and area under normal path deviation t.p.d. in motion capture experiments using Follow the Thread.}
+\captionof{table}{Max trajectory deviation m.t.d. and area under normal trajectory deviation t.t.d. in motion capture experiments using Follow the thread.}
 \label{tab:cap_ft_npd_table}
@@
 \hline
-$c_v$ & $c_r$ & $Dt_{off}$ & m.p.d. [m] & t.p.d. \\
+$c_v$ & $c_r$ & $Dt_{off}$ & m.t.d. [m] & t.t.d. \\
 \hline
@@
 \end{tabular}
-\captionof{table}{Max path deviation m.p.d. and area under normal path deviation t.p.d. in motion capture experiments using Rotate and Go.}
+\captionof{table}{Max trajectory deviation m.t.d. and area under normal trajectory deviation t.p.d. in motion capture experiments using Rotate and go.}
 \label{tab:cap_rg_npd_table}
@@
 
-From the leader and robot trajectories in Figure~\ref{fig:simulationresults}(a,c), both strategies exhibits basic following behaviour.
-However, Rotate-And-Go on Figure~\ref{fig:simulationresults}(b,d) generates a more irregular path, due to the two stage movement algorithm.
-
-Regarding algorithms parameters, the simulation shows that for Follow-The-Thread, a low $c_{\alpha}$  means that the robot is slow to turn and makes wider turns, providing a smoother trajectory.
+From the leader and robot trajectories in Figure~\ref{fig:simulationresults}(a to d), both strategies exhibits basic following behaviour.
+However, Rotate and go on Figure~\ref{fig:simulationresults}(b,d) generates a more irregular trajectory, due to the two stage movement algorithm.
+
+Regarding algorithms parameters, the simulation shows that for Follow the thread, a low $c_{\alpha}$  means that the robot is slow to turn and makes wider turns, providing a smoother trajectory.
 On the other hand, for low $c_v$ values, the robot tends to lag behind the leader when it is going in a straight line.
 The highlighted values on Table~\ref{tab:simulationmetricsrg} show the best configuration found.
-For the Rotate-And-Go configuration, a high $c_r$ was important so the robot can turn quickly to point to the leader, but high values of $c_r$ and $c_v$ should be avoided, which may lead to dynamical issues with the inertia affecting the forward direction and missing the leader.
+For the Rotate and go configuration, a high $c_r$ was important so the robot can turn quickly to point to the leader, but high values of $c_r$ and $c_v$ should be avoided, which may lead to dynamical issues with the inertia affecting the forward direction and missing the leader.
 Lastly, increasing the $Dt_{off}$ from 0.05 to 0.10 made the vehicle less prone to fall behind and results in a better following profile.
 
-%From the leader and robot trajectories in Figure \ref{fig:sim_ft_cv15_cr15}, we can see that the robot using Follow the Thread control exhibits basic following behaviour.
-Deviation occurs when the leader path turns with a tight radius.
+%From the leader and robot trajectories in Figure \ref{fig:sim_ft_cv15_cr15}, we can see that the robot using Follow the thread control exhibits basic following behaviour.
+Deviation occurs when the leader trajectory turns with a tight radius.
 Comparing different simulations using different values of $c_{\alpha}$, we can also see that the robot tends to miss or cut corners when the leader turns, depending on the value of this constant (figures \ref{fig:sim_ft_cv15_cr5}, \ref{fig:sim_ft_cv15_cr10} and \ref{fig:sim_ft_cv15_cr15}).
@@
 %When we measured the robot-leader distance over time, we see another significant between the two strategies.
-As mentioned before, using \textit{Rotate and Go}, as the robot rotates, the leader can get further away.
+As mentioned before, using \textit{Rotate and go}, as the robot rotates, the leader can get further away.
 Figures \ref{fig:ft_vs_rg_dist} and \ref{fig:ft_vs_rg_dist_hist} show the distance over time and the distribution of distances for each strategy, respectively.
@@
 
-%For the \textit{Rotate and Go} strategy, we compared how different values of $Dt_{off}$ behaved.
+%For the \textit{Rotate and go} strategy, we compared how different values of $Dt_{off}$ behaved.
 As we can see in figure \ref{fig:ft2_cap_leader_robot}, a higher $Dt_{off}$ makes the robot less sensitive to leader turns, so the robot has time to move forward and follow the shape of the curve more accurately.
 
-As expected, the \textit{Rotate and Go} strategy performs a \textit{stop and go} movement, since the vehicle completely stops when rotating to face the leader.
-This is shown on~\ref{fig:simulationresults}(c,d) as well as on the real experiment on Figure~\ref{fig:realworldresults}(c,d).
-The smoother movement of the \textit{Follow the Thread} algorithm can be desired by the followed subject, since it can be perceived as less violent or unexpected.
+As expected, the \textit{Rotate and go} strategy performs a \textit{stop and go} movement, since the vehicle completely stops when rotating to face the leader.
+This is shown on the speed profiles in Figure~\ref{fig:simulationresults}(c,d) as well as on the real experiment on Figure~\ref{fig:realworldresults}(c,d).
+The smoother movement of the \textit{Follow the thread} algorithm can be desired by the followed subject, since it can be perceived as less violent or unexpected.
 %Additionally, when the vehicle stops to rotate, if the leader continues moving, when going forward, the robot might need to accelerate sharply to catch up with the leader, resulting in a sudden movement that can be disturbing for a medical patient or a doctor in the middle of a treatment.
 
-%As it was mentioned before, a smoother movement, with no sudden accelerations, can be desired in the environment where the robot is expected to function, so the \textit{Follow and Turn} strategy is better in this respect.
-
-%Another thing to note is that, since the robot has inertia when it tries to stop and rotate (in \textit{Rotate and Go}), its speed never gets to  zero as it rotates on its axis.
+%As it was mentioned before, a smoother movement, with no sudden accelerations, can be desired in the environment where the robot is expected to function, so the \textit{Rotate and go} strategy is better in this respect.
+
+%Another thing to note is that, since the robot has inertia when it tries to stop and rotate (in \textit{Rotate and go}), its speed never gets to  zero as it rotates on its axis.
 This can also be an issue when following the leader, since the robot can be pushed off its course due to its own movement.
@@
 %As in the simulation, we also see differences in the robot-leader distance over time.
-The \textit{Rotate and Go} strategy performs worse in this respect, reaching higher maximum distances and a higher average distance to the leader (table \ref{tab:cap_ft_vs_rg_dist_table}).
-
-Finally, regarding the Robot-leader distance r.l.d., the \textit{Rotate and Go} strategy results in a less stable distance (higher standard deviation), with higher maximum values, both on the simulated (Figure~\ref{fig:simulationresults}(e)) and on the real world scenario (Figure~\ref{fig:realworldresults}(e)).
+The \textit{Rotate and go} strategy performs worse in this respect, reaching higher maximum distances and a higher average distance to the leader (table \ref{tab:cap_ft_vs_rg_dist_table}).
+
+Finally, regarding the Robot-leader distance r.l.d., the \textit{Rotate and go} strategy results in a less stable distance (higher standard deviation), with higher maximum values, both on the simulated (Figure~\ref{fig:simulationresults}(e)) and on the real world scenario (Figure~\ref{fig:realworldresults}(e)).
 These values are also compared in Table \ref{tab:cap_ft_npd_table} and Table~\ref{tab:cap_rg_npd_table}.
@@
 
-No amount of metrics are enough to evaluate if the robot is a viable solution for this problem or not, without the input and the evaluation of the people that are going to physically make use of it.
+No amount of metrics are enough to evaluate if the robot is a viable solution for this problem or not, without the feedback and the evaluation of the people that are going to physically make use of it.
 
@@
 Afterwards, health care professionals were invited to use the robot themselves, simulating they were the patient being followed.
-They were allowed to switch between the two control strategies to evaluate both of them, and made different tests, one walking along standardized path marked on the floor and another one walking freely along the available space of the Motion Capture Lab.
+They were allowed to switch between the two control strategies to evaluate both of them, and made different tests, one walking along standardized trajectory marked on the floor and another one walking freely along the available space of the Motion Capture Lab.
 They used the robot freely to get a general idea of how it behaved, and how it could be used in the rehabilitation process.
@@
 \hline
-How would you qualify, from 1 to 5, your overall experience with Follow the Thread? (1:Bad, 5:Excellent)  & 5.0\\
-\hline
-How safe would a patient be, from 1 to 5, being followed by the robot using the Follow the Thread strategy? (1:Very unsafe, 5:Very safe)  & 4.25\\
-\hline
-How would you qualify, from 1 to 5, your overall experience with Rotate and Go? (1:Bad, 5:Excellent)  & 3.5\\
-\hline
-How safe would a patient be, from 1 to 5, being followed by the robot using the Rotate and Go strategy? (1:Very unsafe, 5:Very safe) & 3.25\\
+How would you qualify, from 1 to 5, your overall experience with Follow the thread? (1:Bad, 5:Excellent)  & 5.0\\
+\hline
+How safe would a patient be, from 1 to 5, being followed by the robot using the Follow the thread strategy? (1:Very unsafe, 5:Very safe)  & 4.25\\
+\hline
+How would you qualify, from 1 to 5, your overall experience with Rotate and go? (1:Bad, 5:Excellent)  & 3.5\\
+\hline
+How safe would a patient be, from 1 to 5, being followed by the robot using the Rotate and go strategy? (1:Very unsafe, 5:Very safe) & 3.25\\
 \hline
@@
 
-According to their answers, and the discussion we had after testing the robot, the general opinion was that the \textit{Follow the Thread} strategy was safer and more convenient for the task.
-In the survey, when asked \textit{Which of the two strategies is more effective at following the patient in a rehabilitation exercise?}, all 4 people responded that \textit{Follow the Thread} is "much better".
-
-The main concern with the \textit{Rotate and Go} strategy was that having to wait for the robot to rotate before moving forward might be unsafe, as the patient could move away from it and compromise the cannula connecting him or her to the oxygen tank.
+According to their answers, and the discussion we had after testing the robot, the general opinion was that the \textit{Follow the thread} strategy was safer and more convenient for the task.
+In the survey, when asked \textit{Which of the two strategies is more effective at following the patient in a rehabilitation exercise?}, all 4 people responded that \textit{Follow the thread} is "much better".
+
+The main concern with the \textit{Rotate and go} strategy was that having to wait for the robot to rotate before moving forward might be unsafe, as the patient could move away from it and compromise the cannula connecting him or her to the oxygen tank.
 This issue was identified during our own tests, and was not mentioned when explaining the following mechanism to the doctors, to avoid skewing them.
@@
 
-Another aspect that was value from the \textit{Follow the Thread} strategy is that since its smoother movement, with no sudden stops or accelerations, was favorable for the stability of the robot if it were to carry the heavy oxygen tank.
+Another aspect that was value from the \textit{Follow the thread} strategy is that since its smoother movement, with no sudden stops or accelerations, was favorable for the stability of the robot if it were to carry the heavy oxygen tank.
 
@@
 
-From the practical experiments, it is verified that the both algorithms gave reasonable following performance in the task of following the leader along a lemniscate-shaped path.
+From the practical experiments, it is verified that the both algorithms gave reasonable following performance in the task of following the leader along a lemniscate-shaped trajectory.
 This reasonable following behaviour is accomplished with a simple mechanism, a characteristic that significantly keep the price of the device low, putting it within reach of many medical institutions.
 
-Each control strategy has its advantages, but according various metrics described in this work, the \textit{Follow and Turn} strategy had a more desirable behaviour, as it tended to follow the leader from a closer distance at all times, while moving in a smooth and predictable way.
+Each control strategy has its advantages, but according various metrics described in this work, the \textit{Follow the thread} strategy had a more desirable behaviour, as it tended to follow the leader from a closer distance at all times, while moving in a smooth and predictable way.
 
 Insightful feedback is gathered from healthcare professionals from ALPI, who provided invaluable data to evaluate the solution.
-Over all, they highlight the Follow and Turn strategy as being the safer and more effective one.
+Over all, they highlight the \textit{Follow the thread} strategy as being the safer and more effective one.
 Most importantly, they also validated the research and were enthusiast about the direction of the project.
@@
 
-The next steps for this project is to scale the design to be able to carry safely the oxygen tank.
-
-Further research is needed to develop a full fledged solution: 
+The next steps for this project is to scale and iterate the design towards the desired solution, using the data obtained from this experiments and the feedback from care-givers.
+
 \begin{itemize}
     \item Power the motor wheels with a \textit{closed-loop and more efficient controller}, to have a more precise control over their speed.
-    \item Redesign the active spring mechanism in the reels to avoid overheating.
-In our tests, the small electric motors in each reel got very hot after just 20 minutes of use.
-The active reel algorithm is enough to keep them in manageable temperatures (under $70^0C$), which is enough for testing the prototype, but it will not be appropriate for a useful robot.
-Not only because the heat produced can be dangerous, but also because the motors can overheat and break their insulation in regular use.
+    \item Redesign the active spring control mechanism in order to hold the motor temperature in its operational range.
     \item An easy and safe interaction between the patient, the operator and the robot.

### 2020-05-01 alpibot.tex
Relectura y ajustes.  Eliminé algunas partes.

@@
 Patients that suffer Chronic Obstructive Pulmonary Disease (COPD) undergo a procedure called Pulmonary Rehabilitation that helps them to improve disease prognosis.
-During Pulmonary Rehabilitation procedures patients require oxygen assistance.
-The oxygen tank cannot be carried by the patient and external assistance is required.
+During Pulmonary Rehabilitation programs patients perform different exercises and walking activities and may require oxygen assistance.
+The oxygen tank cannot be carried by the patient due to the eskeletomuscular atrophy that characterize this pathology and external assistance is required.
+The assistance to transport the bulky oxygen tank can be provided by a robotic device that follows the patient while performing the physical activities.
 This work provides an initial study on the controlling mechanism of a differential tethered robot that implements a leader-follower configuration to carry the oxygen tank for these procedures.
@@
 \begin{IEEEkeywords}
-SAR, IoRT, tethered, leader-follower, pulmonary rehabilitation, COPD
+COPD, PR, SAR, IoRT, tethered, leader-follower
 \end{IEEEkeywords}
@@
 In order to carve these after effects a Pulmonary Rehabilitation procedure is a viable treatment for patients.
-During these procedures patients undergo controlled walking activities and physical exercises under the supervision of a physical therapist.
+During this rehabilitation, patients undergo controlled walking activities and physical exercises under the supervision of a physical therapist.
 Patients present a severe low oxygen saturation illness and they require effective oxygen supply, particularly when performing physical activity~\cite{Celli2014}.
@@
 However, their own condition prevent them with the ability to precisely carry the often bulky external tank.
-This situation entails to find a pragmatic solution to avoid an additional physical therapist has to carry the oxygen tank.
-Alternatively, they are required to improvise a customized wheelchair to allow them to carry the oxygen tank on top of it, leading to a cumbersome situation for the physical therapist because s/he had to push and maneuver the wheelchair, and at the same time, take care of the patient and perform the rehabilitation treatment.
-
-The scenario may be aggravated by the fact that this procedure is performed on a rehabilitation gym that could be potentially crowded with several patients, physiologist and physical therapists.
+This situation entails to find a pragmatic solution to avoid an additional physical therapist to carry the oxygen tank.
+This scenario may be aggravated by the fact that this procedure is performed on a rehabilitation gym that could be potentially crowded with several patients, physiologist and physical therapists.
+
+%Alternatively, they are required to improvise a customized wheelchair to allow them to carry the oxygen tank on top of it, leading to a cumbersome situation for the physical therapist because s/he had to push and maneuver the wheelchair, and at the same time, take care of the patient and perform the rehabilitation treatment.
 
@@
 Section~\ref{experimental} documents the experimental protocol to perform the solution assessment on a simulation and on a real world scenario.
-Results and Discussions are described in~\ref{results} and~\ref{discussion}.
-The Clinical Assessment performed jointly with medical personnel is tackled in Section~\ref{clinical}.
-Finally, conclusions are exposed in the remaining section~\ref{conclusion}.
+Results and discussions are described in Sections ~\ref{results} and~\ref{discussion}.
+The clinical assessment performed jointly with medical personnel is tackled in Section~\ref{clinical}.
+Finally, conclusions are exposed in the remaining Section~\ref{conclusion}.
 
@@
     % Text Node
-    \draw (302,1344) node    {$left\ reel$};
+    \draw (302,1344) node    {$left\ reel\  \ $};
     % Text Node
-    \draw (299,1425) node    {$right\ reel$};
+    \draw (299,1425) node    {$right\ reel\  \ $};
     % Text Node
@@
     \end{tikzpicture}
-    \caption{Components of the robotic vehicle and following mechanism}
+    \caption{Components of the robotic vehicle and the tether mechanism}
     \label{fig:components}
@@
 \noindent where pulses obtained for each reel encoder are $pulses_{l}$ and $pulses_{r}$ for the left and right encoder.
-This equation provides the estimated values for the left $D_l$ and right $D_r$  thread distance.
+These equations provide the estimated values for the left $D_l$ and right $D_r$  thread distance.
 The variable $r$ is the radius of the reel and $ppr$ is the pulses per revolution (resolution) of the encoder.
@@
 Frames are constructed from aluminum extrusions produced by Makeblock (Shenzhen, China).
-The prototype is a three-wheeled robot with two frontal differential drive wheels and a free castor wheel as a third point of contact on the back, as can be seen on Figure~\ref{fig:alpibot2}.
+The prototype is a three-wheeled robot with two frontal differential drive wheels and a free castor wheel as a third point of contact on the back, as can be seen on Figures~\ref{fig:alpibot1} and~\ref{fig:alpibot2}.
 
@@
 On top of it an Adafruit (Adafruit, New York City) Motor Shield v2 bridge is used to drive the four DC motors, one for each wheel and one on each reel.
-The Arduino board is also connected to a Single Board Computer (SBC) Raspberry Pi (Raspberry Pi Foundation, United Kingdom) 3B+ through serial connection on the USB port.
+The Arduino board is also connected to a Single Board Computer (SBC) Raspberry Pi (Raspberry Pi Foundation, United Kingdom) 3B+ through serial connection on one of the USB port.
 
@@
 The algorithm requires three parameters, $c_v$ and $c_r$ that regulates the coefficients for the forward and rotation movement, and an additional parameter $Dt_off$ that regulates the sensibility of the rotation movement.
-The constant parameter $Dm_off$ is similar to $d_0$, and is used to customize the length of the thread where the robot does not move.
+The constant parameter $Dm_off$ is similar to $d_0$, and is used to customize the length of the thread where the robot does not move at all.
 
@@
 
-A tracked marker was placed in each side of the robot (on top of each thread reel).
+A tracked marker was placed on each side of the robot (on top of each thread reel).
 The human leader used his hand to grab the tip at which the two tethers were tied together.
@@
 The lemniscate of gerone was marked on the floor.
-The subject follows this track on the experiments performed.}
+The subject follows this track on the performed experiments.}
 \label{fig:distance_sim}
@@
 Simulation results for both control strategies are shown on Figure~\ref{fig:simulationresults}.
+Subfigures (a) and (b) expound the trajectories of the leader and the follower for each strategy, while (c) and (d) describe their speed profiles.
+Subfigure (e) show the distance between the robot and the patient for both strategies.
 Results metrics for the simulations are shown on Table~\ref{tab:simulationmetricsftt} for the Follow the thread strategy, whereas metrics for Rotate and go are shown on Table~\ref{tab:simulationmetricsrg}.
@@
 \end{center}
-
-Additionally, speed profiles can be observed from both trajectories.
 
@@
 \hline
-25  &   20  & 0.3876 & 1.9761\\
+\textbf{25}  &   \textbf{20}  & \textbf{0.3876} & \textbf{1.9761}\\
 25  &   35  & 0.4672 & 2.3528\\
@@
 30  &   35  & 0.04  & 0.4116 & 2.8309\\
-30  &   35  & 0.08 & 0.3739 & 2.0367\\
+\textbf{30}  &   \textbf{35}  & \textbf{0.08 }& \textbf{0.3739} & \textbf{2.0367}\\
 \hline
@@
 Finally, regarding the Robot-leader distance r.l.d., the \textit{Rotate and go} strategy results in a less stable distance (higher standard deviation), with higher maximum values, both on the simulated (Figure~\ref{fig:simulationresults}(e)) and on the real world scenario (Figure~\ref{fig:realworldresults}(e)).
-These values are also compared in Table \ref{tab:cap_ft_npd_table} and Table~\ref{tab:cap_rg_npd_table}.
 This can be specially problematic if we consider the cannula connecting the oxygen tank to the patient, as the cannula has a limited length.
+
+%These values are also compared in Table \ref{tab:cap_ft_npd_table} and Table~\ref{tab:cap_rg_npd_table}.
 
@@
 
-Another aspect that was value from the \textit{Follow the thread} strategy is that since its smoother movement, with no sudden stops or accelerations, was favorable for the stability of the robot if it were to carry the heavy oxygen tank.
+Another aspect that was remarked from the \textit{Follow the thread} strategy is that since it had a smoother movement, with no sudden stops or accelerations, it was favorable for the stability of the robot in order to carry the heavy oxygen tank.
 
@@
 
-The staff at ALPI was enthusiastic about helping in the development of the robot, and their encouragement and support are key reasons to take this project forward.
+%The staff at ALPI was enthusiastic about helping in the development of the robot, and their encouragement and support are key reasons to take this project forward.
 There are yet many challenges ahead, but these initial results are promising about the viability to develop a real world solution that can improve the quality of life of many people through the use of technology and engineering.

### 2020-05-01 alpibot.tex
Actualizo los gráficos que me pasó Luciano.

@@
 Patients that suffer Chronic Obstructive Pulmonary Disease (COPD) undergo a procedure called Pulmonary Rehabilitation that helps them to improve disease prognosis.
-During Pulmonary Rehabilitation programs patients perform different exercises and walking activities and may require oxygen assistance.
+During Pulmonary Rehabilitation patients perform different exercises and walking activities and may require oxygen assistance.
 The oxygen tank cannot be carried by the patient due to the eskeletomuscular atrophy that characterize this pathology and external assistance is required.
@@
     \item \textit{Normal trajectory deviation, n.t.d.}: the subject trajectory is divided into small segments and then the normal distance to the robot trajectory is calculated for each of those segments.
-Trajectory deviation is relevant to evaluate how closely the robot mimics the leader trajectory, which is the ultimate goal of the robotic vehicle.
+Trajectory deviation curve is relevant to evaluate how closely the robot mimics the leader trajectory, which is the ultimate goal of the robotic vehicle.
+    \item \textit{Robot-leader distance, r.l.d.}: The euclidean distance between the robot and the leader, at any point in time.
+This curve is particularly important since the robot has a limited amount of thread available, so if the leader uses all the available thread, it will start dragging the robot and damaging the following mechanis, and overall it may rise the possibility of disconnecting the breathing oxygen cannula.
+This is a scenario that must not happen under any circumstance, as it can also be dangerous for a potential patient using the device.
     \item \textit{Total trajectory deviation, t.t.d.}: The area under the curve resulting from the the \textit{normal trajectory deviation} over the length traveled by the leader.
     \item \textit{Maximum trajectory deviation, m.t.d.}: the maximum \textit{normal trajectory deviation} registered during an experiment.    
-    \item \textit{Robot-leader distance, r.l.d.}: The euclidean distance between the robot and the leader, at any point in time.
-This is particularly important since the robot has a limited amount of thread available, so if the leader uses all the available thread, it will start dragging the robot and damaging the following mechanis, and overall it may rise the possibility of disconnecting the breathing oxygen cannula.
-This is a scenario that must not happen under any circumstance, as it can also be dangerous for a potential patient using the device.
 \end{itemize}{}
@@
   \subfloat[ \label{1a}]{%
-       \includegraphics[width=0.49\linewidth]{images/ft_a2_n1500_cv15_cr15.eps}}
+       \includegraphics[width=0.49\linewidth]{images/simulation_follow-thread.eps}}
     \hspace*{-1.5em}
   \subfloat[ \label{1b}]{%
-        \includegraphics[width=0.49\linewidth]{images/rgv_cv20_cr20_dr10.eps}}
+        \includegraphics[width=0.49\linewidth]{images/simulation_rotate-go.eps}}
     \\
@@
   \subfloat[ \label{1c}]{%
-        \includegraphics[width=0.49\linewidth]{images/ft_cv15_cr10_leader_robot_speed.eps}}
+        \includegraphics[width=0.49\linewidth]{images/simulation_speed_profile_follow-thread.eps}}
     \hspace*{-1.5em}
   \subfloat[ \label{1d}]{%
-        \includegraphics[width=0.49\linewidth]{images/rg_cv20_cr20_bv10_dtoff10_leader_robot_speed.eps}}
+        \includegraphics[width=0.49\linewidth]{images/simulation_speed_profile_rotate-go.eps}}
     \\
@@
         Speed profiles of the leader and follower for Follow the thread (c) and Rotate and go (d). 
-        (e) Separation distance between robot and leader for both strategies [m].}
+        (e) Separation distance between robot and leader for both strategies.}
   \label{fig:simulationresults} 
@@
   \subfloat[ \label{1a}]{%
-       \includegraphics[width=0.49\linewidth]{images/ft1cap.eps}}
+       \includegraphics[width=0.49\linewidth]{images/capture_follow-turn.eps}}
     \hspace*{-1.5em}
   \subfloat[ \label{1b}]{%
-        \includegraphics[width=0.49\linewidth]{images/rg2cap.eps}}
+        \includegraphics[width=0.49\linewidth]{images/capture_rotate-go.eps}}
     \\
@@
   \subfloat[ \label{1c}]{%
-        \includegraphics[width=0.49\linewidth]{images/ft1cap_leader_robot_speed.eps}}
+        \includegraphics[width=0.49\linewidth]{images/capture_speed_profile_follow-thread.eps}}
     \hspace*{-1.5em}
   \subfloat[ \label{1d}]{%
-        \includegraphics[width=0.49\linewidth]{images/rg2cap_leader_robot_speed.eps}}
+        \includegraphics[width=0.49\linewidth]{images/capture_speed_profile_rotate-go.eps}}
     \\

### 2020-05-12 alpibot.tex
Ajustes varios para enviar a la revista.

@@
 % The paper headers
-\markboth{IEEE Latin American Transactions}%
-{Bianchi \MakeLowercase{\textit{et al.}} Controller Design for a Tethered Follower Robot for Pulmonary Rehabilitation}
+\markboth{IEEE Transactions on Neural Systems and Rehabilitation Engineering}%
+{Bianchi \MakeLowercase{\textit{et al.}} A Control Strategy for a Tethered Follower Robot
+for Pulmonary Rehabilitation}
 % The only time the second header will appear is for the odd numbered pages
@@
 \begin{IEEEkeywords}
-COPD, PR, SAR, IoRT, tethered, leader-follower
+COPD, PR, SAR, IoRT, tethered
 \end{IEEEkeywords}
@@
 This situation entails to find a pragmatic solution to avoid an additional physical therapist to carry the oxygen tank.
-This scenario may be aggravated by the fact that this procedure is performed on a rehabilitation gym that could be potentially crowded with several patients, physiologist and physical therapists.
+This scenario may be aggravated by the fact that this procedure is performed on a rehabilitation gym that could be potentially crowded with several patients, physiologists and physical therapists.
 
@@
 
-Regarding metrics, three are proposed to evaluate the performance.  They are:
+Regarding metrics, four are proposed to evaluate the performance.  They are:
 
@@
 
-A tracked marker was placed on each side of the robot (on top of each thread reel).
+As shown in Figure~\ref{fig:capturesystem}, a tracking marker was placed on each side of the robot (on top of each thread reel).
 The human leader used his hand to grab the tip at which the two tethers were tied together.
@@
 The subject follows this track on the performed experiments.}
-\label{fig:distance_sim}
+\label{fig:capturesystem}
 \end{figure}
@@
     \centering
-  \subfloat[ \label{1a}]{%
+  \subfloat[ \label{sr1a}]{%
        \includegraphics[width=0.49\linewidth]{images/simulation_follow-thread.eps}}
     \hspace*{-1.5em}
-  \subfloat[ \label{1b}]{%
+  \subfloat[ \label{sr1b}]{%
         \includegraphics[width=0.49\linewidth]{images/simulation_rotate-go.eps}}
     \\
-    \vspace*{-1.5em}
-  \subfloat[ \label{1c}]{%
+    \vspace*{-1.1em}
+  \subfloat[ \label{sr1c}]{%
         \includegraphics[width=0.49\linewidth]{images/simulation_speed_profile_follow-thread.eps}}
     \hspace*{-1.5em}
-  \subfloat[ \label{1d}]{%
+  \subfloat[ \label{sr1d}]{%
         \includegraphics[width=0.49\linewidth]{images/simulation_speed_profile_rotate-go.eps}}
@@
     \vspace*{-2.5em}
-  \subfloat[ \label{1e}]{%
+  \subfloat[ \label{sr1e}]{%
         \includegraphics[width=0.99\linewidth]{images/ft_vs_rg_dist.eps}}
@@
     \centering
-  \subfloat[ \label{1a}]{%
+  \subfloat[ \label{rw1a}]{%
        \includegraphics[width=0.49\linewidth]{images/capture_follow-turn.eps}}
     \hspace*{-1.5em}
-  \subfloat[ \label{1b}]{%
+  \subfloat[ \label{rw1b}]{%
         \includegraphics[width=0.49\linewidth]{images/capture_rotate-go.eps}}
     \\
-    \vspace*{-1.5em}
-  \subfloat[ \label{1c}]{%
+    \vspace*{-1.1em}
+  \subfloat[ \label{rw1c}]{%
         \includegraphics[width=0.49\linewidth]{images/capture_speed_profile_follow-thread.eps}}
     \hspace*{-1.5em}
-  \subfloat[ \label{1d}]{%
+  \subfloat[ \label{rw1d}]{%
         \includegraphics[width=0.49\linewidth]{images/capture_speed_profile_rotate-go.eps}}
@@
     \vspace*{-2.5em}
-  \subfloat[ \label{1e}]{%
+  \subfloat[ \label{rw1e}]{%
         \includegraphics[width=0.99\linewidth]{images/cap_ft_vs_rg_dist.eps}}
@@
 
-From the practical experiments, it is verified that the both algorithms gave reasonable following performance in the task of following the leader along a lemniscate-shaped trajectory.
+From the practical experiments, it is verified that both algorithms gave reasonable following performance in the task of following the leader along a lemniscate-shaped trajectory.
 This reasonable following behaviour is accomplished with a simple mechanism, a characteristic that significantly keep the price of the device low, putting it within reach of many medical institutions.

### 2020-05-12 coverletter.tex
Subo la coverletter.

Dear Editor,

We are submitting a Full Paper, entitled "A Control Strategy for a Tethered Follower Robot for Pulmonary Rehabilitation" which we believe is within the scope of IEEE Transactions on Neural Systems and Rehabilitation Engineering .  This article encompass work performed at CiC Laboratory of the ITBA University in Buenos Aires, Argentina, in the context of research on Rehabilitation Robotics.  This is an original production, which has not been submitted for publication elsewhere.

This work is the first approach to implement a solution that was brought to us from ALPI organization.  ALPI is a non-profit civil association located in Buenos Aires, Argentina, that provides neuromotor rehabilitation for pediatric and adult patients. It was founded in 1943 with the main focus of treating children with poliomyelitis, and has since expanded to deal with all kinds of related diseases.  In the context of a translational collaboration agreement with them, they identified the problem stated in this work and we started to work together on a technological and engineering solution.  We describe in these pages a basic scheme that we design to drive a robotic platform that can carry the oxygen tank of emphysema patients that are under Pulmonary Rehabilitation treatments.  Our objective was to implement a gradual iterative design strategy and produce a prototyping platform that tackled the problem of the patient following mechanism while keeping affordable costs and that can receive quick feedback from ALPI care-givers.  We show that the very simple design and controlling scheme is able to implement a following strategy and that the proposed solution is viable.

We believe this is an interesting work that is in response to a practical call received from the medical community.  One of the goals of worldwide rehabilitation engineering is to provide solutions to society health care problems, and we understood that in order to tackle these problems effectively, we need to attain design strategies that allow quick feedback from medical personnel.  We hope this project to be the cornerstone of a fruitful collaboration.

Yours truly
Rodrigo Ramele
ITBA University 
Argentina

### 2020-05-13 alpibot.tex
Version submitteada.

@@
     \item \textit{Robot-leader distance, r.l.d.}: The euclidean distance between the robot and the leader, at any point in time.
-This curve is particularly important since the robot has a limited amount of thread available, so if the leader uses all the available thread, it will start dragging the robot and damaging the following mechanis, and overall it may rise the possibility of disconnecting the breathing oxygen cannula.
+This curve is particularly important since the robot has a limited amount of thread available, so if the leader uses all the available thread, it will start dragging the robot and damaging the following mechanism, and overall it may rise the possibility of disconnecting the breathing oxygen cannula.
 This is a scenario that must not happen under any circumstance, as it can also be dangerous for a potential patient using the device.
@@
     \\
-    \vspace*{-2.5em}
+    \vspace*{-1.8em}
   \subfloat[ \label{rw1e}]{%
@@
 
-From the leader and robot trajectories in Figure~\ref{fig:simulationresults}(a to d), both strategies exhibits basic following behaviour.
+From the leader and robot trajectories in Figure~\ref{fig:simulationresults}(a to d), both strategies exhibits basic following behavior.
 However, Rotate and go on Figure~\ref{fig:simulationresults}(b,d) generates a more irregular trajectory, due to the two stage movement algorithm.
@@
 
-%From the leader and robot trajectories in Figure \ref{fig:sim_ft_cv15_cr15}, we can see that the robot using Follow the thread control exhibits basic following behaviour.
+%From the leader and robot trajectories in Figure \ref{fig:sim_ft_cv15_cr15}, we can see that the robot using Follow the thread control exhibits basic following behavior.
 Deviation occurs when the leader trajectory turns with a tight radius.
@@
 
-%The rotateand go We see that this strategy also exhibits a following behaviour, although in a more irregular path, as expected, due to the two stage movement algorithm.
+%The rotateand go We see that this strategy also exhibits a following behavior, although in a more irregular path, as expected, due to the two stage movement algorithm.
 In this case we evaluated the effects of the constants $c_v$, $c_r$ and $Dt_{off}$.
@@
 
-%Both behaviours are described by figures \ref{fig:ft_speed_1} and \ref{fig:rg_speed_1}, which show the speed of the leader and the robot over time.
+%Both behaviors are described by figures \ref{fig:ft_speed_1} and \ref{fig:rg_speed_1}, which show the speed of the leader and the robot over time.
 
@@
 
-In line with the simulated results, a similar behaviour was found on the experiments performed inside the Motion Capture Lab, and the robot exhibits following behaviour for both control strategies as shown on Figure~\ref{fig:realworldresults}.
+In line with the simulated results, a similar behavior was found on the experiments performed inside the Motion Capture Lab, and the robot exhibits following behavior for both control strategies as shown on Figure~\ref{fig:realworldresults}.
 Parameters had to be readjusted for the real world case, but the same situation occurs for both strategies.
 
-%We can see that, again, the robot exhibits following behaviour.
-The trajectories in the figures mentioned above show a similar behaviour to the one in the software simulation, for both control strategies.
+%We can see that, again, the robot exhibits following behavior.
+The trajectories in the figures mentioned above show a similar behavior to the one in the software simulation, for both control strategies.
 
@@
 
-From the practical experiments, it is verified that both algorithms gave reasonable following performance in the task of following the leader along a lemniscate-shaped trajectory.
-This reasonable following behaviour is accomplished with a simple mechanism, a characteristic that significantly keep the price of the device low, putting it within reach of many medical institutions.
-
-Each control strategy has its advantages, but according various metrics described in this work, the \textit{Follow the thread} strategy had a more desirable behaviour, as it tended to follow the leader from a closer distance at all times, while moving in a smooth and predictable way.
+From the practical experiments, it is verified that both algorithms ensuing the \textit{following behavior}  in the task of tracking the patient along a lemniscate-shaped trajectory.
+This following behavior is accomplished with a simple mechanism, a characteristic that significantly keep the price of the device low, putting it within reach of many medical institutions on developing nations.
+
+Each control strategy has its advantages, but according various metrics described in this work, the \textit{Follow the thread} strategy had a more desirable behavior, as it tended to follow the leader from a closer distance at all times, while moving in a smooth and predictable way.
 

### 2020-07-15 review/response.letter.tex
Primera version despues de las revisiones.

\documentclass[journal,onecolumn,12pt]{IEEEtran}

\usepackage{amsmath,amssymb,bm}
\usepackage{amsthm, amsfonts}	
\usepackage{bm,bbm}
\usepackage[normalem]{ulem}
\usepackage{color}
\usepackage{fancybox}
\usepackage{url,booktabs}
\usepackage[round]{natbib}

\usepackage{xr}

\title{Reply to Reviewer's Comments on\\
``A Control Strategy for a Tethered Follower Robot for Pulmonary Rehabilitation''}
\author{}

\begin{document}

\maketitle
\pagenumbering{roman}
\setcounter{page}{1}

We are grateful to the reviewer for pointing out relevant issues in our manuscript.

In the following, we discuss how we dealt with each raised issue.

\vskip+1ex
\noindent \dotfill

\section*{\fbox{Associate Editor Transcript:}}

Comments to the Author:
Authors addressed an important problem, providing a simple yet effective solution. There is no doubt that the device presented in the paper is of great interest for both patients and clinicians. To publish the paper on a robotics journal, though, the technical content has to be deepened so to increase the interest of researchers.
As better detailed by the the two reviewers, the Authors are encouraged to improve the manuscript along two main directions:

\begin{itemize}
\item Clarifying the novelties wrt to state of the art (robot design and robot control), to highlight the original and novel contributions;
\item  Improving the modelling, which should take into account realistic working conditions (e.g. weight and inertial properties of the tank).
\end{itemize}

\section*{\fbox{Reviewer \#1 Transcript:}}

I found the work interesting and tackling an important problem, which can have applications outside support in this specific task. The need to carry oxygen tanks around patients with respiratory problems is high in hospitals and rehabilitation facilities. This solution frees nurses or other medical staff that would be occupied with such tasks and even patients that could carry the tank themselves with such solution can have an easier life in these places.

\vspace{2em}

Overall the article is very well written, I've found some typos to which I'll refer in a momment. The idea is transmitted clearly and the authors did an excellent job motivating the problem and covering the relevant literature. The prototype defined solves the problem in an elegant and simple way and I find the documentation on how the robot was designed sufficient. However, Section II.D, the control strategy section, could have explored more the strategies as some variables and expressions are not entirely clear regarding their aim; despite that the reader can understand how each strategy works and their differences.

\vspace{2em}

Regarding the evaluation and discussion I don't have much to add, I believe you conducted two solid validations of the system that showed the system performs adequately and the conclusions from those studies seem in line with both the results and the clinical assessment performed. My only question is why were the configurations used in the real world study not part of the set of configurations used in the simulation? I understand why you reduced the sets of configurations used in the real world study, but I would expect that those sets would come from the simulations ran not entirely new sets of configurations. I think this part should have been better explained.

\vspace{2em}

Finally, the typos I found:
      in page 2, left side, line 22, at the begin of section II - "Hence, design methodologies that allow rapid prototyping can bring quickly feedback from...", I think it sould be "quick" instead of "quickly"
      in page 3, left side, line 46, at the begin of section II.C - "... and to save battery,
an algorithm to activate and deactivate the motors is developed.", it should be "motors was developed" and not "is"
      in page 7, left side, line 19 - "Both control strategies were expounded, along with the main superficial differences between them.", did you mean "expounded" or "explained"

\vspace{2em}

\section*{\fbox{Reviewer \#2 Transcript:}}

The authors propose a differential tethered robot to follow a COPD patient during rehabilitation. The robot carries the heavy oxygen tank that the patient cannot carry on their own. The tethered robot uses two threads attached to motorized reels to control the robots position w.r.t the patient as the patient performs various rehabilitation exercises. The authors propose two control strategies and conduct both simulated and real-world experiments.

\vspace{2em}

Overall, the reviewer found the application to be interesting; it seems like a great opportunity for a robotic solution. The experiments conducted appear to be reasonably selected. It is unclear, however, what the significance of the work is, particularly in the context of previous work [1] which also investigates a tethered robot solution for COPD rehabilitation. The discernible differences between the proposed work and that within [1] appear to be that the proposed robot has two tethers instead of one, and that the solution can be used in medical facilities instead of home therapy environments (the latter of which would probably have even more complicated navigation environments). There is no comparison between this existing work and the proposed work which, given they are essentially tackling the same application, is problematic: what are the advantages/disadvantages of one design over another? Is there something specifically wrong with the design in [1] that prompted the authors to develop their own approach?

\vspace{2em}

[1]  Endo, Gen, et al. "Mobile follower robot as an assistive device for home oxygen therapy–evaluation of tether control algorithms." ROBOMECH Journal 2.1 (2015): 1-9.

\vspace{2em}

There were also a number of technical elements the reviewer had a hard time understanding:
\begin{itemize}
\item How are "pulses" defined (equation (1)), and how is this value used in an equation?
\item How does a distance multiplied by a constant yield a velocity? (equation (2)); It would seem there is some element of time missing here.  
\item How does a velocity, plus a distance multiplied by a constant yield a power value? (equations (3) and (4)); additionally, using "v" to denote both velocity and power seems like a poor notation choice
\end{itemize}

\vspace{2em}

In addition to these comments, it would seem that a robot being designed to carry a heavy oxygen tank would be designed with said tank in mind. The current calibration of constants and experimental results are not done with the weight of a heavy oxygen tank, and thus it is hard to determine if the design and control techniques would actually succeed in their intended application. Is there a reason that a weight of some kind was not used for the experiments? Can non-weighted experiments be justified somehow?

\vspace{2em}

Given the issues of significance w.r.t. existing work and technical questions raised, the reviewer recommends major revisions.

\vspace{2em}

Detailed minor comments:

\begin{itemize}
\item - The abstract and introduction (specifically) suffer from English/grammar issues. (e.g. "eskeletomuscular" to musculoskeletal); the reviewer recommends the paper be peer-edited by a native speaker if it has not been already.
\item - The footer says the paper has been "revised August 26, 2020", which is of course a date that has not happened yet.
\item - The term $Dt{off}$ is often spelled $Dt{o}ff$ and should be fixed (same with $Dm{off}$).
\item - The caption for Table 1 should be with the actual table, not in a separate column.
\item - In Figure 5 it is not visible that a marker is placed on the side of the robot, nor on the hand of the patient (also the patient is not wearing a glove which is mentioned in the paper).
\end{itemize}

Page 6, first paragraph: "a high cr was important so the robot can turn quickly to point to the leader, but high values of cr should be avoided"; this seems contradictory.

\section*{\fbox{Reviewer \#1 Responses:}}

%\subsection*{\ovalbox{Reviewer 1 General Comments}}

I found the work interesting and tackling an important problem, which can have applications outside support in this specific task. The need to carry oxygen tanks around patients with respiratory problems is high in hospitals and rehabilitation facilities. This solution frees nurses or other medical staff that would be occupied with such tasks and even patients that could carry the tank themselves with such solution can have an easier life in these places.

Overall the article is very well written, I've found some typos to which I'll refer in a momment. The idea is transmitted clearly and the authors did an excellent job motivating the problem and covering the relevant literature. The prototype defined solves the problem in an elegant and simple way and I find the documentation on how the robot was designed sufficient. However, Section II.D, the control strategy section, could have explored more the strategies as some variables and expressions are not entirely clear regarding their aim; despite that the reader can understand how each strategy works and their differences.

\begin{quotation}
{\color{blue}
We appreciate the Reviewers comments.  We revised the notation on Section II.D per reviewers comments.  We also added a section to establish the parallels between a single and double thread configuration.  We clarified more clearly all the variables and their physicial meaning and included an extra Figure to aid the explanation provided.
}
\end{quotation}

Regarding the evaluation and discussion I don't have much to add, I believe you conducted two solid validations of the system that showed the system performs adequately and the conclusions from those studies seem in line with both the results and the clinical assessment performed. My only question is why were the configurations used in the real world study not part of the set of configurations used in the simulation? I understand why you reduced the sets of configurations used in the real world study, but I would expect that those sets would come from the simulations ran not entirely new sets of configurations. I think this part should have been better explained.

\begin{quotation}
{\color{blue}
During the simulation we explored different sets of values for the controller parameters.  Once we arrived to values where we obtain a range of satisfactory results in terms of the metrics that we defined, we moved forward to the real world testing of the prototype.  We found that we had to fine-tunning the parameters to adapt to the real world scenario but not search through the whole space of configurations.  This was one of the values that the simulation step provided us.
}
\end{quotation}

Finally, the typos I found:
      in page 2, left side, line 22, at the begin of section II - "Hence, design methodologies that allow rapid prototyping can bring quickly feedback from...", I think it sould be "quick" instead of "quickly"

 \begin{quotation}
{\color{blue}
Fixed
}
\end{quotation}

      in page 3, left side, line 46, at the begin of section II.C - "... and to save battery, an algorithm to activate and deactivate the motors is developed.", it should be "motors was developed" and not "is"

\begin{quotation}
{\color{blue}
Fixed
}
\end{quotation}

      in page 7, left side, line 19 - "Both control strategies were expounded, along with the main superficial differences between them.", did you mean "expounded" or "explained"

      \begin{quotation}
{\color{blue}
We changed to "explained", which we believe is clearer.  Thanks for the suggestion.
}
\end{quotation}

\section*{\fbox{Reviewer \#2 Responses:}}
%\subsection*{\ovalbox{Reviewer 2 General Comments}}

The authors propose a differential tethered robot to follow a COPD patient during rehabilitation. The robot carries the heavy oxygen tank that the patient cannot carry on their own. The tethered robot uses two threads attached to motorized reels to control the robots position w.r.t the patient as the patient performs various rehabilitation exercises. The authors propose two control strategies and conduct both simulated and real-world experiments.

\vspace{2em}

Overall, the reviewer found the application to be interesting; it seems like a great opportunity for a robotic solution. The experiments conducted appear to be reasonably selected. It is unclear, however, what the significance of the work is, particularly in the context of previous work [1] which also investigates a tethered robot solution for COPD rehabilitation. The discernible differences between the proposed work and that within [1] appear to be that the proposed robot has two tethers instead of one, and that the solution can be used in medical facilities instead of home therapy environments (the latter of which would probably have even more complicated navigation environments). There is no comparison between this existing work and the proposed work which, given they are essentially tackling the same application, is problematic: what are the advantages/disadvantages of one design over another? Is there something specifically wrong with the design in [1] that prompted the authors to develop their own approach?

\vspace{2em}

[1]  Endo, Gen, et al. "Mobile follower robot as an assistive device for home oxygen therapy–evaluation of tether control algorithms." ROBOMECH Journal 2.1 (2015): 1-9.

\begin{quotation}
{\color{blue}
Este es el primer punto con que vamos a tratar de responder, que es el que revive también el Editor asociado.
}
\end{quotation}

There were also a number of technical elements the reviewer had a hard time understanding:
- How are "pulses" defined (equation (1)), and how is this value used in an equation?

\begin{quotation}
{\color{blue}
Pulses are a discrete number obtained directly from the rotary encoder KY-040.  The values are incremented in one direction and decremented in the other direction.  The encoder is connected to the Arduino board and it generates an interruption each time the encoder produces a rotation movement.  PODEMOS ACA DAR DETALLES DE LA CONEXION ELECTRONICA SI QUIERE O DE COMO LO LEE EL CODIGO.
}
\end{quotation}

- How does a distance multiplied by a constant yield a velocity? (equation (2)); It would seem there is some element of time missing here.

\begin{quotation}
{\color{blue}
The parameter $c_v$ is the proportional gain.  As long as the average distance obtained from both threads deviates from the resting distance $d_0$ the desired velocity is increased.
EN EL PAPER DE ENDO USAN EL MISMO ESQUEMA LLAMANDOLO $V_r$.  ES CIERTO QUE NO ES LA "VELOCIDAD" REAL DEL ROBOT SINO MAS BIEN UNA ESTIMACION DEL MOVIMIENTO DEL TARGET ?  ACA NO ESTOY SEGURO HAY QUE PENSARLO MAS.
}
\end{quotation}

- How does a velocity, plus a distance multiplied by a constant yield a power value? (equations (3) and (4)); additionally, using "v" to denote both velocity and power seems like a poor notation choice

\begin{quotation}
{\color{blue}
In this case we shifted the notation to follow the one used in [15] where power assigned to each weels is denoted $\omega$.
}
\end{quotation}

In addition to these comments, it would seem that a robot being designed to carry a heavy oxygen tank would be designed with said tank in mind. The current calibration of constants and experimental results are not done with the weight of a heavy oxygen tank, and thus it is hard to determine if the design and control techniques would actually succeed in their intended application. Is there a reason that a weight of some kind was not used for the experiments? Can non-weighted experiments be justified somehow?

\begin{quotation}
{\color{blue}
@Luciano esta es la simulación que vos estás haciendo.  ESPEREMOS TENER BUENOS RESULTADOS
}
\end{quotation}

Given the issues of significance w.r.t. existing work and technical questions raised, the reviewer recommends major revisions.

Detailed minor comments:

- The abstract and introduction (specifically) suffer from English/grammar issues. (e.g. "eskeletomuscular" to musculoskeletal); the reviewer recommends the paper be peer-edited by a native speaker if it has not been already.

\begin{quotation}
{\color{blue}
We replaced the term with musculoskeletal and verified the English grammar on the abstract and introduction.

@Luciano agarra un canadiense x ahi
@Esteban si querés darle alguna leída para ver si detectás algún error gramatical concreto.
}
\end{quotation}

- The footer says the paper has been "revised August 26, 2020", which is of course a date that has not happened yet.

\begin{quotation}
{\color{blue}
We are very surprised the fast review process.  We truly thank Reviewers for their valuable time.  Dates has now been fixed.
}
\end{quotation}

- The term $Dt{off}$ is often spelled $Dt{o}ff$ and should be fixed (same with $Dm{off}$).

\begin{quotation}
{\color{blue}
We review all the constants definitions to be consistent along the manuscript.
}
\end{quotation}

- The caption for Table 1 should be with the actual table, not in a separate column.

\begin{quotation}
{\color{blue}
This is an issue with Latex reformatting.  WE WILL FIX IT AT THE END.
}
\end{quotation}

- In Figure 5 it is not visible that a marker is placed on the side of the robot, nor on the hand of the patient (also the patient is not wearing a glove which is mentioned in the paper).

\begin{quotation}
{\color{blue}
@Luciano, tenés alguna foto donde sí tengas el guante ?
}
\end{quotation}

Page 6, first paragraph: "a high cr was important so the robot can turn quickly to point to the leader, but high values of cr should be avoided"; this seems contradictory.

\begin{quotation}
{\color{blue}
@Luciano, alguna sugerencia ?
}
\end{quotation}

\vskip+1ex
\noindent \dotfill
\vskip+1ex
\bibliographystyle{mdpi}
\bibliography{article}

\end{document}

### 2020-07-15 alpibot.tex
Primera version despues de las revisiones.

@@
 of  Computer Engineering, Instituto Tecnológico de Buenos Aires (ITBA), Ciudad de Buenos Aires,  Argentina e-mail: rramele@itba.edu.ar}% <-this % stops a space
-\thanks{Manuscript received April 19, 2020; revised August 26, 2020.}}
+\thanks{Manuscript received May 16, 2020; revised July 6, 2020.}}
 
@@
 % The paper headers
-\markboth{IEEE Transactions on Neural Systems and Rehabilitation Engineering}%
+\markboth{IEEE Transactions on Medical Robotics and Bionics}%
 {Bianchi \MakeLowercase{\textit{et al.}} A Control Strategy for a Tethered Follower Robot
@@
 During Pulmonary Rehabilitation patients perform different exercises and walking activities and may require oxygen assistance.
-The oxygen tank cannot be carried by the patient due to the eskeletomuscular atrophy that characterize this pathology and external assistance is required.
+The oxygen tank cannot be carried by the patient due to the musculoskeletal atrophy that characterize this pathology and external assistance is required.
 The assistance to transport the bulky oxygen tank can be provided by a robotic device that follows the patient while performing the physical activities.
@@
 They are characterized as a slowly progressive condition marked by airflow limitation, being cigarette smoking the main etiologic factor~\cite{MacNee2005}.
-This pathology presents an eskeletomuscular atrophy~\cite{Kocsis2016,Wu2012}.
+This pathology presents an musculoskeletal atrophy~\cite{Kocsis2016,Wu2012}.
 In order to carve these after effects a Pulmonary Rehabilitation procedure is a viable treatment for patients.
@@
 To be effective, any technological solution for the medical community requires active involvement of key stakeholders: physicians, care-givers, patients and their families~\cite{Gaggioli2006,Fasola2012,Cherubini2008,Wolpaw2012BraincomputerIP,Salvini2015}.
-Hence, design methodologies that allow rapid prototyping can bring quickly feedback from real users about drawbacks or opportunities for improvements.
+Hence, design methodologies that allow rapid prototyping can bring quick feedback from real users about drawbacks or opportunities for improvements.
 
@@
 \begin{array}{ll}  
-D_l = pulses_{l} \; \frac{2\pi r}{ppr}  \\
+l_l = pulses_{l} \; \frac{2\pi r}{ppr}  \\
 \\
-D_r = pulses_{r} \; \frac{2\pi r}{ppr}
+l_r = pulses_{r} \; \frac{2\pi r}{ppr}
 \end{array}
@@
 \noindent where pulses obtained for each reel encoder are $pulses_{l}$ and $pulses_{r}$ for the left and right encoder.
-These equations provide the estimated values for the left $D_l$ and right $D_r$  thread distance.
+These equations provide the estimated values for the left $l_l$ and right $l_r$  thread distance.
 The variable $r$ is the radius of the reel and $ppr$ is the pulses per revolution (resolution) of the encoder.
@@
 As previously mentioned, an active spring mechanism is also put in place to keep the threads tense.
-However, in order to extend the useful life of the reel motors, and to save battery, an algorithm to  activate and deactivate the motors is developed.
+However, in order to extend the useful life of the reel motors, and to save battery, an algorithm to  activate and deactivate the motors was developed.
 
@@
 \begin{equation}
-v_{tar} = c_v (\frac{D_l + D_r}{2} - d_0)
+v_{tar} = c_v (\frac{l_l + l_r}{2} - l_d)
 \label{eq:lemniscata}
@@
 \begin{equation}
-vleft_{tar} = v_{tar} + c_{\alpha} (D_l - D_r)
+\omega_{L} = v_{tar} + c_{\alpha} (l_l - l_r)
 \label{eq:lemniscata2}
@@
 \begin{equation}
-vright_{tar} = v_{tar} - c_{\alpha} (D_l - D_r)
+\omega_{R} = v_{tar} - c_{\alpha} (l_l - l_r)
 \label{eq:lemniscata3}
@@
 \noindent where  $c_v$ and $c_{\alpha}$ are constant coefficients used for calibration.
-As shown in Figure~\ref{fig:follow_params},  $D_l$ and $D_r$ are thread distances obtained from encoder information from Equation~\ref{eq:encoder_len}, $v_{tar}$ is the estimated forward velocity for the target and finally $vleft_{tar}$ are $vright_{tar}$ are the power values that are directly used to drive power to each wheel motor.
+As shown in Figure~\ref{fig:follow_params},  $l_l$ and $l_r$ are thread distances obtained from encoder information from Equation~\ref{eq:encoder_len}, $v_{tar}$ is the estimated forward velocity for the target and finally $\omega_{L}$ are $\omega_{R}$ are the power values that are directly used to drive power to each wheel motor.
 
@@
 \begin{equation}
-\textit{if  }{\frac{D_l + D_r}{2} < d_0}\textit{ then } vleft_{tar} = vright_{tar} = 0 
+\textit{if  }{\frac{l_l + l_r}{2} < l_d}\textit{ then } \omega_{L} = \omega_{R} = 0 
 \end{equation}
 
-\noindent where $d_0$ is a constant offset that is used to customize the length of the thread where the robot does not move.
+\noindent where $l_d$ is a constant offset that is used to customize the length of the thread where the robot does not move.
 
@@
     \draw (524,1960) node  [font=\large,color={rgb, 255:red, 74; green, 144; blue, 226 }  ,opacity=1 ]  {$ \begin{array}{l}
-    D_{r}\\
+    l_{r}\\
     \end{array}$};
@@
     \draw (510,1910) node  [font=\large,color={rgb, 255:red, 74; green, 144; blue, 226 }  ,opacity=1 ,rotate=-358.27]  {$ \begin{array}{l}
-    D_{l}\\
+    l_{l}\\
     \end{array}$};
@@
     % Text Node
-    \draw (390.58,1925.93) node [anchor=west] [inner sep=0.75pt]  [font=\normalsize,color={rgb, 255:red, 0; green, 0; blue, 0 }  ,opacity=1 ]  {$vleft_{tar}$};
-    % Text Node
-    \draw (407.58,2100.93) node [anchor=west] [inner sep=0.75pt]  [font=\normalsize,color={rgb, 255:red, 0; green, 0; blue, 0 }  ,opacity=1 ]  {$vright_{tar}$};
+    \draw (390.58,1925.93) node [anchor=west] [inner sep=0.75pt]  [font=\normalsize,color={rgb, 255:red, 0; green, 0; blue, 0 }  ,opacity=1 ]  {$\omega_{L}$};
+    % Text Node
+    \draw (407.58,2100.93) node [anchor=west] [inner sep=0.75pt]  [font=\normalsize,color={rgb, 255:red, 0; green, 0; blue, 0 }  ,opacity=1 ]  {$\omega_{R}$};
 
@@
 The variable $V_r$ is the speed at which the vehicle will rotate on its axis, and $V_f$ is the speed at which the vehicle will move forward once it can move on the subject's direction.
-The algorithm requires three parameters, $c_v$ and $c_r$ that regulates the coefficients for the forward and rotation movement, and an additional parameter $Dt_off$ that regulates the sensibility of the rotation movement.
-The constant parameter $Dm_off$ is similar to $d_0$, and is used to customize the length of the thread where the robot does not move at all.
+The algorithm requires three parameters, $c_v$ and $c_r$ that regulates the coefficients for the forward and rotation movement, and an additional parameter $Dt_{off}$ that regulates the sensibility of the rotation movement.
+The constant parameter $Dm_{off}$ is similar to $l_d$, and is used to customize the length of the thread where the robot does not move at all.
 
@@
 \begin{algorithmic}
-\STATE $Dt \gets D_l - D_r$
-\STATE $Dm \gets \frac{D_r+D_l}{2}$
-\STATE $V_r \gets c_r * (abs(Dt) - Dt_off) + base_{vr}$ 
-\STATE $V_f \gets c_v * (Dm - Dm_off)$
+\STATE $Dt \gets l_l - l_r$
+\STATE $Dm \gets \frac{l_r+l_l}{2}$
+\STATE $V_r \gets c_r * (abs(Dt) - Dt_{off}) + base_{vr}$ 
+\STATE $V_f \gets c_v * (Dm - Dm_{off})$
 
 \IF{$abs(Dt) > Dt_{off}$}
-\IF{$D_l > D_r$}
-\STATE $vleft_{tar} \gets V_r$
-\STATE $vright_{tar} \gets -V_r$
+\IF{$l_l > l_r$}
+\STATE $\omega_{L} \gets V_r$
+\STATE $\omega_{R} \gets -V_r$
 \ELSE
-\STATE $vleft_{tar} \gets -V_r$
-\STATE $vright_{tar} \gets V_r$
+\STATE $\omega_{L} \gets -V_r$
+\STATE $\omega_{R} \gets V_r$
 \ENDIF
@@
 \IF{$D_m > Dm_{off}$}
-\STATE $vright_{tar} \gets V_f$
-\STATE $vleft_{tar} \gets V_f$
+\STATE $\omega_{R} \gets V_f$
+\STATE $\omega_{L} \gets V_f$
 \ELSE
-\STATE $vright_{tar} \gets 0$
-\STATE $vleft_{tar} \gets 0$
+\STATE $\omega_{R} \gets 0$
+\STATE $\omega_{L} \gets 0$
 \ENDIF
 \ENDIF
-\STATE Return $vright_{tar}$ and $vleft_{tar}$.
+\STATE Return $\omega_{R}$ and $\omega_{L}$.
 \end{algorithmic}
@@
 \end{algorithm}
+
+\section{Equivalence between the single and double tethered system} \label{development}
+
+\begin{figure}[]
+    \centering
+    \tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt
+
+\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
+%uncomment if require: \path (0,235); %set diagram left start at 0, and has height of 235
+
+%Shape: Axis 2D [id:dp7543865369052545] 
+\draw  (312,162.89) -- (521,162.89)(332.9,2.69) -- (332.9,180.69) (514,157.89) -- (521,162.89) -- (514,167.89) (327.9,9.69) -- (332.9,2.69) -- (337.9,9.69)  ;
+%Rounded Same Side Corner Rect [id:dp31787525316857024] 
+\draw   (297.9,150.89) ..
+controls (297.9,146.47) and (301.48,142.89) ..
+(305.9,142.89) -- (359.9,142.89) ..
+controls (364.32,142.89) and (367.9,146.47) ..
+(367.9,150.89) -- (367.9,182.89) ..
+controls (367.9,182.89) and (367.9,182.89) ..
+(367.9,182.89) -- (297.9,182.89) ..
+controls (297.9,182.89) and (297.9,182.89) ..
+(297.9,182.89) -- cycle ;
+%Shape: Square [id:dp8026938389823051] 
+\draw   (434,151.69) -- (457,151.69) -- (457,174.69) -- (434,174.69) -- cycle ;
+%Shape: Square [id:dp06857182701706144] 
+\draw   (207.66,149.66) -- (231.34,149.66) -- (231.34,173.34) -- (207.66,173.34) -- cycle ;
+%Flowchart: Summing Junction [id:dp2757072161924359] 
+\draw   (317.9,162.89) ..
+controls (317.9,155.16) and (324.62,148.89) ..
+(332.9,148.89) ..
+controls (341.18,148.89) and (347.9,155.16) ..
+(347.9,162.89) ..
+controls (347.9,170.62) and (341.18,176.89) ..
+(332.9,176.89) ..
+controls (324.62,176.89) and (317.9,170.62) ..
+(317.9,162.89) -- cycle ; \draw   (322.29,152.99) -- (343.51,172.79) ; \draw   (343.51,152.99) -- (322.29,172.79) ;
+%Straight Lines [id:da9717478160302964] 
+\draw [line width=2.25]    (405,12.69) -- (332.9,162.89) ;
+%Straight Lines [id:da30114532827552765] 
+\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=2.25]    (405,12.69) -- (219.5,161.5) ;
+%Straight Lines [id:da0582330345196298] 
+\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=2.25]    (405,12.69) -- (445.5,163.19) ;
+%Flowchart: Terminator [id:dp11720735085928635] 
+\draw   (393.1,2.69) -- (416.9,2.69) ..
+controls (419.99,2.69) and (422.5,7.16) ..
+(422.5,12.69) ..
+controls (422.5,18.21) and (419.99,22.69) ..
+(416.9,22.69) -- (393.1,22.69) ..
+controls (390.01,22.69) and (387.5,18.21) ..
+(387.5,12.69) ..
+controls (387.5,7.16) and (390.01,2.69) ..
+(393.1,2.69) -- cycle ;
+%Straight Lines [id:da6471016295019487] 
+\draw  [dash pattern={on 0.84pt off 2.51pt}]  (405,12.69) -- (406,162.69) ;
+%Shape: Circle [id:dp5568563915916713] 
+\draw   (403.5,162.69) ..
+controls (403.5,161.31) and (404.62,160.19) ..
+(406,160.19) ..
+controls (407.38,160.19) and (408.5,161.31) ..
+(408.5,162.69) ..
+controls (408.5,164.07) and (407.38,165.19) ..
+(406,165.19) ..
+controls (404.62,165.19) and (403.5,164.07) ..
+(403.5,162.69) -- cycle ;
+%Straight Lines [id:da6605293994654671] 
+\draw    (188,161.69) -- (332.9,162.89) ;
+%Shape: Brace [id:dp034761867026847026] 
+\draw   (221,180.69) ..
+controls (220.93,185.36) and (223.22,187.73) ..
+(227.88,187.8) -- (302.39,189.02) ..
+controls (309.06,189.13) and (312.35,191.52) ..
+(312.27,196.19) ..
+controls (312.35,191.52) and (315.72,189.24) ..
+(322.38,189.35)(319.38,189.3) -- (396.89,190.57) ..
+controls (401.56,190.65) and (403.93,188.36) ..
+(404,183.69) ;
+%Straight Lines [id:da6885647931399865] 
+\draw [line width=0.75]  [dash pattern={on 0.84pt off 2.51pt}]  (333,13.69) -- (405,12.69) ;
+%Curve Lines [id:da2681192753910786] 
+\draw    (351.04,132.38) .. controls (360.75,134.84) and (374.54,141.11) .. (375.89,158.8) ;
+\draw [shift={(376,161.69)}, rotate = 270] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;
+\draw [shift={(348,131.69)}, rotate = 11.31] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;
+%Curve Lines [id:da7846164710785513] 
+\draw    (334.02,102.11) .. controls (342.64,102.92) and (351.11,101.09) .. (351.93,117.88) ;
+\draw [shift={(352,120.69)}, rotate = 270] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;
+\draw [shift={(331,101.69)}, rotate = 11.31] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;
+%Shape: Brace [id:dp8659813161256731] 
+\draw   (220,203.69) ..
+controls (219.98,208.36) and (222.3,210.7) ..
+(226.97,210.72) -- (322.47,211.13) ..
+controls (329.14,211.16) and (332.46,213.51) ..
+(332.44,218.18) ..
+controls (332.46,213.51) and (335.8,211.19) ..
+(342.47,211.22)(339.47,211.21) -- (437.97,211.64) ..
+controls (442.64,211.66) and (444.98,209.34) ..
+(445,204.67) ;
+
+% Text Node
+\draw (399,166.09) node [anchor=north west][inner sep=0.75pt]    {$x$};
+% Text Node
+\draw (305,12.4) node [anchor=north west][inner sep=0.75pt]    {$y$};
+% Text Node
+\draw (438,173.4) node [anchor=north west][inner sep=0.75pt]    {$x'$};
+% Text Node
+\draw (308,193.4) node [anchor=north west][inner sep=0.75pt]    {$x''$};
+% Text Node
+\draw (357,52.4) node [anchor=north west][inner sep=0.75pt]    {$l_{m}$};
+% Text Node
+\draw (439,90.4) node [anchor=north west][inner sep=0.75pt]    {$l_{R}$};
+% Text Node
+\draw (261,90.4) node [anchor=north west][inner sep=0.75pt]    {$l_{L}$};
+% Text Node
+\draw (333,114.4) node [anchor=north west][inner sep=0.75pt]    {$\theta $};
+% Text Node
+\draw (372,122.4) node [anchor=north west][inner sep=0.75pt]    {$\gamma $};
+% Text Node
+\draw (334,215.4) node [anchor=north west][inner sep=0.75pt]    {$b$};
+
+\end{tikzpicture}
+\label{fig:dibujito}
+\caption{Following mechanism parameters}
+\end{figure}
+
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[width=8cm]{images/dibujito.jpg}
+%\caption{Scheme of the equivalence between a single and dual thread systems.}
+%\label{fig:dibujito}
+%\end{figure}
+
+In~\cite{Endo2015} authors propose a similar design based on only one thread.
+They propose two control strategies based on two input parameters that are obtained from a linear and circular potentiometer that determines the length of the thread $l_m$ and the orientation angle $\theta$.
+We show here that they are equivalent to the approach presented based on $l_l$ and  $l_r$.
+
+Based on a frame reference with the robot on the center of coordinates as shown in Figure~\ref{fig:dibujito}, from the input values $l_m$ and $\theta$ we can derive the position of the leader as
+
+\begin{align*}
+   y &= l_m  \cos(\theta)\\
+   x &= l_m  \sin(\theta) \\
+\end{align*}
+
+%From the figure, it can be obtained that
+%
+%\begin{align*}
+%x'            	&= \frac{b}{2}  - x \\
+%x'' 			&= \frac{b}{2}  + x \\
+%\end{align*}
+%
+
+\noindent where $x$ is the follower distance on the horizontal direction, and $y$ the follower distance on the vertical direction.
+
+In this frame of reference, the lengths of the threads, $l_r$ and $l_l$ can be determined as
+
+\begin{equation}
+\begin{array}{ll}  
+l_r &= \sqrt[2]{y^2 + {x'}^2} \\
+l_l &= \sqrt[2]{y^2 + {x''}^2}. \\
+\end{array}
+\label{eq:threadsincoordinates}
+\end{equation}
+
+From Figure~\ref{fig:dibujito} we can see that ${x''} + x' =  b$. Combining this with~\ref{eq:threadsincoordinates}, we can form a system of equations to determine $x'$, $x''$ and $y$:
+
+\begin{align*}
+{x''} + x'					&=  b &&\text{(a)}  \\
+{x'}^2 +  y^2                &= l_r &&\text{(b)} \\
+{x''}^2 + y^2                &= l_l  &&\text{(c)}. \\
+\end{align*}
+
+From the first equation, rearranging and squaring both terms leads to
+
+\begin{equation}
+\begin{array}{ll}  
+{x'}^2   	&=  ( b - x'')^2   \\
+ 						&= b^2 - 2 b x''+ {x''}^2. \\
+\end{array}
+\label{eq:thirdstep}
+\end{equation}
+
+By substracting (b) and (c), it can be obtained
+
+\begin{align*}
+{x'}^2 - {x'' }^2 = {l_r}^2 - {l_l}^2 \\
+{x'}^2 - {x'' }^2 = {l_r}^2 - {l_l}^2, \\
+\end{align*}
+
+and replacing ${x'}^2$ from Equation~\ref{eq:thirdstep} in it
+
+\begin{align*}
+b^2 - 2 b {x''}= {l_r}^2 - {l_l}^2. \\
+\end{align*}
+
+From this equation $x''$ can be determined as
+
+\begin{align*}
+x'' = \frac{({l_d}^2 - {l_i}^2 - b^2)}{-2 \; b}.    \\
+\end{align*}
+
+From Figure~\ref{fig:dibujito} it can also be seen that $\frac{b}{2}$ is equals to $x'' - x$.
+Hence this can be used to finally determine $x$ and $y$ values based on the threads length $l_l$ and $l_r$ and the axle length $b$.
+The $y$ value comes from Equation~\ref{eq:threadsincoordinates}.
+
+\begin{align*}
+y &= \sqrt[2]{{l_l}^2 - {x''}^2 } \\
+x &= x'' - \frac{b}{2}\\
+\end{align*}
+
+Finally, from trigonometry, it can be seen that
+
+\begin{align*}
+\sin{\gamma'} = \frac{x}{l_m}\\
+\end{align*}
+
+and
+
+\begin{align*}
+\cos{\gamma'} = \frac{y}{l_m}\\
+\end{align*}
+
+\noindent and as $\theta = \gamma'$, we can obtain the values of $\theta$ and $l_m$:
+
+\begin{align*}
+\theta &= \arctan{ \frac{x}{y}} \\
+\l_m    &= \frac{y}{\cos{\theta}}. \\
+\end{align*}
+
+There is no loss of information and both systems are equivalent.
+
+Additionally in the proposed system based on circular spring, the spring tension depends on the released thread length, hence the recovery force will vary accordingly.
+The proposed scheme allows a more controlled situation and depends exclusively on the software that controls the reel motor, which can be regulated.
+Another issue that can be presented is when $l_m$ is equals to $l_d$.
+If the patient performs a circular movement where these values are equal, the robot will not change direction at all, because $\omega_{r}$ will be also zero.
+If the patient restart the movement, $\omega_{r}$ could be very low and the $K_p$ constant could not be enough to match the person movement rapidly.
+
+Regarding the second approach, the dead-reckoning could be challenging and requires very precise measurements to reduce the error-prone characteristic of this solution~\cite{DurrantWhyte1994}.
+
+%Otro problema que encontré es en el segundo método de control.
+Éste es un poco más complicado de explicar con palabras así que haré un gráfico y se los envío.
+Como adelanto, está en la ecuación 17 del paper que es cuando toman como medida de la diferencia de distancias (l_m - l_d) para el cálculo de la V_r.
+
+%Por último, un problema del algoritmo de control de los tipos (el de follow-the-leader) es que está basado en "dead-reckoning" el cual es un método impreciso cuyo error crece a lo largo del tiempo.
+Esto es, la estrategia de los tipos es ir estimando la posición del paciente en base a la posición del robot en un sistema de coordenadas universal (no en el sistema de coordenadas cuyo centro es el robot).
+Esta estimación la obtienen, entiendo yo, basado en la información de los encoders de los motores del robot (ahí es donde meten la pata) y sumándole un vector que obtienen al rotar (l_m, 0) mediante una matriz de rotación con el ángulo expresado en el sistema de referencia universal).
+Es importante remarcar que su estrategia se basa en esto: en poder estimar la posición del paciente e ir guardándola (para luego estimar T(s_t const) y para eso usan dead reckoning, es decir, a medida que pasa el tiempo (en rigor, a medida que mayor longitud recorren) peor será la estimación y peor andará el algoritmo.
+Es cierto que en las curvas no se vé eso, pero también es cierto que, en una sesión terapéutica, recorran más distancia que las mostradas en los test.
 
@@
 In the demonstration, the robot design was outlined and an explanation was given on how the robot worked, how it was built and how to operate it.
-Both control strategies were expounded, along with the main superficial differences between them.
+Both control strategies were explained, along with the main superficial differences between them.
 

### 2020-07-19 alpibot.tex
Subo una nueva version que incorpora los nuevos graficos que hizo Luciano, la explicacion de Juan sobre la equivalencia entre los mecanismos con uno y dos hilos, y varios ajustes que pidieron los revisores.

@@
 
-\IEEEPARstart{C}{hronic} Obstructive Pulmonary Disease (COPD) is an umbrella term that describes several pulmonary affections.
+%\IEEEPARstart{C}{hronic} Obstructive Pulmonary Disease (COPD) is an umbrella term that describes several pulmonary affections.
 They are characterized as a slowly progressive condition marked by airflow limitation, being cigarette smoking the main etiologic factor~\cite{MacNee2005}.
-This pathology presents an musculoskeletal atrophy~\cite{Kocsis2016,Wu2012}.
+This pathology presents a musculoskeletal atrophy~\cite{Kocsis2016,Wu2012}.
 In order to carve these after effects a Pulmonary Rehabilitation procedure is a viable treatment for patients.
 During this rehabilitation, patients undergo controlled walking activities and physical exercises under the supervision of a physical therapist.
-Patients present a severe low oxygen saturation illness and they require effective oxygen supply, particularly when performing physical activity~\cite{Celli2014}.
+Patients present a severe low oxygen saturation illness and they require effective oxygen supply, particularly when performing physical activities~\cite{Celli2014}.
 Hence, patients require to carry with them an oxygen tank for the oxygenotherapy assistance.
@@
 
+\IEEEPARstart{C}{hronic} Obstructive Pulmonary Disease (COPD) is an umbrella term that describes several pulmonary affections.
+They are characterized as a slowly progressive condition marked by airflow limitation, being cigarette smoking the main etiologic factor~\cite{MacNee2005}.
+This pathology presents a musculoskeletal atrophy~\cite{Kocsis2016,Wu2012}.
+In order to carve these after effects a Pulmonary Rehabilitation procedure is a viable treatment for patients.
+Pulmonary Rehabilitation procedures consist of controlled waking activities and physical exercises that patients perform under the supervision of a physical therapist.
+However, COPD patients present a severe low oxygen saturation illness and they require oxygen supply, particularly when performing physical activities~\cite{Celli2014}.
+Hence, they are required to carry an oxygen tank for the oxygenotherapy assistance, but their own condition prevent them with the ability to precisely carry the often bulky external tank.
+This situation entails to find a pragmatic solution to avoid an additional physical therapist to carry the oxygen tank.
+The scenario may be aggravated by the fact that this procedure is performed on a rehabilitation gym that could be potentially crowded with several patients, physiologists and physical therapists.
+
 %Alternatively, they are required to improvise a customized wheelchair to allow them to carry the oxygen tank on top of it, leading to a cumbersome situation for the physical therapist because s/he had to push and maneuver the wheelchair, and at the same time, take care of the patient and perform the rehabilitation treatment.
 
-An alternative solution is to use an assitive ground service robot~\cite{Neto2015} to carry the oxygen tank, while following the patient in a leader-follower configuration.
+An alternative solution is to use an assitive ground service robot~\cite{Neto2015} to carry the oxygen tank while following the patient in a leader-follower configuration.
 There are two reasons that support the initial viability of this idea.
@@
 On the other hand, the range of movements performed by the patient during the Rehabilitation Procedure is highly predictable by the treatment.
-At the same time,  the global robotic research community looks forward for the development of robotic affordable solutions to the social and health-related worldwide problems~\cite{Khamis2019}.
+At the same time,  the global robotic research community looks forward for the development of robotic affordable solutions to social and health-related worldwide problems~\cite{Khamis2019}.
 
@@
     \draw  [draw opacity=0][fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ]  (478.5,1350) -- (535.5,1350) -- (535.5,1374) -- (478.5,1374) -- cycle  ;
-    \draw (507,1362) node    {$subject$};
+    \draw (517,1362) node    {$subject$};
     % Text Node
@@
 Encoders in each reel measure the difference in length of each thread compared to its initial position.
-This difference in length is the input for the control algorithm.
+These differences in length are the input for the control algorithm.
 Using the encoder, the difference in length for each thread can be measured with Equations~\ref{eq:encoder_len}.
@@
 \begin{array}{ll}  
-l_l = pulses_{l} \; \frac{2\pi r}{ppr}  \\
+l_L = pulses_{l} \; \frac{2\pi r}{ppr}  \\
 \\
-l_r = pulses_{r} \; \frac{2\pi r}{ppr}
+l_R = pulses_{r} \; \frac{2\pi r}{ppr}
 \end{array}
@@
 \noindent where pulses obtained for each reel encoder are $pulses_{l}$ and $pulses_{r}$ for the left and right encoder.
-These equations provide the estimated values for the left $l_l$ and right $l_r$  thread distance.
+These equations provide the estimated values for the left $l_L$ and right $l_R$  thread distance.
 The variable $r$ is the radius of the reel and $ppr$ is the pulses per revolution (resolution) of the encoder.
@@
 This control strategy is similar to the one presented in \cite{Ortlieb2016}.
-It is based on the idea that the difference in length between both threads provides a quasi-linear function of the relative angle between the subject and the vehicle orientation and, additionally, the mean of both thread length allows to approximate the relative distance between the robot and the subject.
+It is based on the idea that both the relative angle between the subject and the vehicle orientation and, the relative distance between the robot and the subject, can be computed from the length of the left and right threads.
 They are described by Equations~\ref{eq:lemniscata}, \ref{eq:lemniscata2} and~\ref{eq:lemniscata3}.
@@
 \begin{equation}
-v_{tar} = c_v (\frac{l_l + l_r}{2} - l_d)
+V_t = c_v (\frac{l_L + l_R}{2} - l_D)
 \label{eq:lemniscata}
@@
 \begin{equation}
-\omega_{L} = v_{tar} + c_{\alpha} (l_l - l_r)
+\omega_{L} = V_t + c_{\alpha} (l_L - l_R)
 \label{eq:lemniscata2}
@@
 \begin{equation}
-\omega_{R} = v_{tar} - c_{\alpha} (l_l - l_r)
+\omega_{R} = V_t - c_{\alpha} (l_L - l_R)
 \label{eq:lemniscata3}
@@
 
-\noindent where  $c_v$ and $c_{\alpha}$ are constant coefficients used for calibration.
-As shown in Figure~\ref{fig:follow_params},  $l_l$ and $l_r$ are thread distances obtained from encoder information from Equation~\ref{eq:encoder_len}, $v_{tar}$ is the estimated forward velocity for the target and finally $\omega_{L}$ are $\omega_{R}$ are the power values that are directly used to drive power to each wheel motor.
+\noindent where  $c_v$ and $c_{\alpha}$ are constant coefficients used for calibration whose units are expressed in $[\frac{1}{s}]$, and $l_D$ is a constant offset in $[m]$ that is used to customize the desired length of the thread where the robot does not move.
+As shown on Figure~\ref{fig:follow_params},  $l_L$ and $l_R$ are thread distances in $[m]$  obtained from encoder information from Equation~\ref{eq:encoder_len}, $V_t$ is the estimated forward velocity for the target and finally $\omega_{L}$ are $\omega_{R}$ are the velocity values that are directly used to drive each wheel motor.
 
@@
 \begin{equation}
-\textit{if  }{\frac{l_l + l_r}{2} < l_d}\textit{ then } \omega_{L} = \omega_{R} = 0 
+\textit{if  }{\frac{l_L + l_R}{2} < l_D}\textit{ then } \omega_{L} = \omega_{R} = 0.
 \end{equation}
-
-\noindent where $l_d$ is a constant offset that is used to customize the length of the thread where the robot does not move.
 
@@
     \draw (524,1960) node  [font=\large,color={rgb, 255:red, 74; green, 144; blue, 226 }  ,opacity=1 ]  {$ \begin{array}{l}
-    l_{r}\\
+    l_{R}\\
     \end{array}$};
@@
     \draw (510,1910) node  [font=\large,color={rgb, 255:red, 74; green, 144; blue, 226 }  ,opacity=1 ,rotate=-358.27]  {$ \begin{array}{l}
-    l_{l}\\
+    l_{L}\\
     \end{array}$};
@@
 The algorithm requires three parameters, $c_v$ and $c_r$ that regulates the coefficients for the forward and rotation movement, and an additional parameter $Dt_{off}$ that regulates the sensibility of the rotation movement.
-The constant parameter $Dm_{off}$ is similar to $l_d$, and is used to customize the length of the thread where the robot does not move at all.
+The constant parameter $Dm_{off}$ is similar to $l_D$, and is used to customize the length of the thread where the robot does not move at all.
 
@@
 \begin{algorithmic}
-\STATE $Dt \gets l_l - l_r$
-\STATE $Dm \gets \frac{l_r+l_l}{2}$
+\STATE $Dt \gets l_L - l_R$
+\STATE $Dm \gets \frac{l_R+l_L}{2}$
 \STATE $V_r \gets c_r * (abs(Dt) - Dt_{off}) + base_{vr}$ 
@@
 \IF{$abs(Dt) > Dt_{off}$}
-\IF{$l_l > l_r$}
+\IF{$l_L > l_R$}
 \STATE $\omega_{L} \gets V_r$
@@
 
-\section{Equivalence between the single and double tethered system} \label{development}
+\section{Comparision with alternative methods} \label{development}
+
+\subsection{Equivalence between the single and double tethered system}
+
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[width=8cm]{images/dibujito.jpg}
+%\caption{Scheme of the equivalence between a single and dual thread systems.}
+%\label{fig:dibujito}
+%\end{figure}
+
+In~\cite{Endo2015} authors propose a similar design based on only one thread.
+They propose two control strategies based on two input parameters that are obtained from a linear and circular potentiometer that determines the length of the thread $l_m$ and the orientation angle $\theta$.
+We show here that they are equivalent to the approach presented based on $l_L$ and  $l_R$.
+
+Based on a frame reference with the robot on the center of coordinates as shown on Figure~\ref{fig:dibujitos}, from the input values $l_m$ and $\theta$ we can derive the position of the leader as
+
+\begin{align*}
+   y &= l_m  \cos(\theta)\\
+   x &= l_m  \sin(\theta) \\
+\end{align*}
+
+%From the figure, it can be obtained that
+%
+%\begin{align*}
+%x'            	&= \frac{b}{2}  - x \\
+%x'' 			&= \frac{b}{2}  + x \\
+%\end{align*}
+%
+
+\noindent where $x$ is the follower distance on the horizontal direction, and $y$ the follower distance on the vertical direction.
+
+In this frame of reference, the lengths of the threads, $l_R$ and $l_L$ can be determined as
 
@@
 %Curve Lines [id:da2681192753910786] 
-\draw    (351.04,132.38) .. controls (360.75,134.84) and (374.54,141.11) .. (375.89,158.8) ;
-\draw [shift={(376,161.69)}, rotate = 270] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;
-\draw [shift={(348,131.69)}, rotate = 11.31] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;
+\draw    (384.56,66.54) .. controls (391.15,78.31) and (389.42,81.22) .. (403.17,79.98) ;
+\draw [shift={(406,79.69)}, rotate = 533.6600000000001] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;
+\draw [shift={(382.95,63.79)}, rotate = 58.73] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;
 %Curve Lines [id:da7846164710785513] 
@@
 (445,204.67) ;
+%Shape: Brace [id:dp4829292799998628] 
+\draw   (409,183.69) ..
+controls (409,188.36) and (411.33,190.69) ..
+(416,190.69) -- (418.5,190.69) ..
+controls (425.17,190.69) and (428.5,193.02) ..
+(428.5,197.69) ..
+controls (428.5,193.02) and (431.83,190.69) ..
+(438.5,190.69)(435.5,190.69) -- (441,190.69) ..
+controls (445.67,190.69) and (448,188.36) ..
+(448,183.69) ;
+%Shape: Circle [id:dp5543776560649496] 
+\draw   (330.5,13.69) ..
+controls (330.5,12.31) and (331.62,11.19) ..
+(333,11.19) ..
+controls (334.38,11.19) and (335.5,12.31) ..
+(335.5,13.69) ..
+controls (335.5,15.07) and (334.38,16.19) ..
+(333,16.19) ..
+controls (331.62,16.19) and (330.5,15.07) ..
+(330.5,13.69) -- cycle ;
 
@@
 % Text Node
-\draw (305,12.4) node [anchor=north west][inner sep=0.75pt]    {$y$};
-% Text Node
-\draw (438,173.4) node [anchor=north west][inner sep=0.75pt]    {$x'$};
+\draw (312,2.4) node [anchor=north west][inner sep=0.75pt]    {$y$};
+% Text Node
+\draw (422,194.4) node [anchor=north west][inner sep=0.75pt]    {$x'$};
 % Text Node
@@
 % Text Node
-\draw (333,114.4) node [anchor=north west][inner sep=0.75pt]    {$\theta $};
-% Text Node
-\draw (372,122.4) node [anchor=north west][inner sep=0.75pt]    {$\gamma $};
+\draw (342,86.4) node [anchor=north west][inner sep=0.75pt]    {$\theta $};
+% Text Node
+\draw (379,76.4) node [anchor=north west][inner sep=0.75pt]    {$\gamma $};
 % Text Node
@@
 \end{tikzpicture}
-\label{fig:dibujito}
-\caption{Following mechanism parameters}
-\end{figure}
-
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[width=8cm]{images/dibujito.jpg}
-%\caption{Scheme of the equivalence between a single and dual thread systems.}
-%\label{fig:dibujito}
-%\end{figure}
-
-In~\cite{Endo2015} authors propose a similar design based on only one thread.
-They propose two control strategies based on two input parameters that are obtained from a linear and circular potentiometer that determines the length of the thread $l_m$ and the orientation angle $\theta$.
-We show here that they are equivalent to the approach presented based on $l_l$ and  $l_r$.
-
-Based on a frame reference with the robot on the center of coordinates as shown in Figure~\ref{fig:dibujito}, from the input values $l_m$ and $\theta$ we can derive the position of the leader as
-
-\begin{align*}
-   y &= l_m  \cos(\theta)\\
-   x &= l_m  \sin(\theta) \\
-\end{align*}
-
-%From the figure, it can be obtained that
-%
-%\begin{align*}
-%x'            	&= \frac{b}{2}  - x \\
-%x'' 			&= \frac{b}{2}  + x \\
-%\end{align*}
-%
-
-\noindent where $x$ is the follower distance on the horizontal direction, and $y$ the follower distance on the vertical direction.
-
-In this frame of reference, the lengths of the threads, $l_r$ and $l_l$ can be determined as
+
+\caption{The center of coordinates is on the robot center of mass, where the y positive axis points in the same direction as the robot direction.}
+\label{fig:dibujitos}
+\end{figure}{}
 
@@
 \begin{array}{ll}  
-l_r &= \sqrt[2]{y^2 + {x'}^2} \\
-l_l &= \sqrt[2]{y^2 + {x''}^2}. \\
+l_R &= \sqrt[2]{y^2 + {x'}^2} \\
+l_L &= \sqrt[2]{y^2 + {x''}^2}. \\
 \end{array}
@@
 
-From Figure~\ref{fig:dibujito} we can see that ${x''} + x' =  b$. Combining this with~\ref{eq:threadsincoordinates}, we can form a system of equations to determine $x'$, $x''$ and $y$:
+From Figure~\ref{fig:dibujitos} we can see that ${x''} + x' =  b$, with $b$ the axle length.
+Combining this with~\ref{eq:threadsincoordinates}, we can form a system of equations to determine $x'$, $x''$ and $y$:
 
@@
 {x''} + x'					&=  b &&\text{(a)}  \\
-{x'}^2 +  y^2                &= l_r &&\text{(b)} \\
-{x''}^2 + y^2                &= l_l  &&\text{(c)}. \\
+{x'}^2 +  y^2                &= {l_R}^2 &&\text{(b)} \\
+{x''}^2 + y^2                &= {l_L}^2  &&\text{(c)}. \\
 \end{align*}
 
-From the first equation, rearranging and squaring both terms leads to
+From the first equation (a), rearranging and squaring both terms leads to
 
@@
 \begin{align*}
-{x'}^2 - {x'' }^2 = {l_r}^2 - {l_l}^2 \\
-{x'}^2 - {x'' }^2 = {l_r}^2 - {l_l}^2, \\
+{x'}^2 - {x'' }^2 = {l_R}^2 - {l_L}^2 \\
 \end{align*}
@@
 \begin{align*}
-b^2 - 2 b {x''}= {l_r}^2 - {l_l}^2. \\
+b^2 - 2 b {x''}= {l_R}^2 - {l_L}^2. \\
 \end{align*}
@@
 \begin{align*}
-x'' = \frac{({l_d}^2 - {l_i}^2 - b^2)}{-2 \; b}.    \\
+x'' = \frac{({l_R}^2 - {l_L}^2 - b^2)}{-2 \; b}.    \\
 \end{align*}
 
-From Figure~\ref{fig:dibujito} it can also be seen that $\frac{b}{2}$ is equals to $x'' - x$.
-Hence this can be used to finally determine $x$ and $y$ values based on the threads length $l_l$ and $l_r$ and the axle length $b$.
+From Figure~\ref{fig:dibujitos} it can also be seen that $\frac{b}{2}$ is equals to $x'' - x$.
+Hence this can be used to finally determine $x$ and $y$ values based on the threads length $l_L$ and $l_R$ and the axle length $b$.
 The $y$ value comes from Equation~\ref{eq:threadsincoordinates}.
@@
 \begin{align*}
-y &= \sqrt[2]{{l_l}^2 - {x''}^2 } \\
+y &= \sqrt[2]{{l_L}^2 - {x''}^2 } \\
 x &= x'' - \frac{b}{2}\\
@@
 \begin{align*}
-\sin{\gamma'} = \frac{x}{l_m}\\
+\sin{\gamma} = \frac{x}{l_m}\\
 \end{align*}
@@
 \begin{align*}
-\cos{\gamma'} = \frac{y}{l_m}\\
+\cos{\gamma} = \frac{y}{l_m}\\
 \end{align*}
 
-\noindent and as $\theta = \gamma'$, we can obtain the values of $\theta$ and $l_m$:
+\noindent and as $\theta = \gamma$, we can obtain the values of $\theta$ and $l_m$:
 
@@
 
-Additionally in the proposed system based on circular spring, the spring tension depends on the released thread length, hence the recovery force will vary accordingly.
-The proposed scheme allows a more controlled situation and depends exclusively on the software that controls the reel motor, which can be regulated.
-Another issue that can be presented is when $l_m$ is equals to $l_d$.
+\subsection{Design comparison}
+
+In the scheme proposed by~\cite{Endo2015} a single thread is recovered mechanically by means of a circular flat spring.
+This device works intensively when the robot is following the patient and, therefore, the spring is exposed to wear.
+Instead, the double thread recovery implemented by an active reel spring allows a more controlled situation and depends exclusively on the software that controls the reel motor, which can be regulated.
+
+Additionally with a circular flat spring, the spring tension depends on the released thread length, hence the recovery force will vary accordingly.
+The scheme based on an actuated reel allows a more controlled situation and depends exclusively on the software that controls the reel motor, which can be regulated.
+
+Regarding the control algorithm, authors in \cite{Endo2015} introduced two methods.
+The first of them computes the angular velocity as a  function of the difference between the measured thread length $l_m$ and the desired distance to patient  $l_D$.
+In this way, if the patient moves around the robot with a $l_m$ equals to $l_d$, the robot will not adjust its direction until the patient stops turning and continues moving forward.
+Hence, the robot must correct its direction but the correction angle could be large, which leads the robot to deviate off the desired trajectory.
+Additionally, the second method proposed is based on dead-reckoning to estimate the patient position.
+This strategy requires very precise measurements to reduce the cumulative error, which is characteristic of this estimation method~\cite{DurrantWhyte1994}.
+
+%Another issue that can be presented is when $l_m$ is equals to $l_D$.
 If the patient performs a circular movement where these values are equal, the robot will not change direction at all, because $\omega_{r}$ will be also zero.
-If the patient restart the movement, $\omega_{r}$ could be very low and the $K_p$ constant could not be enough to match the person movement rapidly.
-
-Regarding the second approach, the dead-reckoning could be challenging and requires very precise measurements to reduce the error-prone characteristic of this solution~\cite{DurrantWhyte1994}.
+If the patient restart the movement, $\omega_{r}$ could be very low and the proportional gain $K_p$ could not be enough to match the person movement rapidly.
+
+%Additionally, authors in~\cite{Endo2015} propose an alternative control strategy based on dead-reckoning requires very precise measurements to reduce the error-prone characteristic of this solution~\cite{DurrantWhyte1994}.
 
@@
 Éste es un poco más complicado de explicar con palabras así que haré un gráfico y se los envío.
-Como adelanto, está en la ecuación 17 del paper que es cuando toman como medida de la diferencia de distancias (l_m - l_d) para el cálculo de la V_r.
+Como adelanto, está en la ecuación 17 del paper que es cuando toman como medida de la diferencia de distancias (l_m - l_D) para el cálculo de la V_r.
 
@@
 
-\noindent where $ \phi $ is the free parameter, and $x$ and $y$ determine the shape of the trajectory on the navigation plane.
+\noindent where $ \phi $ is the free parameter, $a$ are the limit length of the arc of the curve, and $x$ and $y$ are the parametric functions that determine the shape of the trajectory on the navigation plane.
 
@@
 
-As shown in Figure~\ref{fig:capturesystem}, a tracking marker was placed on each side of the robot (on top of each thread reel).
+As shown on Figure~\ref{fig:capturesystem}, a tracking marker was placed on each side of the robot (on top of each thread reel).
 The human leader used his hand to grab the tip at which the two tethers were tied together.
@@
 \caption{Hardware prototype on the motion capture system and a testing subject holding the threads.
-The lemniscate of gerone was marked on the floor.
+The lemniscate of Gerono was marked on the floor.
 The subject follows this track on the performed experiments.}
@@
 \end{tabular}
-\captionof{table}{Max trajectory deviation m.t.d. [m] and Total trajectory deviation t.t.d. for different Follow the thread constants.}
+\captionof{table}{Maximum trajectory deviation m.t.d. [m] and Total trajectory deviation t.t.d. for different Follow the thread constants.}
 \label{tab:simulationmetricsftt}
@@
 \end{tabular}
-\captionof{table}{Max trajectory deviation m.t.d. (in [m]) and total trajectory deviation t.p.d. for different Rotate and go constants.}
+\captionof{table}{Maximum trajectory deviation m.t.d. (in [m]) and total trajectory deviation t.p.d. for different Rotate and go constants.}
 \label{tab:simulationmetricsrg}
@@
         Speed profiles of the leader and follower for Follow the thread (c) and Rotate and go (d). 
-        (e) Separation distance between robot and leader for both strategies.}
+        (e) Separation distance between robot and leader for both strategies  $[m]$.}
   \label{fig:simulationresults} 
@@
 \end{tabular}
-\captionof{table}{Max trajectory deviation m.t.d. and area under normal trajectory deviation t.t.d. in motion capture experiments using Follow the thread.}
+\captionof{table}{Maximum trajectory deviation m.t.d. and area under normal trajectory deviation t.t.d. in motion capture experiments using Follow the thread.}
 \label{tab:cap_ft_npd_table}
@@
 \end{tabular}
-\captionof{table}{Max trajectory deviation m.t.d. and area under normal trajectory deviation t.p.d. in motion capture experiments using Rotate and go.}
+\captionof{table}{Maximum trajectory deviation m.t.d. and area under normal trajectory deviation t.p.d. in motion capture experiments using Rotate and go.}
 \label{tab:cap_rg_npd_table}
@@
 The highlighted values on Table~\ref{tab:simulationmetricsrg} show the best configuration found.
-For the Rotate and go configuration, a high $c_r$ was important so the robot can turn quickly to point to the leader, but high values of $c_r$ and $c_v$ should be avoided, which may lead to dynamical issues with the inertia affecting the forward direction and missing the leader.
+
+For the Rotate and go configuration, the parameter $c_r$ affects the dynamic behavior of the robot which needs to be adjusted accordingly.
+Lastly, increasing the $Dt_{off}$ from 0.05 to 0.10 made the vehicle less prone to fall behind and results in a better following profile.
+%For the Rotate and go configuration, a high $c_r$ was important so the robot can turn quickly to point to the leader, but high values of $c_r$ and $c_v$ should be avoided, which may lead to dynamical issues with the inertia affecting the forward direction and missing the leader.
 Lastly, increasing the $Dt_{off}$ from 0.05 to 0.10 made the vehicle less prone to fall behind and results in a better following profile.
@@
 
-\subsection{Conclusion and Future Work}
+\section{Conclusion and Future Work}
 \label{conclusion}
@@
 
+\subsection{Validation carrying the oxygen tank}
+
+In order to advance towards then next step, we performed an additional simulated experiment to validate whether or not the selected algorithm could be extended to  handle the oxygen tank.
+ALPI works with 5 kg oxygen tanks, which are manufactured locally.
+
+We added on the Webots simulation a 5kg tank, and adjusted the coefficients $c_v$ and $c_{\alpha}$ for the  \textit{Follow the thread}  control strategy.
+We verified that very similar results are obtained in this scenario.
+In Figure~\ref{fig:validationwithtank} it can be seen that the shape of the trajectory follows the patient smoothly, the speed profile also follows stringently the leader speed, and finally the distance between the patient and the robot is maintained within the safe range.
+The obtained metrics are THIS and THAT which shows that from the simulation shows the feasibility of the control strategy to be included on the hardware prototype and iterate the project toward the next stage.
+
+\begin{figure} 
+    \centering
+  \subfloat[ \label{rw1a}]{%
+       \includegraphics[width=0.97\linewidth]{images/Wft_5kg.eps}}
+  \\
+    \hspace*{-1.5em}
+  \subfloat[ \label{rw1b}]{%
+        \includegraphics[width=0.49\linewidth]{images/Wft_5.0kg_leader_robot_speed.eps}}
+    \vspace*{-1.1em}
+  \subfloat[ \label{rw1c}]{%
+        \includegraphics[width=0.49\linewidth]{images/Wft_5kg_distance.eps}}
+    \hspace*{-1.5em}
+    \vspace*{1.1em}
+  \caption{Simulation results with the oxygent tank: the lemniscate trajectory is shown on (a) with the selected strategy Follow the thread.
+(b) Speed profiles of the patient-leader and the robot-follower.
+(c) Distance between the patient and the robot while traversing the trajectory.}
+  \label{fig:validationwithtank} 
+\end{figure}
+
 \subsection{Future Work}
@@
 \begin{itemize}
-    \item Power the motor wheels with a \textit{closed-loop and more efficient controller}, to have a more precise control over their speed.
     \item Redesign the active spring control mechanism in order to hold the motor temperature in its operational range.

### 2020-07-24 review/response.letter.tex
Version ready for submission.

@@
 
-We are grateful to the reviewer for pointing out relevant issues in our manuscript.
+We are grateful to all the reviewer for  the time taken to pointing out relevant issues in our manuscript.
 
@@
 
+\section*{\fbox{Associate Editor Responses:}}
+
+Comments to the Author:
+Authors addressed an important problem, providing a simple yet effective solution.
+There is no doubt that the device presented in the paper is of great interest for both patients and clinicians.
+To publish the paper on a robotics journal, though, the technical content has to be deepened so to increase the interest of researchers.
+As better detailed by the the two reviewers, the Authors are encouraged to improve the manuscript along two main directions:
+
+\begin{itemize}
+\item Clarifying the novelties wrt to state of the art (robot design and robot control), to highlight the original and novel contributions;
+\item  Improving the modelling, which should take into account realistic working conditions (e.g. weight and inertial properties of the tank).
+\end{itemize}
+
+\begin{quotation}
+{\color{blue}
+We appreciate for the Associate Editor's comments.  We concentrated our efforts modifying the manuscript in the two important lines that the Editor recommended.
+
+We addressed the first issue by adding a new Section "Comparison with alternative methods" where we compared the tethered controller with one and two threads and compared our proposal with other alternative proposals from the literature.
+We also included an evaluation of technical aspects of the design.
+
+Regarding the second issue, we performed an additional experiment on the simulation environment where we included a 5 kg mass, in a 45 degrees orientation, according to the weight and size of the oxygen tank regularly used at ALPI.
+We verified in the simulation that we obtained results very similar to the those obtained without the tank and that the controlling algorithm worked within the safe boundaries that we defined and that the following behavior was satisfied.
+}
+\end{quotation}
+
 \section*{\fbox{Reviewer \#1 Responses:}}
@@
 {\color{blue}
-We appreciate the Reviewers comments.
-We revised the notation on Section II.D per reviewers comments.
-We also added a section to establish the parallels between a single and double thread configuration.
-We clarified more clearly all the variables and their physicial meaning and included an extra Figure to aid the explanation provided.
+We thanks very much, the Reviewer, for the positive appraisal.
+We revised the notation on Section II.D.
+We also added a section to establish the parallels between a single and double thread configuration, comparing the solution to other alternatives found in the literature.
+We clarified all the variables and their physical meaning and included an extra Figure to aid in the provided explanation.
 }
@@
 During the simulation we explored different sets of values for the controller parameters.
-Once we arrived to values where we obtain a range of satisfactory results in terms of the metrics that we defined, we moved forward to the real world testing of the prototype.
-We found that we had to fine-tunning the parameters to adapt to the real world scenario but not search through the whole space of configurations.
-This was one of the values that the simulation step provided us.
+Once we arrived to parameter values where we obtain a range of satisfactory results in terms of the metrics that we defined, we moved forward to the real world testing of the prototype.
+We found that we had to fine-tune the parameters to adapt them to the real world scenario.
+However, the relative relation between parameters was consistent on both the simulation and in the real world, and that helped us to avoid searching extensively the parameter space for the real world scenario.
+We added in the manuscript a clarification of this point in Section VI.
 }
@@
 {\color{blue}
-Este es el primer punto con que vamos a tratar de responder, que es el que revive también el Editor asociado.
+We added a new specific section to address this issue.
+Now in Section III "Comparison with alternative methods", we show that the proposed algorithm based on two threads is equivalent to the one presented in the work [1].
+We added a new Figure 5 to aid in the explanation.
+Additionally, we included an additional "Design Analysis" section comparing both alternatives (two threads, and one thread and an angle).
+Moreover, we modified the notation to allow an improved comparison with previous proposals.
 }
@@
 {\color{blue}
-Pulses are a discrete number obtained directly from the rotary encoder KY-040.
-The values are incremented in one direction and decremented in the other direction.
-The encoder is connected to the Arduino board and it generates an interruption each time the encoder produces a rotation movement.
-PODEMOS ACA DAR DETALLES DE LA CONEXION ELECTRONICA SI QUIERE O DE COMO LO LEE EL CODIGO.
+We refer to pulses as a discrete number obtained directly from the rotary encoder KY-040.
+The encoder generates two digital pulses, one that indicates the presence of movement and the other is the direction of movement.
+This digital encoder is connected to the microcontroller and it generates an interruption each time the encoder produces a rotational movement.
+By reading the additional pulse, the direction can be determined and the value for pulses represents the relative circular distance traveled by the reel wheel.
+We added the definition of pulses in the manuscript on Section II.A.
 }
@@
 {\color{blue}
-The parameter $c_v$ is the proportional gain.  As long as the average distance obtained from both threads deviates from the resting distance $d_0$ the desired velocity is increased.
-EN EL PAPER DE ENDO USAN EL MISMO ESQUEMA LLAMANDOLO $V_r$.
-ES CIERTO QUE NO ES LA "VELOCIDAD" REAL DEL ROBOT SINO MAS BIEN UNA ESTIMACION DEL MOVIMIENTO DEL TARGET ?
-ACA NO ESTOY SEGURO HAY QUE PENSARLO MAS.
+We appreciate a lot this important issue that the Reviewer is pointing out here.
+We modified the notation trying to make it more consistent across the paper and more consistent with previous works as well.
+We clarify the units in the proportional gain $c_v$ and $c_{\alpha}$ to make equations physically meaningful.
 }
@@
 {\color{blue}
-In this case we shifted the notation to follow the one used in [15] where power assigned to each weels is denoted $\omega$.
+We added proper units for the constants and modify the equations clarifying that left-hand values reference speeds for each wheel.
 }
@@
 {\color{blue}
-@Luciano esta es la simulación que vos estás haciendo.  ESPEREMOS TENER BUENOS RESULTADOS
+We performed an additional simulation and added the obtained results in Section VII.B, at the end of the manuscript.
+We modified our simulated prototype and added a $5$ kg oxygen tank at an inclination of 45 degrees, and confirmed that the oxygen supply is not affected by this inclination, and that is under safe usage guidelines for the tank.
+The proposed algorithm \textit{Follow the thread} worked in a very similar way as shown in Figure 7, and only constant coefficients had to be readjusted for this purpose.
+Results show that the behavior of the robot is similar.
+This is a very important feasibility test that allows and encourage us to modify the hardware prototype to include the tank holder, and to move forward to this project's next step.
 }
@@
 {\color{blue}
-We replaced the term with musculoskeletal and verified the English grammar on the abstract and introduction.
-
-@Luciano agarra un canadiense x ahi
-@Esteban si querés darle alguna leída para ver si detectás algún error gramatical concreto.
+We replaced the term with musculoskeletal and verified the English grammar on the Abstract and Introduction section. 
 }
@@
 {\color{blue}
-We are very surprised the fast review process.  We truly thank Reviewers for their valuable time.  Dates has now been fixed.
+We are very surprised the the fast review process!   We truly thank all the Reviewers for their valuable time.  Dates have now been fixed.
 }
@@
 {\color{blue}
-We review all the constants definitions to be consistent along the manuscript.
+We reviewed all the constants definitions and naming conventions to be consistent along the manuscript.
 }
@@
 {\color{blue}
-This is an issue with Latex reformatting.  WE WILL FIX IT AT THE END.
+This is an issue with Latex reformatting.  We fixed it by modifying the layout of the manuscript.
 }
@@
 {\color{blue}
-@Luciano, tenés alguna foto donde sí tengas el guante ?
+There are two markers on top of the robot, just above each reel.  They are quite small and the picture lighting does not help.  We modified the Figure caption clarifying all this information.
 }
@@
 
-Page 6, first paragraph: "a high cr was important so the robot can turn quickly to point to the leader, but high values of cr should be avoided"; this seems contradictory.
+Page 6, first paragraph: "a high $c_r$ was important so the robot can turn quickly to point to the leader, but high values of $c_r$ should be avoided"; this seems contradictory.
 
@@
 {\color{blue}
-@Luciano, alguna sugerencia ?
+We change the phrase to emphasize that the value of $c_r$ was related with the dynamical behavior of the robot.  We appreciate the Reviewers comment.
 }

### 2020-07-24 alpibot.tex
Version ready for submission.

@@
 Patients that suffer Chronic Obstructive Pulmonary Disease (COPD) undergo a procedure called Pulmonary Rehabilitation that helps them to improve disease prognosis.
-During Pulmonary Rehabilitation patients perform different exercises and walking activities and may require oxygen assistance.
-The oxygen tank cannot be carried by the patient due to the musculoskeletal atrophy that characterize this pathology and external assistance is required.
+Pulmonary Rehabilitation consists of different physical exercises and walking activities performed at medical facilities under supervision of a physical therapist.
+In order to perform these procedures, patients require oxygen assistance, but the oxygen tank cannot be carried by the patient due to the musculoskeletal atrophy that characterize this pathology and external assistance is required.
 The assistance to transport the bulky oxygen tank can be provided by a robotic device that follows the patient while performing the physical activities.
@@
 Two alternative control strategies are proposed.
-Results are simulated and tested on a real prototype in a Motion Capture System.
+Results on a simulated and on a real prototype confirms the feasibility of the proposed solution.
 \end{abstract}
+
+%Results on a simulated and on a real prototype, and external validation from professional healthcare givers, confirm the feasibility of the proposed solution.
+%While performing Pulmonary Rehabilitation patients perform different exercises and walking activities and may require oxygen assistance.
+
+%Pulmonary Rehabilitation procedures consist of controlled waking activities and physical exercises that patients perform under the supervision of a physical therapist.
 
@@
 To do so, this document unrolls as follows.
-Section~\ref{materials} poses the problem and the solution design.
+Section~\ref{materials} and~\ref{Comparison} poses the problem and the solution design.
 Section~\ref{experimental} documents the experimental protocol to perform the solution assessment on a simulation and on a real world scenario.
@@
 
-\noindent where pulses obtained for each reel encoder are $pulses_{l}$ and $pulses_{r}$ for the left and right encoder.
+\noindent where $pulses_{l}$ and $pulses_{r}$ are the pulses obtained for the left and right encoder.
+Pulses refer to natural discrete numbers that represents the circular movement of the encoder shaft.
+The variable $r$ is the radius of the reel and $ppr$ is the pulses per revolution (resolution) of the encoder.
 These equations provide the estimated values for the left $l_L$ and right $l_R$  thread distance.
-The variable $r$ is the radius of the reel and $ppr$ is the pulses per revolution (resolution) of the encoder.
 The initial position of the threads is configurable.
@@
 The control algorithms run in this board, which continuously communicates with the Arduino board to receive sensor data and to issue commands to move the motors or retract the reels.
-The algorithms are programmed in python 3.7, which allows the exact same code to run both the simulated and real world prototype.
+The algorithms are programmed in Python 3.7, which allows the exact same code to run both the simulated and real world prototype.
 
@@
 They are attached to regular FA-12350 DC motors scavenged from old compact discs.
-Each reel is axially locked to inexpensive Ky-040 rotary encoder which provides around 20 pulses per revolution.
+Each reel is axially locked to inexpensive Ky-040 rotary encoder~\cite{SugaharaJunOnoKoji1998} which provides around 20 pulses per revolution.
 %Hardware specifications are described on table \ref{tab:specifications}.
@@
 Two simple algorithmic control strategies are proposed and evaluated.
-The first one is called Follow-the-thread and the second strategy is Rotate and go.
+The first one is called \textit{Follow the thread} and the second strategy is \textit{Rotate and go}.
 %The Adafruit Shield controller provides a very stable output signal, provided battery are kept above reference.
@@
 
-\section{Comparision with alternative methods} \label{development}
+\section{Comparison with alternative methods} \label{Comparison}
 
@@
 l_R &= \sqrt[2]{y^2 + {x'}^2} \\
-l_L &= \sqrt[2]{y^2 + {x''}^2}. \\
+l_L &= \sqrt[2]{y^2 + {x''}^2} \\
 \end{array}
@@
 
+\noindent where $x'$ and $x''$ are the distances from the leader horizontal position to the right and left wheel respectively.
+
 From Figure~\ref{fig:dibujitos} we can see that ${x''} + x' =  b$, with $b$ the axle length.
-Combining this with~\ref{eq:threadsincoordinates}, we can form a system of equations to determine $x'$, $x''$ and $y$:
+Combining this with Equation~\ref{eq:threadsincoordinates}, we can form a system of equations to determine $x'$, $x''$ and $y$:
 
@@
 
-From the first equation (a), rearranging and squaring both terms leads to
+From the first Equation (a), rearranging and squaring both terms leads to
 
@@
 
-By substracting (b) and (c), it can be obtained
+By subtracting (b) and (c), it can be obtained
 
@@
 
-From this equation $x''$ can be determined as
+From this equation, $x''$ can be determined as
 
@@
 This device works intensively when the robot is following the patient and, therefore, the spring is exposed to wear.
-Instead, the double thread recovery implemented by an active reel spring allows a more controlled situation and depends exclusively on the software that controls the reel motor, which can be regulated.
-
 Additionally with a circular flat spring, the spring tension depends on the released thread length, hence the recovery force will vary accordingly.
-The scheme based on an actuated reel allows a more controlled situation and depends exclusively on the software that controls the reel motor, which can be regulated.
+Instead, the double thread tethered design implemented with an active reel spring allows a more controlled situation and depends exclusively on the software that controls the reel motor, which can be regulated.
 
@@
 The first of them computes the angular velocity as a  function of the difference between the measured thread length $l_m$ and the desired distance to patient  $l_D$.
-In this way, if the patient moves around the robot with a $l_m$ equals to $l_d$, the robot will not adjust its direction until the patient stops turning and continues moving forward.
+In this way, if the patient moves around the robot with a $l_m$ equals to $l_D$, the robot will not adjust its direction until the patient stops turning and restarts moving forward.
 Hence, the robot must correct its direction but the correction angle could be large, which leads the robot to deviate off the desired trajectory.
 Additionally, the second method proposed is based on dead-reckoning to estimate the patient position.
-This strategy requires very precise measurements to reduce the cumulative error, which is characteristic of this estimation method~\cite{DurrantWhyte1994}.
+This strategy requires very precise measurements to reduce the cumulative error, which is characteristic of this estimation procedure~\cite{DurrantWhyte1994}.
 
@@
 
-\noindent where $ \phi $ is the free parameter, $a$ are the limit length of the arc of the curve, and $x$ and $y$ are the parametric functions that determine the shape of the trajectory on the navigation plane.
+\noindent where $ \phi $ is the free parameter, $a$ is the limit length of the arc of the curve, and $x$ and $y$ are the parametric functions that determine the shape of the trajectory on the navigation plane.
 
@@
 
-The simulation is also useful to study the effects of the different constants in each strategy on the movement of the robot.
-The leader starts at the midpoint of the trajectory and completes a full complete circuit getting back to the initial position, while the robot follows its track.
-Four different configuration sets of $c_v$ and  $c_{\alpha}$ were tested for Follow the thread, while 6 different configurations were tested for Rotate and go.
+The simulation is also useful to study the effects of the different constants on the robot movement  for the different strategies.
+The leader starts at the midpoint of the trajectory and completes a full circuit getting back to the initial position, while the robot follows its track.
+Four different configuration sets of $c_v$ and  $c_{\alpha}$ were tested for \textit{Follow the thread}, while 6 different configurations were tested for \textit{Rotate and go}.
 
@@
 A motion capture system is used to track the movement of a human leader along a predetermined trajectory.
-The motion tracking system consists of an array of 16 OptiTrack (NaturalPoint Inc, Oregon, US) Flex 3 cameras, which measure the position of reflective markers with an accuracy of $\pm1$ [cm] at sampling rate of 100 Hz.
+The motion tracking system consists of an array of 16 OptiTrack (NaturalPoint Inc, Oregon, US) Flex 3 cameras, which measure the position of reflective markers with an accuracy of $\pm1$ cm at sampling rate of 100 Hz.
 The calibration and data collection was made using the Motive motion capture software.
@@
 The lemniscate of Gerono, used in the simulation, was marked on the floor, and the human leader tried to move his hand following this shape as close as possible, with stable speed.
-The shape was marked according to the shape described in Equation~\ref{eq:lemniscata}, using $a = 2$ [m].
+The shape was marked according to the shape described in Equation~\ref{eq:lemniscata}, using $a = 2$ m.
 
@@
 \caption{Hardware prototype on the motion capture system and a testing subject holding the threads.
+On top of the device, two markers are placed and an additional marker is on a glove that the user is wearing (not shown on the picture).
 The lemniscate of Gerono was marked on the floor.
@@
 Subfigure (e) show the distance between the robot and the patient for both strategies.
-Results metrics for the simulations are shown on Table~\ref{tab:simulationmetricsftt} for the Follow the thread strategy, whereas metrics for Rotate and go are shown on Table~\ref{tab:simulationmetricsrg}.
+Results metrics for the simulations are shown on Table~\ref{tab:simulationmetricsftt} for the \textit{Follow the thread} strategy, whereas metrics for \textit{Rotate and go} are shown on Table~\ref{tab:simulationmetricsrg}.
 
@@
 \end{tabular}
-\captionof{table}{Maximum trajectory deviation m.t.d. [m] and Total trajectory deviation t.t.d. for different Follow the thread constants.}
+\captionof{table}{Maximum trajectory deviation m.t.d. (m) and Total trajectory deviation t.t.d. for different \textit{Follow the thread} constants.}
 \label{tab:simulationmetricsftt}
@@
 \end{tabular}
-\captionof{table}{Maximum trajectory deviation m.t.d. (in [m]) and total trajectory deviation t.p.d. for different Rotate and go constants.}
+\captionof{table}{Maximum trajectory deviation m.t.d. (m) and total trajectory deviation t.t.d. for different \textit{Rotate and go} constants.}
 \label{tab:simulationmetricsrg}
@@
 %\includegraphics[width=8cm]{images/ft_a2_n1500_cv15_cr15.eps}
-%\caption{Trajectories of the leader and follower for the Follow the thread control strategy.}
+%\caption{Trajectories of the leader and follower for the \textit{Follow the thread} control strategy.}
 %\label{fig:distance_sim}
@@
 %\includegraphics[width=8cm]{images/rgv_cv20_cr20_dr10.eps}
-%\caption{Trajectories of the leader and follower for Rotate and go control strategy.}
+%\caption{Trajectories of the leader and follower for \textit{Rotate and go} control strategy.}
 %\label{fig:distance_sim}
@@
 %\includegraphics[width=8cm]{images/ft_cv15_cr10_leader_robot_speed.eps}
-%\caption{Speed profiles of the leader and follower for Follow the thread control strategy.}
+%\caption{Speed profiles of the leader and follower for \textit{Follow the thread} control strategy.}
 %\label{fig:distance_sim}
@@
 %\includegraphics[width=8cm]{images/rg_cv20_cr20_bv10_dtoff10_leader_robot_speed.eps}
-%\caption{Speed profiles of the leader and follower for Rotate and go control strategy.}
+%\caption{Speed profiles of the leader and follower for \textit{Rotate and go} control strategy.}
 %\label{fig:distance_sim}
@@
         \includegraphics[width=0.99\linewidth]{images/ft_vs_rg_dist.eps}}
-  \caption{Simulation Results: Trajectories of the leader and follower for Follow the thread (a) and Rotate and go (b) . 
-        Speed profiles of the leader and follower for Follow the thread (c) and Rotate and go (d). 
-        (e) Separation distance between robot and leader for both strategies  $[m]$.}
+  \caption{Simulation Results: Trajectories of the leader and follower for \textit{Follow the thread} (a) and \textit{Rotate and go} (b) . 
+        Speed profiles of the leader and follower for \textit{Follow the thread} (c) and \textit{Rotate and go} (d). 
+        (e) Separation distance between robot and leader for both strategies  (m).}
   \label{fig:simulationresults} 
@@
 Results for the real world experiment can be seen on Figure~\ref{fig:realworldresults}.
-Table~\ref{tab:cap_ft_npd_table} show the metrics of the Follow the thread strategy, whereas Table~\ref{tab:cap_rg_npd_table} provides the metrics for the Rotate and go approach.
+Table~\ref{tab:cap_ft_npd_table} shows the metrics of the \textit{Follow the thread} strategy, whereas Table~\ref{tab:cap_rg_npd_table} provides the metrics for the \textit{Rotate and go} approach.
 
@@
 %\includegraphics[width=8cm]{images/ft1cap.eps}
-%\caption{Trajectories for the Follow the thread control strategy.}
+%\caption{Trajectories for the \textit{Follow the thread} control strategy.}
 %\label{fig:distance_sim}
@@
 %\includegraphics[width=8cm]{images/ft1cap_leader_robot_speed.eps}
-%\caption{Follow the thread speed profiles}
+%\caption{\textit{Follow the thread} speed profiles}
 %\label{fig:distance_sim}
@@
 %\includegraphics[width=8cm]{images/cap_ft_vs_rg_dist.eps}
-%\caption{Rotate and go speed profiles}
+%\caption{\textit{Rotate and go} speed profiles}
 %\label{fig:distance_sim}
@@
         \includegraphics[width=0.99\linewidth]{images/cap_ft_vs_rg_dist.eps}}
-  \caption{Experimentation Results: Trajectories of the leader and follower for Follow the thread (a) and Rotate and go (b). 
-        Speed profiles of the leader and follower for Follow the thread (c) and Rotate and go (d). 
-        (e) Separation distance between robot and leader for both strategies [m].}
+  \caption{Experimentation Results: Trajectories of the leader and follower for \textit{Follow the thread} (a) and \textit{Rotate and go} (b). 
+        Speed profiles of the leader and follower for \textit{Follow the thread} (c) and \textit{Rotate and go} (d). 
+        (e) Separation distance between robot and leader for both strategies (m).}
   \label{fig:realworldresults} 
@@
 \hline
-$c_v$ & $c_{\alpha}$ & m.t.d. [m] & t.t.d. \\
+$c_v$ & $c_{\alpha}$ & m.t.d. (m) & t.t.d. \\
 \hline
@@
 \end{tabular}
-\captionof{table}{Maximum trajectory deviation m.t.d. and area under normal trajectory deviation t.t.d. in motion capture experiments using Follow the thread.}
+\captionof{table}{Maximum trajectory deviation m.t.d. and area under normal trajectory deviation t.t.d. in motion capture experiments using \textit{Follow the thread}.}
 \label{tab:cap_ft_npd_table}
@@
 \hline
-$c_v$ & $c_r$ & $Dt_{off}$ & m.t.d. [m] & t.t.d. \\
+$c_v$ & $c_r$ & $Dt_{off}$ & m.t.d. (m) & t.t.d. \\
 \hline
@@
 \end{tabular}
-\captionof{table}{Maximum trajectory deviation m.t.d. and area under normal trajectory deviation t.p.d. in motion capture experiments using Rotate and go.}
+\captionof{table}{Maximum trajectory deviation m.t.d. and area under normal trajectory deviation t.t.d. in motion capture experiments using \textit{Rotate and go}.}
 \label{tab:cap_rg_npd_table}
@@
 From the leader and robot trajectories in Figure~\ref{fig:simulationresults}(a to d), both strategies exhibits basic following behavior.
-However, Rotate and go on Figure~\ref{fig:simulationresults}(b,d) generates a more irregular trajectory, due to the two stage movement algorithm.
-
-Regarding algorithms parameters, the simulation shows that for Follow the thread, a low $c_{\alpha}$  means that the robot is slow to turn and makes wider turns, providing a smoother trajectory.
+However, \textit{Rotate and go} on Figure~\ref{fig:simulationresults}(b,d) generates a more irregular trajectory, due to the two stage movement algorithm.
+
+Regarding algorithms parameters, the simulation shows that for \textit{Follow the thread}, a low $c_{\alpha}$  means that the robot is slow to turn and makes wider turns, providing a smoother trajectory.
 On the other hand, for low $c_v$ values, the robot tends to lag behind the leader when it is going in a straight line.
@@
 
-For the Rotate and go configuration, the parameter $c_r$ affects the dynamic behavior of the robot which needs to be adjusted accordingly.
+For the \textit{Rotate and go} configuration, the parameter $c_r$ affects the dynamic behavior of the robot which needs to be adjusted accordingly.
+Lastly, increasing the $Dt_{off}$ from 0.05 to 0.10 made the vehicle less prone to fall behind and produces a better following profile.
+%For the \textit{Rotate and go} configuration, a high $c_r$ was important so the robot can turn quickly to point to the leader, but high values of $c_r$ and $c_v$ should be avoided, which may lead to dynamical issues with the inertia affecting the forward direction and missing the leader.
 Lastly, increasing the $Dt_{off}$ from 0.05 to 0.10 made the vehicle less prone to fall behind and results in a better following profile.
-%For the Rotate and go configuration, a high $c_r$ was important so the robot can turn quickly to point to the leader, but high values of $c_r$ and $c_v$ should be avoided, which may lead to dynamical issues with the inertia affecting the forward direction and missing the leader.
-Lastly, increasing the $Dt_{off}$ from 0.05 to 0.10 made the vehicle less prone to fall behind and results in a better following profile.
-
-%From the leader and robot trajectories in Figure \ref{fig:sim_ft_cv15_cr15}, we can see that the robot using Follow the thread control exhibits basic following behavior.
+
+%From the leader and robot trajectories in Figure \ref{fig:sim_ft_cv15_cr15}, we can see that the robot using \textit{Follow the thread} control exhibits basic following behavior.
 Deviation occurs when the leader trajectory turns with a tight radius.
@@
 In line with the simulated results, a similar behavior was found on the experiments performed inside the Motion Capture Lab, and the robot exhibits following behavior for both control strategies as shown on Figure~\ref{fig:realworldresults}.
-Parameters had to be readjusted for the real world case, but the same situation occurs for both strategies.
+Although the parameter values obtained from the simulation had to be readjusted for the real world scenario, the relative relation between them was maintained and that helped to narrow the parameter search space.
 
@@
 This is shown on the speed profiles in Figure~\ref{fig:simulationresults}(c,d) as well as on the real experiment on Figure~\ref{fig:realworldresults}(c,d).
-The smoother movement of the \textit{Follow the thread} algorithm can be desired by the followed subject, since it can be perceived as less violent or unexpected.
+The smoother movement of the \textit{Follow the thread} algorithm is an additional desired goal, since it can be perceived as less violent or unexpected.
 %Additionally, when the vehicle stops to rotate, if the leader continues moving, when going forward, the robot might need to accelerate sharply to catch up with the leader, resulting in a sudden movement that can be disturbing for a medical patient or a doctor in the middle of a treatment.
@@
 \hline
-How would you qualify, from 1 to 5, your overall experience with Follow the thread? (1:Bad, 5:Excellent)  & 5.0\\
+How would you qualify, from 1 to 5, your overall experience with \textit{Follow the thread}? (1:Bad, 5:Excellent)  & 5.0\\
 \hline
-How safe would a patient be, from 1 to 5, being followed by the robot using the Follow the thread strategy? (1:Very unsafe, 5:Very safe)  & 4.25\\
+How safe would a patient be, from 1 to 5, being followed by the robot using the \textit{Follow the thread} strategy? (1:Very unsafe, 5:Very safe)  & 4.25\\
 \hline
-How would you qualify, from 1 to 5, your overall experience with Rotate and go? (1:Bad, 5:Excellent)  & 3.5\\
+How would you qualify, from 1 to 5, your overall experience with \textit{Rotate and go}? (1:Bad, 5:Excellent)  & 3.5\\
 \hline
-How safe would a patient be, from 1 to 5, being followed by the robot using the Rotate and go strategy? (1:Very unsafe, 5:Very safe) & 3.25\\
+How safe would a patient be, from 1 to 5, being followed by the robot using the \textit{Rotate and go} strategy? (1:Very unsafe, 5:Very safe) & 3.25\\
 \hline
@@
 
-In order to advance towards then next step, we performed an additional simulated experiment to validate whether or not the selected algorithm could be extended to  handle the oxygen tank.
+In order to advance towards the next step, we performed an additional simulated experiment to validate whether or not the selected algorithm could be extended to  handle the oxygen tank.
 ALPI works with 5 kg oxygen tanks, which are manufactured locally.
 
-We added on the Webots simulation a 5kg tank, and adjusted the coefficients $c_v$ and $c_{\alpha}$ for the  \textit{Follow the thread}  control strategy.
-We verified that very similar results are obtained in this scenario.
-In Figure~\ref{fig:validationwithtank} it can be seen that the shape of the trajectory follows the patient smoothly, the speed profile also follows stringently the leader speed, and finally the distance between the patient and the robot is maintained within the safe range.
-The obtained metrics are THIS and THAT which shows that from the simulation shows the feasibility of the control strategy to be included on the hardware prototype and iterate the project toward the next stage.
+We added on the Webots simulation a 5 kg mass on top of our model, and adjusted the coefficients $c_v$ and $c_{\alpha}$ for the  \textit{Follow the thread}  control strategy.
+Results are consistent with the outcomes shown previously on Figures~\ref{fig:simulationresults} and~\ref{fig:realworldresults}.
+In Figure~\ref{fig:validationwithtank} it can be seen that the shape of the trajectory follows the patient smoothly, the speed profile also follows stringently the leader speed, and finally the distance between the patient and the robot is maintained within the safe boundaries.
+The obtained metrics are 0.2223 for m.t.d.
+and 1.1068 for t.t.d.
+which shows that the following behavior is achieved and verifies, from the simulation perspective, the initial feasibility of the control strategy considering the addition of the oxygen tank.
 
@@
     \vspace*{1.1em}
-  \caption{Simulation results with the oxygent tank: the lemniscate trajectory is shown on (a) with the selected strategy Follow the thread.
+  \caption{Simulation results with the oxygen tank: the lemniscate trajectory is shown on (a) with the selected strategy \textit{Follow the thread}.
 (b) Speed profiles of the patient-leader and the robot-follower.

### 2020-07-24 alpibot.tex
Submitted version.

@@
 Patients that suffer Chronic Obstructive Pulmonary Disease (COPD) undergo a procedure called Pulmonary Rehabilitation that helps them to improve disease prognosis.
-Pulmonary Rehabilitation consists of different physical exercises and walking activities performed at medical facilities under supervision of a physical therapist.
+Pulmonary Rehabilitation consists of different physical exercises and walking activities conducted at medical facilities under supervision of a physical therapist.
 In order to perform these procedures, patients require oxygen assistance, but the oxygen tank cannot be carried by the patient due to the musculoskeletal atrophy that characterize this pathology and external assistance is required.
@@
 The variable $r$ is the radius of the reel and $ppr$ is the pulses per revolution (resolution) of the encoder.
-These equations provide the estimated values for the left $l_L$ and right $l_R$  thread distance.
+These equations provide the estimated values for the left $l_L$ and right $l_R$  thread longitude.
 The initial position of the threads is configurable.
@@
 \noindent where  $c_v$ and $c_{\alpha}$ are constant coefficients used for calibration whose units are expressed in $[\frac{1}{s}]$, and $l_D$ is a constant offset in $[m]$ that is used to customize the desired length of the thread where the robot does not move.
-As shown on Figure~\ref{fig:follow_params},  $l_L$ and $l_R$ are thread distances in $[m]$  obtained from encoder information from Equation~\ref{eq:encoder_len}, $V_t$ is the estimated forward velocity for the target and finally $\omega_{L}$ are $\omega_{R}$ are the velocity values that are directly used to drive each wheel motor.
+As shown on Figure~\ref{fig:follow_params},  $l_L$ and $l_R$ are thread longitudes in $[m]$  obtained from encoder information from Equation~\ref{eq:encoder_len}, $V_t$ is the estimated forward velocity for the target and finally $\omega_{L}$ are $\omega_{R}$ are the velocity values that are directly used to drive each wheel motor.
 
@@
 
-\caption{The center of coordinates is on the robot center of mass, where the y positive axis points in the same direction as the robot direction.}
+\caption{The center of coordinates is the midpoint of the wheel axle, where the y positive axis points in the same direction as the robot direction.}
 \label{fig:dibujitos}

### 2020-10-16 alpibot.tex
Correcciones del revisor puntos 1, 2, 3 y 5

@@
 
-%Pulmonary Rehabilitation procedures consist of controlled waking activities and physical exercises that patients perform under the supervision of a physical therapist.
+%Pulmonary Rehabilitation procedures consist of controlled walking activities and physical exercises that patients perform under the supervision of a physical therapist.
 
@@
 
-%\IEEEPARstart{C}{hronic} Obstructive Pulmonary Disease (COPD) is an umbrella term that describes several pulmonary affections.
-They are characterized as a slowly progressive condition marked by airflow limitation, being cigarette smoking the main etiologic factor~\cite{MacNee2005}.
-This pathology presents a musculoskeletal atrophy~\cite{Kocsis2016,Wu2012}.
-In order to carve these after effects a Pulmonary Rehabilitation procedure is a viable treatment for patients.
-During this rehabilitation, patients undergo controlled walking activities and physical exercises under the supervision of a physical therapist.
-Patients present a severe low oxygen saturation illness and they require effective oxygen supply, particularly when performing physical activities~\cite{Celli2014}.
-Hence, patients require to carry with them an oxygen tank for the oxygenotherapy assistance.
-However, their own condition prevent them with the ability to precisely carry the often bulky external tank.
-This situation entails to find a pragmatic solution to avoid an additional physical therapist to carry the oxygen tank.
-This scenario may be aggravated by the fact that this procedure is performed on a rehabilitation gym that could be potentially crowded with several patients, physiologists and physical therapists.
-
 \IEEEPARstart{C}{hronic} Obstructive Pulmonary Disease (COPD) is an umbrella term that describes several pulmonary affections.
@@
 In order to carve these after effects a Pulmonary Rehabilitation procedure is a viable treatment for patients.
-Pulmonary Rehabilitation procedures consist of controlled waking activities and physical exercises that patients perform under the supervision of a physical therapist.
+Pulmonary Rehabilitation procedures consist of controlled walking activities and physical exercises that patients perform under the supervision of a physical therapist.
 However, COPD patients present a severe low oxygen saturation illness and they require oxygen supply, particularly when performing physical activities~\cite{Celli2014}.
-Hence, they are required to carry an oxygen tank for the oxygenotherapy assistance, but their own condition prevent them with the ability to precisely carry the often bulky external tank.
+Hence, they are required to carry an oxygen tank for the oxygenotherapy assistance, but their own condition hinders their ability to precisely carry the often bulky external tank.
 This situation entails to find a pragmatic solution to avoid an additional physical therapist to carry the oxygen tank.
@@
 \noindent where  $c_v$ and $c_{\alpha}$ are constant coefficients used for calibration whose units are expressed in $[\frac{1}{s}]$, and $l_D$ is a constant offset in $[m]$ that is used to customize the desired length of the thread where the robot does not move.
-As shown on Figure~\ref{fig:follow_params},  $l_L$ and $l_R$ are thread longitudes in $[m]$  obtained from encoder information from Equation~\ref{eq:encoder_len}, $V_t$ is the estimated forward velocity for the target and finally $\omega_{L}$ are $\omega_{R}$ are the velocity values that are directly used to drive each wheel motor.
+As shown on Figure~\ref{fig:follow_params},  $l_L$ and $l_R$ are changes in thread longitudes (the length of thread that was released from each reel) in $[m]$  obtained from encoder information from Equation~\ref{eq:encoder_len}, $V_t$ is the estimated forward velocity for the target and finally $\omega_{L}$ are $\omega_{R}$ are the velocity values that are directly used to drive each wheel motor.
 
@@
 The constant parameter $Dm_{off}$ is similar to $l_D$, and is used to customize the length of the thread where the robot does not move at all.
+Lastly, $base_{vr}$, is another constant used to determine the minimum rotational velocity of the vehicle.
 
@@
 
+\subsection{Validation carrying the oxygen tank weight}
+
+We performed an additional simulated experiment to validate whether or not the selected algorithm could be extended to handle weight of the real oxygen tank.
+A common type of tank used in this type of therapies is a 5kg oxygen tank.
+
+We added on the Webots simulation a 5 kg mass on top of our model, and adjusted the coefficients $c_v$ and $c_{\alpha}$ for the  \textit{Follow the thread}  control strategy.
+Results are consistent with the outcomes shown previously on Figures~\ref{fig:simulationresults} and~\ref{fig:realworldresults}.
+In Figure~\ref{fig:validationwithtank} it can be seen that the shape of the trajectory follows the patient smoothly, the speed profile also follows stringently the leader speed, and finally the distance between the patient and the robot is maintained within the safe boundaries.
+The obtained metrics are 0.2223 for m.t.d.
+and 1.1068 for t.t.d.
+which shows that the following behavior is achieved and verifies, from the simulation perspective, the initial feasibility of the control strategy considering the addition of the oxygen tank.
+
 \section{Discussion}
@@
 
-Each control strategy has its advantages, but according various metrics described in this work, the \textit{Follow the thread} strategy had a more desirable behavior, as it tended to follow the leader from a closer distance at all times, while moving in a smooth and predictable way.
+Each control strategy has its advantages, but according to various metrics described in this work, the \textit{Follow the thread} strategy had a more desirable behavior, as it tended to follow the leader from a closer distance at all times, while moving in a smooth and predictable way.
 
@@
 
-\subsection{Validation carrying the oxygen tank}
-
-In order to advance towards the next step, we performed an additional simulated experiment to validate whether or not the selected algorithm could be extended to  handle the oxygen tank.
+% \subsection{Validation carrying the oxygen tank}
+
+% In order to advance towards the next step, we performed an additional simulated experiment to validate whether or not the selected algorithm could be extended to  handle the oxygen tank.
 ALPI works with 5 kg oxygen tanks, which are manufactured locally.
 
-We added on the Webots simulation a 5 kg mass on top of our model, and adjusted the coefficients $c_v$ and $c_{\alpha}$ for the  \textit{Follow the thread}  control strategy.
+% We added on the Webots simulation a 5 kg mass on top of our model, and adjusted the coefficients $c_v$ and $c_{\alpha}$ for the  \textit{Follow the thread}  control strategy.
 Results are consistent with the outcomes shown previously on Figures~\ref{fig:simulationresults} and~\ref{fig:realworldresults}.

### 2020-10-16 review2/response.letter.tex
Versión casi lista, integrada con los cambios de Luciano.  Solo falta la aclaración de Juan.

\documentclass[journal,onecolumn,12pt]{IEEEtran}

\usepackage{amsmath,amssymb,bm}
\usepackage{amsthm, amsfonts}	
\usepackage{bm,bbm}
\usepackage[normalem]{ulem}
\usepackage{color}
\usepackage{fancybox}
\usepackage{url,booktabs}
\usepackage[round]{natbib}

\usepackage{xr}

\title{2nd Reply to Reviewer's Comments on\\
``A Control Strategy for a Tethered Follower Robot for Pulmonary Rehabilitation''}
\author{}

\begin{document}

\maketitle
\pagenumbering{roman}
\setcounter{page}{1}

We appreciate a lot all the comments provided by the Reviewers and the AE.

In the following, we discuss our responses to pointed out issues.

\vskip+1ex
\noindent \dotfill

\section*{\fbox{Associate Editor Transcript:}}

Associate Editor Comments to Author:

Associate Editor
Comments to the Author:
Most of the comments expressed by the two Reviewers have been successfully addressed. As a result, the clarity of the paper improved significantly. There remain some very minor issues, which I list below for the Authors' convenience.

\begin{itemize}
\item Better define $l_R$ and $l_L$, which probably could be named "changes in length".
\item Please avoid the use of the word "longitude".
\item The content of Section VII-A (Validation experiments) should be part of the "Results" sections, i.e. Sec. IV,V.
\item Amend the remaining typos and link up all the references.
\end{itemize}

\section*{\fbox{Reviewer \#1 Transcript:}}

The work presented in this work is extremely interessant and of paramount importance for help in the medical field, since it frees not needed medical personel to help in other tasks thus improving patient care.

The manuscript is very well written with some typos I will refer in short. Besides those minor issues I did not find any issues the require reviewing, the addition of sections III and VII as well as the rewriting of section II.D have greatly improved the quality of the manuscript, allowing for the readers to better understand the proposed system and the technology behind it making possible for the system to be used to its fullest extent as well as for improvements to the system to be made even by groups outside the authors of the manuscript.

Thus, I would like to congratulate the authors on the work done and on the manuscript.

Finally, there are some typos I feel need to be addressed for this manuscript to be of examplar quality, these are:
on section I, line 11, it reads "waking activities" when it should be "walking activites"
on section I, line 18, it reads "their own condition prevent them" and it should be "their own condition prevents them"
on section VII, beginning of the second paragraph, it reads "Each control strategy has its advantages, but according various" and it should be "Each control strategy has its advantages, but according to various"

\section*{\fbox{Reviewer \#2 Transcript:}}

The revisions the authors have made have improved the manuscript - thank you for implementing those. At this stage, I have a few remaining concerns.

- The word definitions of $l_R$ and $l_L$ are still unclear. They are introduced as differences in length, but then also referred to as "thread longitudes". I find the term "longitude" here to be confusing given that these length differences can have a non-horizontal component if the target is at some angle from the reel (i.e., they are not purely longitudinal measurements, but rather can be seen as two vectors pointing to the same point in 2D space with both x and y components). My recommendation is that these definitions be further clarified.

- In Algorithm 1, as far as I can tell, no definition is provided for $D_t$, $D_m$, nor $base_{vr}$. This should be remedied.

- I believe Section III B. can be improved by being even more explicit about the specific improvements in the current design vs. the existing design. Section III, in general, reads as-if these designs are just equivalent (which begs the question: "why was this work done then?")

- Finally, the validation experiments with the oxygen tank (Section VIIA) should not be in the concluding remarks, but rather more tightly integrated into the experimental results section (Sections IV/V) with accompanying discussion.

\section*{\fbox{Associate Editor Responses:}}

Associate Editor Comments to Author:

Associate Editor
Comments to the Author:
Most of the comments expressed by the two Reviewers have been successfully addressed. As a result, the clarity of the paper improved significantly. There remain some very minor issues, which I list below for the Authors' convenience.

\begin{itemize}
\item Better define $l_R$ and $l_L$, which probably could be named "changes in length".
\item Please avoid the use of the word "longitude".
\item The content of Section VII-A (Validation experiments) should be part of the "Results" sections, i.e. Sec. IV,V.
\item Amend the remaining typos and link up all the references.
\end{itemize}

\begin{quotation}
{\color{blue}
We thank gratefully to the Associate Editor for their comments and suggestions.

\vspace{10pt}

Regarding the first two issues we have fixed them following Reviewer's  suggestions.  We have clarified missing terms definitions and the word "longitude" is not used anymore in the manuscript.  In relation with the third issue, we have modified the manuscript's structure as per Reviewer's recommendation to provide more readability.   Finally, regarding the fourth issue, we have fixed all the pointed out typos and verified all the references.

\vspace{10pt}

We appreciate a lot the time taken by the AE and all the Reviewers.
}
\end{quotation}

\section*{\fbox{Reviewer \#1 Responses:}}

%\subsection*{\ovalbox{Reviewer 1 General Comments}}

The work presented in this work is extremely interessant and of paramount importance for help in the medical field, since it frees not needed medical personel to help in other tasks thus improving patient care.

The manuscript is very well written with some typos I will refer in short. Besides those minor issues I did not find any issues the require reviewing, the addition of sections III and VII as well as the rewriting of section II.D have greatly improved the quality of the manuscript, allowing for the readers to better understand the proposed system and the technology behind it making possible for the system to be used to its fullest extent as well as for improvements to the system to be made even by groups outside the authors of the manuscript.

Thus, I would like to congratulate the authors on the work done and on the manuscript.

Finally, there are some typos I feel need to be addressed for this manuscript to be of examplar quality, these are:
on section I, line 11, it reads "waking activities" when it should be "walking activites"

\vspace{10pt} 
\begin{quotation}
{\color{blue}
We fixed this typo, thanks a lot for pointing out this.
}
\end{quotation}
\vspace{10pt}

on section I, line 18, it reads "their own condition prevent them" and it should be "their own condition prevents them"

\vspace{10pt} 
\begin{quotation}
{\color{blue}
Fixed.
}
\end{quotation}
\vspace{10pt}

on section VII, beginning of the second paragraph, it reads "Each control strategy has its advantages, but according various" and it should be "Each control strategy has its advantages, but according to various"

\vspace{10pt} 
\begin{quotation}
{\color{blue}
We also fixed this grammar error.  We appreciate a lot the careful reading of our manuscript.
}
\end{quotation}
\vspace{10pt}

\section*{\fbox{Reviewer \#2 Responses:}}
%\subsection*{\ovalbox{Reviewer 2 General Comments}}

The revisions the authors have made have improved the manuscript - thank you for implementing those. At this stage, I have a few remaining concerns.

- The word definitions of $l_R$ and $l_L$ are still unclear. They are introduced as differences in length, but then also referred to as "thread longitudes". I find the term "longitude" here to be confusing given that these length differences can have a non-horizontal component if the target is at some angle from the reel (i.e., they are not purely longitudinal measurements, but rather can be seen as two vectors pointing to the same point in 2D space with both x and y components). My recommendation is that these definitions be further clarified.

\vspace{10pt} 
\begin{quotation}
{\color{blue}
We have modified on Section II.D, at the end of page 3, and clarified the terms according to the Reviewer suggestion.  They are indeed changes in the length of the thread released from each reel.  We have eliminated the use of the confusing term "longitude".  
}
\end{quotation}
\vspace{10pt}

- In Algorithm 1, as far as I can tell, no definition is provided for $D_t$, $D_m$, nor $base_{vr}$. This should be remedied.

\vspace{10pt} 
\begin{quotation}
{\color{blue}
We have included in Section II, page 4, the definition of $base_{vr}$ which was missing, and also clarified that $D_t$, $D_m$ are temporary variables used to perform the calculations on Algorithm 1.
}
\end{quotation}
\vspace{10pt}

- I believe Section III B. can be improved by being even more explicit about the specific improvements in the current design vs. the existing design. Section III, in general, reads as-if these designs are just equivalent (which begs the question: "why was this work done then?")

\vspace{10pt} 
\begin{quotation}
{\color{blue}
Underway
}
\end{quotation}
\vspace{10pt}

- Finally, the validation experiments with the oxygen tank (Section VIIA) should not be in the concluding remarks, but rather more tightly integrated into the experimental results section (Sections IV/V) with accompanying discussion.

\vspace{10pt} 
\begin{quotation}
{\color{blue}
We appreciate the Reviewer suggestion.  We have modified the manuscript structure following the Reviewer recommendation.  There is now a new Section V.A which included this experimental simulation.
}
\end{quotation}
\vspace{10pt}

\vskip+1ex
\noindent \dotfill
\vskip+1ex
\bibliographystyle{mdpi}
\bibliography{article}

\end{document}

### 2020-10-16 alpibot.1.tex
Versión casi lista, integrada con los cambios de Luciano.  Solo falta la aclaración de Juan.

--- 2020-07-24 alpibot.tex
@@
 of  Computer Engineering, Instituto Tecnológico de Buenos Aires (ITBA), Ciudad de Buenos Aires,  Argentina e-mail: rramele@itba.edu.ar}% <-this % stops a space
-\thanks{Manuscript received May 16, 2020; revised July 6, 2020.}}
+\thanks{Manuscript received April 19, 2020; revised August 26, 2020.}}
 
@@
 Patients that suffer Chronic Obstructive Pulmonary Disease (COPD) undergo a procedure called Pulmonary Rehabilitation that helps them to improve disease prognosis.
-Pulmonary Rehabilitation consists of different physical exercises and walking activities performed at medical facilities under supervision of a physical therapist.
-In order to perform these procedures, patients require oxygen assistance, but the oxygen tank cannot be carried by the patient due to the musculoskeletal atrophy that characterize this pathology and external assistance is required.
+During Pulmonary Rehabilitation patients perform different exercises and walking activities and may require oxygen assistance.
+The oxygen tank cannot be carried by the patient due to the eskeletomuscular atrophy that characterize this pathology and external assistance is required.
 The assistance to transport the bulky oxygen tank can be provided by a robotic device that follows the patient while performing the physical activities.
@@
 Two alternative control strategies are proposed.
-Results on a simulated and on a real prototype confirms the feasibility of the proposed solution.
+Results are simulated and tested on a real prototype in a Motion Capture System.
 \end{abstract}
-
-%Results on a simulated and on a real prototype, and external validation from professional healthcare givers, confirm the feasibility of the proposed solution.
-%While performing Pulmonary Rehabilitation patients perform different exercises and walking activities and may require oxygen assistance.
-
-%Pulmonary Rehabilitation procedures consist of controlled waking activities and physical exercises that patients perform under the supervision of a physical therapist.
 
@@
 
-%\IEEEPARstart{C}{hronic} Obstructive Pulmonary Disease (COPD) is an umbrella term that describes several pulmonary affections.
+\IEEEPARstart{C}{hronic} Obstructive Pulmonary Disease (COPD) is an umbrella term that describes several pulmonary affections.
 They are characterized as a slowly progressive condition marked by airflow limitation, being cigarette smoking the main etiologic factor~\cite{MacNee2005}.
-This pathology presents a musculoskeletal atrophy~\cite{Kocsis2016,Wu2012}.
+This pathology presents an eskeletomuscular atrophy~\cite{Kocsis2016,Wu2012}.
 In order to carve these after effects a Pulmonary Rehabilitation procedure is a viable treatment for patients.
 During this rehabilitation, patients undergo controlled walking activities and physical exercises under the supervision of a physical therapist.
-Patients present a severe low oxygen saturation illness and they require effective oxygen supply, particularly when performing physical activities~\cite{Celli2014}.
+Patients present a severe low oxygen saturation illness and they require effective oxygen supply, particularly when performing physical activity~\cite{Celli2014}.
 Hence, patients require to carry with them an oxygen tank for the oxygenotherapy assistance.
@@
 
-\IEEEPARstart{C}{hronic} Obstructive Pulmonary Disease (COPD) is an umbrella term that describes several pulmonary affections.
-They are characterized as a slowly progressive condition marked by airflow limitation, being cigarette smoking the main etiologic factor~\cite{MacNee2005}.
-This pathology presents a musculoskeletal atrophy~\cite{Kocsis2016,Wu2012}.
-In order to carve these after effects a Pulmonary Rehabilitation procedure is a viable treatment for patients.
-Pulmonary Rehabilitation procedures consist of controlled waking activities and physical exercises that patients perform under the supervision of a physical therapist.
-However, COPD patients present a severe low oxygen saturation illness and they require oxygen supply, particularly when performing physical activities~\cite{Celli2014}.
-Hence, they are required to carry an oxygen tank for the oxygenotherapy assistance, but their own condition prevent them with the ability to precisely carry the often bulky external tank.
-This situation entails to find a pragmatic solution to avoid an additional physical therapist to carry the oxygen tank.
-The scenario may be aggravated by the fact that this procedure is performed on a rehabilitation gym that could be potentially crowded with several patients, physiologists and physical therapists.
-
 %Alternatively, they are required to improvise a customized wheelchair to allow them to carry the oxygen tank on top of it, leading to a cumbersome situation for the physical therapist because s/he had to push and maneuver the wheelchair, and at the same time, take care of the patient and perform the rehabilitation treatment.
 
-An alternative solution is to use an assitive ground service robot~\cite{Neto2015} to carry the oxygen tank while following the patient in a leader-follower configuration.
+An alternative solution is to use an assitive ground service robot~\cite{Neto2015} to carry the oxygen tank, while following the patient in a leader-follower configuration.
 There are two reasons that support the initial viability of this idea.
@@
 On the other hand, the range of movements performed by the patient during the Rehabilitation Procedure is highly predictable by the treatment.
-At the same time,  the global robotic research community looks forward for the development of robotic affordable solutions to social and health-related worldwide problems~\cite{Khamis2019}.
+At the same time,  the global robotic research community looks forward for the development of robotic affordable solutions to the social and health-related worldwide problems~\cite{Khamis2019}.
 
@@
 To do so, this document unrolls as follows.
-Section~\ref{materials} and~\ref{Comparison} poses the problem and the solution design.
+Section~\ref{materials} poses the problem and the solution design.
 Section~\ref{experimental} documents the experimental protocol to perform the solution assessment on a simulation and on a real world scenario.
@@
 To be effective, any technological solution for the medical community requires active involvement of key stakeholders: physicians, care-givers, patients and their families~\cite{Gaggioli2006,Fasola2012,Cherubini2008,Wolpaw2012BraincomputerIP,Salvini2015}.
-Hence, design methodologies that allow rapid prototyping can bring quick feedback from real users about drawbacks or opportunities for improvements.
+Hence, design methodologies that allow rapid prototyping can bring quickly feedback from real users about drawbacks or opportunities for improvements.
 
@@
     \draw  [draw opacity=0][fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ]  (478.5,1350) -- (535.5,1350) -- (535.5,1374) -- (478.5,1374) -- cycle  ;
-    \draw (517,1362) node    {$subject$};
+    \draw (507,1362) node    {$subject$};
     % Text Node
@@
 Encoders in each reel measure the difference in length of each thread compared to its initial position.
-These differences in length are the input for the control algorithm.
+This difference in length is the input for the control algorithm.
 Using the encoder, the difference in length for each thread can be measured with Equations~\ref{eq:encoder_len}.
@@
 \begin{array}{ll}  
-l_L = pulses_{l} \; \frac{2\pi r}{ppr}  \\
+D_l = pulses_{l} \; \frac{2\pi r}{ppr}  \\
 \\
-l_R = pulses_{r} \; \frac{2\pi r}{ppr}
+D_r = pulses_{r} \; \frac{2\pi r}{ppr}
 \end{array}
@@
 
-\noindent where $pulses_{l}$ and $pulses_{r}$ are the pulses obtained for the left and right encoder.
-Pulses refer to natural discrete numbers that represents the circular movement of the encoder shaft.
+\noindent where pulses obtained for each reel encoder are $pulses_{l}$ and $pulses_{r}$ for the left and right encoder.
+These equations provide the estimated values for the left $D_l$ and right $D_r$  thread distance.
 The variable $r$ is the radius of the reel and $ppr$ is the pulses per revolution (resolution) of the encoder.
-These equations provide the estimated values for the left $l_L$ and right $l_R$  thread distance.
 The initial position of the threads is configurable.
@@
 The control algorithms run in this board, which continuously communicates with the Arduino board to receive sensor data and to issue commands to move the motors or retract the reels.
-The algorithms are programmed in Python 3.7, which allows the exact same code to run both the simulated and real world prototype.
+The algorithms are programmed in python 3.7, which allows the exact same code to run both the simulated and real world prototype.
 
@@
 They are attached to regular FA-12350 DC motors scavenged from old compact discs.
-Each reel is axially locked to inexpensive Ky-040 rotary encoder~\cite{SugaharaJunOnoKoji1998} which provides around 20 pulses per revolution.
+Each reel is axially locked to inexpensive Ky-040 rotary encoder which provides around 20 pulses per revolution.
 %Hardware specifications are described on table \ref{tab:specifications}.
@@
 As previously mentioned, an active spring mechanism is also put in place to keep the threads tense.
-However, in order to extend the useful life of the reel motors, and to save battery, an algorithm to  activate and deactivate the motors was developed.
+However, in order to extend the useful life of the reel motors, and to save battery, an algorithm to  activate and deactivate the motors is developed.
 
@@
 Two simple algorithmic control strategies are proposed and evaluated.
-The first one is called \textit{Follow the thread} and the second strategy is \textit{Rotate and go}.
+The first one is called Follow-the-thread and the second strategy is Rotate and go.
 %The Adafruit Shield controller provides a very stable output signal, provided battery are kept above reference.
@@
 This control strategy is similar to the one presented in \cite{Ortlieb2016}.
-It is based on the idea that both the relative angle between the subject and the vehicle orientation and, the relative distance between the robot and the subject, can be computed from the length of the left and right threads.
+It is based on the idea that the difference in length between both threads provides a quasi-linear function of the relative angle between the subject and the vehicle orientation and, additionally, the mean of both thread length allows to approximate the relative distance between the robot and the subject.
 They are described by Equations~\ref{eq:lemniscata}, \ref{eq:lemniscata2} and~\ref{eq:lemniscata3}.
@@
 \begin{equation}
-V_t = c_v (\frac{l_L + l_R}{2} - l_D)
+v_{tar} = c_v (\frac{D_l + D_r}{2} - d_0)
 \label{eq:lemniscata}
@@
 \begin{equation}
-\omega_{L} = V_t + c_{\alpha} (l_L - l_R)
+vleft_{tar} = v_{tar} + c_{\alpha} (D_l - D_r)
 \label{eq:lemniscata2}
@@
 \begin{equation}
-\omega_{R} = V_t - c_{\alpha} (l_L - l_R)
+vright_{tar} = v_{tar} - c_{\alpha} (D_l - D_r)
 \label{eq:lemniscata3}
@@
 
-\noindent where  $c_v$ and $c_{\alpha}$ are constant coefficients used for calibration whose units are expressed in $[\frac{1}{s}]$, and $l_D$ is a constant offset in $[m]$ that is used to customize the desired length of the thread where the robot does not move.
-As shown on Figure~\ref{fig:follow_params},  $l_L$ and $l_R$ are thread distances in $[m]$  obtained from encoder information from Equation~\ref{eq:encoder_len}, $V_t$ is the estimated forward velocity for the target and finally $\omega_{L}$ are $\omega_{R}$ are the velocity values that are directly used to drive each wheel motor.
+\noindent where  $c_v$ and $c_{\alpha}$ are constant coefficients used for calibration.
+As shown in Figure~\ref{fig:follow_params},  $D_l$ and $D_r$ are thread distances obtained from encoder information from Equation~\ref{eq:encoder_len}, $v_{tar}$ is the estimated forward velocity for the target and finally $vleft_{tar}$ are $vright_{tar}$ are the power values that are directly used to drive power to each wheel motor.
 
@@
 \begin{equation}
-\textit{if  }{\frac{l_L + l_R}{2} < l_D}\textit{ then } \omega_{L} = \omega_{R} = 0.
+\textit{if  }{\frac{D_l + D_r}{2} < d_0}\textit{ then } vleft_{tar} = vright_{tar} = 0 
 \end{equation}
+
+\noindent where $d_0$ is a constant offset that is used to customize the length of the thread where the robot does not move.
 
@@
     \draw (524,1960) node  [font=\large,color={rgb, 255:red, 74; green, 144; blue, 226 }  ,opacity=1 ]  {$ \begin{array}{l}
-    l_{R}\\
+    D_{r}\\
     \end{array}$};
@@
     \draw (510,1910) node  [font=\large,color={rgb, 255:red, 74; green, 144; blue, 226 }  ,opacity=1 ,rotate=-358.27]  {$ \begin{array}{l}
-    l_{L}\\
+    D_{l}\\
     \end{array}$};
@@
     % Text Node
-    \draw (390.58,1925.93) node [anchor=west] [inner sep=0.75pt]  [font=\normalsize,color={rgb, 255:red, 0; green, 0; blue, 0 }  ,opacity=1 ]  {$\omega_{L}$};
+    \draw (390.58,1925.93) node [anchor=west] [inner sep=0.75pt]  [font=\normalsize,color={rgb, 255:red, 0; green, 0; blue, 0 }  ,opacity=1 ]  {$vleft_{tar}$};
     % Text Node
-    \draw (407.58,2100.93) node [anchor=west] [inner sep=0.75pt]  [font=\normalsize,color={rgb, 255:red, 0; green, 0; blue, 0 }  ,opacity=1 ]  {$\omega_{R}$};
+    \draw (407.58,2100.93) node [anchor=west] [inner sep=0.75pt]  [font=\normalsize,color={rgb, 255:red, 0; green, 0; blue, 0 }  ,opacity=1 ]  {$vright_{tar}$};
 
@@
 The variable $V_r$ is the speed at which the vehicle will rotate on its axis, and $V_f$ is the speed at which the vehicle will move forward once it can move on the subject's direction.
-The algorithm requires three parameters, $c_v$ and $c_r$ that regulates the coefficients for the forward and rotation movement, and an additional parameter $Dt_{off}$ that regulates the sensibility of the rotation movement.
-The constant parameter $Dm_{off}$ is similar to $l_D$, and is used to customize the length of the thread where the robot does not move at all.
+The algorithm requires three parameters, $c_v$ and $c_r$ that regulates the coefficients for the forward and rotation movement, and an additional parameter $Dt_off$ that regulates the sensibility of the rotation movement.
+The constant parameter $Dm_off$ is similar to $d_0$, and is used to customize the length of the thread where the robot does not move at all.
 
@@
 \begin{algorithmic}
-\STATE $Dt \gets l_L - l_R$
-\STATE $Dm \gets \frac{l_R+l_L}{2}$
-\STATE $V_r \gets c_r * (abs(Dt) - Dt_{off}) + base_{vr}$ 
-\STATE $V_f \gets c_v * (Dm - Dm_{off})$
+\STATE $Dt \gets D_l - D_r$
+\STATE $Dm \gets \frac{D_r+D_l}{2}$
+\STATE $V_r \gets c_r * (abs(Dt) - Dt_off) + base_{vr}$ 
+\STATE $V_f \gets c_v * (Dm - Dm_off)$
 
 \IF{$abs(Dt) > Dt_{off}$}
-\IF{$l_L > l_R$}
-\STATE $\omega_{L} \gets V_r$
-\STATE $\omega_{R} \gets -V_r$
+\IF{$D_l > D_r$}
+\STATE $vleft_{tar} \gets V_r$
+\STATE $vright_{tar} \gets -V_r$
 \ELSE
-\STATE $\omega_{L} \gets -V_r$
-\STATE $\omega_{R} \gets V_r$
+\STATE $vleft_{tar} \gets -V_r$
+\STATE $vright_{tar} \gets V_r$
 \ENDIF
@@
 \IF{$D_m > Dm_{off}$}
-\STATE $\omega_{R} \gets V_f$
-\STATE $\omega_{L} \gets V_f$
+\STATE $vright_{tar} \gets V_f$
+\STATE $vleft_{tar} \gets V_f$
 \ELSE
-\STATE $\omega_{R} \gets 0$
-\STATE $\omega_{L} \gets 0$
+\STATE $vright_{tar} \gets 0$
+\STATE $vleft_{tar} \gets 0$
 \ENDIF
 \ENDIF
-\STATE Return $\omega_{R}$ and $\omega_{L}$.
+\STATE Return $vright_{tar}$ and $vleft_{tar}$.
 \end{algorithmic}
@@
 \end{algorithm}
-
-\section{Comparison with alternative methods} \label{Comparison}
-
-\subsection{Equivalence between the single and double tethered system}
-
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[width=8cm]{images/dibujito.jpg}
-%\caption{Scheme of the equivalence between a single and dual thread systems.}
-%\label{fig:dibujito}
-%\end{figure}
-
-In~\cite{Endo2015} authors propose a similar design based on only one thread.
-They propose two control strategies based on two input parameters that are obtained from a linear and circular potentiometer that determines the length of the thread $l_m$ and the orientation angle $\theta$.
-We show here that they are equivalent to the approach presented based on $l_L$ and  $l_R$.
-
-Based on a frame reference with the robot on the center of coordinates as shown on Figure~\ref{fig:dibujitos}, from the input values $l_m$ and $\theta$ we can derive the position of the leader as
-
-\begin{align*}
-   y &= l_m  \cos(\theta)\\
-   x &= l_m  \sin(\theta) \\
-\end{align*}
-
-%From the figure, it can be obtained that
-%
-%\begin{align*}
-%x'            	&= \frac{b}{2}  - x \\
-%x'' 			&= \frac{b}{2}  + x \\
-%\end{align*}
-%
-
-\noindent where $x$ is the follower distance on the horizontal direction, and $y$ the follower distance on the vertical direction.
-
-In this frame of reference, the lengths of the threads, $l_R$ and $l_L$ can be determined as
-
-\begin{figure}[]
-    \centering
-    \tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt
-
-\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
-%uncomment if require: \path (0,235); %set diagram left start at 0, and has height of 235
-
-%Shape: Axis 2D [id:dp7543865369052545] 
-\draw  (312,162.89) -- (521,162.89)(332.9,2.69) -- (332.9,180.69) (514,157.89) -- (521,162.89) -- (514,167.89) (327.9,9.69) -- (332.9,2.69) -- (337.9,9.69)  ;
-%Rounded Same Side Corner Rect [id:dp31787525316857024] 
-\draw   (297.9,150.89) ..
-controls (297.9,146.47) and (301.48,142.89) ..
-(305.9,142.89) -- (359.9,142.89) ..
-controls (364.32,142.89) and (367.9,146.47) ..
-(367.9,150.89) -- (367.9,182.89) ..
-controls (367.9,182.89) and (367.9,182.89) ..
-(367.9,182.89) -- (297.9,182.89) ..
-controls (297.9,182.89) and (297.9,182.89) ..
-(297.9,182.89) -- cycle ;
-%Shape: Square [id:dp8026938389823051] 
-\draw   (434,151.69) -- (457,151.69) -- (457,174.69) -- (434,174.69) -- cycle ;
-%Shape: Square [id:dp06857182701706144] 
-\draw   (207.66,149.66) -- (231.34,149.66) -- (231.34,173.34) -- (207.66,173.34) -- cycle ;
-%Flowchart: Summing Junction [id:dp2757072161924359] 
-\draw   (317.9,162.89) ..
-controls (317.9,155.16) and (324.62,148.89) ..
-(332.9,148.89) ..
-controls (341.18,148.89) and (347.9,155.16) ..
-(347.9,162.89) ..
-controls (347.9,170.62) and (341.18,176.89) ..
-(332.9,176.89) ..
-controls (324.62,176.89) and (317.9,170.62) ..
-(317.9,162.89) -- cycle ; \draw   (322.29,152.99) -- (343.51,172.79) ; \draw   (343.51,152.99) -- (322.29,172.79) ;
-%Straight Lines [id:da9717478160302964] 
-\draw [line width=2.25]    (405,12.69) -- (332.9,162.89) ;
-%Straight Lines [id:da30114532827552765] 
-\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=2.25]    (405,12.69) -- (219.5,161.5) ;
-%Straight Lines [id:da0582330345196298] 
-\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=2.25]    (405,12.69) -- (445.5,163.19) ;
-%Flowchart: Terminator [id:dp11720735085928635] 
-\draw   (393.1,2.69) -- (416.9,2.69) ..
-controls (419.99,2.69) and (422.5,7.16) ..
-(422.5,12.69) ..
-controls (422.5,18.21) and (419.99,22.69) ..
-(416.9,22.69) -- (393.1,22.69) ..
-controls (390.01,22.69) and (387.5,18.21) ..
-(387.5,12.69) ..
-controls (387.5,7.16) and (390.01,2.69) ..
-(393.1,2.69) -- cycle ;
-%Straight Lines [id:da6471016295019487] 
-\draw  [dash pattern={on 0.84pt off 2.51pt}]  (405,12.69) -- (406,162.69) ;
-%Shape: Circle [id:dp5568563915916713] 
-\draw   (403.5,162.69) ..
-controls (403.5,161.31) and (404.62,160.19) ..
-(406,160.19) ..
-controls (407.38,160.19) and (408.5,161.31) ..
-(408.5,162.69) ..
-controls (408.5,164.07) and (407.38,165.19) ..
-(406,165.19) ..
-controls (404.62,165.19) and (403.5,164.07) ..
-(403.5,162.69) -- cycle ;
-%Straight Lines [id:da6605293994654671] 
-\draw    (188,161.69) -- (332.9,162.89) ;
-%Shape: Brace [id:dp034761867026847026] 
-\draw   (221,180.69) ..
-controls (220.93,185.36) and (223.22,187.73) ..
-(227.88,187.8) -- (302.39,189.02) ..
-controls (309.06,189.13) and (312.35,191.52) ..
-(312.27,196.19) ..
-controls (312.35,191.52) and (315.72,189.24) ..
-(322.38,189.35)(319.38,189.3) -- (396.89,190.57) ..
-controls (401.56,190.65) and (403.93,188.36) ..
-(404,183.69) ;
-%Straight Lines [id:da6885647931399865] 
-\draw [line width=0.75]  [dash pattern={on 0.84pt off 2.51pt}]  (333,13.69) -- (405,12.69) ;
-%Curve Lines [id:da2681192753910786] 
-\draw    (384.56,66.54) .. controls (391.15,78.31) and (389.42,81.22) .. (403.17,79.98) ;
-\draw [shift={(406,79.69)}, rotate = 533.6600000000001] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;
-\draw [shift={(382.95,63.79)}, rotate = 58.73] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;
-%Curve Lines [id:da7846164710785513] 
-\draw    (334.02,102.11) .. controls (342.64,102.92) and (351.11,101.09) .. (351.93,117.88) ;
-\draw [shift={(352,120.69)}, rotate = 270] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;
-\draw [shift={(331,101.69)}, rotate = 11.31] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;
-%Shape: Brace [id:dp8659813161256731] 
-\draw   (220,203.69) ..
-controls (219.98,208.36) and (222.3,210.7) ..
-(226.97,210.72) -- (322.47,211.13) ..
-controls (329.14,211.16) and (332.46,213.51) ..
-(332.44,218.18) ..
-controls (332.46,213.51) and (335.8,211.19) ..
-(342.47,211.22)(339.47,211.21) -- (437.97,211.64) ..
-controls (442.64,211.66) and (444.98,209.34) ..
-(445,204.67) ;
-%Shape: Brace [id:dp4829292799998628] 
-\draw   (409,183.69) ..
-controls (409,188.36) and (411.33,190.69) ..
-(416,190.69) -- (418.5,190.69) ..
-controls (425.17,190.69) and (428.5,193.02) ..
-(428.5,197.69) ..
-controls (428.5,193.02) and (431.83,190.69) ..
-(438.5,190.69)(435.5,190.69) -- (441,190.69) ..
-controls (445.67,190.69) and (448,188.36) ..
-(448,183.69) ;
-%Shape: Circle [id:dp5543776560649496] 
-\draw   (330.5,13.69) ..
-controls (330.5,12.31) and (331.62,11.19) ..
-(333,11.19) ..
-controls (334.38,11.19) and (335.5,12.31) ..
-(335.5,13.69) ..
-controls (335.5,15.07) and (334.38,16.19) ..
-(333,16.19) ..
-controls (331.62,16.19) and (330.5,15.07) ..
-(330.5,13.69) -- cycle ;
-
-% Text Node
-\draw (399,166.09) node [anchor=north west][inner sep=0.75pt]    {$x$};
-% Text Node
-\draw (312,2.4) node [anchor=north west][inner sep=0.75pt]    {$y$};
-% Text Node
-\draw (422,194.4) node [anchor=north west][inner sep=0.75pt]    {$x'$};
-% Text Node
-\draw (308,193.4) node [anchor=north west][inner sep=0.75pt]    {$x''$};
-% Text Node
-\draw (357,52.4) node [anchor=north west][inner sep=0.75pt]    {$l_{m}$};
-% Text Node
-\draw (439,90.4) node [anchor=north west][inner sep=0.75pt]    {$l_{R}$};
-% Text Node
-\draw (261,90.4) node [anchor=north west][inner sep=0.75pt]    {$l_{L}$};
-% Text Node
-\draw (342,86.4) node [anchor=north west][inner sep=0.75pt]    {$\theta $};
-% Text Node
-\draw (379,76.4) node [anchor=north west][inner sep=0.75pt]    {$\gamma $};
-% Text Node
-\draw (334,215.4) node [anchor=north west][inner sep=0.75pt]    {$b$};
-
-\end{tikzpicture}
-
-\caption{The center of coordinates is on the robot center of mass, where the y positive axis points in the same direction as the robot direction.}
-\label{fig:dibujitos}
-\end{figure}{}
-
-\begin{equation}
-\begin{array}{ll}  
-l_R &= \sqrt[2]{y^2 + {x'}^2} \\
-l_L &= \sqrt[2]{y^2 + {x''}^2} \\
-\end{array}
-\label{eq:threadsincoordinates}
-\end{equation}
-
-\noindent where $x'$ and $x''$ are the distances from the leader horizontal position to the right and left wheel respectively.
-
-From Figure~\ref{fig:dibujitos} we can see that ${x''} + x' =  b$, with $b$ the axle length.
-Combining this with Equation~\ref{eq:threadsincoordinates}, we can form a system of equations to determine $x'$, $x''$ and $y$:
-
-\begin{align*}
-{x''} + x'					&=  b &&\text{(a)}  \\
-{x'}^2 +  y^2                &= {l_R}^2 &&\text{(b)} \\
-{x''}^2 + y^2                &= {l_L}^2  &&\text{(c)}. \\
-\end{align*}
-
-From the first Equation (a), rearranging and squaring both terms leads to
-
-\begin{equation}
-\begin{array}{ll}  
-{x'}^2   	&=  ( b - x'')^2   \\
- 						&= b^2 - 2 b x''+ {x''}^2. \\
-\end{array}
-\label{eq:thirdstep}
-\end{equation}
-
-By subtracting (b) and (c), it can be obtained
-
-\begin{align*}
-{x'}^2 - {x'' }^2 = {l_R}^2 - {l_L}^2 \\
-\end{align*}
-
-and replacing ${x'}^2$ from Equation~\ref{eq:thirdstep} in it
-
-\begin{align*}
-b^2 - 2 b {x''}= {l_R}^2 - {l_L}^2. \\
-\end{align*}
-
-From this equation, $x''$ can be determined as
-
-\begin{align*}
-x'' = \frac{({l_R}^2 - {l_L}^2 - b^2)}{-2 \; b}.    \\
-\end{align*}
-
-From Figure~\ref{fig:dibujitos} it can also be seen that $\frac{b}{2}$ is equals to $x'' - x$.
-Hence this can be used to finally determine $x$ and $y$ values based on the threads length $l_L$ and $l_R$ and the axle length $b$.
-The $y$ value comes from Equation~\ref{eq:threadsincoordinates}.
-
-\begin{align*}
-y &= \sqrt[2]{{l_L}^2 - {x''}^2 } \\
-x &= x'' - \frac{b}{2}\\
-\end{align*}
-
-Finally, from trigonometry, it can be seen that
-
-\begin{align*}
-\sin{\gamma} = \frac{x}{l_m}\\
-\end{align*}
-
-and
-
-\begin{align*}
-\cos{\gamma} = \frac{y}{l_m}\\
-\end{align*}
-
-\noindent and as $\theta = \gamma$, we can obtain the values of $\theta$ and $l_m$:
-
-\begin{align*}
-\theta &= \arctan{ \frac{x}{y}} \\
-\l_m    &= \frac{y}{\cos{\theta}}. \\
-\end{align*}
-
-There is no loss of information and both systems are equivalent.
-
-\subsection{Design comparison}
-
-In the scheme proposed by~\cite{Endo2015} a single thread is recovered mechanically by means of a circular flat spring.
-This device works intensively when the robot is following the patient and, therefore, the spring is exposed to wear.
-Additionally with a circular flat spring, the spring tension depends on the released thread length, hence the recovery force will vary accordingly.
-Instead, the double thread tethered design implemented with an active reel spring allows a more controlled situation and depends exclusively on the software that controls the reel motor, which can be regulated.
-
-Regarding the control algorithm, authors in \cite{Endo2015} introduced two methods.
-The first of them computes the angular velocity as a  function of the difference between the measured thread length $l_m$ and the desired distance to patient  $l_D$.
-In this way, if the patient moves around the robot with a $l_m$ equals to $l_D$, the robot will not adjust its direction until the patient stops turning and restarts moving forward.
-Hence, the robot must correct its direction but the correction angle could be large, which leads the robot to deviate off the desired trajectory.
-Additionally, the second method proposed is based on dead-reckoning to estimate the patient position.
-This strategy requires very precise measurements to reduce the cumulative error, which is characteristic of this estimation procedure~\cite{DurrantWhyte1994}.
-
-%Another issue that can be presented is when $l_m$ is equals to $l_D$.
-If the patient performs a circular movement where these values are equal, the robot will not change direction at all, because $\omega_{r}$ will be also zero.
-If the patient restart the movement, $\omega_{r}$ could be very low and the proportional gain $K_p$ could not be enough to match the person movement rapidly.
-
-%Additionally, authors in~\cite{Endo2015} propose an alternative control strategy based on dead-reckoning requires very precise measurements to reduce the error-prone characteristic of this solution~\cite{DurrantWhyte1994}.
-
-%Otro problema que encontré es en el segundo método de control.
-Éste es un poco más complicado de explicar con palabras así que haré un gráfico y se los envío.
-Como adelanto, está en la ecuación 17 del paper que es cuando toman como medida de la diferencia de distancias (l_m - l_D) para el cálculo de la V_r.
-
-%Por último, un problema del algoritmo de control de los tipos (el de follow-the-leader) es que está basado en "dead-reckoning" el cual es un método impreciso cuyo error crece a lo largo del tiempo.
-Esto es, la estrategia de los tipos es ir estimando la posición del paciente en base a la posición del robot en un sistema de coordenadas universal (no en el sistema de coordenadas cuyo centro es el robot).
-Esta estimación la obtienen, entiendo yo, basado en la información de los encoders de los motores del robot (ahí es donde meten la pata) y sumándole un vector que obtienen al rotar (l_m, 0) mediante una matriz de rotación con el ángulo expresado en el sistema de referencia universal).
-Es importante remarcar que su estrategia se basa en esto: en poder estimar la posición del paciente e ir guardándola (para luego estimar T(s_t const) y para eso usan dead reckoning, es decir, a medida que pasa el tiempo (en rigor, a medida que mayor longitud recorren) peor será la estimación y peor andará el algoritmo.
-Es cierto que en las curvas no se vé eso, pero también es cierto que, en una sesión terapéutica, recorran más distancia que las mostradas en los test.
 
@@
 
-\noindent where $ \phi $ is the free parameter, $a$ is the limit length of the arc of the curve, and $x$ and $y$ are the parametric functions that determine the shape of the trajectory on the navigation plane.
+\noindent where $ \phi $ is the free parameter, and $x$ and $y$ determine the shape of the trajectory on the navigation plane.
 
@@
 
-The simulation is also useful to study the effects of the different constants on the robot movement  for the different strategies.
-The leader starts at the midpoint of the trajectory and completes a full circuit getting back to the initial position, while the robot follows its track.
-Four different configuration sets of $c_v$ and  $c_{\alpha}$ were tested for \textit{Follow the thread}, while 6 different configurations were tested for \textit{Rotate and go}.
+The simulation is also useful to study the effects of the different constants in each strategy on the movement of the robot.
+The leader starts at the midpoint of the trajectory and completes a full complete circuit getting back to the initial position, while the robot follows its track.
+Four different configuration sets of $c_v$ and  $c_{\alpha}$ were tested for Follow the thread, while 6 different configurations were tested for Rotate and go.
 
@@
 A motion capture system is used to track the movement of a human leader along a predetermined trajectory.
-The motion tracking system consists of an array of 16 OptiTrack (NaturalPoint Inc, Oregon, US) Flex 3 cameras, which measure the position of reflective markers with an accuracy of $\pm1$ cm at sampling rate of 100 Hz.
+The motion tracking system consists of an array of 16 OptiTrack (NaturalPoint Inc, Oregon, US) Flex 3 cameras, which measure the position of reflective markers with an accuracy of $\pm1$ [cm] at sampling rate of 100 Hz.
 The calibration and data collection was made using the Motive motion capture software.
 
-As shown on Figure~\ref{fig:capturesystem}, a tracking marker was placed on each side of the robot (on top of each thread reel).
+As shown in Figure~\ref{fig:capturesystem}, a tracking marker was placed on each side of the robot (on top of each thread reel).
 The human leader used his hand to grab the tip at which the two tethers were tied together.
@@
 The lemniscate of Gerono, used in the simulation, was marked on the floor, and the human leader tried to move his hand following this shape as close as possible, with stable speed.
-The shape was marked according to the shape described in Equation~\ref{eq:lemniscata}, using $a = 2$ m.
+The shape was marked according to the shape described in Equation~\ref{eq:lemniscata}, using $a = 2$ [m].
 
@@
 \caption{Hardware prototype on the motion capture system and a testing subject holding the threads.
-On top of the device, two markers are placed and an additional marker is on a glove that the user is wearing (not shown on the picture).
-The lemniscate of Gerono was marked on the floor.
+The lemniscate of gerone was marked on the floor.
 The subject follows this track on the performed experiments.}
@@
 Subfigure (e) show the distance between the robot and the patient for both strategies.
-Results metrics for the simulations are shown on Table~\ref{tab:simulationmetricsftt} for the \textit{Follow the thread} strategy, whereas metrics for \textit{Rotate and go} are shown on Table~\ref{tab:simulationmetricsrg}.
+Results metrics for the simulations are shown on Table~\ref{tab:simulationmetricsftt} for the Follow the thread strategy, whereas metrics for Rotate and go are shown on Table~\ref{tab:simulationmetricsrg}.
 
@@
 \end{tabular}
-\captionof{table}{Maximum trajectory deviation m.t.d. (m) and Total trajectory deviation t.t.d. for different \textit{Follow the thread} constants.}
+\captionof{table}{Max trajectory deviation m.t.d. [m] and Total trajectory deviation t.t.d. for different Follow the thread constants.}
 \label{tab:simulationmetricsftt}
@@
 \end{tabular}
-\captionof{table}{Maximum trajectory deviation m.t.d. (m) and total trajectory deviation t.t.d. for different \textit{Rotate and go} constants.}
+\captionof{table}{Max trajectory deviation m.t.d. (in [m]) and total trajectory deviation t.p.d. for different Rotate and go constants.}
 \label{tab:simulationmetricsrg}
@@
 %\includegraphics[width=8cm]{images/ft_a2_n1500_cv15_cr15.eps}
-%\caption{Trajectories of the leader and follower for the \textit{Follow the thread} control strategy.}
+%\caption{Trajectories of the leader and follower for the Follow the thread control strategy.}
 %\label{fig:distance_sim}
@@
 %\includegraphics[width=8cm]{images/rgv_cv20_cr20_dr10.eps}
-%\caption{Trajectories of the leader and follower for \textit{Rotate and go} control strategy.}
+%\caption{Trajectories of the leader and follower for Rotate and go control strategy.}
 %\label{fig:distance_sim}
@@
 %\includegraphics[width=8cm]{images/ft_cv15_cr10_leader_robot_speed.eps}
-%\caption{Speed profiles of the leader and follower for \textit{Follow the thread} control strategy.}
+%\caption{Speed profiles of the leader and follower for Follow the thread control strategy.}
 %\label{fig:distance_sim}
@@
 %\includegraphics[width=8cm]{images/rg_cv20_cr20_bv10_dtoff10_leader_robot_speed.eps}
-%\caption{Speed profiles of the leader and follower for \textit{Rotate and go} control strategy.}
+%\caption{Speed profiles of the leader and follower for Rotate and go control strategy.}
 %\label{fig:distance_sim}
@@
         \includegraphics[width=0.99\linewidth]{images/ft_vs_rg_dist.eps}}
-  \caption{Simulation Results: Trajectories of the leader and follower for \textit{Follow the thread} (a) and \textit{Rotate and go} (b) . 
-        Speed profiles of the leader and follower for \textit{Follow the thread} (c) and \textit{Rotate and go} (d). 
-        (e) Separation distance between robot and leader for both strategies  (m).}
+  \caption{Simulation Results: Trajectories of the leader and follower for Follow the thread (a) and Rotate and go (b) . 
+        Speed profiles of the leader and follower for Follow the thread (c) and Rotate and go (d). 
+        (e) Separation distance between robot and leader for both strategies.}
   \label{fig:simulationresults} 
@@
 Results for the real world experiment can be seen on Figure~\ref{fig:realworldresults}.
-Table~\ref{tab:cap_ft_npd_table} shows the metrics of the \textit{Follow the thread} strategy, whereas Table~\ref{tab:cap_rg_npd_table} provides the metrics for the \textit{Rotate and go} approach.
+Table~\ref{tab:cap_ft_npd_table} show the metrics of the Follow the thread strategy, whereas Table~\ref{tab:cap_rg_npd_table} provides the metrics for the Rotate and go approach.
 
@@
 %\includegraphics[width=8cm]{images/ft1cap.eps}
-%\caption{Trajectories for the \textit{Follow the thread} control strategy.}
+%\caption{Trajectories for the Follow the thread control strategy.}
 %\label{fig:distance_sim}
@@
 %\includegraphics[width=8cm]{images/ft1cap_leader_robot_speed.eps}
-%\caption{\textit{Follow the thread} speed profiles}
+%\caption{Follow the thread speed profiles}
 %\label{fig:distance_sim}
@@
 %\includegraphics[width=8cm]{images/cap_ft_vs_rg_dist.eps}
-%\caption{\textit{Rotate and go} speed profiles}
+%\caption{Rotate and go speed profiles}
 %\label{fig:distance_sim}
@@
         \includegraphics[width=0.99\linewidth]{images/cap_ft_vs_rg_dist.eps}}
-  \caption{Experimentation Results: Trajectories of the leader and follower for \textit{Follow the thread} (a) and \textit{Rotate and go} (b). 
-        Speed profiles of the leader and follower for \textit{Follow the thread} (c) and \textit{Rotate and go} (d). 
-        (e) Separation distance between robot and leader for both strategies (m).}
+  \caption{Experimentation Results: Trajectories of the leader and follower for Follow the thread (a) and Rotate and go (b). 
+        Speed profiles of the leader and follower for Follow the thread (c) and Rotate and go (d). 
+        (e) Separation distance between robot and leader for both strategies [m].}
   \label{fig:realworldresults} 
@@
 \hline
-$c_v$ & $c_{\alpha}$ & m.t.d. (m) & t.t.d. \\
+$c_v$ & $c_{\alpha}$ & m.t.d. [m] & t.t.d. \\
 \hline
@@
 \end{tabular}
-\captionof{table}{Maximum trajectory deviation m.t.d. and area under normal trajectory deviation t.t.d. in motion capture experiments using \textit{Follow the thread}.}
+\captionof{table}{Max trajectory deviation m.t.d. and area under normal trajectory deviation t.t.d. in motion capture experiments using Follow the thread.}
 \label{tab:cap_ft_npd_table}
@@
 \hline
-$c_v$ & $c_r$ & $Dt_{off}$ & m.t.d. (m) & t.t.d. \\
+$c_v$ & $c_r$ & $Dt_{off}$ & m.t.d. [m] & t.t.d. \\
 \hline
@@
 \end{tabular}
-\captionof{table}{Maximum trajectory deviation m.t.d. and area under normal trajectory deviation t.t.d. in motion capture experiments using \textit{Rotate and go}.}
+\captionof{table}{Max trajectory deviation m.t.d. and area under normal trajectory deviation t.p.d. in motion capture experiments using Rotate and go.}
 \label{tab:cap_rg_npd_table}
@@
 From the leader and robot trajectories in Figure~\ref{fig:simulationresults}(a to d), both strategies exhibits basic following behavior.
-However, \textit{Rotate and go} on Figure~\ref{fig:simulationresults}(b,d) generates a more irregular trajectory, due to the two stage movement algorithm.
-
-Regarding algorithms parameters, the simulation shows that for \textit{Follow the thread}, a low $c_{\alpha}$  means that the robot is slow to turn and makes wider turns, providing a smoother trajectory.
+However, Rotate and go on Figure~\ref{fig:simulationresults}(b,d) generates a more irregular trajectory, due to the two stage movement algorithm.
+
+Regarding algorithms parameters, the simulation shows that for Follow the thread, a low $c_{\alpha}$  means that the robot is slow to turn and makes wider turns, providing a smoother trajectory.
 On the other hand, for low $c_v$ values, the robot tends to lag behind the leader when it is going in a straight line.
 The highlighted values on Table~\ref{tab:simulationmetricsrg} show the best configuration found.
-
-For the \textit{Rotate and go} configuration, the parameter $c_r$ affects the dynamic behavior of the robot which needs to be adjusted accordingly.
-Lastly, increasing the $Dt_{off}$ from 0.05 to 0.10 made the vehicle less prone to fall behind and produces a better following profile.
-%For the \textit{Rotate and go} configuration, a high $c_r$ was important so the robot can turn quickly to point to the leader, but high values of $c_r$ and $c_v$ should be avoided, which may lead to dynamical issues with the inertia affecting the forward direction and missing the leader.
+For the Rotate and go configuration, a high $c_r$ was important so the robot can turn quickly to point to the leader, but high values of $c_r$ and $c_v$ should be avoided, which may lead to dynamical issues with the inertia affecting the forward direction and missing the leader.
 Lastly, increasing the $Dt_{off}$ from 0.05 to 0.10 made the vehicle less prone to fall behind and results in a better following profile.
 
-%From the leader and robot trajectories in Figure \ref{fig:sim_ft_cv15_cr15}, we can see that the robot using \textit{Follow the thread} control exhibits basic following behavior.
+%From the leader and robot trajectories in Figure \ref{fig:sim_ft_cv15_cr15}, we can see that the robot using Follow the thread control exhibits basic following behavior.
 Deviation occurs when the leader trajectory turns with a tight radius.
@@
 In line with the simulated results, a similar behavior was found on the experiments performed inside the Motion Capture Lab, and the robot exhibits following behavior for both control strategies as shown on Figure~\ref{fig:realworldresults}.
-Although the parameter values obtained from the simulation had to be readjusted for the real world scenario, the relative relation between them was maintained and that helped to narrow the parameter search space.
+Parameters had to be readjusted for the real world case, but the same situation occurs for both strategies.
 
@@
 This is shown on the speed profiles in Figure~\ref{fig:simulationresults}(c,d) as well as on the real experiment on Figure~\ref{fig:realworldresults}(c,d).
-The smoother movement of the \textit{Follow the thread} algorithm is an additional desired goal, since it can be perceived as less violent or unexpected.
+The smoother movement of the \textit{Follow the thread} algorithm can be desired by the followed subject, since it can be perceived as less violent or unexpected.
 %Additionally, when the vehicle stops to rotate, if the leader continues moving, when going forward, the robot might need to accelerate sharply to catch up with the leader, resulting in a sudden movement that can be disturbing for a medical patient or a doctor in the middle of a treatment.
@@
 In the demonstration, the robot design was outlined and an explanation was given on how the robot worked, how it was built and how to operate it.
-Both control strategies were explained, along with the main superficial differences between them.
+Both control strategies were expounded, along with the main superficial differences between them.
 
@@
 \hline
-How would you qualify, from 1 to 5, your overall experience with \textit{Follow the thread}? (1:Bad, 5:Excellent)  & 5.0\\
-\hline
-How safe would a patient be, from 1 to 5, being followed by the robot using the \textit{Follow the thread} strategy? (1:Very unsafe, 5:Very safe)  & 4.25\\
-\hline
-How would you qualify, from 1 to 5, your overall experience with \textit{Rotate and go}? (1:Bad, 5:Excellent)  & 3.5\\
-\hline
-How safe would a patient be, from 1 to 5, being followed by the robot using the \textit{Rotate and go} strategy? (1:Very unsafe, 5:Very safe) & 3.25\\
+How would you qualify, from 1 to 5, your overall experience with Follow the thread? (1:Bad, 5:Excellent)  & 5.0\\
+\hline
+How safe would a patient be, from 1 to 5, being followed by the robot using the Follow the thread strategy? (1:Very unsafe, 5:Very safe)  & 4.25\\
+\hline
+How would you qualify, from 1 to 5, your overall experience with Rotate and go? (1:Bad, 5:Excellent)  & 3.5\\
+\hline
+How safe would a patient be, from 1 to 5, being followed by the robot using the Rotate and go strategy? (1:Very unsafe, 5:Very safe) & 3.25\\
 \hline
@@
 
-\section{Conclusion and Future Work}
+\subsection{Conclusion and Future Work}
 \label{conclusion}
@@
 
-\subsection{Validation carrying the oxygen tank}
-
-In order to advance towards the next step, we performed an additional simulated experiment to validate whether or not the selected algorithm could be extended to  handle the oxygen tank.
-ALPI works with 5 kg oxygen tanks, which are manufactured locally.
-
-We added on the Webots simulation a 5 kg mass on top of our model, and adjusted the coefficients $c_v$ and $c_{\alpha}$ for the  \textit{Follow the thread}  control strategy.
-Results are consistent with the outcomes shown previously on Figures~\ref{fig:simulationresults} and~\ref{fig:realworldresults}.
-In Figure~\ref{fig:validationwithtank} it can be seen that the shape of the trajectory follows the patient smoothly, the speed profile also follows stringently the leader speed, and finally the distance between the patient and the robot is maintained within the safe boundaries.
-The obtained metrics are 0.2223 for m.t.d.
-and 1.1068 for t.t.d.
-which shows that the following behavior is achieved and verifies, from the simulation perspective, the initial feasibility of the control strategy considering the addition of the oxygen tank.
-
-\begin{figure} 
-    \centering
-  \subfloat[ \label{rw1a}]{%
-       \includegraphics[width=0.97\linewidth]{images/Wft_5kg.eps}}
-  \\
-    \hspace*{-1.5em}
-  \subfloat[ \label{rw1b}]{%
-        \includegraphics[width=0.49\linewidth]{images/Wft_5.0kg_leader_robot_speed.eps}}
-    \vspace*{-1.1em}
-  \subfloat[ \label{rw1c}]{%
-        \includegraphics[width=0.49\linewidth]{images/Wft_5kg_distance.eps}}
-    \hspace*{-1.5em}
-    \vspace*{1.1em}
-  \caption{Simulation results with the oxygen tank: the lemniscate trajectory is shown on (a) with the selected strategy \textit{Follow the thread}.
-(b) Speed profiles of the patient-leader and the robot-follower.
-(c) Distance between the patient and the robot while traversing the trajectory.}
-  \label{fig:validationwithtank} 
-\end{figure}
-
 \subsection{Future Work}
@@
 \begin{itemize}
+    \item Power the motor wheels with a \textit{closed-loop and more efficient controller}, to have a more precise control over their speed.
     \item Redesign the active spring control mechanism in order to hold the motor temperature in its operational range.

### 2020-10-16 alpibot.tex
Versión casi lista, integrada con los cambios de Luciano.  Solo falta la aclaración de Juan.

@@
 The variable $r$ is the radius of the reel and $ppr$ is the pulses per revolution (resolution) of the encoder.
-These equations provide the estimated values for the left $l_L$ and right $l_R$  thread longitude.
+These equations provide the estimated values for the left $l_L$ and right $l_R$, which are the length of thread released from each reel.
 The initial position of the threads is configurable.
@@
 \centering
-\subfloat[Case I]{\includegraphics[width=15cm]{images/alpibot2.jpg}\label{subject8}}
+\subfloat{\includegraphics[width=15cm]{images/alpibot2.jpg}\label{subject8}}
 \caption{Top view of the robot prototype.
@@
 \noindent where  $c_v$ and $c_{\alpha}$ are constant coefficients used for calibration whose units are expressed in $[\frac{1}{s}]$, and $l_D$ is a constant offset in $[m]$ that is used to customize the desired length of the thread where the robot does not move.
-As shown on Figure~\ref{fig:follow_params},  $l_L$ and $l_R$ are changes in thread longitudes (the length of thread that was released from each reel) in $[m]$  obtained from encoder information from Equation~\ref{eq:encoder_len}, $V_t$ is the estimated forward velocity for the target and finally $\omega_{L}$ are $\omega_{R}$ are the velocity values that are directly used to drive each wheel motor.
+As shown on Figure~\ref{fig:follow_params},  $l_L$ and $l_R$ are changes in the length of thread that was released from each reel in $[m]$  obtained from encoder information from Equation~\ref{eq:encoder_len}, $V_t$ is the estimated forward velocity for the target and finally $\omega_{L}$ are $\omega_{R}$ are the velocity values that are directly used to drive each wheel motor.
 
@@
 The constant parameter $Dm_{off}$ is similar to $l_D$, and is used to customize the length of the thread where the robot does not move at all.
+Temporary variables $Dt$ and $Dm$ are used in the algorithm to calculate the difference and the average of the changes in released thread.
 Lastly, $base_{vr}$, is another constant used to determine the minimum rotational velocity of the vehicle.
@@
 \begin{algorithmic}
-\STATE $Dt \gets l_L - l_R$
-\STATE $Dm \gets \frac{l_R+l_L}{2}$
+\STATE \textbf{Define} $Dt \gets l_L - l_R$
+\STATE \textbf{Define} $Dm \gets \frac{l_R+l_L}{2}$
 \STATE $V_r \gets c_r * (abs(Dt) - Dt_{off}) + base_{vr}$ 
@@
 We performed an additional simulated experiment to validate whether or not the selected algorithm could be extended to handle weight of the real oxygen tank.
-A common type of tank used in this type of therapies is a 5kg oxygen tank.
+A common type of tank used in this type of therapies is a 5 kg oxygen tank.
+
+\begin{figure} 
+    \centering
+  \subfloat[ \label{rw1a}]{%
+       \includegraphics[width=0.97\linewidth]{images/Wft_5kg.eps}}
+  \\
+    \hspace*{-1.5em}
+  \subfloat[ \label{rw1b}]{%
+        \includegraphics[width=0.49\linewidth]{images/Wft_5.0kg_leader_robot_speed.eps}}
+    \vspace*{-1.1em}
+  \subfloat[ \label{rw1c}]{%
+        \includegraphics[width=0.49\linewidth]{images/Wft_5kg_distance.eps}}
+    \hspace*{-1.5em}
+    \vspace*{1.1em}
+  \caption{Simulation results with the oxygen tank: the lemniscate trajectory is shown on (a) with the selected strategy \textit{Follow the thread}.
+(b) Speed profiles of the patient-leader and the robot-follower.
+(c) Distance between the patient and the robot while traversing the trajectory.}
+  \label{fig:validationwithtank} 
+\end{figure}
 
@@
 
-\begin{figure} 
-    \centering
-  \subfloat[ \label{rw1a}]{%
-       \includegraphics[width=0.97\linewidth]{images/Wft_5kg.eps}}
-  \\
-    \hspace*{-1.5em}
-  \subfloat[ \label{rw1b}]{%
-        \includegraphics[width=0.49\linewidth]{images/Wft_5.0kg_leader_robot_speed.eps}}
-    \vspace*{-1.1em}
-  \subfloat[ \label{rw1c}]{%
-        \includegraphics[width=0.49\linewidth]{images/Wft_5kg_distance.eps}}
-    \hspace*{-1.5em}
-    \vspace*{1.1em}
-  \caption{Simulation results with the oxygen tank: the lemniscate trajectory is shown on (a) with the selected strategy \textit{Follow the thread}.
-(b) Speed profiles of the patient-leader and the robot-follower.
-(c) Distance between the patient and the robot while traversing the trajectory.}
-  \label{fig:validationwithtank} 
-\end{figure}
-
 \subsection{Future Work}

### 2020-10-16 review2/response.letter.tex
Subo la versión que mandamos para la 3ra vuelta.

@@
 {\color{blue}
-We have modified on Section II.D, at the end of page 3, and clarified the terms according to the Reviewer suggestion.
-They are indeed changes in the length of the thread released from each reel.
+We have modified on Section II.D, at the end of page 3, and clarified the terms according to the Reviewer's suggestion.
+These terms are indeed changes in the length of the thread released from each reel.
 We have eliminated the use of the confusing term "longitude".
@@
 {\color{blue}
-Underway
+This is an excellent question.
+We have modified Section III.B emphasizing the two aspects that we think justify our proposal.
+As described there, the first one is related with the active spring system that we are proposing.
+We think it is more appropriate for patients that have stringent requirements related to how they move.
+One important aspect of our approach is that the stiffness and other mechanical properties can be regulated directly by software due to the active control mechanism based on motors.
+The second aspect, is that for pulmonary rehabilitation procedures, the design that we propose emphasizes the local calculation of the patient position based on the sensing information obtained from the differential-tethered mechanism.
+However, other approaches, propose dead-reckoning to do so, which may induce cumulative errors that would require frequent interruptions to re-calibrate the system.
+We have added more details regarding these issues in the above-mentioned Section and also included relevant references on Section II.C (Active Reel Spring).
 }

### 2020-10-16 alpibot.tex
Subo la versión que mandamos para la 3ra vuelta.

@@
 
-As previously mentioned, an active spring mechanism is also put in place to keep the threads tense.
+As previously mentioned, an active spring~\cite{Vanderborght2013} mechanism is also put in place to keep the threads tense.
 However, in order to extend the useful life of the reel motors, and to save battery, an algorithm to  activate and deactivate the motors was developed.
@@
 This device works intensively when the robot is following the patient and, therefore, the spring is exposed to wear.
-Additionally with a circular flat spring, the spring tension depends on the released thread length, hence the recovery force will vary accordingly.
+Additionally with a circular flat spring the properties of the spring components are predefined by their structural preconditions, and cannot be altered during spring operation~\cite{wurmthaler2013apparatus}.
+For instance, the spring tension depends on the released thread length, hence the recovery force will vary accordingly.
+If the patient found this tension to be too tight, it is not possible to alter this behavior without structurally modifying the device, or changing the spring component altogether.
 Instead, the double thread tethered design implemented with an active reel spring allows a more controlled situation and depends exclusively on the software that controls the reel motor, which can be regulated.
@@
 Additionally, the second method proposed is based on dead-reckoning to estimate the patient position.
-This strategy requires very precise measurements to reduce the cumulative error, which is characteristic of this estimation procedure~\cite{DurrantWhyte1994}.
+It is well stated~\cite{DurrantWhyte1994} that position estimation using dead-reckoning leads to an increasing error along cumulative distance with continuous changes of angular velocity.
+This presents a limitation to hold therapy sessions with patients who need to cover standard trajectory distances, requiring more frequent interruptions to perform calibration procedures.
 

### 2020-12-07 alpibot.tex
Subo la información final del paper.

@@
 Results are consistent with the outcomes shown previously on Figures~\ref{fig:simulationresults} and~\ref{fig:realworldresults}.
-In Figure~\ref{fig:validationwithtank} it can be seen that the shape of the trajectory follows the patient smoothly, the speed profile also follows stringently the leader speed, and finally the distance between the patient and the robot is maintained within the safe boundaries.
+In Figure~\ref{fig:validationwithtank} it can be seen that the shape of the trajectory follows the patient smoothly, the speed profile also follows closely the leader speed, and finally the distance between the patient and the robot is maintained within the safe boundaries.
 The obtained metrics are 0.2223 for m.t.d.

### 2022-03-21 ICRA2022/alpibot.tex
ICRA2022 draft paper.

--- 2020-12-07 alpibot.tex
@@
-%% bare_jrnl.tex
-%% V1.4b
-%% 2015/08/26
-%% by Michael Shell
-%% see http://www.michaelshell.org/
-%% for current contact information.
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
-%% journal paper.
-%%
-%% Support sites:
-%% http://www.michaelshell.org/tex/ieeetran/
-%% http://www.ctan.org/pkg/ieeetran
-%% and
-%% http://www.ieee.org/
-
-%%*************************************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or
-%% implied; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE! 
-%% User assumes all risk.
-%% In no event shall the IEEE or any contributor to this code be liable for
-%% any damages or losses, including, but not limited to, incidental,
-%% consequential, or any other damages, resulting from the use or misuse
-%% of any information contained here.
-%%
-%% All comments are the opinions of their respective authors and are not
-%% necessarily endorsed by the IEEE.
-%%
-%% This work is distributed under the LaTeX Project Public License (LPPL)
-%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
-%% distributed and modified. A copy of the LPPL, version 1.3, is included
-%% in the base LaTeX documentation of all distributions of LaTeX released
-%% 2003/12/01 or later.
-%% Retain all contribution notices and credits.
-%% ** Modified files should be clearly indicated as such, including  **
-%% ** renaming them and changing author support contact information. **
-%%*************************************************************************
-
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. The IEEE's font choices and paper sizes can   ***
-% *** trigger bugs that do not appear when using other class files.       ***                          ***
-% The testflow support page is at:
-% http://www.michaelshell.org/tex/testflow/
-
-\documentclass[journal]{IEEEtran}
-%
-% If IEEEtran.cls has not been installed into the LaTeX system files,
-% manually specify the path to it like:
-% \documentclass[journal]{../sty/IEEEtran}
-
-\usepackage[utf8]{inputenc}
-\usepackage{hyperref}
+\documentclass[conference]{IEEEtran}
+\IEEEoverridecommandlockouts
+% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
+\usepackage{cite}
+\usepackage{amsmath,amssymb,amsfonts}
+\usepackage{algorithmic}
+\usepackage{graphicx}
+\usepackage{textcomp}
+\usepackage{xcolor}
+\usepackage{tikz}
+\usepackage{subfig}
 \usepackage{algorithm2e} % For algorithms
-\usepackage{amsmath, amssymb, amsthm} % Mathematical packages
-\hypersetup{hidelinks=true}    
-\usepackage{algorithmic}
-
-%\usepackage{fancyhdr}
-%\usepackage{array}
-%\usepackage[rightcaption]{sidecap}
-%\usepackage{wrapfig}
-%\usepackage{float}
-%\usepackage[labelfont=bf]{caption} % bold text for captions
-%\usepackage[para]{threeparttable} % fancy tables, check these before you use them
-%\usepackage{url}
-%\usepackage[table,xcdraw]{xcolor}
-\usepackage{tikz}
-%\usepackage{pgfplots}
-%\pgfplotsset{width=10cm,compat=1.9}
-%\usepgfplotslibrary{external}
-%\tikzexternalize
-%\usepackage{svg}
-
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
-
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/pkg/ifpdf
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
-
-% *** CITATION PACKAGES ***
-%
-%\usepackage{cite}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of the IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off
-% such as if a citation ever needs to be enclosed in parenthesis.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 5.0 (2009-03-20) and later if using hyperref.sty.
-% The latest version can be obtained at:
-% http://www.ctan.org/pkg/cite
-% The documentation is contained in the cite.sty file itself.
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-%\usepackage[pdftex]{graphicx}
-\usepackage{subfig}
-
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../pdf/}{../jpeg/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-\else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-  % \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
-\fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation
-% can be obtained at: 
-% http://www.ctan.org/pkg/graphicx
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found at:
-% http://www.ctan.org/pkg/epslatex
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
-
-% *** MATH PACKAGES ***
-%
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{booktabs}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics.
-%
-% Note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/pkg/amsmath
-
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/pkg/algorithms
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/pkg/algorithmicx
-
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/array
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-% *** SUBFIGURE PACKAGES ***
-%\ifCLASSOPTIONcompsoc
-%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-%\else
-%  \usepackage[caption=false,font=footnotesize]{subfig}
-%\fi
-% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
-% for subfigure.sty, the latter of which is no longer maintained and is
-% incompatible with some LaTeX packages including fixltx2e. However,
-% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
-% which will override IEEEtran.cls' handling of captions and this will result
-% in non-IEEE style figure/table captions. To prevent this problem, be sure
-% and invoke subfig.sty's "caption=false" package option (available since
-% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
-% handling of captions.
-% Note that the Computer Society format requires a larger sans serif font
-% than the serif footnote size font used in traditional IEEE formatting
-% and thus the need to invoke different subfig.sty package options depending
-% on whether compsoc mode has been enabled.
-%
-% The latest version and documentation of subfig.sty can be obtained at:
-% http://www.ctan.org/pkg/subfig
-
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure.
-% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
-% corrections already built into the system in which case a warning will
-% be issued if an attempt is made to load fixltx2e.sty as it is no longer
-% needed.
-% The latest version and documentation can be found at:
-% http://www.ctan.org/pkg/fixltx2e
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/pkg/stfloats
-% Do not use the stfloats baselinefloat ability as the IEEE does not allow
-% \baselineskip to stretch. Authors submitting work to the IEEE should note
-% that the IEEE rarely uses double column equations and that authors should try
-% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
-% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
-% such ways.
-% Do not attempt to use stfloats with fixltx2e as they are incompatible.
-% Instead, use Morten Hogholm'a dblfloatfix which combines the features
-% of both fixltx2e and stfloats:
-%
-% \usepackage{dblfloatfix}
-% The latest version can be found at:
-% http://www.ctan.org/pkg/dblfloatfix
-
-%\ifCLASSOPTIONcaptionsoff
-%  \usepackage[nomarkers]{endfloat}
-% \let\MYoriglatexcaption\caption
-% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
-%\fi
-% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
-% Axel Sommerfeldt. This package may be useful when used in conjunction with 
-% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
-% submissions have lists of figures/tables at the end of the paper and that
-% figures/tables without any captions are placed on a page by themselves at
-% the end of the document. If needed, the draftcls IEEEtran class option or
-% \CLASSINPUTbaselinestretch interface can be used to increase the line
-% spacing as well. Be sure and use the nomarkers option of endfloat to
-% prevent endfloat from "marking" where the figures would have been placed
-% in the text. The two hack lines of code above are a slight modification of
-% that suggested by in the endfloat docs (section 8.4.1) to ensure that
-% the full captions always appear in the list of figures/tables - even if
-% the user used the short optional argument of \caption[]{}.
-% IEEE papers do not typically make use of \caption[]'s optional argument,
-% so this should not be an issue. A similar trick can be used to disable
-% captions of packages such as subfig.sty that lack options to turn off
-% the subcaptions:
-% For subfig.sty:
-% \let\MYorigsubfloat\subfloat
-% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
-% However, the above trick will not work if both optional arguments of
-% the \subfloat command are used. Furthermore, there needs to be a
-% description of each subfigure *somewhere* and endfloat does not add
-% subfigure captions to its list of figures. Thus, the best approach is to
-% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
-% and instead reference/explain all the subfigures within the main caption.
-% The latest version of endfloat.sty and its documentation can obtained at:
-% http://www.ctan.org/pkg/endfloat
-%
-% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
-% later in the document, say, to conditionally put the References on a 
-% page by themselves.
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-%\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/url
-% Basically, \url{my_url_here}.
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-
+\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
+    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
 \begin{document}
-%
-% paper title
-% Titles are generally capitalized except for words such as a, an, and, as,
-% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
-% not capitalized unless they are the first or last word of the title.
-% Linebreaks \\ can be used within to get better formatting as desired.
-% Do not put math or special symbols in the title.
-\title{A Control Strategy for a Tethered Follower Robot for Pulmonary Rehabilitation}
-%
-%
-% author names and IEEE memberships
-% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
-% a structure at a ~ so this keeps an author's name from being broken across
-% two lines.
-% use \thanks{} to gain access to the first footnote area
-% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
-% was not built to handle multiple paragraphs
-%
-
-\author{Bianchi Luciano, Buniak Esteban Alejandro, Ramele~Rodrigo,~\IEEEmembership{Member,~IEEE,}
-        and Santos~Juan~Miguel% <-this % stops a space
+
+\title{Conference Paper Title*\\
+{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
+should not be used}
+\thanks{Identify applicable funding agency here. If none, delete this.}
+}
+
+\author{Luciano Bianchi, Esteban Alejandro Buniak~\IEEEmembership{Member,~IEEE,}, Rodrigo~Ramele,~\IEEEmembership{Member,~IEEE,}
+        and Juan~Miguel~Santos% <-this % stops a space
 \thanks{E. A. Buniak, R. Ramele  and J.M.Santos are with the CIC Laboratory at the Department
@@
 
-% note the % following the last \IEEEmembership and also \thanks - 
-% these prevent an unwanted space from occurring between the last author name
-% and the end of the author line. i.e., if you had this:
-% 
-% \author{....lastname \thanks{...} \thanks{...} }
-%                     ^------------^------------^----Do not want these spaces!
-%
-% a space would be appended to the last name and could cause every name on that
-% line to be shifted left slightly. This is one of those "LaTeX things". For
-% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
-% "AB" then you have to do: "\textbf{A}\textbf{B}"
-% \thanks is no different in this regard, so shield the last } of each \thanks
-% that ends a line with a % and do not let a space in before the next \thanks.
-% Spaces after \IEEEmembership other than the last one are OK (and needed) as
-% you are supposed to have spaces between the names. For what it is worth,
-% this is a minor point as most people would not even notice if the said evil
-% space somehow managed to creep in.
-
-% The paper headers
-\markboth{IEEE Transactions on Medical Robotics and Bionics}%
-{Bianchi \MakeLowercase{\textit{et al.}} A Control Strategy for a Tethered Follower Robot
-for Pulmonary Rehabilitation}
-% The only time the second header will appear is for the odd numbered pages
-% after the title page when using the twoside option.
-% 
-% *** Note that you probably will NOT want to include the author's ***
-% *** name in the headers of peer review papers.                   ***
-% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
-% you desire.
-
-% If you want to put a publisher's ID mark on the page you can do it like
-% this:
-%\IEEEpubid{0000--0000/00\$00.00~\copyright~2015 IEEE}
-% Remember, if you use this you must call \IEEEpubidadjcol in the second
-% column for its text to clear the IEEEpubid mark.
-
-% use for special paper notices
-%\IEEEspecialpapernotice{(Invited Paper)}
-
-% make the title area
+\author{\IEEEauthorblockN{1\textsuperscript{st} Luciano Bianchi}
+\IEEEauthorblockA{\textit{CIC Laboratory at the Department
+of  Computer Engineering} \\
+\textit{Instituto Tecnológico de Buenos Aires (ITBA)}\\
+Buenos Aires, Argentina \\
+lbianchi@itba.edu.ar}
+\and
+\IEEEauthorblockN{2\textsuperscript{nd} Esteban Alejandro Buniak~\IEEEmembership{Member,~IEEE,}}
+\IEEEauthorblockA{\textit{CIC Laboratory at the Department
+of  Computer Engineering} \\
+\textit{Instituto Tecnológico de Buenos Aires (ITBA)}\\
+Buenos Aires, Argentina \\
+lbianchi@itba.edu.ar}
+\and
+\IEEEauthorblockN{3\textsuperscript{rd} Rodrigo~Ramele,~\IEEEmembership{Member,~IEEE,}}
+\IEEEauthorblockA{\textit{CIC Laboratory at the Department
+of  Computer Engineering} \\
+\textit{Instituto Tecnológico de Buenos Aires (ITBA)}\\
+Buenos Aires, Argentina \\
+lbianchi@itba.edu.ar}
+\and
+\IEEEauthorblockN{4\textsuperscript{th} Juan~Miguel~Santos}
+\IEEEauthorblockA{\textit{CIC Laboratory at the Department
+of  Computer Engineering} \\
+\textit{Instituto Tecnológico de Buenos Aires (ITBA)}\\
+Buenos Aires, Argentina \\
+lbianchi@itba.edu.ar}
+}
+
 \maketitle
 
-% As a general rule, do not put math, special symbols or citations
-% in the abstract or keywords.
 \begin{abstract}
@@
 
-%Results on a simulated and on a real prototype, and external validation from professional healthcare givers, confirm the feasibility of the proposed solution.
-%While performing Pulmonary Rehabilitation patients perform different exercises and walking activities and may require oxygen assistance.
-
-%Pulmonary Rehabilitation procedures consist of controlled walking activities and physical exercises that patients perform under the supervision of a physical therapist.
-
-% Note that keywords are not normally used for peerreview papers.
 \begin{IEEEkeywords}
-COPD, PR, SAR, IoRT, tethered
+component, formatting, style, styling, insert
 \end{IEEEkeywords}
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
-\IEEEpeerreviewmaketitle
 
@@
 
-% Can use something like this to put references on a page
-% by themselves when using endfloat and the captionsoff option.
-\ifCLASSOPTIONcaptionsoff
-  \newpage
-\fi
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-% references section
-
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
 \bibliographystyle{IEEEtran}
@@
 
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
-%\begin{thebibliography}{1}
-
-%\bibitem{IEEEhowto:kopka}
-%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
-%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
-
-%\end{thebibliography}
-
-% biography section
-% 
-% If you have an EPS/PDF photo (graphicx package needed) extra braces are
-% needed around the contents of the optional argument to biography to prevent
-% the LaTeX parser from getting confused when it sees the complicated
-% \includegraphics command within an optional argument. (You could create
-% your own custom macro containing the \includegraphics command to make things
-% simpler here.)\begin{IEEEbiographynophoto}
-%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
-% or if you just want to reserve a space for a photo:
-
-%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{ramele}}]{Rodrigo Ramele}
-%currently works at the Centro de Inteligencia Computacional (CIC), ITBA University, Buenos Aires, Argentina.
-Rodrigo does research in Brain Computer Interfaces, Artificial Intelligence, Signal Processing and Robotics.
-%\end{IEEEbiography}
-%
-%% if you will not have a photo at all:
-%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{santos}}]{Juan Miguel Santos}
-%is the Director of the Centro de Inteligencia Computacional (CiC) of ITBA University in Buenos Aires, Argentina, where he works in three different areas: robotics development, reinforcement learning and brain-computer interfaces.
-In the first two lines he works since 1993 developing and constructing robots, national and international robotics conferences and international peer-review journals.
-Currently developing a new development line based on SLAM to autonomous robots.
-%\end{IEEEbiography}
-
-% insert where needed to balance the two columns on the last page with
-% biographies
-%\newpage
-
-% You can push biographies down or up by placing
-% a \vfill before or after them. The appropriate
-% use of \vfill depends on what kind of text is
-% on the last page and whether or not the columns
-% are being equalized.
-
-%\vfill
-
-% Can be used to pull up biographies so that the bottom of the last one
-% is flush with the other column.
-%\enlargethispage{-5in}
-
-% that's all folks
 \end{document}

### 2022-03-21 alpibot.tex
ICRA2022 draft paper.

@@
 
-\author{Bianchi Luciano, Buniak Esteban Alejandro, Ramele~Rodrigo,~\IEEEmembership{Member,~IEEE,}
-        and Santos~Juan~Miguel% <-this % stops a space
+\author{Luciano Bianchi, Esteban Alejandro Buniak~\IEEEmembership{Member,~IEEE,}, Rodrigo~Ramele,~\IEEEmembership{Member,~IEEE,}
+        and Juan~Miguel~Santos% <-this % stops a space
 \thanks{E. A. Buniak, R. Ramele  and J.M.Santos are with the CIC Laboratory at the Department
